### 2010-06-16 pkg/paper/Draft/draft-Z-W.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@11 b679b2f7-6539-45d0-a605-5ddcb976cde9

\documentclass[titlepage,a4paper, fleqn, flalign]{article} \usepackage[pdftex]{graphicx}
\usepackage[multiple]{footmisc} \pagestyle{myheadings}
\usepackage[latin1]{inputenc}
\usepackage{fancyhdr}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{ltxtable}
\usepackage{tabularx}
\usepackage{rotating}
\usepackage{array}
\usepackage{amssymb}
\usepackage{pstricks}
\usepackage{qtree}
\usepackage{lscape}
\usepackage{setspace}
\usepackage[numbers,round]{natbib}
\usepackage{harvard}

\begin{document}

%\title{Draft  \\[2cm]
%\textbf{Structural Breaks in Inflation Rates}}
%
%\author{ \scriptsize by \\
%Mag. Thomas WINDBERGER\\
%Prof. Dr. Achim ZEILEIS \\
%\scriptsize Doctoral Advisors: \\
%\scriptsize Univ.-Prof.Dr. Jesus Crespo-Cuaresma \\
%\scriptsize A.Univ.-Prof.Dr. Janette Walde \\ [1cm]}
%
%
%
%\maketitle

% anderthalb Zeilen Abstand inne

\setstretch{1,5}

\section{Abstract}

The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)  on changing the volatility of the inflation rates in its Member States. 
%We concentrate on the volatility of the inflation rate rather than its mean since \cite{fried} conjectured that the harmful effect of inflation on growth is driven by inflation volatility.

\newpage
\section{Introduction}

The ECB defines price stability ``as a year-on-year increase in the Harmonised Index of Consumer Prices (HICP) for the euro area of below 2 \%'' \cite{castel}. \cite{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22). Thus to achieve a low inflation rate and low variability of inflation must be a key issue for the central bank. \\
To further justify the use of inflation volatility, we refer to \cite{rother}: ``Among the harmful effects of inflation, the negative consequences of inflation volatility are of particular concern. High variability of inflation over time makes expectations over the future price level more uncertain. In a world with nominal contracts, this induces risk premia for long-term arrangements, raises costs for hedging against inflation risks and leads to unanticipated redistribution of wealth. Thus, inflation volatility can impede growth even if inflation on average remains restrained.'' \\
The question of interest centers around the way in which a countries decision to join the EMU changed its inflation volatility. There are a number of reasons why the volatility of  one of these nations should change indeed. Given that the country experienced quite volatile inflation rates, its efforts to meet the convergence criterias are likely to lead to a change in the volatility of their respective inflation rates, as has indeed been the case for a number of EMU countries, like Spain, Portugal, Italy and Greece.

\section{Literature Overview}

%Sinn: Zusammenfassung der gegebenen Literatur. Darstellen, was in der Richtung publiziert wurde und herausarbeiten, warum mein Ansatz hier jetzt neu sein soll und warum die Frage noch nicht beantwortet ist.

Theory unfortunately is quite unsure about whether or not the creation of a monetary union between two or more states is likely to reduce/increase the variability or even the level of the inflation rate.  \cite{cooper} points out, that "a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a "central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation. It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
An interesting approach to this question is taken by \cite{holte}. He creates a two country model for monetary policy analysis along the line of two models by \cite{mc1} and \cite{mc2} as well as \cite{gali}. The inflation is modeled by a hybrid Phillips curve (NKPC) specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union. What he finds out via simulations of these different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union. But monetary policy in a monetary union does not explicitly stabilize the output gap and inflation rate in case of national economic shocks. The effects of joining on inflation rate variability depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations. Due to the fact that not all of these parameters are known and that their interaction as well has to be guessed, theory has some troubles answering the question of this paper. \\
\cite{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation, its degree of persistence and both types of uncertainty. They use a time-varying model with a GARCH specification for the unconditional volatility of inflation and find some instability in the conditional volatility. \\
In a paper examining structural convergence of the inflation rates in EU counries, \cite{sarno} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar. They find that convergence in time of inflation dynamics was only partly observeable. The same is also found by \cite{palomba}, they too find little evidence of similarity of short-run inflation dynamics. \\
In a paper studying core inflation and using an aggregated Euro Area inflation rate, \cite{morana} finds three regimes (roughly 1980-1984, 1984-1993 and 1993-2000) governing the core inflation rate.  \\
Inquiering into the convergence properties of inflation rates among countries of the EMU, \cite{busetti} find that from 1980-1997 there was convergence of inflation rates, but afterwards there is some diverging behavior. \\
%A otherwise very interesting paper by \cite{maheu} can not be used due to its defintion of a structural break as "`an unpredictable event in which the relationship among the variables in a model change"', which differs from the definition of a sructural break used in this paper.

\section{Data}

Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot ln(HICP_{t}/HICP_{t-1})$. Countries included are Austria,	Belgium,	Czech Republic,	Denmark,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,		Spain,	Sweden,		United Kingdom and the	Euro area.  The data are obtained from the OECD Statistics. \\
The countries can be divided into three different groups: the EMU countries (Portugal, Spain, Italy, Greece, France, Belgium, Netherlands, Germany, Austria, Ireland, Slovenia, Luxembourg and Finland), EU members without ERM~II (Exchange Rate Mechanism): United Kingdom, Sweden, Poland, Czech Republic, Hungary. Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.

% slovenien statt slovakei?
% Deutschland: CPI statt HVPI
% vielleicht ein paar Länder aus dem Sample entfernen: 
% slovakei, tschechien, ungarn, polen
% Nachschauen, wann in den jeweiligen Ländern die Entscheidung getroffen wurde, den Euro einzuführen. --> viel zu aufwendig

%/*  1 = austria = 240
%    2 = belgien = 228 (erst ab 91)
%    3 = czechrepublic = 180 (erst ab 95)
%    4 = denmark = 240
%    5 = finland = 240
%    6 = france = 240
%    7 = germany = 180
%    8 = greece = 180
%    9 = hungary = 180
%    10 = iceland = 180
%    11 = ireland = 180
%    12 = italy = 240
%    13 = luxembourg = 180
%    14 = netherlands = 240
%    15 = poland = 168
%    16 = portugal = 240
%    17 = slovakei = 180
%    18 = spanien = 216
%    19 = schweden = 240
%    20 = uk = 239 (fehlt 12.2009)
%    21 = euroarea = 168
%*/
%
%/* CPI Daten von:
%    22 = Japan = 240
%    23 = Norway = 240
%    24 = Switzerland = 240
%    25 = USA = 240
%/*

\newpage
\section{Methodology}

\subsection{Using a generalized logistic distribution (GL)}

As allready stated, we now apply the general framework, as developed in \cite{z08} to a more specific model, in this case by means of a GL distribution.Prior to that, we would like to give a short justification for our using of the GL distribution. Regarding the data at hand, it was not possible to use the already existing method developed in \cite{z07}, since allmost all inflation rates, with the noteable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties. Therefore, a somewhat more flexibel distribution had to be used. We needed a distribution exhibiting rather strong kurtosis and the property to be both left and right skewed. To this end we use a generalization of the logistic distribution as defined in \cite{johnson}. Its probability density is given by:

\begin{eqnarray}
& & f(\pi | \theta, \sigma, \delta) = \frac{\frac{\delta}{\sigma}*\exp^{-\frac{\pi_i-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi_i-\theta}{\sigma}})^{(\delta+1)}}
\end{eqnarray}

with location ($\theta$), scale ($\sigma$) and shape ($\delta$). For b=1 the distribution simplifies to the logistic distribution, for b$<$1 it is skewed to the left and for b$>$1 it is skewed to the right. The moments are given by:

\begin{eqnarray}
& & E(\pi_i) = \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
& & Var(\pi_i)  = \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
& & Skew(\pi_i) =  \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
\end{eqnarray}

where $\gamma()$ is the digamma function and its derivatives. \\
The log-likelihood is given by:

\begin{eqnarray}
& & l(\delta,\theta,\sigma, \pi) = ln(\delta) - ln(\sigma) - \frac{1}{\sigma} (\pi-\theta) - (\delta+1) ln(1+\exp^{-\frac{\pi-\theta}{\sigma}})
\end{eqnarray}

The resulting score function ($\psi()$) for the parameters (the derivatives of the log-likelihood) are given by:

\begin{eqnarray}
& & \psi(\pi_i,\delta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta} = \frac{1}{\delta} - ln(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \\
& & \psi(\pi_i,\theta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta} = \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})} 
\end{eqnarray}

\begin{equation}
\begin{split}
&& \psi(\pi_i,\sigma) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma} = -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) * \\
&& \quad \frac{\frac{1}{\sigma^2}(\pi-\theta)\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
\end{split}
\end{equation}

An enhancement to the allready existing R package "`strucchange"', which currently does not support a generalized logistic distribution (GL) is provided with this paper. The asymptotic testing theory still holds for this generalization. [Beweis] \\
The first part of the results we are going to present here, i.e. the tests and the graphical illustration of the empirical fluctuation process can be found in \cite{z07}. the second part of the results - the dating procedure and the illustration of the denisties fitted for the subsamples (divided by the breaks) - ca be found in \cite{z10}.

\subsection{Tests}

\subsubsection{Cramer-von Mises statistic - Nyblom Hansen Test}

We use the Cramér von Mises type test as given in \cite{z07}. The test statistic is given by:

\begin{eqnarray}
& & \frac{1}{n} \sum_{i=1}^n \Vert efp(\frac{i}{n}) \Vert_2^2
\end{eqnarray}

i.e. first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically as developed in 
%[wann war die Erstpublikation der graphischen Methode?].

\subsubsection{Kolmogorov-Smirnov Test}

As a goodness of fit test, we use the Kolmogorov Smirnov Test for as given in \cite{massey}.

\subsection{Densities}

Another thing we wish to look at is the change in the moments of the distribution after a break occured and whether or not the subsample fit of the GL-distribution is prefereable to another. Allthough our interest centers on the variance of the respective inflation rate, skewness should not alltogether be ignored. If we  observe - for example - a change in skewness from positive to negative from one regime to another, we could conclude that now we will observe higher values in the inflation rate since it shifted.

\newpage
\subsection{Expected Results}

To give an idea what we would expect, we give a short overview over the history of the EMU. Following the Delors Plan with his 3 stages, we have: stage I (1990-1994), stage II (1995-1998) and stage III (1999-2002) which ended with the introduction of the Euro as legal tender. If the EMU had a significant effect upon inflation rate volatility we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular "`Teuro"' argument (i.e. the claim that the Euro lead to a significant increase in prices following the year of its introduction). [Teuro besser formulieren, gibt es dafür einen internationalen Term?]

%
%
%Geschichte für Interpretation 
%- verwende Daten seit 1990
%- 1. Stufe (ab 1990) – Vorbereitung durch Liberalisierung Geld- und Kapitalverkehr
%- 2. Stufe (ab 1994) – Konvergenz:
%- 3. Stufe (ab 1998) – Übergang zum Euro:
%- Euro Währung Buchgeld seit 1.1.1999
%- Euro Währung Bargeld seit 1.1.2002?

\section{Results-very preliminary}

\begin{longtable}{|l|p{4cm}|p{3cm}|}
\hline
{\bf Country} & {\bf glogist Breaks (dating unsure)} & {\bf fxregimes Breaks}   \\
\hline
Austria & 1999,2008 & no break \\
\hline
Belgium & 2000 & December 2000 \\
\hline
Czech Republic & 1998,2002 & no break \\
\hline
Denmark & 2001 & no break \\
\hline
Finland & 2000 & no break \\
\hline
France & 2005 & no break \\
\hline
Germany & 2000, 2005 & no break \\
\hline
Greece & 2005 & no break \\
\hline
Hungary & 1998,2000,2003 (not very clear) & April 1998 \\
\hline
Ireland & 2005 & no break \\
\hline
Italy & 1996,2001 & April 1996, December 2000 \\
\hline
Luxembourg & 1999 & December 1998 \\
\hline
Netherlands & no break & no break\\
\hline
Poland & 2001 & April 2001 \\
\hline
Portugal & 1995 & March 1994 \\
\hline
Slovenia & 2003,2005 & no break \\
\hline
Spain & 2000 & December 2000 \\
\hline
Sweden & 1994 & no break \\
\hline
United Kingdom & 1994 & no break 
\hline
\end{longtable}

We find some similarity of a group of countries, Group I (Belgium, Spain and Italy), Group II (France, Greece and Ireland), Group III (Hungary and the Czech Republic) and Group IV (UK and Sweden). The other countries seem to be a group of their own.

\subsection{Austria}

The Double max test for Austria is insignificant, but the Cramér-von Mises test and the Supremum of LM statistcs test are both significant. This can be interpreted as a sign, that there was some change in the series as a whole, but that we are unsure wheter this happend with respect to location, scale or shape.

\begin{figure}[h!tp]
\includegraphics[width=80mm]{aut-meanL2BB.png}
\caption{Cramér-von Mises test for Austria}
\end{figure}

Regarding the figure, we have two clear breaks, one in 1999 and another in 2008. [dating very unsure] The fxregimes procedure gives no breakdate. To give at least some intuition about the changes in variance [actually this is just scale, but can be adapted quickly] we forced 5 breaks into the Austrian inflation rates.

\begin{figure}[h!tp]
\includegraphics[width=80mm]{aut-morebreak.png}
\caption{Variance Estimation in Austrian Inflation Rates with 5 Breaks}
\end{figure}

Only with some phantasy we can se some slight V shape (a decline in inflation volatility up to 2000, then an increase).

\subsection{Belgium}

Regarding the belgian inflation rates, all tests clearly indicate a break.

\begin{figure}[h!tp]
\includegraphics[width=80mm]{bel-meanL2BB.png}
\caption{Cramér-von Mises test for Belgium}
\end{figure}

The fxregimes procedure indicates a breakpoint in December 2000.

%The impact of skewness: negative skew (rechtssteil), ie. higher values of y are more probable; positive skew (linkssteil): lower values of y are more probable
%so if it is the case like 6-2-1 we can say that the distribution gets less of a positive skew meaning a lower inflation on average (abklären, sobald das mit den Momenten hinhaut)
%

\subsection{Italy}

In the italian case, there is clear evidence for two breaks.

\begin{figure}[h!tp]
\includegraphics[width=80mm]{ita-meanL2BB.png}
\caption{Cramér-von Mises test for Italy}
\end{figure}

If we compare the density of a whole and the denisties of the subsets of the data, we see a clear picture:

\begin{figure}[h!tp]
\includegraphics[width=80mm]{ita-split.png}
\caption{Parameters and Densities for subsamples}
\end{figure}

As we can see regarding the moments of the subsamples, we see a clear increase in variance (0.04, 0.01, 0.32), some smaller increase in the mean (0.41, 0.16, 0.18)  and a interesting increas in decrease in skewness (0.96, 0.71, $-$0.26), which means that lower values of the inflation rate are quickly becoming less probable.

%\section{Ideas, Suggestions, Questions}
%
%Is there a method to get the dates in $"`plot(aut_gefp, functional=meanL2BB)"'$ (ie. return the date when we have the maximum value (or the two or more maximum values))? \\
%How to get more focus on variance ? \\
%How to force more than 5 breaks, so we end up with something as a "`smooth"' variance estimation. Good for graphical interpretation in the cases where there is no clear cut break. \\
%Use a GARCH model with a GL density, if considered useful \\
%If we find a break, how can we be sure that its the EMU whos responsible? The case if we find nothing is simpler, we can state that EMU had no effect (because otherwise we would see something, and the argument that EMU would be counteracted by something else is very very difficult to win) \\
%Goodness of fit test could produce a graph showing the fit in cdf and ecdf and some stars showing significance or a message.... 
%
%
%
%

\setstretch{1,3}

\newpage
\section{Literature}

\renewcommand*{\refname}{}
\bibliography{papers}
\begin{thebibliography}{99}
\bibliographystyle{apsr}

\end{document}

### 2010-06-16 pkg/R/breakpoints.glogisfit.R
documented better starting values, hence increased version
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@13 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@ -8,8 +8,8 @@ breakpoints.glogisfit <- function(obj, h = 0.15, breaks = NULL, ic = "LWZ", hpc 
 {
   stopifnot(require("fxregime"))
   dat <- data.frame(x = as.vector(obj$x))
-  myfit <- function(formula, data, ...) glogisfit.default(data$x, fixed = obj$fixed, hessian = FALSE, ...)
-  rval <- fxregime:::gbreakpoints(x ~ 1, data = dat, order.by = time(obj$x), fit = myfit,
+  glogisfit0 <- function(formula, data, ...) glogisfit.default(data$x, fixed = obj$fixed, hessian = FALSE, ...)
+  rval <- fxregime:::gbreakpoints(x ~ 1, data = dat, order.by = time(obj$x), fit = glogisfit0,
     h = h, breaks = breaks, ic = ic, hpc = hpc)
   rval$null <- obj
   class(rval) <- c("breakpoints.glogisfit", class(rval))

### 2010-06-16 pkg/R/glogisfit.R

@@ -74,15 +74,28 @@ glogisfit.default <- function(x, weights = NULL,
   }
 
   ## choose starting values
-  if(is.null(start)) start <- c(mean(x), log(sd(x)), 0)[is.na(fixed)]
+  if(is.null(start)) start <- c(0, 0, 0)[is.na(fixed)]
   
   ## call optim
-  opt <- if(method == "L-BFGS-B") {
-    optim(par = start, fn = llfun, gr = gradfun, hessian = TRUE,
-      lower = c(-Inf, 0, 0), upper = rep(Inf, 3), method = method, control = list(...))
-  } else {
-    optim(par = start, fn = llfun, gr = gradfun, hessian = TRUE,
-      method = method, control = list(...))  
+  opt <- optim(par = start, fn = llfun, gr = gradfun, hessian = hessian, method = method, control = list(...))
+
+  ## if first attempt fails, try to get better starting values
+  ## (employing simple Nelder-Mead on original parameter scale)
+  if(opt$convergence > 0 | sum(abs(gradfun(opt$par))) > 1) {
+    llfun2 <- function(par) {
+      p1 <- if(is.na(fixed[1])) par[1] else fixed[1]
+      p2 <- if(is.na(fixed[2])) par[is.na(fixed[1]) + 1] else exp(fixed[2])
+      p3 <- if(is.na(fixed[3])) par[sum(is.na(fixed[1:2])) + 1] else exp(fixed[3])
+      if(p2 < 0) return(Inf)
+      if(p3 < 0) return(Inf)
+      -sum(w * dglogis(x, location = par[1], scale = p2, shape = p3, log = TRUE))
+    }
+    start2 <- start
+    start2[which(is.na(fixed)) %in% 2:3] <- exp(start2[which(is.na(fixed)) %in% 2:3])
+    opt2 <- optim(par = start2, fn = llfun2)
+    start <- opt2$par
+    start[2:3] <- log(start[2:3])
+    opt <- optim(par = start, fn = llfun, gr = gradfun, hessian = hessian, method = method, control = list(...))
   }
   
   ## post-process optim result

### 2010-06-17 pkg/paper/Draft/draft-Z-W.tex
some preliminary information for GL use in literature, more to come
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@14 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \subsection{Using a generalized logistic distribution (GL)}
+
+% something more about the reason to use GL
+
+% Book Johnson 95 on logistic distribution
+
+- income distributions, otherwise often used as growth models, or rainfall models
+- definition of the distribution Abberger uses here is a generalization of 23.7 in Johnson 95 (shape ist halt hier zustätzlich noch dabei)
+- longer tails than normal distribution, nonsymmetric (which is a clear net gain)
+- more peaked in the center (net gain)
+- the things above apply to the logistic distribution, not implicitely the GL
+
+% searchterm for these ``generalized logistic distribution''
+% found via Google Scholars
+- in 2005-kjm-olp-gl2 it is shown, that the distribution used here is; cdf is given by: $1-\frac{\exp{-bx}}{(1+\exp{-x})^b}$ (given in this paper)
+- 2005-spl-won-parametersarmitgld: here the GL is used (they find for Autoregressions that mostly data simply is not normal or student, so they use the GL-distribution (which is only somewhat slightly differently specified)
+- 2000-ael-har-stockreturns: with almost the same reasoning as above used to model stock exchange data (not exactly gl as here but can be simplified to GL)
+- 1999-cis-won-asyminnogl: almost the same as the 2005 paper (won.. is also the author :)
+- 2007-tejof-kon-vatriskgstock: same use
+- 2001-joae-wa-garchgl: similar (not exactly using gl but very similar)
+
+% Econ Papers
+- often in financial return data (which has high peakes)
+- in gene expression data (since gl hast longer tails)
+
+% eu database
+% ideas
+% citeceer
+% econometric society
+% questia
+% ecb
+% jstor
+% springerlink
+% ebsco
+% wiley
+% sciencedirect
+% informaworld
+
+% Abberger Publikation: Abberger, K., Heiler, S.
+	Simultaneous estimation of parameters for a generalized logistic distribution and application to time series models 	2000 	84 	041-049 	A Allgemeines Statistisches Archiv/Journal of the German Statistical Society
+
+- in 1987-eco-bat-inflationexpectation: the expected inflation is modelled as coming from a logistic distribution
 
@@
 
+Some History for easier interpretation:
+
+% this one from history-emu.pdf
+
+eleven countries had met the Maastricht Criteria by 1997, allowing the
+European monetary union to be launched in 1999. Greece was further behind in
+meeting the criteria and only joined the monetary union in 2001.
+First, entry into the euro was a much bigger structural shock for "peripheral"
+member countries such as Greece, Ireland, Portugal, and Spain than for "core"
+member countries like Germany and France. In particular, while all euro member
+countries have enjoyed lower real interest rates under monetary union, the relative
+decline was much greater for the countries on the periphery.
+Second, membership in a monetary union can amplify the asymmetric impact
+of certain shocks. A common nominal interest rate implies that persistent differences
+in national inflation rates translate into differences in real interest rates across member countries: countries with relatively higher medium-term inflation
+enjoy lower real interest rates than those with below-average inflation-stimulating
+demand, credit growth, and housing markets in the former group
+The Netherlands has lived though a
+boom-bust cycle since the euro started, with a credit boom and high inflation in
+1999-2001 followed by a contraction in economic activity that has reversed some of
+the country's real appreciation against partner countries.
+Of course, the destabilizing features of European monetary union for the
+member countries do not mean that currency union has led to a net increase in
+macroeconomic instability. The European Central Bank has been successful in
+anchoring medium-term inflation expectations across the euro area at an annual
+rate around 2 percent. At least for some countries, the attainment of such stability
+outside the euro framework may have been a more costly process.
+
+\section{Results}
+
+% in money-growth regime switchin:
+Germany: early 90ies: low inflation/ mid-end 90ies: high / and afterwards low
+UK: low regime
+
+just as a help
+
+ wenn ma was siat dass as co-movement git: \cite{vandi} "`euro area countries move from essentially idiosyncratic inflation to co-movement in the mid-1980s"' (p.
+2); "`price shocks in one country not transmitted to another"'; "`correlations remain stable until 1999 but then jumps sharpely"'; (2009-mwp-dijk-internationalinfl)
+
+wenn was passiert: zitieren von Carporale: in their estimate \cite{cap} find that each country has at least one of the parameters of their variance estimates significantly different from zero, which is a clear indication of change in volatility as well.
+
+falls Volatilität erhöht wird, verweise auf die Debatte in: $2004-wp-minford-ukjoiningemu$, die von einer Erhöhung der britischen Inflationsrate sprechen, sollte UK der EMU beitreten.
+
+falls kleine Länder auffälliges Verhalten zeigen: $2008-wp-arnold-inflationuncertainty$, "`This suggests that within EMU, inflation uncertainty may increase in countries that have a smaller influence on ECB policy."' Ihre Begründung: "`For economic agents in small regions which are hit by positive inflation shocks, it will be uncertain how and when the increase
+in inflation will be reversed.
+This results from imperfect information on the macroeconomic adjustment mechanism combined with the absence of a central bank dedicated to stabilizing their regional inflation rate.
+This does not hold true for the larger regions in the union, who have a largerweight in aggregate inflation and may therefore count on the price stability objective of the union’s central bank."'
+vergleichen mit den Ergebnissen auf S.343 (country size and inflation uncertainty).
+
+berücksichtige die DM Zone (Deutschland, Österreich, Luxemburg, Belgien und Holland), eventuell ähnlich
+
+Caporale (ausgedruckt im Diss-Aktuell Ordner) finds that in the post-Euro area there are still significant inflation differentials across countries.
+
+Wenns grosse Unterschiede zwischen den EMU Ländern gibt: ba Caporale stoht: "`the ECb itself admits that monetary policy can only affect price developements in the Euro area as a whole and cannot influence inflation differentials across regions (The Monetary Policy of the ECB, 2004)."' Und dort findet sich auch: "`some short-term volatility is inevitable given the fact that monetary policy can only affect prices wiht long and uncertain lags, hence the focus of the ECB on medium-term price stabilisation."'
+
+Um den Anstieg in der Volatilität zu verteidigen: Caporale finds that steady state uncertainty has increased in eight out of twelve countries (p.966) (ie.
+Germany, France, Italy, Spain, Ireland, Finland, Luxembourg and Austria).
+
+Zum Abgleich für Break dates: Caporale Table 5.
+
+Wenn Ergebnisse da sind: abgleichen bei Caporale mit seiner Fig. 2, Fig. 3
+
+falls Österreich und Spanien komisch: Wilfing findet, dass Schilling und Peseta nicht zu ihrer fixen Parität konvergiert sind
+
+falls man grosse Unterschiede in der Volatilität sehen würde: $2008-wp-campolmi-laborinflvol$ point out that this can be explained by country specific labor market institutions.
+
+ morana paper: S.10 untersucht convergence, je nachdem, was sich findet herausfinden, was wie ähnlich ist
+
+ bei grossen Unterschieden: Busetti S.6: 3 Clustergruppen: Ger, Fra, Bel, Aut, Fin; Spa, Gre, Por, Ireland; Ned und It
+
 %\section{Ideas, Suggestions, Questions}

### 2010-06-18 pkg/paper/Draft/draft-Z-W.tex
some more text
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@15 b679b2f7-6539-45d0-a605-5ddcb976cde9
some  more changes on economic history
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@16 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2010-06-16 pkg/paper/Draft/draft-Z-W.tex
@@
 
+% something more about the reason to use GL
+
+% Book Johnson 95 on logistic distribution
+
+- income distributions, otherwise often used as growth models, or rainfall models
+- definition of the distribution Abberger uses here is a generalization of 23.7 in Johnson 95 (shape ist halt hier zustätzlich noch dabei)
+- longer tails than normal distribution, nonsymmetric (which is a clear net gain)
+- more peaked in the center (net gain)
+- the things above apply to the logistic distribution, not implicitely the GL
+
+% searchterm for these ``generalized logistic distribution''
+% found via Google Scholars
+- in 2005-kjm-olp-gl2 it is shown, that the distribution used here is; cdf is given by: $1-\frac{\exp{-bx}}{(1+\exp{-x})^b}$ (given in this paper)
+- 2005-spl-won-parametersarmitgld: here the GL is used (they find for Autoregressions that mostly data simply is not normal or student, so they use the GL-distribution (which is only somewhat slightly differently specified)
+- 2000-ael-har-stockreturns: with almost the same reasoning as above used to model stock exchange data (not exactly gl as here but can be simplified to GL)
+- 1999-cis-won-asyminnogl: almost the same as the 2005 paper (won.. is also the author :)
+- 2007-tejof-kon-vatriskgstock: same use
+- 2001-joae-wa-garchgl: similar (not exactly using gl but very similar)
+
+% Econ Papers
+- often in financial return data (which has high peakes)
+- in gene expression data (since gl hast longer tails)
+
+% eu database
+% ideas
+% citeceer
+% econometric society
+% questia
+% ecb
+% jstor
+% springerlink
+% ebsco
+% wiley
+% sciencedirect
+% informaworld
+
+% Abberger Publikation: Abberger, K., Heiler, S.
+	Simultaneous estimation of parameters for a generalized logistic distribution and application to time series models 	2000 	84 	041-049 	A Allgemeines Statistisches Archiv/Journal of the German Statistical Society
+
+- in 1987-eco-bat-inflationexpectation: the expected inflation is modelled as coming from a logistic distribution
+
 As allready stated, we now apply the general framework, as developed in \cite{z08} to a more specific model, in this case by means of a GL distribution.Prior to that, we would like to give a short justification for our using of the GL distribution.
@@
 \hline
-{\bf Country} & {\bf glogist Breaks (dating unsure)} & {\bf fxregimes Breaks}   \\
+{\bf Country} & {\bf Breakpoints}    \\
 \hline
-Austria & 1999,2008 & no break \\
+Austria & 9.2007\\
 \hline
-Belgium & 2000 & December 2000 \\
+Belgium & 12.1999  \\
 \hline
-Czech Republic & 1998,2002 & no break \\
+Czech Republic &  7.1998  \\
 \hline
-Denmark & 2001 & no break \\
+Denmark & 6.2000 \\
 \hline
-Finland & 2000 & no break \\
+Finland & no \\
 \hline
-France & 2005 & no break \\
+France & 12.2004  \\
 \hline
-Germany & 2000, 2005 & no break \\
+Germany & 5.2000 \\
 \hline
-Greece & 2005 & no break \\
+Greece & no \\
 \hline
-Hungary & 1998,2000,2003 (not very clear) & April 1998 \\
+Hungary & 5.1998 \\
 \hline
-Ireland & 2005 & no break \\
+Ireland & 3.2008 \\
 \hline
-Italy & 1996,2001 & April 1996, December 2000 \\
+Italy & 5.1996; 12.2000  \\
 \hline
-Luxembourg & 1999 & December 1998 \\
+Luxembourg & 12.1998 \\
 \hline
-Netherlands & no break & no break\\
+Netherlands & no \\
 \hline
-Poland & 2001 & April 2001 \\
+Poland & 5.2001  \\
 \hline
-Portugal & 1995 & March 1994 \\
+Portugal & 7.1992 \\
 \hline
-Slovenia & 2003,2005 & no break \\
+Slovenia & 7.2003\\
 \hline
-Spain & 2000 & December 2000 \\
+Spain & 12.2000  \\
 \hline
-Sweden & 1994 & no break \\
+Sweden & 1.1993 \\
 \hline
-United Kingdom & 1994 & no break 
+United Kingdom & 4.1992 \\
 \hline
@@
 
-We find some similarity of a group of countries, Group I (Belgium, Spain and Italy), Group II (France, Greece and Ireland), Group III (Hungary and the Czech Republic) and Group IV (UK and Sweden).
-The other countries seem to be a group of their own.
-
-\subsection{Austria}
-
-The Double max test for Austria is insignificant, but the Cramér-von Mises test and the Supremum of LM statistcs test are both significant.
+%
+%\subsection{Austria}
+%
+%The Double max test for Austria is insignificant, but the Cramér-von Mises test and the Supremum of LM statistcs test are both significant.
 This can be interpreted as a sign, that there was some change in the series as a whole, but that we are unsure wheter this happend with respect to location, scale or shape.
-
-\begin{figure}[h!tp]
-\includegraphics[width=80mm]{aut-meanL2BB.png}
-\caption{Cramér-von Mises test for Austria}
-\end{figure}
-
-Regarding the figure, we have two clear breaks, one in 1999 and another in 2008.
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{aut-meanL2BB.png}
+%\caption{Cramér-von Mises test for Austria}
+%\end{figure}
+%
+%Regarding the figure, we have two clear breaks, one in 1999 and another in 2008.
 [dating very unsure] The fxregimes procedure gives no breakdate.
 To give at least some intuition about the changes in variance [actually this is just scale, but can be adapted quickly] we forced 5 breaks into the Austrian inflation rates.
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{aut-morebreak.png}
+%\caption{Variance Estimation in Austrian Inflation Rates with 5 Breaks}
+%\end{figure}
+%
+%Only with some phantasy we can se some slight V shape (a decline in inflation volatility up to 2000, then an increase). 
+%
+%
+%
+%
+%\subsection{Belgium}
+%
+%Regarding the belgian inflation rates, all tests clearly indicate a break. 
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{bel-meanL2BB.png}
+%\caption{Cramér-von Mises test for Belgium}
+%\end{figure}
+%
+%The fxregimes procedure indicates a breakpoint in December 2000.
+%
+%%The impact of skewness: negative skew (rechtssteil), ie. higher values of y are more probable; positive skew/right skew (linkssteil): lower values of y are more probable
+%%so if it is the case like 6-2-1 we can say that the distribution gets less of a positive skew meaning a lower inflation on average (abklären, sobald das mit den Momenten hinhaut)
+%%
+%
+%\subsection{Italy}
+%
+%In the italian case, there is clear evidence for two breaks. 
+%
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{ita-meanL2BB.png}
+%\caption{Cramér-von Mises test for Italy}
+%\end{figure}
+%
+%If we compare the density of a whole and the denisties of the subsets of the data, we see a clear picture:
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{ita-split.png}
+%\caption{Parameters and Densities for subsamples}
+%\end{figure}
+%
+%
+%As we can see regarding the moments of the subsamples, we see a clear increase in variance (0.04, 0.01, 0.32), some smaller increase in the mean (0.41, 0.16, 0.18)  and a interesting increas in decrease in skewness (0.96, 0.71, $-$0.26), which means that lower values of the inflation rate are quickly becoming less probable.
+%
+%
+%
+%Some History for easier interpretation:
+%
+%% this one from history-emu.pdf
+%
+%eleven countries had met the Maastricht Criteria by 1997, allowing the
+%European monetary union to be launched in 1999. Greece was further behind in
+%meeting the criteria and only joined the monetary union in 2001.
+%First, entry into the euro was a much bigger structural shock for "peripheral"
+%member countries such as Greece, Ireland, Portugal, and Spain than for "core"
+%member countries like Germany and France. In particular, while all euro member
+%countries have enjoyed lower real interest rates under monetary union, the relative
+%decline was much greater for the countries on the periphery.
+%Second, membership in a monetary union can amplify the asymmetric impact
+%of certain shocks. A common nominal interest rate implies that persistent differences
+%in national inflation rates translate into differences in real interest rates across member countries: countries with relatively higher medium-term inflation
+%enjoy lower real interest rates than those with below-average inflation-stimulating
+%demand, credit growth, and housing markets in the former group
+%The Netherlands has lived though a
+%boom-bust cycle since the euro started, with a credit boom and high inflation in
+%1999-2001 followed by a contraction in economic activity that has reversed some of
+%the country's real appreciation against partner countries.
+%Of course, the destabilizing features of European monetary union for the
+%member countries do not mean that currency union has led to a net increase in
+%macroeconomic instability. The European Central Bank has been successful in
+%anchoring medium-term inflation expectations across the euro area at an annual
+%rate around 2 percent. At least for some countries, the attainment of such stability
+%outside the euro framework may have been a more costly process.
+%
+%
+%\section{Results}
+%
+%
+%% in money-growth regime switchin:
+%Germany: early 90ies: low inflation/ mid-end 90ies: high / and afterwards low
+%UK: low regime
+%
+%
+%just as a help
+%
+% wenn ma was siat dass as co-movement git: \cite{vandi} "`euro area countries move from essentially idiosyncratic inflation to co-movement in the mid-1980s"' (p.
+2); "`price shocks in one country not transmitted to another"'; "`correlations remain stable until 1999 but then jumps sharpely"'; (2009-mwp-dijk-internationalinfl)
+%
+%wenn was passiert: zitieren von Carporale: in their estimate \cite{cap} find that each country has at least one of the parameters of their variance estimates significantly different from zero, which is a clear indication of change in volatility as well.
+%
+%falls Volatilität erhöht wird, verweise auf die Debatte in: $2004-wp-minford-ukjoiningemu$, die von einer Erhöhung der britischen Inflationsrate sprechen, sollte UK der EMU beitreten.
+%
+%falls kleine Länder auffälliges Verhalten zeigen: $2008-wp-arnold-inflationuncertainty$, "`This suggests that within EMU, inflation uncertainty may increase in countries that have a smaller influence on ECB policy."' Ihre Begründung: "`For economic agents in small regions which are hit by positive inflation shocks, it will be uncertain how and when the increase
+%in inflation will be reversed.
+This results from imperfect information on the macroeconomic adjustment mechanism combined with the absence of a central bank dedicated to stabilizing their regional inflation rate.
+This does not hold true for the larger regions in the union, who have a largerweight in aggregate inflation and may therefore count on the price stability objective of the union’s central bank."'
+%vergleichen mit den Ergebnissen auf S.343 (country size and inflation uncertainty).
+%
+%berücksichtige die DM Zone (Deutschland, Österreich, Luxemburg, Belgien und Holland), eventuell ähnlich
+%
+%Caporale (ausgedruckt im Diss-Aktuell Ordner) finds that in the post-Euro area there are still significant inflation differentials across countries. 
+%
+%Wenns grosse Unterschiede zwischen den EMU Ländern gibt: ba Caporale stoht: "`the ECb itself admits that monetary policy can only affect price developements in the Euro area as a whole and cannot influence inflation differentials across regions (The Monetary Policy of the ECB, 2004)."' Und dort findet sich auch: "`some short-term volatility is inevitable given the fact that monetary policy can only affect prices wiht long and uncertain lags, hence the focus of the ECB on medium-term price stabilisation."'
+%
+%Um den Anstieg in der Volatilität zu verteidigen: Caporale finds that steady state uncertainty has increased in eight out of twelve countries (p.966) (ie.
+Germany, France, Italy, Spain, Ireland, Finland, Luxembourg and Austria).
+%
+%Zum Abgleich für Break dates: Caporale Table 5.
+%
+%
+%Wenn Ergebnisse da sind: abgleichen bei Caporale mit seiner Fig. 2, Fig. 3
+%
+%falls Österreich und Spanien komisch: Wilfing findet, dass Schilling und Peseta nicht zu ihrer fixen Parität konvergiert sind
+%
+%falls man grosse Unterschiede in der Volatilität sehen würde: $2008-wp-campolmi-laborinflvol$ point out that this can be explained by country specific labor market institutions.
+%
+% morana paper: S.10 untersucht convergence, je nachdem, was sich findet herausfinden, was wie ähnlich ist
+%
+% bei grossen Unterschieden: Busetti S.6: 3 Clustergruppen: Ger, Fra, Bel, Aut, Fin; Spa, Gre, Por, Ireland; Ned und It
+%
+%
+%%%%% still some thought necessary for that
+%
+%% OIL: peaked in 4.7.2008 at 144, fell until 25.12.2008 to 38, afterwards increases again but at slower rate
+%% http://markets.ftd.de/commodities/factsheet_overview.html?ID_NOTATION=4062566
+%
+%% the information is saved in a folder called "`Erläuterungen"' which has the country name
+%% http://www.statistik.at/web_de/statistiken/preise/verbraucherpreisindex_vpi_hvpi/index.html
+%% 1 Austria:
+%Sept. 2007: after Sept. 2007 we see a sharp rise in the yearly inflation rate (from 2 to almost 3.5\% in December 2007); 
+%economic reason: this rise is mostly due to oil, which increased (starting in Sept.
+2007 from -4\% to almost 13\% plus on a yearly basis (p.14); a rise in oil prices from 70 to 94 Dollars; rise of mineral taxes; statistik austria: significant increas in inflation in 2007 4th quarter
+%
+%
+%% 2 Belgium
+%% http://www.indexmundi.com/belgium/inflation_rate_%28consumer_prices%29.html
+%Dez. 1999
+%a very sharp fall in yearly inflation in this time (from 2\% in December 1999 to 0.2\% in January 2000); a very clear increase in volatility; from 1999 to 2000 we had a 137\% increase in the yearly inflation rate
+%economic reason:
+%
+%% 3 Cze
+%% http://www.plzen.czso.cz/eng/redakce.nsf/i/inflation_rate
+%% http://en.wikipedia.org/wiki/Economy_of_the_Czech_Republic#1995-2000
+%Aug. 1998
+%by the end of 1998 we see a strong drop in yearly inflation rates 
+%economic reason: political and financial crisis in 1997, government spending cut, growth dropped, pegged currency forced into floating system, lead to more stabilisation
+%% http://www.allbusiness.com/finance/302725-1.html
+%In terms of economic performance in eastern Europe, 1997 can be best described as the year for the return to realism, or, perhaps, the return to fundamentals. 
+% recession until 1998
 
-\begin{figure}[h!tp]
-\includegraphics[width=80mm]{aut-morebreak.png}
-\caption{Variance Estimation in Austrian Inflation Rates with 5 Breaks}
-\end{figure}
-
-Only with some phantasy we can se some slight V shape (a decline in inflation volatility up to 2000, then an increase).
-
-\subsection{Belgium}
-
-Regarding the belgian inflation rates, all tests clearly indicate a break.
-
-\begin{figure}[h!tp]
-\includegraphics[width=80mm]{bel-meanL2BB.png}
-\caption{Cramér-von Mises test for Belgium}
-\end{figure}
-
-The fxregimes procedure indicates a breakpoint in December 2000.
-
-%The impact of skewness: negative skew (rechtssteil), ie. higher values of y are more probable; positive skew (linkssteil): lower values of y are more probable
-%so if it is the case like 6-2-1 we can say that the distribution gets less of a positive skew meaning a lower inflation on average (abklären, sobald das mit den Momenten hinhaut)
+%% 4 Denmark:
+%Jun. 2000
+%some change in yearly inflation rate; strong increase in variance
+%economic reason: 
 %
-
-\subsection{Italy}
-
-In the italian case, there is clear evidence for two breaks.
-
-\begin{figure}[h!tp]
-\includegraphics[width=80mm]{ita-meanL2BB.png}
-\caption{Cramér-von Mises test for Italy}
-\end{figure}
-
-If we compare the density of a whole and the denisties of the subsets of the data, we see a clear picture:
-
-\begin{figure}[h!tp]
-\includegraphics[width=80mm]{ita-split.png}
-\caption{Parameters and Densities for subsamples}
-\end{figure}
-
-As we can see regarding the moments of the subsamples, we see a clear increase in variance (0.04, 0.01, 0.32), some smaller increase in the mean (0.41, 0.16, 0.18)  and a interesting increas in decrease in skewness (0.96, 0.71, $-$0.26), which means that lower values of the inflation rate are quickly becoming less probable.
+%% 6 Finland
+%no
+%
+%% 7 France
+%Dez. 2004
+%nothing visible in yearly inflation
+%economic reason:
+%
+%% 8 Germany
+%Mar. 2000
+%strong increase in mean and variance; 
+%economic reason: strong increases in oil price, not enough
+%
+%% 9 Greece
+%no
+%but there is severe doubt about the greek rate (looks almost like a uniform distribution)
+%
+%% 10 Hungary
+%May 1998
+%yearly inflation rate sharply decreasing till 1999
+%perhapts same explanation as for czech republic, ie. financial crisis
+%economic reasons: import restrictions have been lifted, trade liberalization; disinflationary impulses (http://findarticles.com/p/articles/mi_m4456/is_n63/ai_20792401/)
+%
+%% 11 Ireland
+%Mar. 2008
+%some fall in yearly inflation after 2008
+%economic reasons: in recession since second quarter of 2008 (http://en.wikipedia.org/wiki/Economy_of_the_Republic_of_Ireland)
+%
+%
+%% 12 Italy
+%May 1996
+%Dez. 2000
+%drop in yearly inflation rate towards lower mean around 1996; in December 2000 nothing really visible; in 2000 beginning of a much more volatile regime
+%economic reasons: in november 1996 Italy rejoined the european monetary system (left it in 1992)
+%September 1996, the new Cabinet announced Italy’s decision to take part in the Monetary Union (http://users.ox.ac.uk/~hine/seminarpapers/Visco.htm)
+%in 1999 italy qualifies for EMU
+%strong economic growth from 1996-2001
+%2000: economic crisis
+%
+%
+%% 13 Lux
+%Dez. 1998
+%in January 1999 deflation with almost 2\%; strongly increased volatility
+%economic reasons: creation of the Banque Centrale du Luxembourg (1998), founded 22.4.1998 :)
+%
+%
+%% 14 Netherlands
+%no
+%
+%% 15 Poland
+%May 2001
+%decrease in yearly inflation 
+%economic reasons: growth has slowed considerably in Q2 2001 (http://www.knowledgerush.com/kr/encyclopedia/Economy_of_Poland/); or interpret it just as a transition to a more stable path
+%
+%
+%% 16 Portugal
+%Jul. 1992
+%decrease in yearly inflation rate from 1990-2000; after 1992 or so yearly inflation stabilized;
+%economic reasons: following a transition period lasting until 1992; so this marks the end of the transition period (http://www.voyagesphotosmanu.com/portugal_economy.html)
+%
+%
+%% 17 Slovenia
+%Jul. 2003
+%economic reasons: EU member in 2004 (perhaps some sign of pre-EU convergenc); mean and volatility fell afterwards; Government announced it plans to meet Maastricht Criteria
+%In 2003, several measures were introduced to make both portfolio and direct investments easier and more transparent in Slovenia and to conduct many financial operations, including banking, securities brokering, and undertaking various credit transactions.
+(http://www.historycentral.com/nationbynation/Slovenia/Economy.html); 1 January 2007 Slovenia adopted the euro as the national currency
+%
+%% 18 Spain
+%Dez. 2000
+%extreme rise in inflation
+%economic reasons: 
+%
+%% 19 Sweden
+%Jan. 1993
+%until 1993 sharp decline in yearly inflation, giving away to sharp rise until 1994, then almost stable
+%economic reasons: 1992 there was a run on the currency, the central bank briefly jacking up interest to 500\% in an unsuccessful effort to defend the currency's fixed exchange rate
+%had a crisis of the 1990s; decided in 1997 against euro introduction; had a banking crises at that time and succesfully ended it in 1993
+%
+%% 20 UK 
+%Apr. 1992
+%sharp decline in yearly inflation rate
+%similar series to sweden
+%recession  hit the British economy at the start of the 1990s, beginning in the summer of 1990 and lasting until the end of 1992
+%
+%
+%
 

### 2010-06-24 pkg/paper/Draft/draft-Z-W.tex
moved figures to separate directory
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@20 b679b2f7-6539-45d0-a605-5ddcb976cde9
GL and some economic interpretation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@23 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \usepackage{harvard}
+
+\graphicspath{{Figures/}}
 
@@
 
-% something more about the reason to use GL
-
-% Book Johnson 95 on logistic distribution
-
-- income distributions, otherwise often used as growth models, or rainfall models
-- definition of the distribution Abberger uses here is a generalization of 23.7 in Johnson 95 (shape ist halt hier zustätzlich noch dabei)
-- longer tails than normal distribution, nonsymmetric (which is a clear net gain)
-- more peaked in the center (net gain)
-- the things above apply to the logistic distribution, not implicitely the GL
-
-% searchterm for these ``generalized logistic distribution''
-% found via Google Scholars
-- in 2005-kjm-olp-gl2 it is shown, that the distribution used here is; cdf is given by: $1-\frac{\exp{-bx}}{(1+\exp{-x})^b}$ (given in this paper)
-- 2005-spl-won-parametersarmitgld: here the GL is used (they find for Autoregressions that mostly data simply is not normal or student, so they use the GL-distribution (which is only somewhat slightly differently specified)
-- 2000-ael-har-stockreturns: with almost the same reasoning as above used to model stock exchange data (not exactly gl as here but can be simplified to GL)
-- 1999-cis-won-asyminnogl: almost the same as the 2005 paper (won.. is also the author :)
-- 2007-tejof-kon-vatriskgstock: same use
-- 2001-joae-wa-garchgl: similar (not exactly using gl but very similar)
-
-% Econ Papers
-- often in financial return data (which has high peakes)
-- in gene expression data (since gl hast longer tails)
-
-% eu database
-% ideas
-% citeceer
-% econometric society
-% questia
-% ecb
-% jstor
-% springerlink
-% ebsco
-% wiley
-% sciencedirect
-% informaworld
-
-% Abberger Publikation: Abberger, K., Heiler, S.
-	Simultaneous estimation of parameters for a generalized logistic distribution and application to time series models 	2000 	84 	041-049 	A Allgemeines Statistisches Archiv/Journal of the German Statistical Society
-
-- in 1987-eco-bat-inflationexpectation: the expected inflation is modelled as coming from a logistic distribution
-
+In econometric issues, the logistic distribution is often used in income distributions and growth models.
+Its use stems from the usefulness of its longer tails and its higher peak, which fits these problems somewhat better.
+The use of the generalized logistic distribution as we will use it in this paper is somewhat scarcer.
+\cite{won} use a GL distribution in an regressio model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that oftentimes is leptokurtic and severely left or right skewed.
+A similar GL distribution is also used in \cite{tolikas} who analyse extreme risk and value--at--risk in the German stock market, allthough they don't use a shape parameter.
+Regarding inflation rates, the GL distribution is --to our best knowledge-- only used in relationship with expected inflation.
+\cite{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
+\\
 As allready stated, we now apply the general framework, as developed in \cite{z08} to a more specific model, in this case by means of a GL distribution.Prior to that, we would like to give a short justification for our using of the GL distribution.
@@
 
-\subsubsection{Kolmogorov-Smirnov Test}
-
-As a goodness of fit test, we use the Kolmogorov Smirnov Test for as given in \cite{massey}.
+%\subsubsection{Kolmogorov-Smirnov Test}
+%
+%As a goodness of fit test, we use the Kolmogorov Smirnov Test for as given in \cite{massey}.
 
@@
 %
-%
 %Geschichte für Interpretation 
@@
 
-\section{Results-very preliminary}
+\section{Results}
 
-\begin{longtable}{|l|p{4cm}|p{3cm}|}
+The results presented here are for the countries within the Euro--zone (Austria, Belgium, Finland, France, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).
+Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due to data scarcity.
+The only ERM~II country included is Denmark, Latvia and Lithunia are excluded due to lack of data.
+The other EU countries included are the Czech Republic, Hungary, Poland, Sweden and the United Kingdom.
+Of these countries Sweden and the UK opted out of the Euro, whereas the other countries are opted for a future adoption of the Euro.
+
+% http://www.euro-anwaerter.de/anwaerter/daenemark.html
+
+[Estonia should be included]
+
+\begin{longtable}{|l|p{4cm}|p{3cm}|p{3cm}|}
 \hline
-{\bf Country} & {\bf Breakpoints}    \\
+{\bf Country} & {\bf 1. breakpoint}  & {\bf 2. breakpoint  \\
 \hline
-Austria & 9.2007\\
+Austria (EURO) & Sep. 2007 & none\\
 \hline
-Belgium & 12.1999  \\
+Belgium (EURO) & Dec. 1999 & none  \\
 \hline
-Czech Republic &  7.1998  \\
+Czech Republic &  Jul. 1998  & none \\
 \hline
-Denmark & 6.2000 \\
+Denmark (ERM~II)& Jun. 2000 & none \\
 \hline
-Finland & no \\
+Finland (EURO) & none & none  \\
 \hline
-France & 12.2004  \\
+France (EURO) & Dec. 2004  & none\\
 \hline
-Germany & 5.2000 \\
+Germany (EURO) & May 2000 & Dec. 2004 \\
 \hline
-Greece & no \\
+Greece (EURO) & none & none \\
 \hline
-Hungary & 5.1998 \\
+Hungary & May 1998 & none\\
 \hline
-Ireland & 3.2008 \\
+Ireland (EURO) & Mar. 2008 & none\\
 \hline
-Italy & 5.1996; 12.2000  \\
+Italy (EURO) & May 1996 &  Dec. 2000  \\
 \hline
-Luxembourg & 12.1998 \\
+Luxembourg (EURO) & Dec. 1998 & none\\
 \hline
-Netherlands & no \\
+Netherlands  (EURO) & none & none \\
 \hline
-Poland & 5.2001  \\
+Poland & May 2001 & none\\
 \hline
-Portugal & 7.1992 \\
+Portugal (EURO) & Jul. 1992 & Mar. 2004\\
 \hline
-Slovenia & 7.2003\\
+Slovakia (EURO) & Apr. 1997 & Feb. 2004 \\
 \hline
-Spain & 12.2000  \\
+Slovenia (EURO) & Jul. 2003 & none\\
 \hline
-Sweden & 1.1993 \\
+Spain (EURO) & May 1996  & Dec. 2000\\
 \hline
-United Kingdom & 4.1992 \\
+Sweden & Jan. 1993 & none\\
+\hline
+United Kingdom & Apr. 1992 & none\\
 \hline
@@
 
-%
-%\subsection{Austria}
-%
-%The Double max test for Austria is insignificant, but the Cramér-von Mises test and the Supremum of LM statistcs test are both significant.
-This can be interpreted as a sign, that there was some change in the series as a whole, but that we are unsure wheter this happend with respect to location, scale or shape.
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{aut-meanL2BB.png}
-%\caption{Cramér-von Mises test for Austria}
-%\end{figure}
-%
-%Regarding the figure, we have two clear breaks, one in 1999 and another in 2008.
-[dating very unsure] The fxregimes procedure gives no breakdate.
-To give at least some intuition about the changes in variance [actually this is just scale, but can be adapted quickly] we forced 5 breaks into the Austrian inflation rates.
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{aut-morebreak.png}
-%\caption{Variance Estimation in Austrian Inflation Rates with 5 Breaks}
-%\end{figure}
-%
-%Only with some phantasy we can se some slight V shape (a decline in inflation volatility up to 2000, then an increase). 
-%
-%
-%
-%
-%\subsection{Belgium}
-%
-%Regarding the belgian inflation rates, all tests clearly indicate a break. 
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{bel-meanL2BB.png}
-%\caption{Cramér-von Mises test for Belgium}
-%\end{figure}
-%
-%The fxregimes procedure indicates a breakpoint in December 2000.
-%
-%%The impact of skewness: negative skew (rechtssteil), ie. higher values of y are more probable; positive skew/right skew (linkssteil): lower values of y are more probable
-%%so if it is the case like 6-2-1 we can say that the distribution gets less of a positive skew meaning a lower inflation on average (abklären, sobald das mit den Momenten hinhaut)
-%%
-%
-%\subsection{Italy}
-%
-%In the italian case, there is clear evidence for two breaks. 
-%
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{ita-meanL2BB.png}
-%\caption{Cramér-von Mises test for Italy}
-%\end{figure}
-%
-%If we compare the density of a whole and the denisties of the subsets of the data, we see a clear picture:
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{ita-split.png}
-%\caption{Parameters and Densities for subsamples}
-%\end{figure}
-%
-%
-%As we can see regarding the moments of the subsamples, we see a clear increase in variance (0.04, 0.01, 0.32), some smaller increase in the mean (0.41, 0.16, 0.18)  and a interesting increas in decrease in skewness (0.96, 0.71, $-$0.26), which means that lower values of the inflation rate are quickly becoming less probable.
-%
-%
+On this basis we trace out some groups of countries.
+The two biggest south european countries Italy and Spain form one group with identical break dates.
+Another group may be formed of the United Kingdom and Denmark.
+Another of the eastern european countries Czech Repbulic, Hungary and Poland.
+Another one of the eastern european countries that allready adopted the Euro like Slovakia and Slovenia.
+Another one of Germany and France.
+And Belgium and Luxembourg, reflecting the fact that Luxembourg and Belgium used their currencies as legal tender in one another.
+
 %
@@
 %
+
+\subsection{Austria}
 %% the information is saved in a folder called "`Erläuterungen"' which has the country name
 %% http://www.statistik.at/web_de/statistiken/preise/verbraucherpreisindex_vpi_hvpi/index.html
-%% 1 Austria:
-%Sept. 2007: after Sept. 2007 we see a sharp rise in the yearly inflation rate (from 2 to almost 3.5\% in December 2007); 
-%economic reason: this rise is mostly due to oil, which increased (starting in Sept.
-2007 from -4\% to almost 13\% plus on a yearly basis (p.14); a rise in oil prices from 70 to 94 Dollars; rise of mineral taxes; statistik austria: significant increas in inflation in 2007 4th quarter
-%
-%
-%% 2 Belgium
+In 2007 we saw a sharp rise in the yearly inflation rate (from 2 to almost 3.5\% in December 2007).
+This rise was mostly due to oil, which increased (starting in Sept.
+2007 from -4\% to almost 13\% plus on a yearly basis a rise in oil prices from 70 to 94 Dollars.
+This was further amplified by a  rise of mineral taxes.
+Also the Statistic Austria official documents report a significant rise in inflation in the fourth quarter of 2007.
+
+\subsection{Belgium}
 %% http://www.indexmundi.com/belgium/inflation_rate_%28consumer_prices%29.html
-%Dez. 1999
-%a very sharp fall in yearly inflation in this time (from 2\% in December 1999 to 0.2\% in January 2000); a very clear increase in volatility; from 1999 to 2000 we had a 137\% increase in the yearly inflation rate
-%economic reason:
-%
-%% 3 Cze
+There was a  very sharp fall in yearly inflation in this time (from 2\% in December 1999 to 0.2\% in January 2000).
+Volatility  a very clear increase in volatility; from 1999 to 2000 we had a 137\% increase in the yearly inflation rate.
+Economic reason?
+
+\subsection{Czech Republic}
 %% http://www.plzen.czso.cz/eng/redakce.nsf/i/inflation_rate
 %% http://en.wikipedia.org/wiki/Economy_of_the_Czech_Republic#1995-2000
-%Aug. 1998
-%by the end of 1998 we see a strong drop in yearly inflation rates 
-%economic reason: political and financial crisis in 1997, government spending cut, growth dropped, pegged currency forced into floating system, lead to more stabilisation
-%% http://www.allbusiness.com/finance/302725-1.html
+By the end of 1998 we see a strong drop in yearly inflation rates.
+Economic reason: political and financial crisis in 1997, government spending cut, growth dropped, pegged currency forced into floating system, lead to more stabilisation.
+% http://www.allbusiness.com/finance/302725-1.html
 %In terms of economic performance in eastern Europe, 1997 can be best described as the year for the return to realism, or, perhaps, the return to fundamentals. 
@@
 
-%% 4 Denmark:
-%Jun. 2000
-%some change in yearly inflation rate; strong increase in variance
-%economic reason: 
-%
-%% 6 Finland
-%no
-%
-%% 7 France
-%Dez. 2004
-%nothing visible in yearly inflation
-%economic reason:
-%
-%% 8 Germany
-%Mar. 2000
-%strong increase in mean and variance; 
-%economic reason: strong increases in oil price, not enough
-%
-%% 9 Greece
-%no
-%but there is severe doubt about the greek rate (looks almost like a uniform distribution)
-%
-%% 10 Hungary
-%May 1998
-%yearly inflation rate sharply decreasing till 1999
-%perhapts same explanation as for czech republic, ie. financial crisis
-%economic reasons: import restrictions have been lifted, trade liberalization; disinflationary impulses (http://findarticles.com/p/articles/mi_m4456/is_n63/ai_20792401/)
-%
-%% 11 Ireland
-%Mar. 2008
-%some fall in yearly inflation after 2008
-%economic reasons: in recession since second quarter of 2008 (http://en.wikipedia.org/wiki/Economy_of_the_Republic_of_Ireland)
-%
-%
-%% 12 Italy
-%May 1996
-%Dez. 2000
-%drop in yearly inflation rate towards lower mean around 1996; in December 2000 nothing really visible; in 2000 beginning of a much more volatile regime
-%economic reasons: in november 1996 Italy rejoined the european monetary system (left it in 1992)
-%September 1996, the new Cabinet announced Italy’s decision to take part in the Monetary Union (http://users.ox.ac.uk/~hine/seminarpapers/Visco.htm)
+\subsection{Denmark}
+Some change in yearly inflation rate; strong increase in variance; Economic reason?
+
+\subsection{France}
+Nothing visible in yearly inflation; Economic reason?
+
+\subsection{Germany}
+Strong increase in mean and variance; Economic reason: strong increases in oil price, not enough. Could be traced back to the poor economic performance of Germany since 2000.
+
+\subsection{Greece}
+There is severe doubt about the credibility greek rate (looks almost like a uniform distribution); extremely high variance, almost 50\% above the next hightes (Belgium).
+
+\subsection{Hungary}
+Yearly inflation rate sharply decreasing till 1999; perhaps same explanation as for Czech republic, import restrictions have been lifted, trade liberalization; disinflationary impulses 
+% (http://findarticles.com/p/articles/mi_m4456/is_n63/ai_20792401/)
+
+\subsection{Ireland}
+In recession since second quarter of 2008. The European country most severely hit by 
+% (http://en.wikipedia.org/wiki/Economy_of_the_Republic_of_Ireland)
+
+\subsection{Italy}
+Drop in yearly inflation rate towards lower mean around 1996; in December 2000 nothing really visible; in 2000 beginning of a much more volatile regime.
+Economic reasons: in November 1996 Italy rejoined the european monetary system (left it in 1992); in September 1996, the new Cabinet announced Italy’s decision to take part in the Monetary Union.
+After 2000 Italy entered an econmic crisis [auf Auffälligkeiten eingehen]
+%(http://users.ox.ac.uk/~hine/seminarpapers/Visco.htm)
 %in 1999 italy qualifies for EMU
@@
 %2000: economic crisis
-%
-%
-%% 13 Lux
-%Dez. 1998
-%in January 1999 deflation with almost 2\%; strongly increased volatility
-%economic reasons: creation of the Banque Centrale du Luxembourg (1998), founded 22.4.1998 :)
-%
-%
-%% 14 Netherlands
-%no
-%
-%% 15 Poland
-%May 2001
-%decrease in yearly inflation 
-%economic reasons: growth has slowed considerably in Q2 2001 (http://www.knowledgerush.com/kr/encyclopedia/Economy_of_Poland/); or interpret it just as a transition to a more stable path
-%
-%
-%% 16 Portugal
-%Jul. 1992
-%decrease in yearly inflation rate from 1990-2000; after 1992 or so yearly inflation stabilized;
-%economic reasons: following a transition period lasting until 1992; so this marks the end of the transition period (http://www.voyagesphotosmanu.com/portugal_economy.html)
-%
-%
-%% 17 Slovenia
-%Jul. 2003
-%economic reasons: EU member in 2004 (perhaps some sign of pre-EU convergenc); mean and volatility fell afterwards; Government announced it plans to meet Maastricht Criteria
-%In 2003, several measures were introduced to make both portfolio and direct investments easier and more transparent in Slovenia and to conduct many financial operations, including banking, securities brokering, and undertaking various credit transactions.
-(http://www.historycentral.com/nationbynation/Slovenia/Economy.html); 1 January 2007 Slovenia adopted the euro as the national currency
-%
-%% 18 Spain
-%Dez. 2000
-%extreme rise in inflation
-%economic reasons: 
-%
-%% 19 Sweden
-%Jan. 1993
-%until 1993 sharp decline in yearly inflation, giving away to sharp rise until 1994, then almost stable
-%economic reasons: 1992 there was a run on the currency, the central bank briefly jacking up interest to 500\% in an unsuccessful effort to defend the currency's fixed exchange rate
-%had a crisis of the 1990s; decided in 1997 against euro introduction; had a banking crises at that time and succesfully ended it in 1993
-%
-%% 20 UK 
-%Apr. 1992
-%sharp decline in yearly inflation rate
-%similar series to sweden
-%recession  hit the British economy at the start of the 1990s, beginning in the summer of 1990 and lasting until the end of 1992
-%
-%
-%
+
+\subsection{Luxembourg}
+In January 1999 deflation with almost 2\%; strongly increased volatility; Economic reasons: creation of the Banque Centrale du Luxembourg (1998), founded 22.4.1998 :)
+
+\subsection{Poland}
+Decrease in yearly inflation; Economic reasons: growth has slowed considerably in Q2 2001; or interpret it just as a transition to a more stable path
+%(http://www.knowledgerush.com/kr/encyclopedia/Economy_of_Poland/);
+
+\subsection{Portugal}
+Decrease in yearly inflation rate from 1990-2000; after 1992 or so yearly inflation stabilized; Economic reasons: following a transition period lasting until 1992; so this marks the end of the transition period
+% (http://www.voyagesphotosmanu.com/portugal_economy.html)
+
+\subsection{Slovakia}
+Economic reasons?
+
+\subsection{Slovenia}
+Economic reasons: EU member in 2004 (perhaps some sign of pre-EU convergenc); mean and volatility fell afterwards; Government announced it plans to meet Maastricht Criteria; In 2003, several measures were introduced to make both portfolio and direct investments easier and more transparent in Slovenia and to conduct many financial operations, including banking, securities brokering, and undertaking various credit transactions.
+In January 2007 Slovenia adopted the euro as the national currency
+% (http://www.historycentral.com/nationbynation/Slovenia/Economy.html);
+
+\subsection{Spain}
+Extreme rise in inflation; Economic reasons?
+
+\subsection{Sweden}
+Until 1993 sharp decline in yearly inflation, giving away to sharp rise until 1994, then almost stable; Economic reasons: 1992 there was a run on the currency, the central bank briefly jacking up interest to 500\% in an unsuccessful effort to defend the currency's fixed exchange rate; had a crisis of the 1990s; decided in 1997 against euro introduction; had a banking crises at that time and succesfully ended it in 1993
+
+\subsection{United Kingdom}
+Sharp decline in yearly inflation rate; recession  hit the British economy at the start of the 1990s, beginning in the summer of 1990 and lasting until the end of 1992
 
@@
 
+% something more about the reason to use GL
+
+% Book Johnson 95 on logistic distribution
+%
+%- income distributions, otherwise often used as growth models, or rainfall models
+%- definition of the distribution Abberger uses here is a generalization of 23.7 in Johnson 95 (shape ist halt hier zustätzlich noch dabei)
+%- longer tails than normal distribution, nonsymmetric (which is a clear net gain)
+%- more peaked in the center (net gain)
+%- the things above apply to the logistic distribution, not implicitely the GL
+
+% searchterm for these ``generalized logistic distribution''
+% found via Google Scholars
+%- in 2005-kjm-olp-gl2 it is shown, that the distribution used here is; cdf is given by: $1-\frac{\exp{-bx}}{(1+\exp{-x})^b}$ (given in this paper)
+%- 2005-spl-won-parametersarmitgld: here the GL is used (they find for Autoregressions that mostly data simply is not normal or student, so they use the GL-distribution (which is only somewhat slightly differently specified)
+%- 2000-ael-har-stockreturns: with almost the same reasoning as above used to model stock exchange data (not exactly gl as here but can be simplified to GL)
+%- 1999-cis-won-asyminnogl: almost the same as the 2005 paper (won.. is also the author :)
+%- 2007-tejof-kon-vatriskgstock: same use
+%- 2001-joae-wa-garchgl: similar (not exactly using gl but very similar)
+%
+%
+%% Econ Papers
+%- often in financial return data (which has high peakes)
+%- in gene expression data (since gl hast longer tails)
+
+% eu database
+% ideas
+% citeceer
+% econometric society
+% questia
+% ecb
+% jstor
+% springerlink
+% ebsco
+% wiley
+% sciencedirect
+% informaworld
+
+% Abberger Publikation: Abberger, K., Heiler, S.
+	Simultaneous estimation of parameters for a generalized logistic distribution and application to time series models 	2000 	84 	041-049 	A Allgemeines Statistisches Archiv/Journal of the German Statistical Society
+
+%- in 1987-eco-bat-inflationexpectation: the expected inflation is modelled as coming from a logistic distribution
+
+%
+%\subsection{Austria}
+%
+%The Double max test for Austria is insignificant, but the Cramér-von Mises test and the Supremum of LM statistcs test are both significant.
+This can be interpreted as a sign, that there was some change in the series as a whole, but that we are unsure wheter this happend with respect to location, scale or shape.
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{aut-meanL2BB.png}
+%\caption{Cramér-von Mises test for Austria}
+%\end{figure}
+%
+%Regarding the figure, we have two clear breaks, one in 1999 and another in 2008.
+[dating very unsure] The fxregimes procedure gives no breakdate.
+To give at least some intuition about the changes in variance [actually this is just scale, but can be adapted quickly] we forced 5 breaks into the Austrian inflation rates.
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{aut-morebreak.png}
+%\caption{Variance Estimation in Austrian Inflation Rates with 5 Breaks}
+%\end{figure}
+%
+%Only with some phantasy we can se some slight V shape (a decline in inflation volatility up to 2000, then an increase). 
+%
+%
+%
+%
+%\subsection{Belgium}
+%
+%Regarding the belgian inflation rates, all tests clearly indicate a break. 
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{bel-meanL2BB.png}
+%\caption{Cramér-von Mises test for Belgium}
+%\end{figure}
+%
+%The fxregimes procedure indicates a breakpoint in December 2000.
+%
+%%The impact of skewness: negative skew (rechtssteil), ie. higher values of y are more probable; positive skew/right skew (linkssteil): lower values of y are more probable
+%%so if it is the case like 6-2-1 we can say that the distribution gets less of a positive skew meaning a lower inflation on average (abklären, sobald das mit den Momenten hinhaut)
+%%
+%
+%\subsection{Italy}
+%
+%In the italian case, there is clear evidence for two breaks. 
+%
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{ita-meanL2BB.png}
+%\caption{Cramér-von Mises test for Italy}
+%\end{figure}
+%
+%If we compare the density of a whole and the denisties of the subsets of the data, we see a clear picture:
+%
+%\begin{figure}[h!tp]
+%\includegraphics[width=80mm]{ita-split.png}
+%\caption{Parameters and Densities for subsamples}
+%\end{figure}
+%
+%
+%As we can see regarding the moments of the subsamples, we see a clear increase in variance (0.04, 0.01, 0.32), some smaller increase in the mean (0.41, 0.16, 0.18)  and a interesting increas in decrease in skewness (0.96, 0.71, $-$0.26), which means that lower values of the inflation rate are quickly becoming less probable.
+%
+%
+
 \setstretch{1,3}

### 2010-06-29 pkg/paper/Draft/draft-Z-W.tex
some papgers and comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@24 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 %% http://www.indexmundi.com/belgium/inflation_rate_%28consumer_prices%29.html
+% Country Report Belgium  bei ebscohost in TX all text
 There was a  very sharp fall in yearly inflation in this time (from 2\% in December 1999 to 0.2\% in January 2000).
 Volatility  a very clear increase in volatility; from 1999 to 2000 we had a 137\% increase in the yearly inflation rate.
-Economic reason?
+Economic reason: belgish franc was pegged to German Mark (since 1990).
+Possibly: Domestic inflation was mainly driven by the surge in world oil prices, with the result that energy prices in Belgium were around 25 \% higher in the first quarter of this year than in the year-earlier period.
+Also a general upward trend in prices.
+% source: Belgium_3206697
+tax reductions in personal income tax
+% source: Belgium_4611003
+inflation turns out to be more volatile in small, open economies such as Belgium (might be the case in some countries, but not in all of them).
+some explanations happen to be: a) a more volatile currency risk premium (which cannot be the case in Belgium, since it happend after 1.1999 and the currency was pegged thereafter, and even before pegged to Germany); b) degree of exchange--rate--passthrough to inflation (not possible as well) and c) size of monetary policy shocks (if there were any, should have to be observed in ALL EMU countries, since monetary policy equal).
+another possible explanation are productivity shocks which boost inflation rate volatility (but this for more than 10 years????)
+
+[some connection with interest rate? maybe not the same until 1.2002 ?]
+[volatility of inflation decreases the rate of growth]
+%(growth_rate_vol_inflation)
 
@@
 \subsection{Ireland}
-In recession since second quarter of 2008. The European country most severely hit by 
+In recession since second quarter of 2008. The European country most severely hit by the recession
 % (http://en.wikipedia.org/wiki/Economy_of_the_Republic_of_Ireland)

### 2010-07-03 pkg/paper/Draft/draft-Z-W.tex
smaller head at draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@25 b679b2f7-6539-45d0-a605-5ddcb976cde9
- Simplified header further. - Reduced spacing and increased textwidth (via a4wide), hence equations   don't need special split treatment anymore. - Put table into a regular {table}/{tabular} environment. Changed formatting   somewhat. - Changed citation style (due to problems with the Harvard style) to   abbrvnat, still can choose something better later on. - Still many problems with citations: some missing, some BibTeX items   are clearly malformatted.
git-svn-id: svn://sv

@@
-\documentclass[titlepage,a4paper, fleqn, flalign]{article} \usepackage[pdftex]{graphicx}
-\usepackage[multiple]{footmisc} \pagestyle{myheadings}
+\documentclass[a4paper]{article}
+\usepackage{graphicx,a4wide,color,hyperref}
 \usepackage[latin1]{inputenc}
-\usepackage{fancyhdr}
-\usepackage{rotating}
-\usepackage{graphicx}
-\usepackage{longtable}
-\usepackage{amsmath}
-\usepackage{ltxtable}
-\usepackage{tabularx}
-\usepackage{rotating}
-\usepackage{array}
-\usepackage{amssymb}
-\usepackage{pstricks}
-\usepackage{qtree}
-\usepackage{lscape}
-\usepackage{setspace}
-\usepackage[numbers,round]{natbib}
-\usepackage{harvard}
+\usepackage[authoryear,round]{natbib}
+\bibpunct{(}{)}{;}{a}{}{,}
+\pagestyle{myheadings}
 
@@
 
-%\title{Draft  \\[2cm]
-%\textbf{Structural Breaks in Inflation Rates}}
-%
-%\author{ \scriptsize by \\
-%Mag. Thomas WINDBERGER\\
-%Prof. Dr. Achim ZEILEIS \\
-%\scriptsize Doctoral Advisors: \\
-%\scriptsize Univ.-Prof.Dr. Jesus Crespo-Cuaresma \\
-%\scriptsize A.Univ.-Prof.Dr. Janette Walde \\ [1cm]}
-%
-%
-%
-%\maketitle
+\title{Still need a good title}
+\author{}
+\date{}
+\maketitle
 
-% anderthalb Zeilen Abstand inne
-
-\setstretch{1,5}
-
-\section{Abstract}
-
+\begin{abstract}
 The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)  on changing the volatility of the inflation rates in its Member States. 
 %We concentrate on the volatility of the inflation rate rather than its mean since \cite{fried} conjectured that the harmful effect of inflation on growth is driven by inflation volatility. 
+\end{abstract}
 
-\newpage
 \section{Introduction}
 
-The ECB defines price stability ``as a year-on-year increase in the Harmonised Index of Consumer Prices (HICP) for the euro area of below 2 \%'' \cite{castel}.
+The ECB defines price stability ``as a year-on-year increase in the Harmonised Index of Consumer Prices (HICP) for the euro area of below 2\%'' \cite{castel}.
 \cite{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22).
@@
 
-Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot ln(HICP_{t}/HICP_{t-1})$.
+Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot \log(HICP_{t}/HICP_{t-1})$.
 Countries included are Austria,	Belgium,	Czech Republic,	Denmark,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,		Spain,	Sweden,		United Kingdom and the	Euro area.
@@
 
-\newpage
 \section{Methodology}
@@
 \begin{eqnarray}
-& & f(\pi | \theta, \sigma, \delta) = \frac{\frac{\delta}{\sigma}*\exp^{-\frac{\pi_i-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi_i-\theta}{\sigma}})^{(\delta+1)}}
+f(\pi | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma}*\exp^{-\frac{\pi_i-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi_i-\theta}{\sigma}})^{(\delta+1)}}
 \end{eqnarray}
@@
 \begin{eqnarray}
-& & E(\pi_i) = \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
-& & Var(\pi_i)  = \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
-& & Skew(\pi_i) =  \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
+E(\pi_i) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
+Var(\pi_i)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
+Skew(\pi_i) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
 \end{eqnarray}
@@
 \begin{eqnarray}
-& & l(\delta,\theta,\sigma, \pi) = ln(\delta) - ln(\sigma) - \frac{1}{\sigma} (\pi-\theta) - (\delta+1) ln(1+\exp^{-\frac{\pi-\theta}{\sigma}})
+l(\delta,\theta,\sigma, \pi) & = & \log(\delta) - \log(\sigma) - \frac{1}{\sigma} (\pi-\theta) - (\delta+1) \log(1+\exp^{-\frac{\pi-\theta}{\sigma}})
 \end{eqnarray}
@@
 \begin{eqnarray}
-& & \psi(\pi_i,\delta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta} = \frac{1}{\delta} - ln(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \\
-& & \psi(\pi_i,\theta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta} = \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
+\psi(\pi_i,\delta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta} = \frac{1}{\delta} - \log(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \\
+\psi(\pi_i,\theta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta} = \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
 \end{eqnarray}
 
-\begin{equation}
-\begin{split}
-&& \psi(\pi_i,\sigma) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma} = -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) * \\
+\begin{eqnarray}
+\psi(\pi_i,\sigma) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma} = -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) * \\
 && \quad \frac{\frac{1}{\sigma^2}(\pi-\theta)\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
-\end{split}
-\end{equation}
+\end{eqnarray}
 
@@
 %[wann war die Erstpublikation der graphischen Methode?].
-
+%
+%
 %\subsubsection{Kolmogorov-Smirnov Test}
+%% information from: Heiberger/Holland: Statistical Analysis and Data Display
 %
-%As a goodness of fit test, we use the Kolmogorov Smirnov Test for as given in \cite{massey}.
+%The one sample test tests whether a random sample is coming from a hypothesized null distribution.
+It compares the maximum discrepancy between the empirical cumulative distribution of the data, defined as S(y) = fraction of the data that is less than or equal to y and the cdf of the hypothesized population being sampled.
+The null hypothesis (ie.
+that a dataset has a distribution described by a fully determined (theoretical) distribution function) is rejected at a given significance level if the testing statistic exceeds the critical value.
+%It is not possible to use the Kolmogorov–Smirnov test when the parameters of the distribution function are estimated from the same dataset onwhich they are tested.
+This, however, is usually the case in any practical application.
+This approach leads to type II error (failing to reject the null hypothesiswhen it is false).
+%% Bemerkung: der ganze obige Absatz stammt aus: gamma-ksmirnov im Goodness-of-Fit Ordner; daher etwas umformulieren
+%
+%% this text: Kendal's Advanced Theory of Statistics, Sixth Edition, Volume 2A, by Alan Stuard et al.
+%The measure of deviation used is the maximum difference between $S_n(x)$ (the cdf given by the observed sample) and $F_0(x)$ (the cdf hypothesized under $H_0$).
+Where $S_n(x)$ is r/n where r is the number of observations less than or equal x and n is the sample size.
+%
+%\begin{eqnarray}
+%& & D_n = \underset{x}{sup}|S_n(x)-F_0(x)|
+%\end{eqnarray}
+%
+%When the hypothesized distribution is composite, say $F_0(x|\theta_1,\theta_2,...\theta_s)$, with s paramters $\theta_j$, we must select s functions $t_1,...t_s$ of the $x_i$ for this purpose.
+[in the easy case $F_0$ is uniformly distributed on (0,1)] The transformed variables are now:
+%
+%\begin{eqnarray}
+%& & y_i = \int_{-\infty}^{x_i} f_0(t_1,t_2,..t_s) du
+%\end{eqnarray}
+%
+%but they are neither independent nor uniformly distributed, and their disribution depend on general both on the hypothetical distribution $F_0$ and on the true values of its parameters.
+So in our 3--parameter case there is some problem.
+%Perhaps the most serious limitation is that the distribution must be fully specified.
+That is, if location, scale, and shape parameters are estimated from the data, the critical region of the K-S test is no longer valid.
+It typically must be determined by simulation.
+%
+%% http://www.itl.nist.gov/div898/handbook/eda/section3/eda35g.htm
+% 
+%% sup ist hier das Maximum
+%% ist der Wert grösser als der kritische Wert, wird die H0 abgelehnt
+%% http://de.wikipedia.org/wiki/Kolmogorow-Smirnow-Test
+%
+%
+%% There is some more refined distribution theory for the KS test with estimated parameters (see Durbin, 1973), but that is not implemented in ks.test.
+%% Durbin, J. (1973) Distribution theory for tests based on the sample distribution function. SIAM. 
+%
+%
+%
+%% A Kolmogorov - Smirnov goodness-of-fit test for the two-parameter weibull distribution when the parameters are estimated from the data 
+%
+%http://www.sciencedirect.com/science?_ob=ArticleURL&_udi=B6V47-4D452FR-4P&_user=10&_coverDate=12%2F31%2F1982&_rdoc=1&_fmt=high&_orig=search&_sort=d&_docanchor=&view=c&_searchStrId=1385134404&_rerunOrigin=google&_acct=C000050221&_version=1&_urlVersion=0&_userid=10&md5=3102a52af9d8937550065045f0b94ee3
+%
+%
+%
+%%Handbook of parametric and nonparametric statistical procedures von Sheskin David wäre nützlich
+%
+%
+%\subsubsection{The $\chi^2$ Test}
+%
+%The main problem is to determine the number of classes k (data is grouped into k classes where we then calculate the difference between observed and expected frequencies).
+In a continouus distribution case, there are no natural boundaries.
+One procedure advocated by Mann and Gumbel was to choose the classes so that - given k - the hypothetical probabilites are all equal to 1/k.
+Recommendations: use classes with equal or nearly equal probabilities.
+Determine number of classes by:
+%
+%\begin{eqnarray}
+%& & k = b (\frac{2^{1/2}(n-1)}{\lambda_{\alpha} + G^{-1}P_0})
+%\end{eqnarray}
+%
+%with b between 2 and 4; where $\lambda_{\alpha}$ is the critical value (ie. 1.64 in normal distribution case); P is the value of the power function, and $G^{-1}$ comes from ????.
+%In case of ML estimation, critical values should be adjusted upwards.
+%The test statistic follows, approximately, a chi-square distribution with (k - c) degrees of freedom where k is the number of non-empty cells and c = the number of estimated parameters (including location and scale parameters  and shape parameters) for the distribution + 1.
+%
+%\begin{eqnarray}
+%& & \chi^2 = \sum_{i=1}^k \frac{O_i - E_i}{E_i}
+%\end{eqnarray}
+%
+%where the $O_i$ are the observed frequencies and the $E_i$ the expected frequencies. Null hypothesis is rejected if: $\chi^2 $>$ \chi^2_{\alpha, k-c}$.
+%One rule would be to let k be equal to  $3.765(N-1)^{2/5}$ for a 5\% significance level, when using intervals with equal probability under $H_0$.
+But the statistical community still does not agree on the optimal choice of k.
+[see 15715341- goodness of fit folder, untere und obere grenze sind +- infinity]
+%
+%
+%
+%
+%
+%
 
@@
 
-\newpage
 \subsection{Expected Results}
@@
 
-[Estonia should be included]
-
-\begin{longtable}{|l|p{4cm}|p{3cm}|p{3cm}|}
+\begin{table}{tbh}
+\begin{center}
+\begin{tabular}{llll}
 \hline
-{\bf Country} & {\bf 1. breakpoint}  & {\bf 2. breakpoint  \\
-\hline
-Austria (EURO) & Sep. 2007 & none\\
-\hline
-Belgium (EURO) & Dec. 1999 & none  \\
-\hline
-Czech Republic &  Jul. 1998  & none \\
-\hline
-Denmark (ERM~II)& Jun. 2000 & none \\
-\hline
-Finland (EURO) & none & none  \\
-\hline
-France (EURO) & Dec. 2004  & none\\
-\hline
-Germany (EURO) & May 2000 & Dec. 2004 \\
-\hline
-Greece (EURO) & none & none \\
-\hline
-Hungary & May 1998 & none\\
-\hline
-Ireland (EURO) & Mar. 2008 & none\\
-\hline
-Italy (EURO) & May 1996 &  Dec. 2000  \\
-\hline
-Luxembourg (EURO) & Dec. 1998 & none\\
-\hline
-Netherlands  (EURO) & none & none \\
-\hline
-Poland & May 2001 & none\\
-\hline
-Portugal (EURO) & Jul. 1992 & Mar. 2004\\
-\hline
-Slovakia (EURO) & Apr. 1997 & Feb. 2004 \\
-\hline
-Slovenia (EURO) & Jul. 2003 & none\\
-\hline
-Spain (EURO) & May 1996  & Dec. 2000\\
-\hline
-Sweden & Jan. 1993 & none\\
-\hline
-United Kingdom & Apr. 1992 & none\\
-\hline
-\end{longtable}
+Country        & Type   & \multicolumn{2}{l}{Breakpoints}  \\ \hline
+Austria        & EURO   & Sep 2007 &          \\
+Belgium        & EURO   & Dec 1999 &          \\
+Czech Republic &        & Jul 1998 &          \\
+Denmark        & ERM~II & Jun 2000 &          \\
+Estonia        &        & TODO     &          \\
+Finland        & EURO   & none     &          \\
+France         & EURO   & Dec 2004 &          \\
+Germany        & EURO   & May 2000 & Dec 2004 \\
+Greece         & EURO   & none     &          \\
+Hungary        &        & May 1998 &          \\
+Ireland        & EURO   & Mar 2008 &          \\
+Italy          & EURO   & May 1996 & Dec 2000 \\
+Luxembourg     & EURO   & Dec 1998 &          \\
+Netherlands    & EURO   & none     &          \\
+Poland         &        & May 2001 &          \\
+Portugal       & EURO   & Jul 1992 & Mar 2004 \\
+Slovakia       & EURO   & Apr 1997 & Feb 2004 \\
+Slovenia       & EURO   & Jul 2003 &          \\
+Spain          & EURO   & May 1996 & Dec 2000 \\
+Sweden         &        & Jan 1993 &          \\
+United Kingdom &        & Apr 1992 &          \\ \hline
+\end{tabular}
+\caption{\label{tab:breakpoints} Caption needed.}
+\end{center}
+\end{table}
 
@@
 Economic reason: belgish franc was pegged to German Mark (since 1990).
-Possibly: Domestic inflation was mainly driven by the surge in world oil prices, with the result that energy prices in Belgium were around 25 \% higher in the first quarter of this year than in the year-earlier period.
+Possibly: Domestic inflation was mainly driven by the surge in world oil prices, with the result that energy prices in Belgium were around 25\% higher in the first quarter of this year than in the year-earlier period.
 Also a general upward trend in prices.
@@
 Drop in yearly inflation rate towards lower mean around 1996; in December 2000 nothing really visible; in 2000 beginning of a much more volatile regime.
-Economic reasons: in November 1996 Italy rejoined the european monetary system (left it in 1992); in September 1996, the new Cabinet announced Italy’s decision to take part in the Monetary Union.
+Economic reasons: in November 1996 Italy rejoined the european monetary system (left it in 1992); in September 1996, the new Cabinet announced Italy's decision to take part in the Monetary Union.
 After 2000 Italy entered an econmic crisis [auf Auffälligkeiten eingehen]
@@
 
-\setstretch{1,3}
-
-\newpage
-\section{Literature}
-
-\renewcommand*{\refname}{}
+\bibliographystyle{abbrvnat}
 \bibliography{papers}
-\begin{thebibliography}{99}
-\bibliographystyle{apsr}
 

### 2010-07-17 pkg/paper/Draft/draft-Z-W.tex
Estonia included in Dataset Draft now Elsevier Design papers.bib adapted
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@27 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2010-06-24 pkg/paper/Draft/draft-Z-W.tex
@@
-\documentclass[titlepage,a4paper, fleqn, flalign]{article} \usepackage[pdftex]{graphicx}
-\usepackage[multiple]{footmisc} \pagestyle{myheadings}
+\documentclass[a4paper, 5p, preprint]{elsarticle}
+\usepackage{graphicx,a4wide,color,hyperref}
 \usepackage[latin1]{inputenc}
-\usepackage{fancyhdr}
-\usepackage{rotating}
-\usepackage{graphicx}
-\usepackage{longtable}
-\usepackage{amsmath}
-\usepackage{ltxtable}
-\usepackage{tabularx}
-\usepackage{rotating}
-\usepackage{array}
-\usepackage{amssymb}
-\usepackage{pstricks}
-\usepackage{qtree}
-\usepackage{lscape}
-\usepackage{setspace}
-\usepackage[numbers,round]{natbib}
-\usepackage{harvard}
+\usepackage[authoryear, round]{natbib}
+\bibpunct{(}{)}{;}{a}{}{,}
+\pagestyle{myheadings}
+
+\biboptions{longnamesfirst,angle,semicolon}
 
@@
 
-%\title{Draft  \\[2cm]
-%\textbf{Structural Breaks in Inflation Rates}}
-%
-%\author{ \scriptsize by \\
-%Mag. Thomas WINDBERGER\\
-%Prof. Dr. Achim ZEILEIS \\
-%\scriptsize Doctoral Advisors: \\
-%\scriptsize Univ.-Prof.Dr. Jesus Crespo-Cuaresma \\
-%\scriptsize A.Univ.-Prof.Dr. Janette Walde \\ [1cm]}
-%
-%
-%
-%\maketitle
+\title{Still need a good title}
 
-% anderthalb Zeilen Abstand inne
+% Verkaufsstrategien
+% 1. Widerlegung des Teuro Effektes
+% 2. Keine Veränderung d. EMU sichtbar; nach 2002 kein globaler Trend aller EMU Staaten Richtung Wechsel
+% 3. Nutzen zeigen, den die osteuropäischen Staaten haben, dadurch, dass sie dem EURO beitreten müssen  (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
+% 	 so etwa in Slowenien und der Slovakei ein Bruch bevor diese Staaten dem ERMII/EURO beitreten
+% 4. Finanzkrise: schwieriger, ausser Irland keine Änderung > 2008
 
-\setstretch{1,5}
+\author[rvt]{A.˜Zeileis}
+\author[focal]{T.˜Windberger}
+\address[rvt]{University of Innsbruck, Institute for Statistics, Universitätsstraße 15, 6020 Innsbruck, Austria}
+\address[focal]{University of Innsbruck, Austria}
 
-\section{Abstract}
+\date{}
+\maketitle
 
+\begin{abstract}
 The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)  on changing the volatility of the inflation rates in its Member States. 
-%We concentrate on the volatility of the inflation rate rather than its mean since \cite{fried} conjectured that the harmful effect of inflation on growth is driven by inflation volatility.
+\end{abstract}
 
-\newpage
+\begin{keyword}
+inflation rate \sep structural break \sep EMU 
+\end{keyword}
+
 \section{Introduction}
 
-The ECB defines price stability ``as a year-on-year increase in the Harmonised Index of Consumer Prices (HICP) for the euro area of below 2 \%'' \cite{castel}.
-\cite{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22).
+The ECB defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of below 2\%'' \citet{castel}.
+\citet{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22).
 Thus to achieve a low inflation rate and low variability of inflation must be a key issue for the central bank.
 \\
-To further justify the use of inflation volatility, we refer to \cite{rother}: ``Among the harmful effects of inflation, the negative consequences of inflation volatility are of particular concern.
-High variability of inflation over time makes expectations over the future price level more uncertain.
-In a world with nominal contracts, this induces risk premia for long-term arrangements, raises costs for hedging against inflation risks and leads to unanticipated redistribution of wealth.
-Thus, inflation volatility can impede growth even if inflation on average remains restrained.'' \\
-The question of interest centers around the way in which a countries decision to join the EMU changed its inflation volatility.
-There are a number of reasons why the volatility of  one of these nations should change indeed.
-Given that the country experienced quite volatile inflation rates, its efforts to meet the convergence criterias are likely to lead to a change in the volatility of their respective inflation rates, as has indeed been the case for a number of EMU countries, like Spain, Portugal, Italy and Greece.
+The question of interest centers around the way in which a countries decision to join the EMU changed its inflation rate dynamics.
+There are a number of reasons why these should change indeed.
+Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria are likely to lead to a change at least in the mean and possible in the volatility of their respective inflation rates as well.
+This has indeed been the case for a number of EMU countries, like Spain, Italy and Portugal, reflecting their effort to meet the Maastricht Criteria for inflation rates.
 
@@
 
-%Sinn: Zusammenfassung der gegebenen Literatur.
-Darstellen, was in der Richtung publiziert wurde und herausarbeiten, warum mein Ansatz hier jetzt neu sein soll und warum die Frage noch nicht beantwortet ist.
-
-Theory unfortunately is quite unsure about whether or not the creation of a monetary union between two or more states is likely to reduce/increase the variability or even the level of the inflation rate.
-\cite{cooper} points out, that "a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a "central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation.
+Theory unfortunately is quite unsure about whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
+\citet{cooper} points out, that "a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a "central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation.
 It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
-An interesting approach to this question is taken by \cite{holte}.
-He creates a two country model for monetary policy analysis along the line of two models by \cite{mc1} and \cite{mc2} as well as \cite{gali}.
+An interesting approach to this question is taken by \citet{holte}.
+He creates a two country model for monetary policy analysis along the line of two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}.
 The inflation is modeled by a hybrid Phillips curve (NKPC) specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
@@
 \\
-\cite{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation, its degree of persistence and both types of uncertainty.
+\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation, its degree of persistence and both types of uncertainty.
 They use a time-varying model with a GARCH specification for the unconditional volatility of inflation and find some instability in the conditional volatility.
 \\
-In a paper examining structural convergence of the inflation rates in EU counries, \cite{sarno} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
-They find that convergence in time of inflation dynamics was only partly observeable.
-The same is also found by \cite{palomba}, they too find little evidence of similarity of short-run inflation dynamics.
+In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
+They find that convergence in time of inflation dynamics was only partly observable.
+In a paper studying core inflation and using an aggregated Euro Area inflation rate, \citet{morana} finds three regimes (roughly 1980-1984, 1984-1993 and 1993-2000) governing the core inflation rate.
 \\
-In a paper studying core inflation and using an aggregated Euro Area inflation rate, \cite{morana} finds three regimes (roughly 1980-1984, 1984-1993 and 1993-2000) governing the core inflation rate.
+Inquiring into the convergence properties of inflation rates among countries of the EMU, \citet{busetti} find that from 1980-1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
 \\
-Inquiering into the convergence properties of inflation rates among countries of the EMU, \cite{busetti} find that from 1980-1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
-\\
-%A otherwise very interesting paper by \cite{maheu} can not be used due to its defintion of a structural break as "`an unpredictable event in which the relationship among the variables in a model change"', which differs from the definition of a sructural break used in this paper.
 
@@
 
-Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot ln(HICP_{t}/HICP_{t-1})$.
-Countries included are Austria,	Belgium,	Czech Republic,	Denmark,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,		Spain,	Sweden,		United Kingdom and the	Euro area.
+Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot \log(HICP_{t}/HICP_{t-1})$.
+Countries included are Austria,	Belgium,	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Spain,	Sweden,		United Kingdom and the	Euro area.
 The data are obtained from the OECD Statistics.
 \\
-The countries can be divided into three different groups: the EMU countries (Portugal, Spain, Italy, Greece, France, Belgium, Netherlands, Germany, Austria, Ireland, Slovenia, Luxembourg and Finland), EU members without ERM~II (Exchange Rate Mechanism): United Kingdom, Sweden, Poland, Czech Republic, Hungary.
+The countries can be divided into three different groups: the EURO countries (Portugal, Spain, Italy, Greece, France, Belgium, Estonia\footnote{although Estonia enters in 2011}, Netherlands, Germany, Austria, Ireland, Slovenia, Luxembourg and Finland), EU members without ERM~II (Exchange Rate Mechanism): United Kingdom, Sweden, Poland, Czech Republic, Hungary.
 Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
 
-% slovenien statt slovakei?
-% Deutschland: CPI statt HVPI
-% vielleicht ein paar Länder aus dem Sample entfernen: 
-% slovakei, tschechien, ungarn, polen
-% Nachschauen, wann in den jeweiligen Ländern die Entscheidung getroffen wurde, den Euro einzuführen. --> viel zu aufwendig
+\section{Methods}
 
-%/*  1 = austria = 240
-%    2 = belgien = 228 (erst ab 91)
-%    3 = czechrepublic = 180 (erst ab 95)
-%    4 = denmark = 240
-%    5 = finland = 240
-%    6 = france = 240
-%    7 = germany = 180
-%    8 = greece = 180
-%    9 = hungary = 180
-%    10 = iceland = 180
-%    11 = ireland = 180
-%    12 = italy = 240
-%    13 = luxembourg = 180
-%    14 = netherlands = 240
-%    15 = poland = 168
-%    16 = portugal = 240
-%    17 = slovakei = 180
-%    18 = spanien = 216
-%    19 = schweden = 240
-%    20 = uk = 239 (fehlt 12.2009)
-%    21 = euroarea = 168
-%*/
-%
-%/* CPI Daten von:
-%    22 = Japan = 240
-%    23 = Norway = 240
-%    24 = Switzerland = 240
-%    25 = USA = 240
-%/*
-
-\newpage
-\section{Methodology}
-
-\subsection{Using a generalized logistic distribution (GL)}
+\subsection{A generalized logistic distribution (GL)}
 
@@
 The use of the generalized logistic distribution as we will use it in this paper is somewhat scarcer.
-\cite{won} use a GL distribution in an regressio model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that oftentimes is leptokurtic and severely left or right skewed.
-A similar GL distribution is also used in \cite{tolikas} who analyse extreme risk and value--at--risk in the German stock market, allthough they don't use a shape parameter.
+\citet{won} uses a GL distribution in a regression model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  leptokurtic and severely left or right skewed.
+A similar GL distribution is also used in \citet{tolikas} who analyses extreme risk and value--at--risk in the German stock market, all though they don't use a shape parameter.
 Regarding inflation rates, the GL distribution is --to our best knowledge-- only used in relationship with expected inflation.
-\cite{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
+\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
 \\
-As allready stated, we now apply the general framework, as developed in \cite{z08} to a more specific model, in this case by means of a GL distribution.Prior to that, we would like to give a short justification for our using of the GL distribution.
-Regarding the data at hand, it was not possible to use the already existing method developed in \cite{z07}, since allmost all inflation rates, with the noteable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties.
-Therefore, a somewhat more flexibel distribution had to be used.
+As already stated, we now apply the general framework, as developed in \citet{z07} to a more specific model, in this case by means of a GL distribution.
+Prior to that, we would like to give a short justification for our using of the GL distribution.
+Regarding the data at hand, it was not possible to use the already existing method developed in \citet{z07}, since almost all inflation rates, with the notable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties.
+\\
+Therefore, a somewhat more flexible distribution had to be used.
 We needed a distribution exhibiting rather strong kurtosis and the property to be both left and right skewed.
-To this end we use a generalization of the logistic distribution as defined in \cite{johnson}.
+To this end we use a generalization of the logistic distribution as defined in \citet{johnson}.
 Its probability density is given by:
@@
 \begin{eqnarray}
-& & l(\delta,\theta,\sigma, \pi) = ln(\delta) - ln(\sigma) - \frac{1}{\sigma} (\pi-\theta) - (\delta+1) ln(1+\exp^{-\frac{\pi-\theta}{\sigma}})
+l(\delta,\theta,\sigma, \pi) & = & \log(\delta) - \log(\sigma)   \\
+& - &  \frac{1}{\sigma} (\pi-\theta) - (\delta+1) \nonumber \\
+& * & \log (1+\exp^{-\frac{\pi-\theta}{\sigma}}) \nonumber
 \end{eqnarray}
@@
 \begin{eqnarray}
-& & \psi(\pi_i,\delta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta} = \frac{1}{\delta} - ln(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \\
-& & \psi(\pi_i,\theta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta} = \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
+\psi(\pi_i,\delta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta}  \\
+& = &  \frac{1}{\delta} - \log(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \nonumber \\
+\psi(\pi_i,\theta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta}  \\
+& = & \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}  \nonumber
 \end{eqnarray}
 
-\begin{equation}
-\begin{split}
-&& \psi(\pi_i,\sigma) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma} = -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) * \\
-&& \quad \frac{\frac{1}{\sigma^2}(\pi-\theta)\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
-\end{split}
-\end{equation}
+\begin{eqnarray}
+\psi(\pi_i,\sigma) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma}  \\
+& = & -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) \nonumber \\
+& * &  \quad \frac{\frac{1}{\sigma^2}(\pi-\theta)\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})} \nonumber
+\end{eqnarray}
 
-An enhancement to the allready existing R package "`strucchange"', which currently does not support a generalized logistic distribution (GL) is provided with this paper.
+An enhancement to the already existing R package "`strucchange"', which currently does not support a generalized logistic distribution (GL) is provided with this paper.
 The asymptotic testing theory still holds for this generalization.
@@
 The first part of the results we are going to present here, i.e.
-the tests and the graphical illustration of the empirical fluctuation process can be found in \cite{z07}.
-the second part of the results - the dating procedure and the illustration of the denisties fitted for the subsamples (divided by the breaks) - ca be found in \cite{z10}.
+the tests and the graphical illustration of the empirical fluctuation process can be found in \citet{z07}.
+the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - ca be found in \citet{z10}.
 
@@
 
-\subsubsection{Cramer-von Mises statistic - Nyblom Hansen Test}
+\subsubsection{Cramér-von Mises statistic - \\ Nyblom Hansen Test}
 
-We use the Cramér von Mises type test as given in \cite{z07}. The test statistic is given by:
+We use the Cramér von Mises type test as given in \citet{z07}. The test statistic is given by:
 
@@
 
-i.e.
-first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic.
-This can also be shown graphically as developed in
-%[wann war die Erstpublikation der graphischen Methode?].
+i.e. first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically.
 
-%\subsubsection{Kolmogorov-Smirnov Test}
-%
-%As a goodness of fit test, we use the Kolmogorov Smirnov Test for as given in \cite{massey}.
+\subsubsection{The $\chi^2$ Test}
+
+The main problem is to determine the number of classes k (data is grouped into k classes where we then calculate the difference between observed and expected frequencies).
+In a continuous distribution case, there are no natural boundaries.
+As given in \citet{kendal} the maximum likelihood estimation of the parameters is not a big problem if k is large enough.
+The classes are taken to cover equal ranges of the variate, which can be done easily once k is determined.
+This procedure was advocated by Mann and Gumberl.
+One rule would be to let k be equal to  $3.765(N-1)^{2/5}$ for a 5\% significance level, when using intervals with equal probability under $H_0$, as given in \citet{boreo}.
+We use this rule with a range of [3,2*k-3] for the tests and report any pvalues below 0.05.
 
@@
 
-Another thing we wish to look at is the change in the moments of the distribution after a break occured and whether or not the subsample fit of the GL-distribution is prefereable to another.
-Allthough our interest centers on the variance of the respective inflation rate, skewness should not alltogether be ignored.
-If we  observe - for example - a change in skewness from positive to negative from one regime to another, we could conclude that now we will observe higher values in the inflation rate since it shifted.
+Another thing we wish to look at is the change in the moments of the distribution after a break occurred and whether or not the subsample fit of the GL-distribution is preferable to another.
+Although our interest centers on the variance of the respective inflation rate, skewness should not altogether be ignored.
+If we  observe - for example - a change in skewness from positive to negative from one regime to another, we could conclude that now we will observe higher values in the inflation rate since its density shifted to the right.
 
-\newpage
 \subsection{Expected Results}
@@
 Following the Delors Plan with his 3 stages, we have: stage I (1990-1994), stage II (1995-1998) and stage III (1999-2002) which ended with the introduction of the Euro as legal tender.
-If the EMU had a significant effect upon inflation rate volatility we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular "`Teuro"' argument (i.e.
-the claim that the Euro lead to a significant increase in prices following the year of its introduction).
-[Teuro besser formulieren, gibt es dafür einen internationalen Term?]
-
-%
-%Geschichte für Interpretation 
-%- verwende Daten seit 1990
-%- 1. Stufe (ab 1990) – Vorbereitung durch Liberalisierung Geld- und Kapitalverkehr
-%- 2. Stufe (ab 1994) – Konvergenz:
-%- 3. Stufe (ab 1998) – Übergang zum Euro:
-%- Euro Währung Buchgeld seit 1.1.1999
-%- Euro Währung Bargeld seit 1.1.2002?
+If the EMU had a significant effect upon inflation rate volatility we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of EURO coins and paper money led to a considerable price increase.
 
@@
 
-The results presented here are for the countries within the Euro--zone (Austria, Belgium, Finland, France, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).
-Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due to data scarcity.
-The only ERM~II country included is Denmark, Latvia and Lithunia are excluded due to lack of data.
+The results presented here are for the countries within the Euro--zone (Austria, Belgium, Estonia, Finland, France, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).
+Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.
+The only ERM~II country included is Denmark--- Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.
 The other EU countries included are the Czech Republic, Hungary, Poland, Sweden and the United Kingdom.
-Of these countries Sweden and the UK opted out of the Euro, whereas the other countries are opted for a future adoption of the Euro.
+Of these countries Sweden and the UK opted against the Euro, whereas the other countries are in  for a future adoption of the Euro.
 
-% http://www.euro-anwaerter.de/anwaerter/daenemark.html
+% http://en.wikipedia.org/wiki/European_Exchange_Rate_Mechanism#Replacement_with_the_euro_and_ERM_II
+% http://en.wikipedia.org/wiki/Eurozone
 
-[Estonia should be included]
+\begin{table}[h!tp]\scriptsize
+\begin{center}
+\begin{tabular}{llll}
+\hline
+Country        & Dates  & \multicolumn{2}{l}{Breakpoints}  \\ \hline
+Austria        & 1999--2002   	& Sep 2007 &          \\
+Belgium        & 1999--2002   	& Dec 1999 &          \\
+Czech Republic & no--no		 			& Jul 1998 &          \\
+Denmark        & 1999--no 			& Jun 2000 &          \\
+Estonia        & 2004--2011   	& Mar 1998 &          \\
+Finland        & 1999--2002   	& none     &          \\
+France         & 1999--2002   	& Dec 2004 &          \\
+Germany        & 1999--2002   	& May 2000 & Dec 2004 \\
+Greece         & 2001--2002   	& none     &          \\
+Hungary        & no--no       	& May 1998 &          \\
+Ireland        & 1999--2002   	& Mar 2008 &          \\
+Italy          & 1999--2002   	& May 1996 & Dec 2000 \\
+Luxembourg     & 1999--2002   	& Dec 1998 &          \\
+Netherlands    & 1999--2002   	& none     &          \\
+Poland         & no--no      		& May 2001 &          \\
+Portugal       & 1999--2002   	& Jul 1992 & Mar 2004 \\
+Slovakia       & 2005--2009   & Apr 1997 & Feb 2004 \\
+Slovenia       & 2004--2007   & Jul 2003 &          \\
+Spain          & 1999--2002   & May 1996 & Dec 2000 \\
+Sweden         & no--no      & Jan 1993 &          \\
+United Kingdom & no--no       & Apr 1992 &          \\ \hline
+\end{tabular}
+\caption{\label{tab:breakpoints} Dating of break points.}
+\subcaption{First date: entry to ERM~II, second date: EURO introduction.}
+\end{center}
+\end{table}
 
-\begin{longtable}{|l|p{4cm}|p{3cm}|p{3cm}|}
-\hline
-{\bf Country} & {\bf 1. breakpoint}  & {\bf 2. breakpoint  \\
-\hline
-Austria (EURO) & Sep. 2007 & none\\
-\hline
-Belgium (EURO) & Dec. 1999 & none  \\
-\hline
-Czech Republic &  Jul. 1998  & none \\
-\hline
-Denmark (ERM~II)& Jun. 2000 & none \\
-\hline
-Finland (EURO) & none & none  \\
-\hline
-France (EURO) & Dec. 2004  & none\\
-\hline
-Germany (EURO) & May 2000 & Dec. 2004 \\
-\hline
-Greece (EURO) & none & none \\
-\hline
-Hungary & May 1998 & none\\
-\hline
-Ireland (EURO) & Mar. 2008 & none\\
-\hline
-Italy (EURO) & May 1996 &  Dec. 2000  \\
-\hline
-Luxembourg (EURO) & Dec. 1998 & none\\
-\hline
-Netherlands  (EURO) & none & none \\
-\hline
-Poland & May 2001 & none\\
-\hline
-Portugal (EURO) & Jul. 1992 & Mar. 2004\\
-\hline
-Slovakia (EURO) & Apr. 1997 & Feb. 2004 \\
-\hline
-Slovenia (EURO) & Jul. 2003 & none\\
-\hline
-Spain (EURO) & May 1996  & Dec. 2000\\
-\hline
-Sweden & Jan. 1993 & none\\
-\hline
-United Kingdom & Apr. 1992 & none\\
-\hline
-\end{longtable}
+\normalsize
 
-On this basis we trace out some groups of countries.
-The two biggest south european countries Italy and Spain form one group with identical break dates.
-Another group may be formed of the United Kingdom and Denmark.
-Another of the eastern european countries Czech Repbulic, Hungary and Poland.
-Another one of the eastern european countries that allready adopted the Euro like Slovakia and Slovenia.
-Another one of Germany and France.
-And Belgium and Luxembourg, reflecting the fact that Luxembourg and Belgium used their currencies as legal tender in one another.
+On this basis we can trace out 6 groups of countries with rather distinct behavior.
+Group 1 consists of Italy and Spain (both very similar) and possibly Belgium, Luxembourg and Portugal.
+Group 2 is an Eastern European group comprising Poland, the Czech Republic, Slovenia, Slovakia and Hungary.
+Group 3 is a Non-ERM~II member group including Sweden and the United Kingdom.
+The Netherlands, Greece and Finland build group 4 which exhibits no changes.
+Group 5 is composed of France, Germany, Austria and Denmark.
+Ireland stands on its own and is the only country which changed its inflation dynamics as an aftermarch of the financial crises of 2008.
+\\
+What we can observe for almost all the EURO countries is a increase in variance and mean, with the exception of Ireland.
+Since inflation volatility increased in all countries, we find no proof of a common claim that inflation uncertainty may increase in countries that have a smaller influence on ECB policy.
+\\
 
+%The AR distances table, measuring the similarity of short--run inflation dynamics as in \citet{palomba} more or less supports this group building, although it is much more accurate.
+The similarity of the structural breaks thus is interesting topic for future research.
+
+%give the 3 most similar in AR distances: in Palomba Paper (table 2)
 %
-%Some History for easier interpretation:
-%
-%% this one from history-emu.pdf
-%
-%eleven countries had met the Maastricht Criteria by 1997, allowing the
-%European monetary union to be launched in 1999. Greece was further behind in
-%meeting the criteria and only joined the monetary union in 2001.
-%First, entry into the euro was a much bigger structural shock for "peripheral"
-%member countries such as Greece, Ireland, Portugal, and Spain than for "core"
-%member countries like Germany and France. In particular, while all euro member
-%countries have enjoyed lower real interest rates under monetary union, the relative
-%decline was much greater for the countries on the periphery.
-%Second, membership in a monetary union can amplify the asymmetric impact
-%of certain shocks. A common nominal interest rate implies that persistent differences
-%in national inflation rates translate into differences in real interest rates across member countries: countries with relatively higher medium-term inflation
-%enjoy lower real interest rates than those with below-average inflation-stimulating
-%demand, credit growth, and housing markets in the former group
-%The Netherlands has lived though a
-%boom-bust cycle since the euro started, with a credit boom and high inflation in
-%1999-2001 followed by a contraction in economic activity that has reversed some of
-%the country's real appreciation against partner countries.
-%Of course, the destabilizing features of European monetary union for the
-%member countries do not mean that currency union has led to a net increase in
-%macroeconomic instability. The European Central Bank has been successful in
-%anchoring medium-term inflation expectations across the euro area at an annual
-%rate around 2 percent. At least for some countries, the attainment of such stability
-%outside the euro framework may have been a more costly process.
+%Italy: Greece, France, Portugal
+%Poland: Ireland, Spain, UK
+%Sweden: Denmark, Netherlands, UK
+%Netherlands: Sweden, UK, Portugal
+%France: Italy, Greece, Belgium
 %
 %
-%\section{Results}
-%
-%
-%% in money-growth regime switchin:
-%Germany: early 90ies: low inflation/ mid-end 90ies: high / and afterwards low
-%UK: low regime
-%
-%
-%just as a help
-%
-% wenn ma was siat dass as co-movement git: \cite{vandi} "`euro area countries move from essentially idiosyncratic inflation to co-movement in the mid-1980s"' (p.
-2); "`price shocks in one country not transmitted to another"'; "`correlations remain stable until 1999 but then jumps sharpely"'; (2009-mwp-dijk-internationalinfl)
-%
-%wenn was passiert: zitieren von Carporale: in their estimate \cite{cap} find that each country has at least one of the parameters of their variance estimates significantly different from zero, which is a clear indication of change in volatility as well.
-%
-%falls Volatilität erhöht wird, verweise auf die Debatte in: $2004-wp-minford-ukjoiningemu$, die von einer Erhöhung der britischen Inflationsrate sprechen, sollte UK der EMU beitreten.
-%
-%falls kleine Länder auffälliges Verhalten zeigen: $2008-wp-arnold-inflationuncertainty$, "`This suggests that within EMU, inflation uncertainty may increase in countries that have a smaller influence on ECB policy."' Ihre Begründung: "`For economic agents in small regions which are hit by positive inflation shocks, it will be uncertain how and when the increase
-%in inflation will be reversed.
-This results from imperfect information on the macroeconomic adjustment mechanism combined with the absence of a central bank dedicated to stabilizing their regional inflation rate.
-This does not hold true for the larger regions in the union, who have a largerweight in aggregate inflation and may therefore count on the price stability objective of the union’s central bank."'
-%vergleichen mit den Ergebnissen auf S.343 (country size and inflation uncertainty).
-%
-%berücksichtige die DM Zone (Deutschland, Österreich, Luxemburg, Belgien und Holland), eventuell ähnlich
-%
-%Caporale (ausgedruckt im Diss-Aktuell Ordner) finds that in the post-Euro area there are still significant inflation differentials across countries. 
-%
-%Wenns grosse Unterschiede zwischen den EMU Ländern gibt: ba Caporale stoht: "`the ECb itself admits that monetary policy can only affect price developements in the Euro area as a whole and cannot influence inflation differentials across regions (The Monetary Policy of the ECB, 2004)."' Und dort findet sich auch: "`some short-term volatility is inevitable given the fact that monetary policy can only affect prices wiht long and uncertain lags, hence the focus of the ECB on medium-term price stabilisation."'
-%
-%Um den Anstieg in der Volatilität zu verteidigen: Caporale finds that steady state uncertainty has increased in eight out of twelve countries (p.966) (ie.
-Germany, France, Italy, Spain, Ireland, Finland, Luxembourg and Austria).
-%
-%Zum Abgleich für Break dates: Caporale Table 5.
-%
-%
-%Wenn Ergebnisse da sind: abgleichen bei Caporale mit seiner Fig. 2, Fig. 3
-%
-%falls Österreich und Spanien komisch: Wilfing findet, dass Schilling und Peseta nicht zu ihrer fixen Parität konvergiert sind
-%
-%falls man grosse Unterschiede in der Volatilität sehen würde: $2008-wp-campolmi-laborinflvol$ point out that this can be explained by country specific labor market institutions.
-%
-% morana paper: S.10 untersucht convergence, je nachdem, was sich findet herausfinden, was wie ähnlich ist
-%
-% bei grossen Unterschieden: Busetti S.6: 3 Clustergruppen: Ger, Fra, Bel, Aut, Fin; Spa, Gre, Por, Ireland; Ned und It
-%
-%
-%%%%% still some thought necessary for that
-%
-%% OIL: peaked in 4.7.2008 at 144, fell until 25.12.2008 to 38, afterwards increases again but at slower rate
-%% http://markets.ftd.de/commodities/factsheet_overview.html?ID_NOTATION=4062566
+%either support or not support group 1
+%AR distance: the lower the more similar
 %
 
-\subsection{Austria}
-%% the information is saved in a folder called "`Erläuterungen"' which has the country name
-%% http://www.statistik.at/web_de/statistiken/preise/verbraucherpreisindex_vpi_hvpi/index.html
-In 2007 we saw a sharp rise in the yearly inflation rate (from 2 to almost 3.5\% in December 2007).
-This rise was mostly due to oil, which increased (starting in Sept.
-2007 from -4\% to almost 13\% plus on a yearly basis a rise in oil prices from 70 to 94 Dollars.
-This was further amplified by a  rise of mineral taxes.
-Also the Statistic Austria official documents report a significant rise in inflation in the fourth quarter of 2007.
+\section{Conlcusion}
 
-\subsection{Belgium}
-%% http://www.indexmundi.com/belgium/inflation_rate_%28consumer_prices%29.html
-There was a  very sharp fall in yearly inflation in this time (from 2\% in December 1999 to 0.2\% in January 2000).
-Volatility  a very clear increase in volatility; from 1999 to 2000 we had a 137\% increase in the yearly inflation rate.
-Economic reason?
-
-\subsection{Czech Republic}
-%% http://www.plzen.czso.cz/eng/redakce.nsf/i/inflation_rate
-%% http://en.wikipedia.org/wiki/Economy_of_the_Czech_Republic#1995-2000
-By the end of 1998 we see a strong drop in yearly inflation rates.
-Economic reason: political and financial crisis in 1997, government spending cut, growth dropped, pegged currency forced into floating system, lead to more stabilisation.
-% http://www.allbusiness.com/finance/302725-1.html
-%In terms of economic performance in eastern Europe, 1997 can be best described as the year for the return to realism, or, perhaps, the return to fundamentals. 
-% recession until 1998
-
-\subsection{Denmark}
-Some change in yearly inflation rate; strong increase in variance; Economic reason?
-
-\subsection{France}
-Nothing visible in yearly inflation; Economic reason?
-
-\subsection{Germany}
-Strong increase in mean and variance; Economic reason: strong increases in oil price, not enough. Could be traced back to the poor economic performance of Germany since 2000.
-
-\subsection{Greece}
-There is severe doubt about the credibility greek rate (looks almost like a uniform distribution); extremely high variance, almost 50\% above the next hightes (Belgium).
-
-\subsection{Hungary}
-Yearly inflation rate sharply decreasing till 1999; perhaps same explanation as for Czech republic, import restrictions have been lifted, trade liberalization; disinflationary impulses 
-% (http://findarticles.com/p/articles/mi_m4456/is_n63/ai_20792401/)
-
-\subsection{Ireland}
-In recession since second quarter of 2008. The European country most severely hit by 
-% (http://en.wikipedia.org/wiki/Economy_of_the_Republic_of_Ireland)
-
-\subsection{Italy}
-Drop in yearly inflation rate towards lower mean around 1996; in December 2000 nothing really visible; in 2000 beginning of a much more volatile regime.
-Economic reasons: in November 1996 Italy rejoined the european monetary system (left it in 1992); in September 1996, the new Cabinet announced Italy’s decision to take part in the Monetary Union.
-After 2000 Italy entered an econmic crisis [auf Auffälligkeiten eingehen]
-%(http://users.ox.ac.uk/~hine/seminarpapers/Visco.htm)
-%in 1999 italy qualifies for EMU
-%strong economic growth from 1996-2001
-%2000: economic crisis
-
-\subsection{Luxembourg}
-In January 1999 deflation with almost 2\%; strongly increased volatility; Economic reasons: creation of the Banque Centrale du Luxembourg (1998), founded 22.4.1998 :)
-
-\subsection{Poland}
-Decrease in yearly inflation; Economic reasons: growth has slowed considerably in Q2 2001; or interpret it just as a transition to a more stable path
-%(http://www.knowledgerush.com/kr/encyclopedia/Economy_of_Poland/);
-
-\subsection{Portugal}
-Decrease in yearly inflation rate from 1990-2000; after 1992 or so yearly inflation stabilized; Economic reasons: following a transition period lasting until 1992; so this marks the end of the transition period
-% (http://www.voyagesphotosmanu.com/portugal_economy.html)
-
-\subsection{Slovakia}
-Economic reasons?
-
-\subsection{Slovenia}
-Economic reasons: EU member in 2004 (perhaps some sign of pre-EU convergenc); mean and volatility fell afterwards; Government announced it plans to meet Maastricht Criteria; In 2003, several measures were introduced to make both portfolio and direct investments easier and more transparent in Slovenia and to conduct many financial operations, including banking, securities brokering, and undertaking various credit transactions.
-In January 2007 Slovenia adopted the euro as the national currency
-% (http://www.historycentral.com/nationbynation/Slovenia/Economy.html);
-
-\subsection{Spain}
-Extreme rise in inflation; Economic reasons?
-
-\subsection{Sweden}
-Until 1993 sharp decline in yearly inflation, giving away to sharp rise until 1994, then almost stable; Economic reasons: 1992 there was a run on the currency, the central bank briefly jacking up interest to 500\% in an unsuccessful effort to defend the currency's fixed exchange rate; had a crisis of the 1990s; decided in 1997 against euro introduction; had a banking crises at that time and succesfully ended it in 1993
-
-\subsection{United Kingdom}
-Sharp decline in yearly inflation rate; recession  hit the British economy at the start of the 1990s, beginning in the summer of 1990 and lasting until the end of 1992
-
-%\section{Ideas, Suggestions, Questions}
-%
-%Is there a method to get the dates in $"`plot(aut_gefp, functional=meanL2BB)"'$ (ie. return the date when we have the maximum value (or the two or more maximum values))? \\
-%How to get more focus on variance ? \\
-%How to force more than 5 breaks, so we end up with something as a "`smooth"' variance estimation. Good for graphical interpretation in the cases where there is no clear cut break. \\
-%Use a GARCH model with a GL density, if considered useful \\
-%If we find a break, how can we be sure that its the EMU whos responsible?
-The case if we find nothing is simpler, we can state that EMU had no effect (because otherwise we would see something, and the argument that EMU would be counteracted by something else is very very difficult to win) \\
-%Goodness of fit test could produce a graph showing the fit in cdf and ecdf and some stars showing significance or a message.... 
-%
-%
-%
-%
-
-% something more about the reason to use GL
-
-% Book Johnson 95 on logistic distribution
-%
-%- income distributions, otherwise often used as growth models, or rainfall models
-%- definition of the distribution Abberger uses here is a generalization of 23.7 in Johnson 95 (shape ist halt hier zustätzlich noch dabei)
-%- longer tails than normal distribution, nonsymmetric (which is a clear net gain)
-%- more peaked in the center (net gain)
-%- the things above apply to the logistic distribution, not implicitely the GL
-
-% searchterm for these ``generalized logistic distribution''
-% found via Google Scholars
-%- in 2005-kjm-olp-gl2 it is shown, that the distribution used here is; cdf is given by: $1-\frac{\exp{-bx}}{(1+\exp{-x})^b}$ (given in this paper)
-%- 2005-spl-won-parametersarmitgld: here the GL is used (they find for Autoregressions that mostly data simply is not normal or student, so they use the GL-distribution (which is only somewhat slightly differently specified)
-%- 2000-ael-har-stockreturns: with almost the same reasoning as above used to model stock exchange data (not exactly gl as here but can be simplified to GL)
-%- 1999-cis-won-asyminnogl: almost the same as the 2005 paper (won.. is also the author :)
-%- 2007-tejof-kon-vatriskgstock: same use
-%- 2001-joae-wa-garchgl: similar (not exactly using gl but very similar)
-%
-%
-%% Econ Papers
-%- often in financial return data (which has high peakes)
-%- in gene expression data (since gl hast longer tails)
-
-% eu database
-% ideas
-% citeceer
-% econometric society
-% questia
-% ecb
-% jstor
-% springerlink
-% ebsco
-% wiley
-% sciencedirect
-% informaworld
-
-% Abberger Publikation: Abberger, K., Heiler, S.
-	Simultaneous estimation of parameters for a generalized logistic distribution and application to time series models 	2000 	84 	041-049 	A Allgemeines Statistisches Archiv/Journal of the German Statistical Society
-
-%- in 1987-eco-bat-inflationexpectation: the expected inflation is modelled as coming from a logistic distribution
-
-%
-%\subsection{Austria}
-%
-%The Double max test for Austria is insignificant, but the Cramér-von Mises test and the Supremum of LM statistcs test are both significant.
-This can be interpreted as a sign, that there was some change in the series as a whole, but that we are unsure wheter this happend with respect to location, scale or shape.
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{aut-meanL2BB.png}
-%\caption{Cramér-von Mises test for Austria}
-%\end{figure}
-%
-%Regarding the figure, we have two clear breaks, one in 1999 and another in 2008.
-[dating very unsure] The fxregimes procedure gives no breakdate.
-To give at least some intuition about the changes in variance [actually this is just scale, but can be adapted quickly] we forced 5 breaks into the Austrian inflation rates.
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{aut-morebreak.png}
-%\caption{Variance Estimation in Austrian Inflation Rates with 5 Breaks}
-%\end{figure}
-%
-%Only with some phantasy we can se some slight V shape (a decline in inflation volatility up to 2000, then an increase). 
-%
-%
-%
-%
-%\subsection{Belgium}
-%
-%Regarding the belgian inflation rates, all tests clearly indicate a break. 
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{bel-meanL2BB.png}
-%\caption{Cramér-von Mises test for Belgium}
-%\end{figure}
-%
-%The fxregimes procedure indicates a breakpoint in December 2000.
-%
-%%The impact of skewness: negative skew (rechtssteil), ie. higher values of y are more probable; positive skew/right skew (linkssteil): lower values of y are more probable
-%%so if it is the case like 6-2-1 we can say that the distribution gets less of a positive skew meaning a lower inflation on average (abklären, sobald das mit den Momenten hinhaut)
-%%
-%
-%\subsection{Italy}
-%
-%In the italian case, there is clear evidence for two breaks. 
-%
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{ita-meanL2BB.png}
-%\caption{Cramér-von Mises test for Italy}
-%\end{figure}
-%
-%If we compare the density of a whole and the denisties of the subsets of the data, we see a clear picture:
-%
-%\begin{figure}[h!tp]
-%\includegraphics[width=80mm]{ita-split.png}
-%\caption{Parameters and Densities for subsamples}
-%\end{figure}
-%
-%
-%As we can see regarding the moments of the subsamples, we see a clear increase in variance (0.04, 0.01, 0.32), some smaller increase in the mean (0.41, 0.16, 0.18)  and a interesting increas in decrease in skewness (0.96, 0.71, $-$0.26), which means that lower values of the inflation rate are quickly becoming less probable.
-%
-%
-
-\setstretch{1,3}
-
-\newpage
-\section{Literature}
-
-\renewcommand*{\refname}{}
+\section{References}
+\bibliographystyle{elsarticle-harv}
 \bibliography{papers}
-\begin{thebibliography}{99}
-\bibliographystyle{apsr}
 

### 2010-07-23 pkg/paper/Draft/draft-Z-W.tex
I couldn't compile the paper at all, even with the latest Elsevier styles. Also, the .bib was broken in numerous places.
I changed the style to JSS and fixed the worst problems so that it can at
least be compiled...



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@29 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
-\documentclass[a4paper, 5p, preprint]{elsarticle}
-\usepackage{graphicx,a4wide,color,hyperref}
+\documentclass[nojss]{jss}
 \usepackage[latin1]{inputenc}
-\usepackage[authoryear, round]{natbib}
-\bibpunct{(}{)}{;}{a}{}{,}
-\pagestyle{myheadings}
-
-\biboptions{longnamesfirst,angle,semicolon}
 
@@
 
-\begin{document}
-
-\title{Still need a good title}
+\title{Something about Euro, EMU, and Inflation}
 
@@
 % 1. Widerlegung des Teuro Effektes
-% 2. Keine Veränderung d. EMU sichtbar; nach 2002 kein globaler Trend aller EMU Staaten Richtung Wechsel
-% 3. Nutzen zeigen, den die osteuropäischen Staaten haben, dadurch, dass sie dem EURO beitreten müssen  (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
+% 2. Keine Ver‰nderung d. EMU sichtbar; nach 2002 kein globaler Trend aller EMU Staaten Richtung Wechsel
+% 3. Nutzen zeigen, den die osteurop‰ischen Staaten haben, dadurch, dass sie dem EURO beitreten m¸ssen  (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
 % 	 so etwa in Slowenien und der Slovakei ein Bruch bevor diese Staaten dem ERMII/EURO beitreten
-% 4. Finanzkrise: schwieriger, ausser Irland keine Änderung > 2008
+% 4. Finanzkrise: schwieriger, ausser Irland keine ƒnderung > 2008
 
-\author[rvt]{A.˜Zeileis}
-\author[focal]{T.˜Windberger}
-\address[rvt]{University of Innsbruck, Institute for Statistics, Universitätsstraße 15, 6020 Innsbruck, Austria}
-\address[focal]{University of Innsbruck, Austria}
+\author{Thomas Windberger\\Universit\"at Innsbruck \And
+        Achim Zeileis\\Universit\"at Innsbruck
+}
+\Plainauthor{Thomas Windberger, Achim Zeileis}
 
-\date{}
-\maketitle
+\Abstract{
+  The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)
+  on changing the volatility of the inflation rates in its Member States. 
+}
 
-\begin{abstract}
-The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)  on changing the volatility of the inflation rates in its Member States. 
-\end{abstract}
+\Keywords{inflation rate, structural break, EMU}
 
-\begin{keyword}
-inflation rate \sep structural break \sep EMU 
-\end{keyword}
+\Address{
+  Thomas Windberger, Achim Zeileis\\
+  Department of Statistics\\
+  Universit\"at Innsbruck\\
+  Universit\"atsstraﬂe 15\\
+  6020 Innsbruck, Austria\\
+  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org}
+}
+
+\begin{document}
 
@@
 
-The ECB defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of below 2\%'' \citet{castel}.
-\citet{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22).
+The European Central Bank (ECB) defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of
+below 2\%'' \citep{castel}.
+\citet{emerson} make it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22).
 Thus to achieve a low inflation rate and low variability of inflation must be a key issue for the central bank.
 \\
-The question of interest centers around the way in which a countries decision to join the EMU changed its inflation rate dynamics.
+The question of interest centers around the way in which a countries decision to join the European Monetary Union (EMU) changed its inflation rate dynamics.
 There are a number of reasons why these should change indeed.
@@
 
-\section{Literature Overview}
+\section{Literature overview}
 
 Theory unfortunately is quite unsure about whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
-\citet{cooper} points out, that "a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a "central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation.
+\citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation.
 It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
@@
 
-Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot \log(HICP_{t}/HICP_{t-1})$.
+Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990--3.2010, so $x_t = 100\cdot \log(\mathit{HICP}_{t}/\mathit{HICP}_{t-1})$.
 Countries included are Austria,	Belgium,	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Spain,	Sweden,		United Kingdom and the	Euro area.
@@
 
-\subsection{A generalized logistic distribution (GL)}
+\subsection{A generalized logistic distribution}
 
-In econometric issues, the logistic distribution is often used in income distributions and growth models.
-Its use stems from the usefulness of its longer tails and its higher peak, which fits these problems somewhat better.
-The use of the generalized logistic distribution as we will use it in this paper is somewhat scarcer.
-\citet{won} uses a GL distribution in a regression model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  leptokurtic and severely left or right skewed.
+In econometric issues, the logistic distribution is often used in income distributions and growth models. Its use stems from the usefulness of its longer
+tails and its higher peak, which fits these problems somewhat better. The use of the generalized logistic distribution as we will use it in this paper is
+somewhat scarcer.
+\citet{won} uses a generalized logistic (GL) distribution in a regression model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  leptokurtic and severely left or right skewed.
 A similar GL distribution is also used in \citet{tolikas} who analyses extreme risk and value--at--risk in the German stock market, all though they don't use a shape parameter.
@@
 
-An enhancement to the already existing R package "`strucchange"', which currently does not support a generalized logistic distribution (GL) is provided with this paper.
+An enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a generalized logistic distribution (GL) is provided with this paper.
 The asymptotic testing theory still holds for this generalization.
 [Beweis] \\
-The first part of the results we are going to present here, i.e.
-the tests and the graphical illustration of the empirical fluctuation process can be found in \citet{z07}.
+The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process can be found in \citet{z07}.
 the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - ca be found in \citet{z10}.
@@
 
-\subsubsection{Cramér-von Mises statistic - \\ Nyblom Hansen Test}
+\subsubsection{Cram{\'e}r-von Mises statistic -- Nyblom-Hansen test}
 
-We use the Cramér von Mises type test as given in \citet{z07}. The test statistic is given by:
+We use the Cram{\'e}r von Mises type test as given in \citet{z07}. The test statistic is given by:
 
@@
 
-i.e. first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically.
+i.e., first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically.
 
-\subsubsection{The $\chi^2$ Test}
+\subsubsection[The chi-squared test]{The $\chi^2$ test}
 
@@
 
-\subsection{Expected Results}
+\subsection{Expected results}
 
@@
 
-\begin{table}[h!tp]\scriptsize
+\begin{table}[t!]
 \begin{center}
@@
 \end{tabular}
-\caption{\label{tab:breakpoints} Dating of break points.}
-\subcaption{First date: entry to ERM~II, second date: EURO introduction.}
+\caption{\label{tab:breakpoints} Dating of break points. First date: entry to ERM~II, second date: EURO introduction.}
 \end{center}
 \end{table}
-
-\normalsize
 
@@
 
-\section{References}
-\bibliographystyle{elsarticle-harv}
 \bibliography{papers}

### 2010-07-23 pkg/paper/Draft/draft-Z-W.tex
cleaned bib changes in gloglisfit: estimate moments with exp(par) due to log estimation in ML
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@30 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 This procedure was advocated by Mann and Gumberl.
-One rule would be to let k be equal to  $3.765(N-1)^{2/5}$ for a 5\% significance level, when using intervals with equal probability under $H_0$, as given in \citet{boreo}.
+One rule would be to let k be equal to  $3.765(N-1)^{2/5}$ for a 5\% significance level, when using intervals with equal probability under $H_0$, as given in \citet{boero}.
 We use this rule with a range of [3,2*k-3] for the tests and report any pvalues below 0.05.
@@
 
+\newpage
+
 On this basis we can trace out 6 groups of countries with rather distinct behavior.

### 2010-07-23 pkg/R/glogisfit.R

@@ -110,11 +110,12 @@ glogisfit.default <- function(x, weights = NULL,
   names(par) <- c("location", "scale", "shape")
 
   ## moments
+	## calculate exp(par)[2] and exp(par)[3] due to log estimation
   mom <- c(  
-    "mean"      = as.vector(par[1] + (digamma(par[3]) - digamma(1)) * par[2]),
-    "variance"  = as.vector((psigamma(par[3], deriv = 1) + psigamma(1, deriv = 1)) * par[2]^2),
-    "skewness"  = as.vector((psigamma(par[3], deriv = 2) - psigamma(1, deriv = 2)) / 
-                  (psigamma(par[3], deriv = 1) + psigamma(1, deriv = 1))^(3/2))
+    "mean"      = as.vector(par[1] + (digamma(exp(par)[3]) - digamma(1)) * exp(par)[2]),
+    "variance"  = as.vector((psigamma(exp(par)[3], deriv = 1) + psigamma(1, deriv = 1)) * exp(par)[2]^2),
+    "skewness"  = as.vector((psigamma(exp(par)[3], deriv = 2) - psigamma(1, deriv = 2)) / 
+                  (psigamma(exp(par)[3], deriv = 1) + psigamma(1, deriv = 1))^(3/2))
   )
 
   ## collect and return everything

### 2010-07-23 pkg/R/glogisfit.R
changes undone
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@31 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@ -110,12 +110,12 @@ glogisfit.default <- function(x, weights = NULL,
   names(par) <- c("location", "scale", "shape")
 
   ## moments
-	## calculate exp(par)[2] and exp(par)[3] due to log estimation
+	
   mom <- c(  
-    "mean"      = as.vector(par[1] + (digamma(exp(par)[3]) - digamma(1)) * exp(par)[2]),
-    "variance"  = as.vector((psigamma(exp(par)[3], deriv = 1) + psigamma(1, deriv = 1)) * exp(par)[2]^2),
-    "skewness"  = as.vector((psigamma(exp(par)[3], deriv = 2) - psigamma(1, deriv = 2)) / 
-                  (psigamma(exp(par)[3], deriv = 1) + psigamma(1, deriv = 1))^(3/2))
+    "mean"      = as.vector(par[1] + (digamma(par[3]) - digamma(1)) * par[2]),
+    "variance"  = as.vector((psigamma(par[3], deriv = 1) + psigamma(1, deriv = 1)) * par[2]^2),
+    "skewness"  = as.vector((psigamma(par[3], deriv = 2) - psigamma(1, deriv = 2)) / 
+                  (psigamma(par[3], deriv = 1) + psigamma(1, deriv = 1))^(3/2))
   )
 
   ## collect and return everything

### 2010-07-26 pkg/R/glogisfit.R
incorporated goodness-of-fit test
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@32 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@ -109,8 +109,7 @@ glogisfit.default <- function(x, weights = NULL,
   par[2:3] <- exp(par[2:3])
   names(par) <- c("location", "scale", "shape")
 
-  ## moments
-	
+  ## moments	
   mom <- c(  
     "mean"      = as.vector(par[1] + (digamma(par[3]) - digamma(1)) * par[2]),
     "variance"  = as.vector((psigamma(par[3], deriv = 1) + psigamma(1, deriv = 1)) * par[2]^2),
@@ -192,7 +191,7 @@ print.glogisfit <- function(x, digits = max(3, getOption("digits") - 3), ...)
   invisible(x)
 }
 
-summary.glogisfit <- function(object, log = TRUE, ...)
+summary.glogisfit <- function(object, log = TRUE, breaks = NULL, ...)
 {
   ## extend coefficient table
   cf <- coef(object, log = log)
@@ -204,6 +203,19 @@ summary.glogisfit <- function(object, log = TRUE, ...)
   
   ## number of iterations
   object$iterations <- as.vector(tail(na.omit(object$optim$count), 1))
+
+  ## chi-squared goodness-of-fit test
+  if(!is.null(object$x)) {
+    if(is.null(breaks)) breaks <- min(max(3, ceiling(length(object$x)/5) - 1), floor(3.765 * length(object$x)^(2/5)))
+    if(length(breaks) == 1) breaks <- qglogis(seq(0, 1, length = breaks + 1),
+      location = object$parameters[1], scale = object$parameters[2], shape = object$parameters[3])
+    breaks <- as.numeric(breaks)
+    ct <- chisq.test(table(cut(object$x, breaks = breaks, include.lowest = TRUE)),
+      p = diff(pglogis(breaks, location = object$parameters[1], scale = object$parameters[2], shape = object$parameters[3])))
+    ct$method <- "Chi-squared goodness-of-fit test"
+    ct$data.name <- "discretized observations"
+    object$chisq.test <- ct
+  }
   
   ## return
   class(object) <- "summary.glogisfit"
@@ -222,6 +234,10 @@ print.summary.glogisfit <- function(x, digits = max(3, getOption("digits") - 3),
   
     cat("\nLog-likelihood:", formatC(x$loglik, digits = digits),
       "on", sum(sapply(x$coefficients, NROW)), "Df")
[…]

### 2010-07-27 pkg/paper/Draft/draft-Z-W.tex
updated bib; language changes in draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@33 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 % 1. Widerlegung des Teuro Effektes
-% 2. Keine Ver‰nderung d. EMU sichtbar; nach 2002 kein globaler Trend aller EMU Staaten Richtung Wechsel
-% 3. Nutzen zeigen, den die osteurop‰ischen Staaten haben, dadurch, dass sie dem EURO beitreten m¸ssen  (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
+% 2. Keine Veränderung d. EMU sichtbar; nach 2002 kein globaler Trend aller EMU Staaten Richtung Wechsel
+% 3. Nutzen zeigen, den die osteuropäischen Staaten haben, dadurch, dass sie dem EURO beitreten müssen  (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
 % 	 so etwa in Slowenien und der Slovakei ein Bruch bevor diese Staaten dem ERMII/EURO beitreten
-% 4. Finanzkrise: schwieriger, ausser Irland keine ƒnderung > 2008
+% 4. Finanzkrise: schwieriger, ausser Irland keine Änderung > 2008
+
+% Loswerden der Working Papers
+% auf Zitat mit castelnuovo kann man verzichten, dass die EZB Preisstabilität so definiert ist kein Geheimenis und nichts, dass man mit Paper zitieren müsste
+% holemöller geht nicht weg, immer noch Discussion Paper (siehe sein lebenslauf)http://www.holtem.de/Dokumente/PublicCV-Holtemoeller.pdf
+% morana detto
 
@@
   Universit\"at Innsbruck\\
-  Universit\"atsstraﬂe 15\\
+  Universit\"atsstraße 15\\
   6020 Innsbruck, Austria\\
@@
 The European Central Bank (ECB) defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of
-below 2\%'' \citep{castel}.
-\citet{emerson} make it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22).
-Thus to achieve a low inflation rate and low variability of inflation must be a key issue for the central bank.
+below 2\%''.
+\citet{emerson} emphasizes the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism (p.~22).
 \\
-The question of interest centers around the way in which a countries decision to join the European Monetary Union (EMU) changed its inflation rate dynamics.
-There are a number of reasons why these should change indeed.
-Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria are likely to lead to a change at least in the mean and possible in the volatility of their respective inflation rates as well.
-This has indeed been the case for a number of EMU countries, like Spain, Italy and Portugal, reflecting their effort to meet the Maastricht Criteria for inflation rates.
-
-\section{Literature overview}
-
-Theory unfortunately is quite unsure about whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
+%Therefore a vital aim of every central bank is to achieve a low inflation rate with negligible variation. \\
+The question of interest centers around the way in which a countries decision to join the EMU changed its inflation rate dynamics. 
+%There are a number of reasons why these should have changed indeed.
+Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria were likely to lead to an alternation at least in the mean (since this was required for a number of countries) and possible in the volatility of their respective inflation rates as well.
+This has indeed been the case for a host of EMU countries, like Spain, Italy and Portugal, reflecting their effort to meet the Maastricht Criteria for inflation rates.
+\\
+% no literature overview
+Theory unfortunately is quite unsure  whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
 \citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation.
@@
 An interesting approach to this question is taken by \citet{holte}. 
-He creates a two country model for monetary policy analysis along the line of two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}.
-The inflation is modeled by a hybrid Phillips curve (NKPC) specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
-What he finds out via simulations of these different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
-But monetary policy in a monetary union does not explicitly stabilize the output gap and inflation rate in case of national economic shocks.
-The effects of joining on inflation rate variability depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-Due to the fact that not all of these parameters are known and that their interaction as well has to be guessed, theory has some troubles answering the question of this paper.
+%He creates a two country model for monetary policy analysis along the line of two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}.
+The inflation is modeled by a hybrid Phillips curve  specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
+What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union. \\
+%But monetary policy in a monetary union does not explicitly stabilize the output gap and inflation rate in case of national economic shocks. 
+The effects of joining a monetary union on inflation rate variability depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
+Due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, theory has some troubles establishing a sound theoretical model.
 \\
-\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation, its degree of persistence and both types of uncertainty.
-They use a time-varying model with a GARCH specification for the unconditional volatility of inflation and find some instability in the conditional volatility.
-\\
+\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
+%They use a time-varying model with a GARCH specification for the unconditional volatility of inflation and find some instability in the conditional volatility. \\
 In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
 They find that convergence in time of inflation dynamics was only partly observable.
-In a paper studying core inflation and using an aggregated Euro Area inflation rate, \citet{morana} finds three regimes (roughly 1980-1984, 1984-1993 and 1993-2000) governing the core inflation rate.
+In a paper studying core inflation and using an aggregated Euro Area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
 \\
-Inquiring into the convergence properties of inflation rates among countries of the EMU, \citet{busetti} find that from 1980-1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
+Inquiring into the convergence properties of inflation rates among countries of the EMU, \citet{busetti} find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
 \\
@@
 
-Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990--3.2010, so $x_t = 100\cdot \log(\mathit{HICP}_{t}/\mathit{HICP}_{t-1})$.
-Countries included are Austria,	Belgium,	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Spain,	Sweden,		United Kingdom and the	Euro area.
-The data are obtained from the OECD Statistics.
-\\
-The countries can be divided into three different groups: the EURO countries (Portugal, Spain, Italy, Greece, France, Belgium, Estonia\footnote{although Estonia enters in 2011}, Netherlands, Germany, Austria, Ireland, Slovenia, Luxembourg and Finland), EU members without ERM~II (Exchange Rate Mechanism): United Kingdom, Sweden, Poland, Czech Republic, Hungary.
+Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010.
+Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Spain,	Sweden, and	the	United Kingdom.
+The data are obtained from the OECD Statistics.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} \\
+The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia enters in 2011}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Spain and Slovenia)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members without ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
 Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
 
+\newpage
 \section{Methods}
 
-\subsection{A generalized logistic distribution}
+\subsection{A Generalized Logistic Distribution}
 
-In econometric issues, the logistic distribution is often used in income distributions and growth models. Its use stems from the usefulness of its longer
-tails and its higher peak, which fits these problems somewhat better. The use of the generalized logistic distribution as we will use it in this paper is
-somewhat scarcer.
-\citet{won} uses a generalized logistic (GL) distribution in a regression model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  leptokurtic and severely left or right skewed.
+In econometrics, the logistic distribution is often used in income distributions and growth models.
+This is due to  its longer tails and  higher peak, which fits these problems somewhat better.
+The use of the generalized logistic distribution as we will use it in this paper is rather rare.
+\citet{won} uses a GL distribution in a regression model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  leptokurtic and severely left or right skewed.
 A similar GL distribution is also used in \citet{tolikas} who analyses extreme risk and value--at--risk in the German stock market, all though they don't use a shape parameter.
-Regarding inflation rates, the GL distribution is --to our best knowledge-- only used in relationship with expected inflation.
+Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in relationship with expected inflation.
 \citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
 \\
-As already stated, we now apply the general framework, as developed in \citet{z07} to a more specific model, in this case by means of a GL distribution.
-Prior to that, we would like to give a short justification for our using of the GL distribution.
+We apply the general framework, as developed in \citet{z07}, to a more specific model, in this case by means of a GL distribution.
+Prior to that, we would like to give a short justification for the usage of the GL distribution.
 Regarding the data at hand, it was not possible to use the already existing method developed in \citet{z07}, since almost all inflation rates, with the notable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties.
@@
 We needed a distribution exhibiting rather strong kurtosis and the property to be both left and right skewed.
-To this end we use a generalization of the logistic distribution as defined in \citet{johnson}.
+To serve this purpose, we use a generalization of the logistic distribution as defined in \citet{johnson}.
 Its probability density is given by:
@@
 
-\subsubsection{Cram{\'e}r-von Mises statistic -- Nyblom-Hansen test}
+%\subsubsection{Cram{\'e}r-von Mises statistic -- Nyblom-Hansen test}
 
-We use the Cram{\'e}r von Mises type test as given in \citet{z07}. The test statistic is given by:
-
-\begin{eqnarray}
-& & \frac{1}{n} \sum_{i=1}^n \Vert efp(\frac{i}{n}) \Vert_2^2
-\end{eqnarray}
-
-i.e., first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically.
-
-\subsubsection[The chi-squared test]{The $\chi^2$ test}
-
-The main problem is to determine the number of classes k (data is grouped into k classes where we then calculate the difference between observed and expected frequencies).
+%We use the Cram{\'e}r von Mises type test as given in \citet{z07}. The test statistic is given by:
+%
+%\begin{eqnarray}
+%& & \frac{1}{n} \sum_{i=1}^n \Vert efp(\frac{i}{n}) \Vert_2^2
+%\end{eqnarray}
+%
+%i.e., first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically.
+%
+%
+%
+%\subsubsection[The chi-squared test]{The $\chi^2$ test}
+%
+%The main problem is to determine the number of classes k (data is grouped into k classes where we then calculate the difference between observed and expected frequencies).
 In a continuous distribution case, there are no natural boundaries.
-As given in \citet{kendal} the maximum likelihood estimation of the parameters is not a big problem if k is large enough.
+%As given in \citet{kendal} the maximum likelihood estimation of the parameters is not a big problem if k is large enough.
 The classes are taken to cover equal ranges of the variate, which can be done easily once k is determined.
@@
 
-\subsection{Densities}
-
-Another thing we wish to look at is the change in the moments of the distribution after a break occurred and whether or not the subsample fit of the GL-distribution is preferable to another.
+For the determination of the numbers of breakpoints, we use the Supremum of LM statistics test and we further investigate into the optimal number of breakpoints using the LWZ criterion, see \citet{z10} for further details.
+To get an idea about the goodness of fit of the GL distribution for a given subsample we also report values of the $\chi^2$ goodness of fit test.
+%
+%
+%\subsection{Densities}
+%
+%Another thing we wish to look at is the change in the moments of the distribution after a break occurred and whether or not the subsample fit of the GL-distribution is preferable to another.
 Although our interest centers on the variance of the respective inflation rate, skewness should not altogether be ignored.
@@
 
-\subsection{Expected results}
-
-To give an idea what we would expect, we give a short overview over the history of the EMU.
-Following the Delors Plan with his 3 stages, we have: stage I (1990-1994), stage II (1995-1998) and stage III (1999-2002) which ended with the introduction of the Euro as legal tender.
-If the EMU had a significant effect upon inflation rate volatility we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of EURO coins and paper money led to a considerable price increase.
-
+\newpage
 \section{Results}
 
-The results presented here are for the countries within the Euro--zone (Austria, Belgium, Estonia, Finland, France, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).
-Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.
-The only ERM~II country included is Denmark--- Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.
-The other EU countries included are the Czech Republic, Hungary, Poland, Sweden and the United Kingdom.
-Of these countries Sweden and the UK opted against the Euro, whereas the other countries are in  for a future adoption of the Euro.
-
+To help with the interpretation, we give a short overview over the history of the EMU.
+Following the Delors Plan with his 3 stages, we have: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender.
+If the EMU had a significant effect upon inflation rate volatility we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of Euro paper money led to a considerable price increase.
 % http://en.wikipedia.org/wiki/European_Exchange_Rate_Mechanism#Replacement_with_the_euro_and_ERM_II
@@
 
-\begin{table}[t!]
+\begin{table}[ht!p]
 \begin{center}
@@
 
-On this basis we can trace out 6 groups of countries with rather distinct behavior.
+On the basis of the results, we can trace out 6 groups of countries with rather distinct behavior.
 Group 1 consists of Italy and Spain (both very similar) and possibly Belgium, Luxembourg and Portugal.
@@
 Group 5 is composed of France, Germany, Austria and Denmark.
-Ireland stands on its own and is the only country which changed its inflation dynamics as an aftermarch of the financial crises of 2008.
+Ireland stands on its own and is the only country which changed its inflation dynamics as an after march of the financial crises of 2008.
 \\
-What we can observe for almost all the EURO countries is a increase in variance and mean, with the exception of Ireland.
+What we can observe for almost all the EURO countries is an increase in variance and mean, with the exception of Ireland.
 Since inflation volatility increased in all countries, we find no proof of a common claim that inflation uncertainty may increase in countries that have a smaller influence on ECB policy.
 \\
+Lets consider a view countries in broader detail to get an idea about what happened.
+In Austria we see evidence of a change in the inflation rate, when the old regime ended in September 2007.
+The new regime is much more volatile, with an anlmost threefold increase in variance.
+
+% sizes !!!
+
+\begin{figure}[ht!p]
+  \centering
+    \includegraphics{results-012.pdf}
+  \caption{Series and supLM test for Austria}
+  \label{aut}
+\end{figure}
+
+The reason for this is a strong increase in the yearly Austrian inflation rate following a heavy increase in oil price which was further amplified by a rise of minearl taxes.
+However, we fail to find any evidence for an incrase in mean or volatility of the Austrian rate following the introduction of the Euro paper money.
+\\
+Lets consider Slovenia as the first Eastern European country that introduced the Euro.
+In August 2003 the new regime started with a much lower mean value (only a third now of its previous value) and interestingly enough, a higher volatility.
+
+\begin{figure}[ht!p]
+  \centering
+    \includegraphics{results-131.pdf}
+  \caption{Series and supLM test for Slovenia}
+  \label{aut}
+\end{figure}
+
+%
+%The LWZ criteria favors one break. 
+%
+%
+%\begin{figure}[ht!p]
+%  \centering
+%    \includegraphics{results-132.pdf}
+%  \caption{LWZ and Negative Log-Likelihood for Break Point Selection}
+%  \label{aut}
+%\end{figure}
+
+Slovenia is a good example of the benefits of an early Euro adoption.
+Its efforts to meet the criterias as set forth in the Maastricht treaty were successfull, actually Slovenia reached them in 2005.
+So from 2003 onwards inflation was lower but with a much higher variance.
+Regarding the dating in 2003, we have to keep in mind that this was the year where most of the financial reforms were introduced.
+Slovenia seems to be the only case where the introduction of the Euro was followed by an increase in yearly inflation, which faded away following the financial crises and its deflationary effects.
+\\
+As a last example we focus on Spain.
 

### 2010-08-13 pkg/paper/Draft/draft-Z-W.tex
some improvements and fixups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@34 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \usepackage[latin1]{inputenc}
+\graphicspath{{Figures/}}
 
-\graphicspath{{Figures/}}
+%% for internal use
+\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
+\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
 
@@
 % 2. Keine Veränderung d. EMU sichtbar; nach 2002 kein globaler Trend aller EMU Staaten Richtung Wechsel
-% 3. Nutzen zeigen, den die osteuropäischen Staaten haben, dadurch, dass sie dem EURO beitreten müssen  (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
+% 3. Nutzen zeigen, den die osteuropäischen Staaten haben, dadurch, dass sie dem EURO beitreten müssen
+%    (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
 %    so etwa in Slowenien und der Slovakei ein Bruch bevor diese Staaten dem ERMII/EURO beitreten
@@
 % Loswerden der Working Papers
-% auf Zitat mit castelnuovo kann man verzichten, dass die EZB Preisstabilität so definiert ist kein Geheimenis und nichts, dass man mit Paper zitieren müsste
-% holemöller geht nicht weg, immer noch Discussion Paper (siehe sein lebenslauf)http://www.holtem.de/Dokumente/PublicCV-Holtemoeller.pdf
+% auf Zitat mit castelnuovo kann man verzichten, dass die EZB Preisstabilität so
+% definiert ist kein Geheimenis und nichts, dass man mit Paper zitieren müsste
+% Z: Bitte das offizielle ECB-Statement zitieren
+
+% holemöller geht nicht weg, immer noch Discussion Paper (siehe sein
+% lebenslauf) http://www.holtem.de/Dokumente/PublicCV-Holtemoeller.pdf
 % morana detto
+% Z: Das ist ok, ein paar Working Papers koennen schon bleiben.
+%    Aber: @TechReport nicht @Article benutzen (schon erledigt)
 
@@
 \Abstract{
-  The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)
-  on changing the volatility of the inflation rates in its Member States. 
+  The aim of this paper is to shed some light on the effect of the economic and monetary union
+  on changing the volatility of the inflation rates in the member states of the European Union. 
 }
@@
 
-The European Central Bank (ECB) defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of
-below 2\%''.
-\citet{emerson} emphasizes the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism (p.~22).
-\\
-%Therefore a vital aim of every central bank is to achieve a low inflation rate with negligible variation. \\
-The question of interest centers around the way in which a countries decision to join the EMU changed its inflation rate dynamics. 
-%There are a number of reasons why these should have changed indeed.
-Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria were likely to lead to an alternation at least in the mean (since this was required for a number of countries) and possible in the volatility of their respective inflation rates as well.
-This has indeed been the case for a host of EMU countries, like Spain, Italy and Portugal, reflecting their effort to meet the Maastricht Criteria for inflation rates.
-\\
+The European Central Bank (ECB) defines price stability ``as a year-on-year
+increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of
+below 2\%''. \fixme{some official ECB reference needed for quote}
+\citet{emerson} emphasize the fact that a high inflation rate is
+also more variable and uncertain and thus causes more relative price
+variability, leading to a less efficient price mechanism (p.~22).
+% Therefore a vital aim of every central bank is to achieve a low inflation rate
+% with negligible variation.
+The question of interest centers around the way in which a country's decision to
+join the economic and monetary union (EMU) \readme{Note that the E in EMU stands for economic, not European}
+changed its inflation rate dynamics.
+
+% There are a number of reasons why these should have changed indeed. Given
+% that a country experienced quite volatile inflation rates, its efforts to meet the
+% convergence criteria were likely to lead to an alternation at least in the mean
+% (since this was required for a number of countries) and possible in the
+% volatility of their respective inflation rates as well. This has indeed been the
+% case for a host of EMU countries, like Spain, Italy and Portugal, reflecting
+% their effort to meet the Maastricht Criteria for inflation rates.
 % no literature overview
-Theory unfortunately is quite unsure  whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
-\citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation.
-It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
+
+Theory unfortunately is quite unsure  whether or not the creation of a monetary
+union between two or more states is likely to reduce or increase the variability
+or even the level of the inflation rate.  \citet{cooper} point out, that ``a
+central bank under a monetary union will internalize the interdependence between
+countries and optimally choose a lower inflation rate" and argue that a
+``central bank governing the growth of money supply will optimally choose zero
+inflation." This is not the case with the ECB which targets 2\% so as to
+avoid the risk of deflation. Thus, it is not quite clear how a monetary union
+will affect the volatility and the level of the inflation rate.
+
 An interesting approach to this question is taken by \citet{holte}. 
-%He creates a two country model for monetary policy analysis along the line of two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}.
-The inflation is modeled by a hybrid Phillips curve  specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
-What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union. \\
-%But monetary policy in a monetary union does not explicitly stabilize the output gap and inflation rate in case of national economic shocks. 
-The effects of joining a monetary union on inflation rate variability depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-Due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, theory has some troubles establishing a sound theoretical model.
-\\
-\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
-%They use a time-varying model with a GARCH specification for the unconditional volatility of inflation and find some instability in the conditional volatility. \\
-In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
-They find that convergence in time of inflation dynamics was only partly observable.
-In a paper studying core inflation and using an aggregated Euro Area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
-\\
-Inquiring into the convergence properties of inflation rates among countries of the EMU, \citet{busetti} find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
-\\
+% He creates a two country model for monetary policy analysis along the line of
+% two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}. The inflation
+% is modeled by a hybrid Phillips curve  specification and there are a number of
+% different home country interest rate rules, like strict inflation targeting,
+% flexible targeting, pegging to a currency and --  monetary union. 
+What he finds out via simulations of  different interest rate rules is that the
+standard deviation of the home consumer price index (CPI) inflation rate can be substantially reduced
+by joining a monetary union.
+% But monetary policy in a monetary union does not explicitly stabilize the
+% output gap and inflation rate in case of national economic shocks. 
+The effects of joining a monetary union on inflation rate variability depend on
+structural parameters like risk aversion, price flexibility, export demand
+elasticity, openness and shock correlations. Due to the fact that not all of
+these parameters are known and that their interaction as well has to be
+estimated, theory has some troubles establishing a sound theoretical model.
+
+\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU
+countries and find a considerable degree of heterogeneity across EMU countries
+in terms of average inflation and its degree of persistence.
+% They use a time-varying model with a GARCH specification for the unconditional
+% volatility of inflation and find some instability in the conditional volatility.
+
+In a paper examining structural convergence of the inflation rates in EU
+countries, \citet{palomba} try to answer the question whether during the 1990s the
+inflation rate dynamics of EU countries became more similar. They find that
+there is only partial evidence for convergence in time of inflation dynamics.  In a
+paper studying core inflation and using an aggregated Euro area inflation rate,
+\citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and
+1993---2000) governing the core inflation rate.
+
+Inquiring into the
+convergence properties of inflation rates among countries of the EMU,
+\citet{busetti} find that from 1980---1997 there was convergence of inflation
+rates, but afterwards there is some diverging behavior.
 
@@
 
-Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010.
-Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Spain,	Sweden, and	the	United Kingdom.
-The data are obtained from the OECD Statistics.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} \\
-The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia enters in 2011}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Spain and Slovenia)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members without ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
-Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
+Inflation is measured as the logarithm of the monthly change in the HICP from
+January 1990 to March 2010. Countries included are Austria, Belgium, the Czech
+Republic, Denmark, Estonia, Finland, France, Germany, Greece, Hungary, Ireland,
+Italy, Luxembourg, The Netherlands, Poland, Portugal, Spain, Sweden, and the
+United Kingdom.  The data are obtained from the OECD Statistics \fixme{It would be
+good to cite something here. Is there some stable reference/link for the OECD data source?}.
+Latvia, Lithuania, Bulgaria, and Romania are excluded due to data scarcity.
 
-\newpage
+The countries can be divided into three different groups. (1)~The Euro countries:
+Austria, Belgium, Estonia (which will enter in 2011), France,
+Finland, Germany, Greece, Ireland, Italy, Luxembourg, The Netherlands, Portugal,
+Spain and Slovenia.\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the
+Vatican are left out due their minor importance.} (2)~EU members without ERM~II
+(exchange rate mechanism): Czech Republic, Hungary, Poland, Sweden, and the United
+Kingdom. Denmark stands on its own as a member of the EU and the
+ERM~II, but not yet a member of the EMU.
+
 \section{Methods}
@@
 \citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
-\\
+
 We apply the general framework, as developed in \citet{z07}, to a more specific model, in this case by means of a GL distribution.
@@
 Regarding the data at hand, it was not possible to use the already existing method developed in \citet{z07}, since almost all inflation rates, with the notable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties.
-\\
+
 Therefore, a somewhat more flexible distribution had to be used.
@@
 Its probability density is given by:
-
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-
+%
 with location ($\theta$), scale ($\sigma$) and shape ($\delta$).
@@
 The moments are given by:
-
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
+%
+where $\gamma()$ is the digamma function and its derivatives.
 
-where $\gamma()$ is the digamma function and its derivatives. \\
 The log-likelihood is given by:
-
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-
+%
 The resulting score function ($\psi()$) for the parameters (the derivatives of the log-likelihood) are given by:
-
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-
+%
 \begin{eqnarray}
@@
 The asymptotic testing theory still holds for this generalization.
-[Beweis] \\
+[Beweis]
+
 The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process can be found in \citet{z07}.
@@
 Ireland stands on its own and is the only country which changed its inflation dynamics as an after march of the financial crises of 2008.
-\\
+
 What we can observe for almost all the EURO countries is an increase in variance and mean, with the exception of Ireland.
 Since inflation volatility increased in all countries, we find no proof of a common claim that inflation uncertainty may increase in countries that have a smaller influence on ECB policy.
-\\
+
 Lets consider a view countries in broader detail to get an idea about what happened.
@@
 However, we fail to find any evidence for an incrase in mean or volatility of the Austrian rate following the introduction of the Euro paper money.
-\\
-Lets consider Slovenia as the first Eastern European country that introduced the Euro.
+
+Let us consider Slovenia as the first Eastern European country that introduced the Euro.
 In August 2003 the new regime started with a much lower mean value (only a third now of its previous value) and interestingly enough, a higher volatility.
@@
 Slovenia seems to be the only case where the introduction of the Euro was followed by an increase in yearly inflation, which faded away following the financial crises and its deflationary effects.
-\\
+
 As a last example we focus on Spain.

### 2010-08-13 pkg/R/glogisfit.R

@@ -90,11 +90,12 @@ glogisfit.default <- function(x, weights = NULL,
       if(p3 < 0) return(Inf)
       -sum(w * dglogis(x, location = par[1], scale = p2, shape = p3, log = TRUE))
     }
+    need_trafo <- which(is.na(fixed)) %in% 2:3
     start2 <- start
-    start2[which(is.na(fixed)) %in% 2:3] <- exp(start2[which(is.na(fixed)) %in% 2:3])
+    start2[need_trafo] <- exp(start2[need_trafo])
     opt2 <- optim(par = start2, fn = llfun2)
     start <- opt2$par
-    start[2:3] <- log(start[2:3])
+    start[need_trafo] <- log(start[need_trafo])
     opt <- optim(par = start, fn = llfun, gr = gradfun, hessian = hessian, method = method, control = list(...))
   }
   

### 2010-10-08 pkg/paper/Draft/draft-Z-W.tex
title suggestion, used for workshop
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@35 b679b2f7-6539-45d0-a605-5ddcb976cde9
updated draft and slides (with graphs) and pdf version
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@37 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2010-07-27 pkg/paper/Draft/draft-Z-W.tex
@@
 
-\title{Something about Euro, EMU, and Inflation}
+\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}
 
@@
   The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)
-  on changing the volatility of the inflation rates in its Member States. 
+  on changing the dynamics of the inflation rates in its Member States. 
 }
@@
   6020 Innsbruck, Austria\\
-  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org}
+  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
 }
@@
 The European Central Bank (ECB) defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of
-below 2\%''.
-\citet{emerson} emphasizes the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism (p.~22).
+below 2\%''\footnote{http://www.ecb.europa.eu/mopo/strategy/pricestab/html/index.en.html}.
+\citet{emerson} emphasizes the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
 \\
 %Therefore a vital aim of every central bank is to achieve a low inflation rate with negligible variation. \\
-The question of interest centers around the way in which a countries decision to join the EMU changed its inflation rate dynamics. 
+The question of interest centers around the way in which a countries decision to join the Economic and Monetary Union (EMU) changed its inflation rate dynamics. 
 %There are a number of reasons why these should have changed indeed.
@@
 Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010.
-Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Spain,	Sweden, and	the	United Kingdom.
-The data are obtained from the OECD Statistics.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} \\
+Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Spain,	Sweden, and	the	United Kingdom.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} The data is obtained from the OECD Statistics.\footnote{http://www.oecd-ilibrary.org/content/data/data-00285-en} \\
 The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia enters in 2011}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Spain and Slovenia)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members without ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
@@
 with location ($\theta$), scale ($\sigma$) and shape ($\delta$).
-For b=1 the distribution simplifies to the logistic distribution, for b$<$1 it is skewed to the left and for b$>$1 it is skewed to the right.
+For $\delta$=1 the distribution simplifies to the logistic distribution, for $\delta$$<$1 it is skewed to the left and for $\delta$$>$1 it is skewed to the right.
 The moments are given by:
@@
 
-%We use the Cram{\'e}r von Mises type test as given in \citet{z07}. The test statistic is given by:
-%
+We use the Cram{\'e}r von Mises type test as given in \citet{z07}. The test statistic is given by:
+
 %\begin{eqnarray}
@@
 
+% mit cecchetti Paper im Unterlagen Ordner abgleichen
+
 \newpage
@@
 \\
-Lets consider a view countries in broader detail to get an idea about what happened.
+Lets consider a few countries in broader detail to get an idea about what happened.
 In Austria we see evidence of a change in the inflation rate, when the old regime ended in September 2007.
-The new regime is much more volatile, with an anlmost threefold increase in variance.
+The new regime is much more volatile, with an almost threefold increase in variance.
 
@@
 
-The reason for this is a strong increase in the yearly Austrian inflation rate following a heavy increase in oil price which was further amplified by a rise of minearl taxes.
+The reason for this is a strong increase in the yearly Austrian inflation rate following a heavy increase in oil price which was further amplified by a rise of mineral taxes.
 However, we fail to find any evidence for an incrase in mean or volatility of the Austrian rate following the introduction of the Euro paper money.
@@
 \\
-As a last example we focus on Spain.
+
+%As a last example we focus on Spain. \\
+
+%Other possible explanations: 
+%
+%- rise in value added tax (VAT; as in Germany in 2007)
+%- change in central bank policy (for example Spain changed to inflation targeting in 1995; Uk, Swe and Finland in 1992)
+%- role of budget deficits
+
+% Quelle für diese Behaupunten: http://www.oenb.at/isaweb/dyna2.do?go=initIndikatoren&definitionLaden=false&hierarchieIdSelected=108
+
+Another possible explanation pattern could be the growth in the  money supply.
+A drawback is the current lack of data\footnote{hopefully to be solved}.
+But for some countries we can find some plausible findings.
+Regarding the case of Denmark, the M1 money supply started contracting in June 2000 (on a monthly base).
+In Estonia we find the highpoint in M3 expansion ending with March 1998, with much lower M3 growth rates afterwards.
+The same happended to Hungary in April and May 1998.
+In the Polnish case we find a severe contraction of the money supply in May 2001.
+Similar findings hold true for Germany as well.
+\\
+In Spain we see a stronger increase in M1 in May and June 1996 (nothing for the other break, data scarceity).
+In Portugal we find a modest increase in year on year M1 money supply growth starting in July 1992.
+We find nothing peculiar for Slovakia and Hungary.
+In Estonia the M1 money supply starts to contract severely on a yearly base following March 1998.
+In Slovenia we find a contraction of the Money Supply in July 2003, with the whole year 2003 witnessing a sharp decrease in the year on year growth.
+Poland starts contracting the year on year M1 money supply after June 2000 with very low growth rates till January 2002.
+\\
+[irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck]
+
+% for denmark: see: C:\Dissertation\Daten\DSI\M1\daten - the denmark analysis, alos: Spain, Portugal, Slovakia, Hungary Estonia noch am absatz für M1
+
+Concerning monetary unions, it is interesting to notice that Belgium and Luxembourg, that constituted a monetary union prior to the European Monetary Union, did not experience the same timing in their break towards a higher mean and a significantly higher volatility regime.
+\\
+
+With regard to the alleged benefit of monetary unions, the import of credibility to former high inflation countries (like Italy and Spain), did nothing to decrease their mean or volatility, rather to the contrary.
+Spain and Italy did have a -- rather -- low mean low volatility regime from 1996:6 to 2000:12 but then things changed towards higher volatility.
+It appears as if both countries tried to decrease inflation prior to the fixing of the exchange rates and let go afterwards.
+\\
+
+Following the reasoning of a study by \cite{duarte}, high volatility of inflation rates within regions of a union may be explained a countries decision to lower the inflation differential by means of fiscal policy.
+This explanation might fit to Spain and Italy, both were above the the average EU inflation.
+\\
+
+% Quelle: für Deutschland M3 bis 1995
+%http://www.bundesbank.de/statistik/statistik_zeitreihen.php?lang=de&open=ewu&func=row&tr=TSD304
+%
+%Deutschland: M3 bis 1995
+%dort ersichtlihc, dass Deutschland zu Bruchzeitpunkten jeweils änderungen im m3 aggregat hat
 
@@
 
+\clearpage
 \section{Conlcusion}
 
+\clearpage
 \bibliography{papers}
 
+\clearpage
+
 \end{document}

### 2010-10-08 pkg/paper/Draft/draft-Z-W.tex
small touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@38 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \title{Structural Breaks in Inflation Dynamics within the European Monetary Union}
+
+%% for internal use
+\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
+\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
 
@@
 Poland starts contracting the year on year M1 money supply after June 2000 with very low growth rates till January 2002.
-\\
-[irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck]
+
+\readme{irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck}
 
@@
 Concerning monetary unions, it is interesting to notice that Belgium and Luxembourg, that constituted a monetary union prior to the European Monetary Union, did not experience the same timing in their break towards a higher mean and a significantly higher volatility regime.
-\\
 
@@
 It appears as if both countries tried to decrease inflation prior to the fixing of the exchange rates and let go afterwards.
-\\
 
@@
 This explanation might fit to Spain and Italy, both were above the the average EU inflation.
-\\
 

### 2010-10-11 pkg/paper/Draft/draft-Z-W.tex
updated draft and bib file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@39 b679b2f7-6539-45d0-a605-5ddcb976cde9
updated draft and slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@45 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 % 4. Finanzkrise: schwieriger, ausser Irland keine Änderung > 2008
-
-% Loswerden der Working Papers
-% auf Zitat mit castelnuovo kann man verzichten, dass die EZB Preisstabilität so definiert ist kein Geheimenis und nichts, dass man mit Paper zitieren müsste
-% holemöller geht nicht weg, immer noch Discussion Paper (siehe sein lebenslauf)http://www.holtem.de/Dokumente/PublicCV-Holtemoeller.pdf
-% morana detto
 
@@
 Country        & Dates  & \multicolumn{2}{l}{Breakpoints}  \\ \hline
-Austria        & 1999--2002   	& Sep 2007 &          \\
-Belgium        & 1999--2002   	& Dec 1999 &          \\
+Austria        & 1995--2002   	& Sep 2007 &          \\
+Belgium        & 1979--2002   	& Dec 1999 &          \\
 Czech Republic & no--no		 			& Jul 1998 &          \\
-Denmark        & 1999--no 			& Jun 2000 &          \\
+Denmark        & 1979/1999--no 			& Jun 2000 &          \\
 Estonia        & 2004--2011   	& Mar 1998 &          \\
-Finland        & 1999--2002   	& none     &          \\
-France         & 1999--2002   	& Dec 2004 &          \\
-Germany        & 1999--2002   	& May 2000 & Dec 2004 \\
-Greece         & 2001--2002   	& none     &          \\
+Finland        & 1996--2002   	& none     &          \\
+France         & 1979--2002   	& Dec 2004 &          \\
+Germany        & 1979--2002   	& May 2000 & Dec 2004 \\
+Greece         & 1998--2002   	& none     &          \\
 Hungary        & no--no       	& May 1998 &          \\
-Ireland        & 1999--2002   	& Mar 2008 &          \\
-Italy          & 1999--2002   	& May 1996 & Dec 2000 \\
-Luxembourg     & 1999--2002   	& Dec 1998 &          \\
-Netherlands    & 1999--2002   	& none     &          \\
+Ireland        & 1979--2002   	& Mar 2008 &          \\
+Italy          & 1979--2002   	& May 1996 & Dec 2000 \\
+Luxembourg     & 1979--2002   	& Dec 1998 &          \\
+Netherlands    & 1979--2002   	& none     &          \\
 Poland         & no--no      		& May 2001 &          \\
-Portugal       & 1999--2002   	& Jul 1992 & Mar 2004 \\
+Portugal       & 1992--2002   	& Jul 1992 & Mar 2004 \\
 Slovakia       & 2005--2009   & Apr 1997 & Feb 2004 \\
 Slovenia       & 2004--2007   & Jul 2003 &          \\
-Spain          & 1999--2002   & May 1996 & Dec 2000 \\
+Spain          & 1989--2002   & May 1996 & Dec 2000 \\
 Sweden         & no--no      & Jan 1993 &          \\
-United Kingdom & no--no       & Apr 1992 &          \\ \hline
+United Kingdom & 1990--no       & Apr 1992 &          \\ \hline
 \end{tabular}
-\caption{\label{tab:breakpoints} Dating of break points. First date: entry to ERM~II, second date: EURO introduction.}
+\caption{\label{tab:breakpoints} Dating of break points. First date: entry to EMS or ERM~II, second date: EURO introduction.}
 \end{center}
@@
 
+% EMS was never very stable
+% careful: ERM II für nicht EURO Mitglieder, EMS war Vorgänger, auch für EURO Mitglieder
+% http://de.wikipedia.org/wiki/Europ%C3%A4isches_W%C3%A4hrungssystem
 % mit cecchetti Paper im Unterlagen Ordner abgleichen
@@
 
-\readme{irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck}
+%\readme{irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck}
 

### 2010-10-12 pkg/paper/Draft/draft-Z-W.tex
updated draft and slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@48 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \usepackage[latin1]{inputenc}
+\usepackage{longtable}
+\usepackage{lscape}
+\usepackage{multirow}
 
@@
 
-\begin{table}[ht!p]
-\begin{center}
-\begin{tabular}{llll}
+\begin{longtable}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|} 
 \hline
-Country        & Dates  & \multicolumn{2}{l}{Breakpoints}  \\ \hline
-Austria        & 1995--2002   	& Sep 2007 &          \\
-Belgium        & 1979--2002   	& Dec 1999 &          \\
-Czech Republic & no--no		 			& Jul 1998 &          \\
-Denmark        & 1979/1999--no 			& Jun 2000 &          \\
-Estonia        & 2004--2011   	& Mar 1998 &          \\
-Finland        & 1996--2002   	& none     &          \\
-France         & 1979--2002   	& Dec 2004 &          \\
-Germany        & 1979--2002   	& May 2000 & Dec 2004 \\
-Greece         & 1998--2002   	& none     &          \\
-Hungary        & no--no       	& May 1998 &          \\
-Ireland        & 1979--2002   	& Mar 2008 &          \\
-Italy          & 1979--2002   	& May 1996 & Dec 2000 \\
-Luxembourg     & 1979--2002   	& Dec 1998 &          \\
-Netherlands    & 1979--2002   	& none     &          \\
-Poland         & no--no      		& May 2001 &          \\
-Portugal       & 1992--2002   	& Jul 1992 & Mar 2004 \\
-Slovakia       & 2005--2009   & Apr 1997 & Feb 2004 \\
-Slovenia       & 2004--2007   & Jul 2003 &          \\
-Spain          & 1989--2002   & May 1996 & Dec 2000 \\
-Sweden         & no--no      & Jan 1993 &          \\
-United Kingdom & 1990--no       & Apr 1992 &          \\ \hline
-\end{tabular}
-\caption{\label{tab:breakpoints} Dating of break points. First date: entry to EMS or ERM~II, second date: EURO introduction.}
-\end{center}
-\end{table}
+Country & Segment~1 & Segment~2 & Segment~3 & ERM~II/EMS & EURO \\
+\hline
+\hline
+\multirow{4}{*}{Austria} & 1990(2)--2007(9)   & 2007(10)--2010(3) & -- & 1995(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.1635 & mean: 0.1731 & &  &  \\
+ & var: 0.05694 & var: 0.16310 & &  &  \\
+ & skew: 0.6056 & skew: 0.1691 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Belgium} & 1991(2)--1999(12)   & 2001(1)--2010(3) & -- & 1979(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.1459 & mean: 0.1768 & &  &  \\
+ & var: 0.06401 & var: 0.95401 & &  &  \\
+ & skew: -0.03708 & skew: 0.50356 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Czech Republic} & 1995(2)--1998(7)   & 1998(8)--2010(3) & -- & no & no \\ 
+\cline{2-4}
+ & mean: 0.6969 & mean: 0.1821& &  &  \\
+ & var: 0.3363 & var: 0.2156 & &  &  \\
+ & skew: 1.139 & skew: 0.990 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Denmark} & 1990(2)--2000(6)   & 2000(7)--2010(3) & -- & 1999(1) & no \\ 
+\cline{2-4}
+ & mean: 0.1664 & mean: 0.1676& &  &  \\
+ & var: 0.09078 & var: 0.18758 & &  &  \\
+ & skew: -0.7425 & skew: 1.0471 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Estonia} & 1996(2)--1998(3)   & 1998(4)--2010(3) & -- & 2004(6) & 2011(1) \\ 
+\cline{2-4}
+ & mean: 0.8649 & mean: 0.3328 & &  &  \\
+ & var: 0.4196 & var: 0.2062 & &  &  \\
+ & skew: 0.4041 & skew: 0.8016 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Finland} & 1990(2)--2010(3)   & -- & -- & 1996(10) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.1653 &  & &  &  \\
+ & var: 0.1321 &  & &  &  \\
+ & skew: 0.2798 &  &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{France} & 1990(2)--2004(12)   & 2005(1)--2010(3) & -- & 1979(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.1588 & mean: 0.1504 & &  &  \\
+ & var: 0.05769 & var: 0.13134 & &  &  \\
+ & skew: 0.1965 & skew: -0.7942 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Germany} & 1995(2)--2000(5)   & 2000(6)--2004(12) & 2005(1)--2010(3) & 1979(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.08799 & mean: 0.14018 & mean: 0.14183 &  &  \\
+ & var: 0.06013 & var: 0.16352 & var: 0.18384 &  &  \\
+ & skew: 0.9219 & skew: 0.9920 & skew: -0.6625 &  &  \\
+\hline
+\newpage 
+\hline
+\multirow{4}{*}{Greece} & 1995(2)--2010(3)   & -- & -- & 1999(1) & 2001(1) \\ 
+\cline{2-4}
+ & mean: 0.3227 &  & &  &  \\
+ & var: 1.48 &  & &  &  \\
+ & skew: 0.4314 &  &   &  &  \\
+\hline
+\hline
+\multirow{4}{*}{Hungary} & 1995(2)--1998(5)   & 1998(6)--2010(3) & -- & no & no \\ 
+\cline{2-4}
+ & mean: 1.6064 & mean: 0.5068 & &  &  \\
+ & var: 1.0243 & var: 0.3161 & &  &  \\
+ & skew: 0.8781 & skew: 0.7095 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Ireland} & 1995(2)--2008(3)   & 2008(4)--2010(3) & -- & 1979(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.2546 & mean: -0.1313 & &  &  \\
+ & var: 0.2045 & var: 0.1836 & &  &  \\
+ & skew: -0.6958 & skew: -0.9947 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Italy} & 1990(2)--1996(5)   & 1996(7)--2000(12) & 2001(1)--2010(3) & 1979(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.4135 & mean: 0.1676 & mean: 0.1819&  &  \\
+ & var: 0.04129 & var: 0.01997 & var: 0.32117 &  &  \\
+ & skew: 0.9627 & skew: 0.7261 & skew: -0.2605 &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Luxembourg} & 1995(2)--1998(12)   & 1999(1)--2010(3) & -- & 1979(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.08761 & mean: 0.22425 & &  &  \\
+ & var: 0.01340 & var: 0.53108 & &  &  \\
+ & skew: 0.2606 & skew: -0.4836 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Netherlands} & 1990(2)--2010(3)   & -- & -- & 1979(1) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.1854 &  & &  &  \\
+ & var: 0.293 &  & &  &  \\
+ & skew: 0.5984 &  &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Poland} & 1996(2)--2001(5)   & 2001(6)--2010(3) & -- & no & no \\ 
+\cline{2-4}
+ & mean: 0.8548 & mean: 0.2024 & &  &  \\
+ & var: 0.4212 & var: 0.1232 & &  &  \\
+ & skew: 0.6675 & skew: -0.3148 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Portugal} & 1990(2)--1992(7)   & 1992(8)--2004(3) & 2004(4)--2010(3) & 1992(4) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.8519 & mean: 0.2700 & mean: 0.1605&  &  \\
+ & var: 0.1718 & var: 0.1052 & var: 0.2559 &  &  \\
+ & skew: 1.1394 & skew: 0.8653 & skew: 0.5690 &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Slovakia} & 1995(2)--1997()   & 1997(5)--2004(2) & 2004(3)--2010(3) & 2005(11) & 2009(1) \\ 
+\cline{2-4}
+ & mean: 0.4799 & mean: 0.5872 & mean: 0.1865 &  &  \\
+ & var: 0.05768 & var: 0.44150 & var: 0.08903 &  &  \\
+ & skew: 1.139 & skew: 1.140 & skew: 1.139 &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Slovenia} & 1996(2)--2003(7)   & 2003(8)--2010(3) & -- & 2004(6) & 2007(1) \\ 
+\cline{2-4}
+ & mean: 0.6309 & mean: 0.2436 & &  &  \\
+ & var: 0.2108 & var: 0.3440 & &  &  \\
+ & skew: 0.5883 & skew: 0.1426 &   &  &  \\
+\hline 
+\newpage
+\hline
+\multirow{4}{*}{Spain} & 1992(2)--1996(5)   & 1996(6)--2000(12) & 2001(1)--2010(3) & 1989(6) & 1999(1) \\ 
+\cline{2-4}
+ & mean: 0.3724 & mean: 0.2001 & mean: 0.2225 &  &  \\
+ & var: 0.06960 & var: 0.03988 & var: 0.34215 &  &  \\
+ & skew: 1.13909 & skew: 0.01874 & skew: -0.36200 &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{Sweden} & 1990(2)--1993(1)   & 1993(2)--2010(3) & -- & no & no \\ 
+\cline{2-4}
+ & mean: 0.4748 & mean: 0.1552 & &  &  \\
+ & var: 0.5715 & var: 0.1848 & &  &  \\
+ & skew: 1.1393 & skew: 0.5344 &   &  &  \\
+\hline 
+\hline
+\multirow{4}{*}{United Kingdom} & 1990(2)--1992(4)   & 1992(5)--2010(3) & -- & 1990(10)\footnote{left after the Black Wednesday 1992} & no \\ 
+\cline{2-4}
+ & mean: 0.5703 & mean: 0.1615 & &  &  \\
+ & var: 0.3869 & var: 0.1490 & &  &  \\
+ & skew: 1.139 & skew: -1.265 &   &  &  \\
+\hline 
+\end{longtable}
+
+% \endfoot beendet ganz
 

### 2010-10-13 pkg/paper/Draft/draft-Z-W.tex
finished slides and table for countries slides.Rnw should be executed within the Analysis Results environment since it needs datasets
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@49 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \hline
-\multirow{4}{*}{Belgium} & 1991(2)--1999(12)   & 2001(1)--2010(3) & -- & 1979(1) & 1999(1) \\ 
+\multirow{4}{*}{Belgium} & 1991(2)--1999(12)   & 2000(1)--2010(3) & -- & 1979(1) & 1999(1) \\ 
 \cline{2-4}
@@
 \hline
-\multirow{4}{*}{Slovakia} & 1995(2)--1997()   & 1997(5)--2004(2) & 2004(3)--2010(3) & 2005(11) & 2009(1) \\ 
+\multirow{4}{*}{Slovakia} & 1995(2)--1997(4)   & 1997(5)--2004(2) & 2004(3)--2010(3) & 2005(11) & 2009(1) \\ 
 \cline{2-4}
@@
 Group 1 consists of Italy and Spain (both very similar) and possibly Belgium, Luxembourg and Portugal.
-Group 2 is an Eastern European group comprising Poland, the Czech Republic, Slovenia, Slovakia and Hungary.
-Group 3 is a Non-ERM~II member group including Sweden and the United Kingdom.
+Group 2 is an Eastern European group comprising Poland, the Czech Republic, Slovenia, Slovakia, Estonia and Hungary.
+Group 3 is a Non-ERM~II member group including Denmark, Sweden and the United Kingdom.
 The Netherlands, Greece and Finland build group 4 which exhibits no changes.
-Group 5 is composed of France, Germany, Austria and Denmark.
+Group 5 is composed of France, Germany, and Austria.
 Ireland stands on its own and is the only country which changed its inflation dynamics as an after march of the financial crises of 2008.

### 2010-10-22 pkg/paper/Paper/draft.Rnw
Some cleanup in SVN, now three directories:   Analysis   Paper   Slides where the latter two rely on the former.
All figures removed because they are now dynamically generated in .Rnw.

Two target files (draft.pdf, slides.pdf) are retained to be linked from
R-Forge project web page.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@60 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2010-10-11 pkg/paper/Draft/draft-Z-W.tex
@@
 \usepackage[latin1]{inputenc}
+\usepackage{longtable}
+\usepackage{lscape}
+\usepackage{multirow}
+%% need no \usepackage{Sweave}
+
+\setkeys{Gin}{width=\textwidth}
+\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}
+
+<<preliminaries>>=
+## packages
+library("glogis")
+library("fxregime")
+
+## additional functions
+moments <- function(object, ...) UseMethod("moments")
+moments.glogisfit <- function(object, ...) object$moments
+moments.breakpoints.glogisfit <- function(object, breaks = NULL, ...)
+  t(sapply(refit(object, breaks = breaks), "[[", "moments"))
+
+## For the creation of the tables
+## <<tabelle, eval=FALSE>>=
+if(!file.exists(file.path("C:/Dissertation/Paper-1-Z", "Analysis", "Austria.rda"))) {
+setwd(file.path("..", "Analysis"))
+Sweave("analysis.Rnw")
+setwd(file.path("..", "Slides"))
+}
+
+## query result file names
+fnam <- Sys.glob(file.path("..", "Analysis", "*.rda"))
+fnam2 <- grep("2", fnam, fixed = TRUE)
+if(length(fnam2) > 0) fnam <- fnam[-fnam2]
+
+## set up named results list
+results <- vector(length = length(fnam), mode = "list")
+names(results) <- sapply(strsplit(sapply(strsplit(fnam, "/", fixed = TRUE), tail, 1), ".", fixed = TRUE), head, 1)
+
+## erm <- yearmon(c(
+##   Austria = 1995,
+##   ...,
+##   France = 1995 + 1/12,
+##   ...))
+
+## create erm (Exchange rate mechanism bzw. exchange rate mechanism 2
+## ERM~II, ERM~I (part of european monetary system expired in 1.1.1999 gave way to ERM II
+## if never entered ""right?
+## http://web.archive.org/web/20080101022356/http://lexikon.meyers.de/meyers/Europ%C3%A4isches_W%C3%A4hrungssystem
+
+erm <- yearmon(c(
+	Austria = 1995,
+	Belgium = 1979 + 2/12,
+	CzechRepubic = NA, 
+	Denmark = 1979 + 2/12,
+	Estonia = NA,
+	Finland = 1996 + 9/12,
+	France = 1979 + 2/12,
+	Germany = 1979 + 2/12,
+	Greece = 1998 + 2/12,
+	Hungary = NA,
+	Ireland = 1979 + 2/12,
+	Italy = 1979 + 2/12,
+	Luxembourg = 1979 + 2/12,
+	Netherlands = 1979 + 2/12,
+	Poland = NA,
+	Portugal = 1992 + 3/12,
+	Slovakia = NA,
+	Slovenia = NA,
+	Spain = 1986 + 5/12,
+	Sweden = NA ,
+	UnitedKingdom = NA
+))
+
+erm2 <- yearmon(c(
+	Austria = NA,
+	Belgium = NA,
+	CzechRepubic = NA, 
+	Denmark = 1999,
+	Estonia = 2004 + 5/12,
+	Finland = NA,
+	France = NA,
+	Germany = NA,
+	Greece = 1999,
+	Hungary = NA,
+	Ireland = NA,
+	Italy = NA,
+	Luxembourg = NA,
+	Netherlands = NA,
+	Poland = NA,
+	Portugal = NA,
+	Slovakia = 2005 + 10/12,
+	Slovenia = 2004 + 5/12,
+	Spain = NA,
+	Sweden = NA ,
+	UnitedKingdom = NA
+))
+
+eur <- yearmon(c(
+	Austria = 1999,
+	Belgium = 1999,
+	CzechRepubic = NA, 
+	Denmark = NA,
+	Estonia = 2011,
+	Finland = 1999,
+	France = 1999,
+	Germany = 1999,
+	Greece = 2001,
+	Hungary = NA,
+	Ireland = 1999,
+	Italy = 1999,
+	Luxembourg = 1999,
+	Netherlands = 1999,
+	Poland = NA,
+	Portugal = 1999,
+	Slovakia = 2009,
+	Slovenia = 2007,
+	Spain = 1999,
+	Sweden = NA ,
+	UnitedKingdom = NA
+))
+
+## sequentially load data
+for(i in seq_along(fnam)) {
+  load(file.path("..", "Analysis", fnam[i]))
+  results[[i]] <- list(
+    data = x,
+    model0 = x_gf,
+    efp = x_efp,
+    nbreaks = x_nbreaks,
+    breakpoints = x_bp,
+    model = x_rf,
+    erm = erm[i],
+    erm2 = erm2[i],
+    eur = eur[i])
+}
+
+## compute moments
+mom <- lapply(results, function(obj) t(sapply(obj$model, moments)))
+
+## LaTeX formatting
+format_moments <- function(i, digits = 3) {
+  rval <- format(round(mom[[i]], digits = digits), nsmall = digits)
+  rval <- cbind("", rownames(rval), matrix(paste("$", rval, "$", sep = ""), ncol = 3), "", "", "")
+  rval[1,1] <- if(is.character(i)) i else names(mom)[i]
+  rval[1,6] <- if(is.na(results[[i]]$erm)) "--" else format(results[[i]]$erm)
+  rval[1,7] <- if(is.na(results[[i]]$erm2)) "--" else format(results[[i]]$erm2)
+  rval[1,8] <- if(is.na(results[[i]]$eur)) "--" else format(results[[i]]$eur)
+  rval <- apply(rval, 1, paste, collapse = " & ")
+  rval <- paste(rval, "\\\\")
+  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
+  return(rval)
+}
+
+## data
+data("HICP", package = "glogis")
+HICP <- HICP[, colnames(HICP) != "EU"]
+@
+%
+%% analysis "macro" (assumes existence of "i", "HICP", and "add_breaks")
+<<analysis, eval=FALSE>>=
+## compute returns
+x <- 100 * diff(log(na.omit(HICP[,i])))
+
+## fit full-sample model
+x_gf <- glogisfit(x)
+
+## assess stability of full-sample model
+x_efp <- gefp(x_gf, fit = NULL)
+
+## estimate breakpoints (with minimal segment size of 2 years)
+x_bp <- breakpoints(x_gf, h = 24)
+
+## (minimal) number of breaks
+x_nbreaks <- if(sctest(x_efp, functional = supLM(0.1))$p.value > 0.05) {
+  0
+} else {
+  max(1, length(breakdates(x_bp))) + add_breaks[i]
+}
+
+## refit segmented model
+x_rf <- refit(x_bp, breaks = x_nbreaks)
+
+## save everything
+save(x, x_gf, x_efp, x_bp, x_nbreaks, x_rf, file = paste(colnames(HICP)[i], ".rda", sep = ""))
+
+<<bla, eval=TRUE>>=
+setwd("C:/Dissertation/Paper-1-Z/R-Forge2/pkg/paper/Analysis")
+load("Belgium.Rda")
+xbel <- x
+xbel_bp <- x_bp
+xbel_efp <- x_efp
+xbel_gf <- x_gf
+xbelnbreaks <- x_nbreaks
+xbel_rf <- x_rf
+load("Luxembourg.Rda")
+
+@
+
+<<plotversuch, eval=FALSE>>=
+par(mfrow = c(2, 1))
+plot(x, main = "Series with Fitted Mean", ylab = "100 * diff(log(HICP))", xlab = "Time")
+lines(x_bp, breaks = x_nbreaks)
+lines(fitted(x_bp, breaks = x_nbreaks, type = "mean"), col = 4)
+plot(xbel, main = "Series with Fitted Mean", ylab = "100 * diff(log(HICP))", xlab = "Time")
+lines(xbel_bp, breaks = xbelnbreaks)
+lines(fitted(xbel_bp, breaks = xbelnbreaks, type = "mean"), col = 4)
+@
+
+% plot "macros" (assuming existence "x", "x_gf", ...)
+<<plot1, eval=FALSE>>=
+par(mfrow = c(1, 2))
+plot(x, main = "Series with Fitted Mean", ylab = "100 * diff(log(HICP))", xlab = "Time")
+lines(x_bp, breaks = x_nbreaks)
+lines(fitted(x_bp, breaks = x_nbreaks, type = "mean"), col = 4)
+plot(x_efp, functional = supLM(0.1), main = "supLM test")
+lines(x_bp, breaks = x_nbreaks)
+@
+
+<<plot1d, eval=FALSE>>=
+plot(x, main = "Slovenia", ylab = "100 * diff(log(HICP))", xlab = "Time")
+@
+
+<<plot1b, eval=FALSE>>=
+plot(x_efp, functional = supLM(0.1), main = "supLM test")
+lines(x_bp, breaks = x_nbreaks)
+@
+
+<<plot2, eval=FALSE>>=
+par(mfrow = c(1, 1))
+plot(x_bp)
+@
+
+<<plot3, eval=FALSE>>=
+par(mfrow = c(1, x_nbreaks + 1))
+for(i in 1:(x_nbreaks+1)) plot(x_rf[[i]], xlab = paste(names(x_rf)[i], "\n", "Goodness-of-fit p-value: ",
+  format.pval(summary(x_rf[[i]])$chisq.test$p.value, digits = 4), sep = ""))
+@
+
+<<plot4, eval=FALSE>>=
+par(mfrow = c(1, 1))
+plot(x_efp, aggregate = FALSE, ylim=c(-2.1,2.1))
+@
 
@@
 \newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
-
-% Verkaufsstrategien
-% 1. Widerlegung des Teuro Effektes
-% 2. Keine Veränderung d. EMU sichtbar; nach 2002 kein globaler Trend aller EMU Staaten Richtung Wechsel
-% 3. Nutzen zeigen, den die osteuropäischen Staaten haben, dadurch, dass sie dem EURO beitreten müssen  (mit Ausnahme von Slovenien fallender Trend in Mittelwert und Varianz)
-% 	 so etwa in Slowenien und der Slovakei ein Bruch bevor diese Staaten dem ERMII/EURO beitreten
-% 4. Finanzkrise: schwieriger, ausser Irland keine Änderung > 2008
 
@@
 In econometrics, the logistic distribution is often used in income distributions and growth models.
-This is due to  its longer tails and  higher peak, which fits these problems somewhat better.
+This is due to  its longer tails, which fits these problems somewhat better.
 The use of the generalized logistic distribution as we will use it in this paper is rather rare.
@@
 Therefore, a somewhat more flexible distribution had to be used.
-We needed a distribution exhibiting rather strong kurtosis and the property to be both left and right skewed.
+We needed a distribution exhibiting rather fatter tails and the property to be both left and right skewed.
 To serve this purpose, we use a generalization of the logistic distribution as defined in \citet{johnson}.
@@
 To help with the interpretation, we give a short overview over the history of the EMU.
-Following the Delors Plan with his 3 stages, we have: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender.
+Following the Delors Plan with his 3 stages, we have: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of exchange rates of Member States.
 If the EMU had a significant effect upon inflation rate volatility we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of Euro paper money led to a considerable price increase.
@@
 
-\begin{table}[ht!p]
-\begin{center}
-\begin{tabular}{llll}
+% automatic longtable provided by R
+
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
 \hline
-Country        & Dates  & \multicolumn{2}{l}{Breakpoints}  \\ \hline
-Austria        & 1995--2002   	& Sep 2007 &          \\
-Belgium        & 1979--2002   	& Dec 1999 &          \\
-Czech Republic & no--no		 			& Jul 1998 &          \\
-Denmark        & 1979/1999--no 			& Jun 2000 &          \\
-Estonia        & 2004--2011   	& Mar 1998 &          \\
-Finland        & 1996--2002   	& none     &          \\
-France         & 1979--2002   	& Dec 2004 &          \\
-Germany        & 1979--2002   	& May 2000 & Dec 2004 \\
-Greece         & 1998--2002   	& none     &          \\
-Hungary        & no--no       	& May 1998 &          \\
-Ireland        & 1979--2002   	& Mar 2008 &          \\
-Italy          & 1979--2002   	& May 1996 & Dec 2000 \\
-Luxembourg     & 1979--2002   	& Dec 1998 &          \\
-Netherlands    & 1979--2002   	& none     &          \\
-Poland         & no--no      		& May 2001 &          \\
-Portugal       & 1992--2002   	& Jul 1992 & Mar 2004 \\
-Slovakia       & 2005--2009   & Apr 1997 & Feb 2004 \\
-Slovenia       & 2004--2007   & Jul 2003 &          \\
-Spain          & 1989--2002   & May 1996 & Dec 2000 \\
-Sweden         & no--no      & Jan 1993 &          \\
-United Kingdom & 1990--no       & Apr 1992 &          \\ \hline
-\end{tabular}
-\caption{\label{tab:breakpoints} Dating of break points. First date: entry to EMS or ERM~II, second date: EURO introduction.}
-\end{center}
-\end{table}
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(seq(1,21)), format_moments)))
+@
+\end{longtable}
 
@@
 \newpage
+\subsection{Belgium and Luxembourg}
 
-On the basis of the results, we can trace out 6 groups of countries with rather distinct behavior.
-Group 1 consists of Italy and Spain (both very similar) and possibly Belgium, Luxembourg and Portugal.
-Group 2 is an Eastern European group comprising Poland, the Czech Republic, Slovenia, Slovakia and Hungary.
-Group 3 is a Non-ERM~II member group including Sweden and the United Kingdom.
-The Netherlands, Greece and Finland build group 4 which exhibits no changes.
-Group 5 is composed of France, Germany, Austria and Denmark.
-Ireland stands on its own and is the only country which changed its inflation dynamics as an after march of the financial crises of 2008.
-\\
-What we can observe for almost all the EURO countries is an increase in variance and mean, with the exception of Ireland.
-Since inflation volatility increased in all countries, we find no proof of a common claim that inflation uncertainty may increase in countries that have a smaller influence on ECB policy.
-\\
+% to show both series with breakpoints and means
+% does not work
+%\begin{center}
+%<<belluxvers, fig=TRUE, height=7, width=8>>=
+%<<plotversuch>>
+%@
+%\end{center}
+
+On the basis of these results, we are able to trace out different groups.
+The first of this group consits of Belgium and Luxembourg.
+This may come as no surprise since both countries formed a monetary union prior even to the start of the European Rate Mechanism (ERM).
+While both countries enjoyed very stable inflation rates with low values of both mean and variance, this changed around the timing of the irrevocable fixing of the exchange rates.
+A causal connection to this events seems likely.
+
+% heavy increase in variance both countries exactly around EURO introduction 
+% interesting thing: both were allready in a currency union (the belgish--luxembourg union)
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(2, 13), format_moments)))
+@
+\end{longtable}
+
+% could be looking much better
+\begin{center}
+<<bellux, fig=TRUE, height=8, width=8>>=
+print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Belgium"]))), 100*diff(log(na.omit(HICP[, "Luxembourg"])))), xlab="Time", ylab="", main="", screens=list("Belgium", "Luxembourg")), ylim=c(-2,2))
+@
+\end{center}
+
+\newpage
+\subsection{The Eastern Countries}
+\normalsize
+
+The second of these groups are the Eastern European Countries, the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
+In all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
+Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
+
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(5,10,15,17,18), format_moments)))
+@
+\end{longtable}
+
+\begin{center}
+<<eastern, fig=TRUE, height=7, width=8>>=
+print(xyplot(cbind(100*diff(log(na.omit(HICP[, "CzechRepublic"]))), 100*diff(log(na.omit(HICP[, "Estonia"]))), 100*diff(log(na.omit(HICP[, "Hungary"]))), 100*diff(log(na.omit(HICP[, "Poland"]))), 100*diff(log(na.omit(HICP[, "Slovakia"]))), 100*diff(log(na.omit(HICP[, "Slovenia"])))), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia")))
+@
+\end{center}
+
+%, ylim=c(-2.7,2.7)
+
+\newpage
+\subsection{Italy and Spain}
+\normalsize
+
+Another combination of countries with -- this time -- identical results are the two biggest southern European countries Italy and Spain.
+Both of them experienced high rates of inflation in the early 90ies.
+Afterwards, in the phase leading up to the monetary union, mean and variance were significantly reduced.
+This trend changed drastically soon after the fixing of the exchange rates leading to higher mean and much higher variances then ever before since the beginning of the HICP series.
+
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(12, 19), format_moments)))
+@
+\end{longtable}
+
+% could be looking much better
+\begin{center}
+<<itaspa, fig=TRUE, height=7, width=8>>=
+print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Italy"]))), 100*diff(log(na.omit(HICP[, "Spain"])))), xlab="Time", ylab="", main="", screens=list("Italy", "Spain"), ylim=c(-2,2)))
+@
+\end{center}
+
+\newpage
+\normalsize
+\subsection{Ireland}
+
+Ireland can be considered a rather special case.
+It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.
+We find a structural break in 2008 following the strong contraction of the Irish housing bubble and a beginning deflation.
+
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(11), format_moments)))
+@
+\end{longtable}
+
+\begin{center}
+<<ireland, fig=TRUE, height=7, width=8>>=
+plot(100*diff(log(na.omit(HICP[, "Ireland"]))), main="Ireland", ylab="", xlab ="Time", ylim=c(-2,2))
+@
+\end{center}
+
+\newpage
+\normalsize
+\subsection{No change countries}
+
+Another interesting group consists of 3 economically very different countries.
+In all of them we find no evidence of a structural change.
+Whereas the Netherlands and Finland are both on the lower side of the inflation margin, Greece exhibits some very strange behaviour.
+It has by far the highest variance of all series and seems to be completely different from all the others, just like some randomly generated series.
+
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(9,14,6), format_moments)))
+@
+\end{longtable}
+
+\begin{center}
+<<nochange, fig=TRUE, height=7, width=8>>=
+print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Finland"]))), 100*diff(log(na.omit(HICP[, "Greece"]))), 100*diff(log(na.omit(HICP[, "Netherlands"])))), xlab="Time", ylab="", main="", screens=list("Finland", "Greece", "Netherlands"), ylim=c(-2.7,2.7)))
+@
+\end{center}
+
+\newpage
+\normalsize
+\subsection{The Non--participating Countries}
+
+The last group with clear patterns can be seen with the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of the United Kingdom in the EMS}, the United Kingdom and Sweden.
+Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom with the "`Black Wednesday"' in September 1992 and the run on the Swedish currency in 1992.
+
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(20, 21), format_moments)))
+@
+\end{longtable}
+
+% could be looking much better
+\begin{center}
+<<sweuk, fig=TRUE, height=7, width=8>>=
+print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Sweden"]))), 100*diff(log(na.omit(HICP[, "UnitedKingdom"])))), xlab="Time", ylab="", main="", screens=list("Sweden", "United Kingdom")), ylim=c(-2,2))
+@
+\end{center}
+
+\newpage
+\normalsize
+\subsection{The Other Countries}
+
+No clear message is available for: Denmark, France, Germany and Portugal.
+
+\subsubsection{Austria}
+
 Lets consider a few countries in broader detail to get an idea about what happened.
@@
 % sizes !!!
-
-\begin{figure}[ht!p]
-  \centering
-    \includegraphics{results-012.pdf}
-  \caption{Series and supLM test for Austria}
-  \label{aut}
-\end{figure}
+<<>>=
+i <- 1
+nam <- colnames(HICP)[i]
+fnam <- paste(nam, ".rda", sep = "")
+@
+%\section*{\Sexpr{nam}}
+<<>>=
+if(file.exists(fnam)) {
+load(fnam)
+} else {
+<<analysis>>
+}
+@
+\begin{center}
+<<autplot, fig=TRUE, height=4, width=8>>=
+<<plot1>>
+@
+\end{center}
 
@@
 \\
+
+\subsubsection{Slovenia}
+
 Lets consider Slovenia as the first Eastern European country that introduced the Euro.
@@
 
-\begin{figure}[ht!p]
-  \centering
-    \includegraphics{results-131.pdf}
-  \caption{Series and supLM test for Slovenia}
-  \label{aut}
-\end{figure}
-
-%
-%The LWZ criteria favors one break. 
-%
-%
-%\begin{figure}[ht!p]
-%  \centering
-%    \includegraphics{results-132.pdf}
-%  \caption{LWZ and Negative Log-Likelihood for Break Point Selection}
-%  \label{aut}
-%\end{figure}
+<<>>=
+i <- 18
+nam <- colnames(HICP)[i]
+fnam <- paste(nam, ".rda", sep = "")
+@
+%\section*{\Sexpr{nam}}
+<<>>=
+if(file.exists(fnam)) {
+load(fnam)
+} else {
+<<analysis>>
+}
+@
+\begin{center}
+<<slovplot, fig=TRUE, height=4, width=8>>=
+<<plot1>>
+@
+\end{center}
 
@@
 
-%As a last example we focus on Spain. \\
-
-%Other possible explanations: 
-%
-%- rise in value added tax (VAT; as in Germany in 2007)
-%- change in central bank policy (for example Spain changed to inflation targeting in 1995; Uk, Swe and Finland in 1992)
-%- role of budget deficits
-
 % Quelle für diese Behaupunten: http://www.oenb.at/isaweb/dyna2.do?go=initIndikatoren&definitionLaden=false&hierarchieIdSelected=108
-
+\newpage
+\subsection{Alternative Explanations}
 Another possible explanation pattern could be the growth in the  money supply.

### 2010-11-04 pkg/paper/Paper/draft.Rnw
Revised draft, some new citations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@69 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \usepackage{multirow}
+\usepackage{amsfonts,amstext,amsmath}
+\usepackage{rotating}
 %% need no \usepackage{Sweave}
@@
 ## packages
+Sys.setlocale("LC_TIME", "English")
 library("glogis")
 library("fxregime")
+library("lattice")
+library("ggplot2")
 
@@
 ## <<tabelle, eval=FALSE>>=
-if(!file.exists(file.path("C:/Dissertation/Paper-1-Z", "Analysis", "Austria.rda"))) {
-setwd(file.path("..", "Analysis"))
-Sweave("analysis.Rnw")
-setwd(file.path("..", "Slides"))
+root_dir <- do.call("file.path", as.list(head(strsplit(getwd(), .Platform$file.sep)[[1]], -1)))
+if(!file.exists(file.path(root_dir, "Analysis", "Austria.rda"))) {
+setwd(file.path(root_dir, "Analysis"))
+Sweave("results.Rnw")
+setwd(file.path(root_dir, "Draft"))
 }
-
 ## query result file names
@@
 
-<<bla, eval=TRUE>>=
-setwd("C:/Dissertation/Paper-1-Z/R-Forge2/pkg/paper/Analysis")
-load("Belgium.Rda")
-xbel <- x
-xbel_bp <- x_bp
-xbel_efp <- x_efp
-xbel_gf <- x_gf
-xbelnbreaks <- x_nbreaks
-xbel_rf <- x_rf
-load("Luxembourg.Rda")
-
-@
-
 <<plotversuch, eval=FALSE>>=
@@
 \Abstract{
-  The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)
-  on changing the dynamics of the inflation rates in its Member States. 
+  The aim of this paper is to investigate how a countries decision to join one or all of the steps leading towards the common European currency changed its inflation rate dynamics.
+Our attempt is to model the inflation rates via a Generalized Logistic Distribution and then test for breaks in the parameters of mean, variance and skewness.
 }
 
-\Keywords{inflation rate, structural break, EMU}
+\Keywords{inflation rate, structural break, EMU, Generalized Logistic Distribution, EURO}
 
@@
 Theory unfortunately is quite unsure  whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
-\citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the risk of deflation.
+\citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the alleged risk of deflation.
 It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
@@
 Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010.
-Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Spain,	Sweden, and	the	United Kingdom.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} The data is obtained from the OECD Statistics.\footnote{http://www.oecd-ilibrary.org/content/data/data-00285-en} \\
-The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia enters in 2011}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Spain and Slovenia)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members without ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
+Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	the	United Kingdom.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} The data is obtained from the OECD Statistics.\footnote{http://www.oecd-ilibrary.org/content/data/data-00285-en} \\
+The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia has not yet introduced the EURO, but will do it in 2011}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members not participating in the ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
 Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
+% alternative link: http://stats.oecd.org/index.aspx
 
 \newpage
-\section{Methods}
+\section{Model}
 
-\subsection{A Generalized Logistic Distribution}
+Our main question of interest is whether or not a countries decision to join one or all of the steps leading towards a common European currency changed its inflation rate dynamics.
+The economic reason why this should be so is rather clear cut for more than half of the countries in our sample.
+We would expect changes in the dynamics of all Eastern European countries due to their transition from previously centrally planned economies towards free market economies.
+All of these countries experienced some problems with very high inflation rates at the beginning and they all were successfull in eliminating excess inflation afterwards, it can be safely assumed that the former member countries of Yugoslavia follow the same pattern.
+In the case of the Southern European countries, with a history of high inflation and less pronounced fiscal discipline, we would also expect inflation dynamics to change towards a regime of lower inflation.
+The case for the Central European countries of course is different.
+There is no reason why the European Central Bank (ECB) should be more bent on fighting inflation than the former Bundesbank.
+Also there is no real theoretical evidence, whether or not a monetary union of former high and low inflation countries would lead to a lower inflation rate in countries with allready low inflation rates.
+This lack of theory of course poses a great challenge.
+\\
+% [allthough it should be added, that the 90ies are perhaps a bad starting point, since we would expect them to be allready rather moderate compared with past values of 60ies and 70ies]
+The approach we use follows from one simple idea, i.e.
+we want to track the evolution of mean, variance and skewness of our series over time.
+Thereby we hope to be able to find and estimate possible breakpoints in the data.
+The correlation and the correlation structure of the time series themselves are of no interest and are treated as a nuisance term.\\
+To do this, we assume that our data is a draw from a Generalized Logistic Distribution (GL) and we proceed by estimating the parameters of the GL--distribution via maximum likelihood.
+We then test for structural change and estimate them using dynamic programming.
+
+\subsection{Generalized Logistic Distribution}
 
 In econometrics, the logistic distribution is often used in income distributions and growth models.
-This is due to  its longer tails, which fits these problems somewhat better.
+This is due to  its fatter tails, which fits these problems somewhat better.
 The use of the generalized logistic distribution as we will use it in this paper is rather rare.
-\citet{won} uses a GL distribution in a regression model with autocorrelated errors and assumes that they follow a GL distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  leptokurtic and severely left or right skewed.
-A similar GL distribution is also used in \citet{tolikas} who analyses extreme risk and value--at--risk in the German stock market, all though they don't use a shape parameter.
+\citet{won} use a GL distribution in a regression model with autocorrelated errors.
+They use this type of distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  nonsymmetric and severely left or right skewed.
+A similar GL distribution is also used in \citet{tolikas} who analyse extreme risk and value--at--risk in the German stock market, all though they don't use a shape parameter.
 Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in relationship with expected inflation.
@@
 To serve this purpose, we use a generalization of the logistic distribution as defined in \citet{johnson}.
-Its probability density is given by:
+For return series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ ($t = 1, \dots, n$) we assume a GL distribution given by:
 
 \begin{eqnarray}
-f(\pi | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma}*\exp^{-\frac{\pi_i-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi_i-\theta}{\sigma}})^{(\delta+1)}}
+f(y | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma} \cdot \exp^{-\frac{y-\theta}{\sigma}}}{(1+\exp^{-\frac{y-\theta}{\sigma}})^{(\delta+1)}} \nonumber
 \end{eqnarray}
@@
 \begin{eqnarray}
-E(\pi_i) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
-Var(\pi_i)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
-Skew(\pi_i) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
+E(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
+Var(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
+Skew(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
 \end{eqnarray}
@@
 \begin{eqnarray}
-l(\delta,\theta,\sigma, \pi) & = & \log(\delta) - \log(\sigma)   \\
-& - &  \frac{1}{\sigma} (\pi-\theta) - (\delta+1) \nonumber \\
-& * & \log (1+\exp^{-\frac{\pi-\theta}{\sigma}}) \nonumber
+\log f(y|\theta,\sigma, \delta) & = & \log(\delta) - \log(\sigma)   \\
+& - &  \frac{1}{\sigma} (y-\theta) - (\delta+1) \nonumber \\
+& \times & \log (1+\exp^{-\frac{y-\theta}{\sigma}}) \nonumber
 \end{eqnarray}
 
-The resulting score function ($\psi()$) for the parameters (the derivatives of the log-likelihood) are given by:
+The resulting score function has 3 components $(s_{\theta}, s_{\sigma}, s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$, where the scores are the derivatives of the log--likelihood function and sum up to 0.
+These are used later on for the construction of the test statistics and are a means of measuring the estimation error in a maximum likelihood framework.
+
+% scores given in dp02_15 p.14
+% bemerkungen zu den herleitungen: siehe herleitung-scores-gen-logistic-neu; wobei vor allem das "`-"' beim term der scale ableitung in mathematica mit seltsamerweise mit "`+"' berechnet wird; beim händischen nachrechnen stellt sich allerdings das - als richtig heraus
 
 \begin{eqnarray}
-\psi(\pi_i,\delta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta}  \\
-& = &  \frac{1}{\delta} - \log(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \nonumber \\
-\psi(\pi_i,\theta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta}  \\
-& = & \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}  \nonumber
-\end{eqnarray}
-
-\begin{eqnarray}
-\psi(\pi_i,\sigma) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma}  \\
-& = & -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) \nonumber \\
-& * &  \quad \frac{\frac{1}{\sigma^2}(\pi-\theta)\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})} \nonumber
+s_{\theta}(y|\theta, \sigma, \delta) & =  &\frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \theta}   \\
+&=&  \frac{1}{\sigma} - (\delta+1)\cdot\frac{\frac{1}{\sigma}\tilde{y}}{(1+\tilde{y})}  \nonumber \\
+s_{\sigma}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \sigma}   \\
+& =&  -\frac{1}{\sigma} + \frac{1}{\sigma^2}(y-\theta) - (\delta+1) \times \frac{\frac{1}{\sigma^2}(y-\theta)\tilde{y}}{(1+\tilde{y})} \nonumber \\
+s_{\delta}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \delta}  \\
+& =&   \frac{1}{\delta} - \log(1+\tilde{y}) \nonumber 
 \end{eqnarray}
@@
 The asymptotic testing theory still holds for this generalization.
-[Beweis] \\
-The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process can be found in \citet{z07}.
+\\
+The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process goes back to \citet{z07}.
 the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - ca be found in \citet{z10}.
 
-\subsection{Tests}
+\subsection{Estimation Framework}
 
-%\subsubsection{Cram{\'e}r-von Mises statistic -- Nyblom-Hansen test}
+We then proceed step by step.
+First, we estimate the prameters of the GL--distribution by means of maximum likelihood, which is possible, if we assume that $\phi = (\theta, \sigma, \delta)$ is stable over time $t$:
 
-We use the Cram{\'e}r von Mises type test as given in \citet{z07}. The test statistic is given by:
+\begin{eqnarray}
+& \hat{\phi} ~=~ \underset{\phi}{argmax} \sum_{t=1}^n \log f(y_t|\phi), \nonumber\\
+& \sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0, \nonumber
+\end{eqnarray}
 
-%\begin{eqnarray}
-%& & \frac{1}{n} \sum_{i=1}^n \Vert efp(\frac{i}{n}) \Vert_2^2
-%\end{eqnarray}
-%
-%i.e., first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically.
-%
-%
-%
-%\subsubsection[The chi-squared test]{The $\chi^2$ test}
-%
-%The main problem is to determine the number of classes k (data is grouped into k classes where we then calculate the difference between observed and expected frequencies).
-In a continuous distribution case, there are no natural boundaries.
-%As given in \citet{kendal} the maximum likelihood estimation of the parameters is not a big problem if k is large enough.
-The classes are taken to cover equal ranges of the variate, which can be done easily once k is determined.
-This procedure was advocated by Mann and Gumberl.
-One rule would be to let k be equal to  $3.765(N-1)^{2/5}$ for a 5\% significance level, when using intervals with equal probability under $H_0$, as given in \citet{boero}.
-We use this rule with a range of [3,2*k-3] for the tests and report any pvalues below 0.05.
+\medskip
 
-For the determination of the numbers of breakpoints, we use the Supremum of LM statistics test and we further investigate into the optimal number of breakpoints using the LWZ criterion, see \citet{z10} for further details.
-To get an idea about the goodness of fit of the GL distribution for a given subsample we also report values of the $\chi^2$ goodness of fit test.
-%
-%
-%\subsection{Densities}
-%
-%Another thing we wish to look at is the change in the moments of the distribution after a break occurred and whether or not the subsample fit of the GL-distribution is preferable to another.
-Although our interest centers on the variance of the respective inflation rate, skewness should not altogether be ignored.
-If we  observe - for example - a change in skewness from positive to negative from one regime to another, we could conclude that now we will observe higher values in the inflation rate since its density shifted to the right.
+Our problem now simplifies to the following question: Is the assumption valid or do the $\phi_t$ vary over time?
+
+\begin{eqnarray}
+& H_0: \phi_t = \phi_0 \quad (t=1,...,n) \nonumber 
+\end{eqnarray}
+
+This can be assessed using the empirical scores $s(y_t|\hat{\phi})$, as given above, as measures of model deviation.
+The empirical scores process $\mathit{efp}(\cdot)$ captures deviations from zero over time and is given by:
+
+\begin{eqnarray}
+\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
+\end{eqnarray}
+
+where $\hat{V}^{-1/2}$ is some consistent estimator for the variance of the scores, it is used to decorrelate the series as to avoid problems due to serial correlation.
+We then employ a Functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ which converges to a 3-dimensional Brownian bridge:
+
+\begin{eqnarray}
+\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot) \nonumber
+\end{eqnarray}
+
+Lets elaborate on this basic idea. Consider some time series given by the following graph. We would not need any test, since we see a clear break at the middle of the sample.
+
+<<fig=TRUE, height=5, width=8>>=
+x <- rnorm(500,0,.1)
+x[251:500] <- rnorm(250,.5,.1) 
+print(plot(x, type="l", ylab = "series", xlab = "Time"))
+abline(lm(x~1))
+@
+
+If we  estimate a simple least squares regression we would get a straight regression line at the mean of the series.
+What we do now is to sum up the scores (which can be viewed as similar to the residuals) by means of the efp (i.e.
+something like a cumulated sum of residuals).
+This summing up of the resiudals would lead to first summing up all the negative residuals down to the lowest point, then  we proceed by summing up the other (positive) residuals.
+The peak of this cumulated sum will be our guess of the breakpoint.
+If this peak crosses some test statistics, then it can be viewed as a significant deviation of our hpyothetized stable parameters.
+
+\subsection{Test}
+
+Actually the $\mathit{efp}(\cdot)$ could be used for aggregated test statistics in various ways.
+In our paper we employ the Supremum of LM (supLM) test by \citet{andrews}.
+This test statistic is given by:
+
+\begin{eqnarray}
+\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
+\end{eqnarray}
+
+where the appropriate p--values can be computed using:
+
+\begin{eqnarray}
+\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert W^0(t) \Vert_2^2}{t(1-t)} \nonumber
+\end{eqnarray}
+
+The supLM test is the single most powerful test for a single abrupt change in unknown time and has the highest power if we suspect a change in more than one parameter of our distribution.
+Throughout the paper we will use just the supLM test.
+
+\subsection{Breakpoint Estimation}
+
+After a break is detected via the supLM test, we proceed by estimating the timing of the break. We estimate $B$ breakpoints $\tau_1,...,\tau_B$ via maximization of the full segmented likelihood:
+
+\begin{eqnarray}
+\sum_{b=1}^{B+1} \sum_{t=\tau_b + 1}^{\tau_b} \log f(y_t | \phi^{(b)}) \nonumber
+\end{eqnarray}
+
+All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming.
 
@@
 Following the Delors Plan with his 3 stages, we have: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of exchange rates of Member States.
-If the EMU had a significant effect upon inflation rate volatility we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of Euro paper money led to a considerable price increase.
+%If the EMU had a significant effect upon inflation rate dynamics, we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of Euro paper money led to a considerable price increase.
+\\
+%[some considerations: 1.
+necessary to include all the results?
+seems to me preferable to deal with them as groups of individual contries rather than all on one page; 2.
+show all the groups?
+3.
+show example like in Slovenian case?]
+
 % http://en.wikipedia.org/wiki/European_Exchange_Rate_Mechanism#Replacement_with_the_euro_and_ERM_II
 % http://en.wikipedia.org/wiki/Eurozone
+% automatic longtable provided by R 
+% rotate table, not sure if this actually looks better
 
-% automatic longtable provided by R
+%\begin{landscape}
 
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(seq(1,21)), format_moments)))
-@
-\end{longtable}
+%\small
+%\begin{longtable}{|l|l|l|l|l|l|l|l|}
+%\hline
+%Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+%\hline
+%<<echo=FALSE, results=tex>>=
+%writeLines(unlist(lapply(c(seq(1,21)), format_moments)))
+%@
+%\end{longtable}
+%\end{landscape}
 
@@
 
-\newpage
 \subsection{Belgium and Luxembourg}
-
-% to show both series with breakpoints and means
-% does not work
-%\begin{center}
-%<<belluxvers, fig=TRUE, height=7, width=8>>=
-%<<plotversuch>>
-%@
-%\end{center}
 
@@
 The first of this group consits of Belgium and Luxembourg.
-This may come as no surprise since both countries formed a monetary union prior even to the start of the European Rate Mechanism (ERM).
+This may come as no surprise since both countries formed a monetary union prior even to the start of the Exchange Rate Mechanism (ERM).
 While both countries enjoyed very stable inflation rates with low values of both mean and variance, this changed around the timing of the irrevocable fixing of the exchange rates.
@@
 
-% heavy increase in variance both countries exactly around EURO introduction 
 % interesting thing: both were allready in a currency union (the belgish--luxembourg union)
+
 \small
@@
 \begin{center}
-<<bellux, fig=TRUE, height=8, width=8>>=
+<<bellux, fig=TRUE, height=5, width=8>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
 print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Belgium"]))), 100*diff(log(na.omit(HICP[, "Luxembourg"])))), xlab="Time", ylab="", main="", screens=list("Belgium", "Luxembourg")), ylim=c(-2,2))
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(5,10,15,17,18), format_moments)))
+writeLines(unlist(lapply(c(3,5,10,15,17,18), format_moments)))
 @
@@
 <<eastern, fig=TRUE, height=7, width=8>>=
-print(xyplot(cbind(100*diff(log(na.omit(HICP[, "CzechRepublic"]))), 100*diff(log(na.omit(HICP[, "Estonia"]))), 100*diff(log(na.omit(HICP[, "Hungary"]))), 100*diff(log(na.omit(HICP[, "Poland"]))), 100*diff(log(na.omit(HICP[, "Slovakia"]))), 100*diff(log(na.omit(HICP[, "Slovenia"])))), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia")))
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(xyplot(cbind(100*diff(log(na.omit(HICP[, "CzechRepublic"]))), 100*diff(log(na.omit(HICP[, "Estonia"]))), 100*diff(log(na.omit(HICP[, "Hungary"]))), 100*diff(log(na.omit(HICP[, "Poland"]))), 100*diff(log(na.omit(HICP[, "Slovakia"]))), 100*diff(log(na.omit(HICP[, "Slovenia"])))), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia"), ylim=c(-2.7,2.7)))
 @
@@
 % could be looking much better
+
 \begin{center}
-<<itaspa, fig=TRUE, height=7, width=8>>=
+<<itaspa, fig=TRUE, height=5, width=8>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
 print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Italy"]))), 100*diff(log(na.omit(HICP[, "Spain"])))), xlab="Time", ylab="", main="", screens=list("Italy", "Spain"), ylim=c(-2,2)))
@@
 \begin{center}
-<<ireland, fig=TRUE, height=7, width=8>>=
+<<ireland, fig=TRUE, height=4, width=8>>=
 plot(100*diff(log(na.omit(HICP[, "Ireland"]))), main="Ireland", ylab="", xlab ="Time", ylim=c(-2,2))
@@
 \begin{center}
-<<nochange, fig=TRUE, height=7, width=8>>=
+<<nochange, fig=TRUE, height=5, width=8>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
 print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Finland"]))), 100*diff(log(na.omit(HICP[, "Greece"]))), 100*diff(log(na.omit(HICP[, "Netherlands"])))), xlab="Time", ylab="", main="", screens=list("Finland", "Greece", "Netherlands"), ylim=c(-2.7,2.7)))
@@
 
-No clear message is available for: Denmark, France, Germany and Portugal.
+No clear message is available for: Austria, Denmark, France, Germany and Portugal.
 
-\subsubsection{Austria}
+\small
+\begin{longtable}{|l|l|l|l|l|l|l|l|}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(1,4,7,8,16), format_moments)))
+@
+\end{longtable}
 
-Lets consider a few countries in broader detail to get an idea about what happened.
-In Austria we see evidence of a change in the inflation rate, when the old regime ended in September 2007.
-The new regime is much more volatile, with an almost threefold increase in variance.
-
-% sizes !!!
-<<>>=
-i <- 1
-nam <- colnames(HICP)[i]
-fnam <- paste(nam, ".rda", sep = "")
-@
-%\section*{\Sexpr{nam}}
-<<>>=
-if(file.exists(fnam)) {
-load(fnam)
-} else {
-<<analysis>>
-}
-@
 \begin{center}
-<<autplot, fig=TRUE, height=4, width=8>>=
-<<plot1>>
+<<nopattern, fig=TRUE, height=7, width=8>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Austria"]))), 100*diff(log(na.omit(HICP[, "Denmark"]))), 100*diff(log(na.omit(HICP[, "France"]))), 100*diff(log(na.omit(HICP[, "Germany"]))), 100*diff(log(na.omit(HICP[, "Portugal"])))), xlab="Time", ylab="", main="", screens=list("Austria", "Denmark", "France", "Germany", "Portugal"), ylim=c(-2,2)))
 @
@@
 
-The reason for this is a strong increase in the yearly Austrian inflation rate following a heavy increase in oil price which was further amplified by a rise of mineral taxes.
-However, we fail to find any evidence for an incrase in mean or volatility of the Austrian rate following the introduction of the Euro paper money.
+\subsection{An example--Slovenia}
+
+Let us go through all the steps of our analysis by means of an example.
+An ideal example for this is the case of Slovenia.
+The economic reason why we suppose to find changing dynamics are clear.
+In the first phase of the consolidation after the break--up of Yugoslavia and after the short war of 1991, Slovenia experienced very high inflation rates,
+with yearly growth rates as high as 90\% in August 1990.
+But gradually Slovenia was able to reduce the inflation rate towards much lower values.
+This first phase is not captured in our data since it starts as late as January 1996 for Slovenia.
+For the period up to today we would expect one or two breaks from an economic point of view.
+The efforts to reach the Maastricht criteria as soon as possible are likely to lead to a further lowering of the mean and possibly also of the variance of the inflation rate.
+Another break point might possibly be found towards the introduction of the EURO coins themselves in 2007.
 \\
-
-\subsubsection{Slovenia}
-
-Lets consider Slovenia as the first Eastern European country that introduced the Euro.
-In August 2003 the new regime started with a much lower mean value (only a third now of its previous value) and interestingly enough, a higher volatility.
+Actually when we look at the data, this is precisely what we get:
 
@@
 
-Slovenia is a good example of the benefits of an early Euro adoption.
-Its efforts to meet the criterias as set forth in the Maastricht treaty were successfull, actually Slovenia reached them in 2005.
-So from 2003 onwards inflation was lower but with a much higher variance.
-Regarding the dating in 2003, we have to keep in mind that this was the year where most of the financial reforms were introduced.
-Slovenia seems to be the only case where the introduction of the Euro was followed by an increase in yearly inflation, which faded away following the financial crises and its deflationary effects.
+To be as precise as possible with break points, we also include a graph showing the values of the LWZ (Liu, Wu and Zidek) criteria as developed in \citet{liu}.
+Whereas the supLM test suggest at least one break, the LWZ criterion favors no break.
+But the decline in the negative log--likelihood from 0 breaks to 1 is the strongest, we concluded that one break point actually would be the best alternative.
+
+\begin{center}
+<<slovplotbreak, fig=TRUE, height=6, width=8>>=
+<<plot2>>
+@
+\end{center}
+
+We find a break in August 2003 d with a much lower mean value (only a third now of its previous value) and interestingly enough, a higher volatility.
+The supLM test graph on the right hand side suggests one break around 2003 and possibly another towards the end of the sample.
+This second break point is not estimated, since we restricted a regime towars a length of at least 24 month.
 \\
+Slovenia did achieve this by two very important measures.
+First it liberalized the financial markets in 2003 and -- most important -- it restricted the year--on--year growth of the monetary base M1 severely starting in 2003.
+Combined these were enough to ensure  Slovenian participation in the EURO--zone.
 
-% Quelle für diese Behaupunten: http://www.oenb.at/isaweb/dyna2.do?go=initIndikatoren&definitionLaden=false&hierarchieIdSelected=108
-\newpage
-\subsection{Alternative Explanations}
-Another possible explanation pattern could be the growth in the  money supply.
+%% Quelle für diese Behaupunten: http://www.oenb.at/isaweb/dyna2.do?go=initIndikatoren&definitionLaden=false&hierarchieIdSelected=108
+%\newpage
+%\subsection{Alternative Explanations}
+%Another possible explanation pattern could be the growth in the  money supply.
 A drawback is the current lack of data\footnote{hopefully to be solved}.
@@
 \\
-In Spain we see a stronger increase in M1 in May and June 1996 (nothing for the other break, data scarceity).
+%In Spain we see a stronger increase in M1 in May and June 1996 (nothing for the other break, data scarceity).
 In Portugal we find a modest increase in year on year M1 money supply growth starting in July 1992.
@@
 Poland starts contracting the year on year M1 money supply after June 2000 with very low growth rates till January 2002.
-
-%\readme{irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck}
-
-% for denmark: see: C:\Dissertation\Daten\DSI\M1\daten - the denmark analysis, alos: Spain, Portugal, Slovakia, Hungary Estonia noch am absatz für M1
-
-Concerning monetary unions, it is interesting to notice that Belgium and Luxembourg, that constituted a monetary union prior to the European Monetary Union, did not experience the same timing in their break towards a higher mean and a significantly higher volatility regime.
-
-With regard to the alleged benefit of monetary unions, the import of credibility to former high inflation countries (like Italy and Spain), did nothing to decrease their mean or volatility, rather to the contrary.
+%
+%%\readme{irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck}
+%
+%
+%% for denmark: see: C:\Dissertation\Daten\DSI\M1\daten - the denmark analysis, alos: Spain, Portugal, Slovakia, Hungary Estonia noch am absatz für M1
+%
+%Concerning monetary unions, it is interesting to notice that Belgium and Luxembourg, that constituted a monetary union prior to the European Monetary Union, did not experience the same timing in their break towards a higher mean and a significantly higher volatility regime.
+%
+%With regard to the alleged benefit of monetary unions, the import of credibility to former high inflation countries (like Italy and Spain), did nothing to decrease their mean or volatility, rather to the contrary.
 Spain and Italy did have a -- rather -- low mean low volatility regime from 1996:6 to 2000:12 but then things changed towards higher volatility.
 It appears as if both countries tried to decrease inflation prior to the fixing of the exchange rates and let go afterwards.
-
-Following the reasoning of a study by \cite{duarte}, high volatility of inflation rates within regions of a union may be explained a countries decision to lower the inflation differential by means of fiscal policy.
+%
+%Following the reasoning of a study by \cite{duarte}, high volatility of inflation rates within regions of a union may be explained a countries decision to lower the inflation differential by means of fiscal policy.
 This explanation might fit to Spain and Italy, both were above the the average EU inflation.
+%
 
@@
 
+We were able to outline changes in the dynamics of the inflation rates in various countries across Europe.
+However, we fail to find any clear pattern relating to the EURO introduction, but find clear evidence that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
+
 \clearpage

### 2010-11-04 pkg/paper/Paper/draft.Rnw
removed superfluous code chunks; added some text
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@70 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-<<plotversuch, eval=FALSE>>=
-par(mfrow = c(2, 1))
-plot(x, main = "Series with Fitted Mean", ylab = "100 * diff(log(HICP))", xlab = "Time")
-lines(x_bp, breaks = x_nbreaks)
-lines(fitted(x_bp, breaks = x_nbreaks, type = "mean"), col = 4)
-plot(xbel, main = "Series with Fitted Mean", ylab = "100 * diff(log(HICP))", xlab = "Time")
-lines(xbel_bp, breaks = xbelnbreaks)
-lines(fitted(xbel_bp, breaks = xbelnbreaks, type = "mean"), col = 4)
-@
-
-% plot "macros" (assuming existence "x", "x_gf", ...)
+## plot "macros" (assuming existence "x", "x_gf", ...)
 <<plot1, eval=FALSE>>=
@@
 
-<<plot1d, eval=FALSE>>=
-plot(x, main = "Slovenia", ylab = "100 * diff(log(HICP))", xlab = "Time")
-@
-
-<<plot1b, eval=FALSE>>=
-plot(x_efp, functional = supLM(0.1), main = "supLM test")
-lines(x_bp, breaks = x_nbreaks)
-@
-
 <<plot2, eval=FALSE>>=
@@
 plot(x_bp)
-@
-
-<<plot3, eval=FALSE>>=
-par(mfrow = c(1, x_nbreaks + 1))
-for(i in 1:(x_nbreaks+1)) plot(x_rf[[i]], xlab = paste(names(x_rf)[i], "\n", "Goodness-of-fit p-value: ",
-  format.pval(summary(x_rf[[i]])$chisq.test$p.value, digits = 4), sep = ""))
-@
-
-<<plot4, eval=FALSE>>=
-par(mfrow = c(1, 1))
-plot(x_efp, aggregate = FALSE, ylim=c(-2.1,2.1))
 @
@@
 Following the Delors Plan with his 3 stages, we have: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of exchange rates of Member States.
+\\
+It should be pointed out however, that these monthly series do still contain all the saisonality and that if we correct for this by using the lag of order 12 instead of order 1, we find somewhat less pronounced changes with the exception of the latest observations, with the financial crises and the overall lowering in inflation rates, which of course leads to a considerable increase in variance for more than half of our sample.
+5 countries in our sample then show a break in October 2007, coinciding with the ongoing financial crises.
+
+% interesting graph to financial crisis: http://de.wikipedia.org/wiki/Datei:Einlagefazilit%C3%A4t.PNG
+
 %If the EMU had a significant effect upon inflation rate dynamics, we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of Euro paper money led to a considerable price increase.
@@
 While both countries enjoyed very stable inflation rates with low values of both mean and variance, this changed around the timing of the irrevocable fixing of the exchange rates.
-A causal connection to this events seems likely.
+A causal connection to this events seems possible.
+However, according to an official of Statec Luxembourg\footnote{www.statec.lu}, these changes can be traced back  to the inclusion of sales (winter and summer) from january 1999 on according to a change in regulation.
+For Belgium we suspect similar change, but our suspicion is not yet confirmed.
+
+% Quelle: e-mail Verkehr mit Jérôme Hury, dem Chef der Statec Price Abteilung
 
 % interesting thing: both were allready in a currency union (the belgish--luxembourg union)
+% Luxembourg and probably in Belgium too, as  you find out sales prices were excluded before the regulation
 
@@
 However, we fail to find any clear pattern relating to the EURO introduction, but find clear evidence that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
+\\
+Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases can be traced back to the strong seasonality in our time series.
 

### 2010-11-07 pkg/paper/Paper/draft.Rnw
updated Draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@71 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 ## packages
-Sys.setlocale("LC_TIME", "English")
+if(Sys.getlocale() != "C") Sys.setlocale("LC_TIME", "English")
 library("glogis")
@@
 library("lattice")
-library("ggplot2")
 
@@
 ## sequentially load data
+## old: load(file.path("..", "Analysis", fnam[i]))
+
 for(i in seq_along(fnam)) {
-  load(file.path("..", "Analysis", fnam[i]))
+  load(file.path(root_dir, "Analysis", fnam[i]))
   results[[i]] <- list(
@@
   return(rval)
+}
+
+xyplot.gefp <- function(x, panel = NULL, ylim = NULL, ...)
+{
+  cval <- maxBB$computeCritval(0.05, ncol(x$process))
+  if(is.null(ylim)) ylim <- range(c(range(x$process), 1.1 * cval, -1.1 * cval))
+
+  gefp_plot <- function(x, y, subscripts, groups, panel = panel.xyplot,
+    col = 1, type = "p", pch = 20, lty = 1, lwd = 1, ...)
+  {
+    col <- rep(as.list(col), length = nlevels(groups))
+    type <- rep(as.list(type), length = nlevels(groups))
+    pch <- rep(as.list(pch), length = nlevels(groups))
+    lty <- rep(as.list(lty), length = nlevels(groups))
+    lwd <- rep(as.list(lwd), length = nlevels(groups))
+
+    for(g in 1:nlevels(groups)) {
+      idx <- levels(groups)[g] == groups[subscripts]
+      if (any(idx)) panel(x[idx], y[idx], ...,
+        col = col[[g]], type = type[[g]], pch = pch[[g]],
+        lty = lty[[g]], lwd = lwd[[g]])
+      grid::grid.lines(y = grid::unit(0, "native"))
+      grid::grid.lines(y = grid::unit(cval, "native"), gp = grid::gpar(col = 2))
+      grid::grid.lines(y = grid::unit(-cval, "native"), gp = grid::gpar(col = 2))
+    }
+  }
+  if(is.null(panel)) panel <- gefp_plot
+
+  xyplot(x$process, panel = panel, ylim = ylim, ...)
 }
@@
   The aim of this paper is to investigate how a countries decision to join one or all of the steps leading towards the common European currency changed its inflation rate dynamics.
-Our attempt is to model the inflation rates via a Generalized Logistic Distribution and then test for breaks in the parameters of mean, variance and skewness.
+Our attempt is to model the inflation rates via a Generalized Logistic Distribution (GL) and then test for breaks in the parameters of the distribution.
 }
@@
 % no literature overview
-Theory unfortunately is quite unsure  whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
+Theory unfortunately does not provide any clear foundation for a proper answer whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
 \citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the alleged risk of deflation.
@@
 %But monetary policy in a monetary union does not explicitly stabilize the output gap and inflation rate in case of national economic shocks. 
-The effects of joining a monetary union on inflation rate variability depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-Due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, theory has some troubles establishing a sound theoretical model.
+The effects of joining a monetary union on inflation rate variability depends on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
+Due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependend on assumptions.
 \\
@@
 In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
-They find that convergence in time of inflation dynamics was only partly observable.
+They find that convergence over time of inflation dynamics was only partly observable.
 In a paper studying core inflation and using an aggregated Euro Area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
@@
 Inquiring into the convergence properties of inflation rates among countries of the EMU, \citet{busetti} find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
+This does not quite go along with our findings, that do not indicate stronger divergence after 1997 by means of the detection of break points.
 \\
+In section 2 we will present the data, in section 3 we will present our approach, in section 4 we will present the results, in section 5 we will discuss our findings and try to bring them into context with the overall economic condition, section 6 concludes.
 
@@
 
-Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010.
+Inflation is measured as the monthly change in the HICP rates from 1.1990---3.2010.
 Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	the	United Kingdom.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} The data is obtained from the OECD Statistics.\footnote{http://www.oecd-ilibrary.org/content/data/data-00285-en} \\
-The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia has not yet introduced the EURO, but will do it in 2011}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members not participating in the ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
+The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia has not yet introduced the EURO, but in 2011.}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members not participating in the ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
 Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
@@
 We would expect changes in the dynamics of all Eastern European countries due to their transition from previously centrally planned economies towards free market economies.
-All of these countries experienced some problems with very high inflation rates at the beginning and they all were successfull in eliminating excess inflation afterwards, it can be safely assumed that the former member countries of Yugoslavia follow the same pattern.
+All of these countries experienced some problems with very high inflation rates at the beginning and they all were successfull in eliminating excess inflation afterwards.
+It can be safely assumed that the former constituent republics of Yugoslavia follow the same pattern.
 In the case of the Southern European countries, with a history of high inflation and less pronounced fiscal discipline, we would also expect inflation dynamics to change towards a regime of lower inflation.
 The case for the Central European countries of course is different.
-There is no reason why the European Central Bank (ECB) should be more bent on fighting inflation than the former Bundesbank.
-Also there is no real theoretical evidence, whether or not a monetary union of former high and low inflation countries would lead to a lower inflation rate in countries with allready low inflation rates.
-This lack of theory of course poses a great challenge.
+There is no reason why the European Central Bank (ECB) should actually  follow the rather stable monetary policies of the former German central bank in a one--to--one fashion.
 \\
-% [allthough it should be added, that the 90ies are perhaps a bad starting point, since we would expect them to be allready rather moderate compared with past values of 60ies and 70ies]
-The approach we use follows from one simple idea, i.e.
+% [although it should be added, that the 90ies are perhaps a bad starting point, since we would expect them to be already rather moderate compared with past values of 60ies and 70ies]
+The approach we use follows one simple idea: i.e.
 we want to track the evolution of mean, variance and skewness of our series over time.
@@
 In econometrics, the logistic distribution is often used in income distributions and growth models.
-This is due to  its fatter tails, which fits these problems somewhat better.
-The use of the generalized logistic distribution as we will use it in this paper is rather rare.
+This is due to  its fatter tails, which fit these kind of data somewhat better.
+The application of the generalized logistic distribution as we will use it in this paper is rather rare.
 \citet{won} use a GL distribution in a regression model with autocorrelated errors.
-They use this type of distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  nonsymmetric and severely left or right skewed.
-A similar GL distribution is also used in \citet{tolikas} who analyse extreme risk and value--at--risk in the German stock market, all though they don't use a shape parameter.
-Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in relationship with expected inflation.
+They use this type of distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  non symmetric and severely left or right skewed.
+A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they don't use a shape parameter.
+Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context to expected inflation.
 \citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
@@
 Prior to that, we would like to give a short justification for the usage of the GL distribution.
-Regarding the data at hand, it was not possible to use the already existing method developed in \citet{z07}, since almost all inflation rates, with the notable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties.
+Regarding the data at hand, it was not possible to utilize the already existing method developed in \citet{z07}, since almost all inflation rates, with the notable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties.
 \\
@@
 To serve this purpose, we use a generalization of the logistic distribution as defined in \citet{johnson}.
-For return series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ ($t = 1, \dots, n$) we assume a GL distribution given by:
+
+\clearpage
+For our inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we assume a GL distribution given by:
 
@@
 For $\delta$=1 the distribution simplifies to the logistic distribution, for $\delta$$<$1 it is skewed to the left and for $\delta$$>$1 it is skewed to the right.
-The moments are given by:
+The moments are:
 
@@
 
-where $\gamma()$ is the digamma function and its derivatives. \\
-The log-likelihood is given by:
+where $\gamma(\cdot)$ is the digamma function and its derivatives. \\
+The log-likelihood is:
 
@@
 
-The resulting score function has 3 components $(s_{\theta}, s_{\sigma}, s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$, where the scores are the derivatives of the log--likelihood function and sum up to 0.
+The resulting score function, where the scores are the derivatives of the log--likelihood function, has 3 components $(s_{\theta}, s_{\sigma}, s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$.
 These are used later on for the construction of the test statistics and are a means of measuring the estimation error in a maximum likelihood framework.
@@
 
-An enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a generalized logistic distribution (GL) is provided with this paper.
+An enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a GL distribution is provided with this paper.
 The asymptotic testing theory still holds for this generalization.
 \\
-The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process goes back to \citet{z07}.
-the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - ca be found in \citet{z10}.
+The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process can be backtracked to \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
 
@@
 
-We then proceed step by step.
-First, we estimate the prameters of the GL--distribution by means of maximum likelihood, which is possible, if we assume that $\phi = (\theta, \sigma, \delta)$ is stable over time $t$:
+In order to analyze our data, we first estimate the parameters of the GL--distribution by means of maximum likelihood, which is possible, if we assume that $\phi = (\theta, \sigma, \delta)$ is stable over time $t$:
 
@@
 
-Our problem now simplifies to the following question: Is the assumption valid or do the $\phi_t$ vary over time?
+Our problem now simplifies to the following question: Is the assumption valid or do the $\phi_t$ change?
 
@@
 
-This can be assessed using the empirical scores $s(y_t|\hat{\phi})$, as given above, as measures of model deviation.
+This can be assessed using the empirical scores $s(y_t|\hat{\phi})$, as presented above, as measures of model deviation.
 The empirical scores process $\mathit{efp}(\cdot)$ captures deviations from zero over time and is given by:
@@
 
-where $\hat{V}^{-1/2}$ is some consistent estimator for the variance of the scores, it is used to decorrelate the series as to avoid problems due to serial correlation.
+where $\hat{V}^{-1/2}$ is some consistent estimator for the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
 We then employ a Functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ which converges to a 3-dimensional Brownian bridge:
@@
 
+\clearpage
 Lets elaborate on this basic idea. Consider some time series given by the following graph. We would not need any test, since we see a clear break at the middle of the sample.
@@
 
-If we  estimate a simple least squares regression we would get a straight regression line at the mean of the series.
+If we  estimate a simple least squares regression we get a straight regression line at the mean of the series.
 What we do now is to sum up the scores (which can be viewed as similar to the residuals) by means of the efp (i.e.
 something like a cumulated sum of residuals).
-This summing up of the resiudals would lead to first summing up all the negative residuals down to the lowest point, then  we proceed by summing up the other (positive) residuals.
-The peak of this cumulated sum will be our guess of the breakpoint.
-If this peak crosses some test statistics, then it can be viewed as a significant deviation of our hpyothetized stable parameters.
+This summing up of the residuals would be constructed by first aggregating all the negative residuals down to the lowest point, then  we proceed by accruing the other (positive) residuals.
+The bottom of this cumulated sum will be our estimation of the breakpoint.
+If this peak crosses some test statistics, then it can be viewed as a significant deviation from the null hypothesis.
 
@@
 
-The supLM test is the single most powerful test for a single abrupt change in unknown time and has the highest power if we suspect a change in more than one parameter of our distribution.
-Throughout the paper we will use just the supLM test.
+The supLM test is the most powerful test for a single abrupt change in unknown time and has the highest power if more than one parameter of the distribution changes over time.
+Throughout the paper we will use the supLM test.
 
@@
 
-All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming.
+All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming. More details for this estimation technique are provided in \cite{z10}.
 
@@
 To help with the interpretation, we give a short overview over the history of the EMU.
-Following the Delors Plan with his 3 stages, we have: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of exchange rates of Member States.
+It basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of the exchange rates of the Member States.
 \\
-It should be pointed out however, that these monthly series do still contain all the saisonality and that if we correct for this by using the lag of order 12 instead of order 1, we find somewhat less pronounced changes with the exception of the latest observations, with the financial crises and the overall lowering in inflation rates, which of course leads to a considerable increase in variance for more than half of our sample.
-5 countries in our sample then show a break in October 2007, coinciding with the ongoing financial crises.
+It should be pointed out however, that these monthly series do still contain all the seasonality\footnote{We would like to mention that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal proquirements.
+In addition to Belgium and Luxembourg, there are a number of other states, where the sales periods are regulated by law.
+France has an individual time span regulated by each department.
+This should pose no problem, since there is no global state level solution.
+In Greece there are also two set sales--periods in winter and summer.
+Given the probelems concerning the Greek data, this does not pose any addional difficulty.
+In Italy the sales periods are fixed each year by the Chambers of Commerce.
+Portugal has two set sales periods as well, but for two month (January and February as well das Mid--July to Mid--September.
+The source for this information is: $http://www.konsumenteuropa.se/en/Shopping-in-Europe$}.
+If we correct for this by using the lag of order 12 instead of order 1, we find somewhat less pronounced changes with the exception of the latest observations, with the financial crises and the overall lowering in inflation rates, which of course leads to a considerable increase in variance for more than half of our sample.
+5 countries change their inflation dynamics after October 2007, coinciding with the ongoing financial crises.
 
@@
 necessary to include all the results?
-seems to me preferable to deal with them as groups of individual contries rather than all on one page; 2.
+seems to me preferable to deal with them as groups of individual countries rather than all on one page; 2.
 show all the groups?
@@
 On the basis of these results, we are able to trace out different groups.
-The first of this group consits of Belgium and Luxembourg.
-This may come as no surprise since both countries formed a monetary union prior even to the start of the Exchange Rate Mechanism (ERM).
+The first of this groups consists of Belgium and Luxembourg.
+This may come as no surprise since both countries formed a monetary union prior to the start of the Exchange Rate Mechanism (ERM).
 While both countries enjoyed very stable inflation rates with low values of both mean and variance, this changed around the timing of the irrevocable fixing of the exchange rates.
-A causal connection to this events seems possible.
-However, according to an official of Statec Luxembourg\footnote{www.statec.lu}, these changes can be traced back  to the inclusion of sales (winter and summer) from january 1999 on according to a change in regulation.
-For Belgium we suspect similar change, but our suspicion is not yet confirmed.
+But in the case of Belgium and Luxembourg, what we actually observe is a change due to the definition of the HICP.
+According to an official of the OECD, these changes can be traced back  to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation.\footnote{for technical details see: http://eur-lex.europa.eu/LexUriServ/LexUriServ.do?uri=OJ:L:2000:300:0016:0017:EN:PDF} This lead to a 6 month seasonality pattern (January and July) afterwards (meaning after the breakpoints induced by sales inclusion), since sales periods in Belgium and Luxembourg are regulated by law and have to occur in the first two weeks of these months and therefor we observe this sharp price decreases in both month.
+Interestingly enough, both these price decreases are immediately followed by increases in February and August of even higher magnitude.
 
-% Quelle: e-mail Verkehr mit Jérôme Hury, dem Chef der Statec Price Abteilung
-
-% interesting thing: both were allready in a currency union (the belgish--luxembourg union)
-% Luxembourg and probably in Belgium too, as  you find out sales prices were excluded before the regulation
+%%%%% Seasonality Problems:
+% for the others: http://www.konsumenteuropa.se/en/Shopping-in-Europe/Luxembourg/
+% we do not find many more regulated sales periods in Europe (given the data above is correct, which I hope :)
+% exemption: France (regulated by each department individually (so changes hopefully leveled out) --> no problem
+% 	         Greece (In Greece there are two set sales-periods. Winter sales: usually during January – February. Summer sales: usually during July – August.)
+%            --> no problem, dont trust series anyway
+%            Italy (There are two set sales periods in Italy, in winter and summer. Dates and duration are fixed by the Chambers of Commerce.)
+%            --> what if these changes???  --> e-mail sent to italian chamber of commerce in austria, awaiting reply
+%            Portugal (There are two set sale periods: Winter sales: 28/12 to 28/02 Summer sales: 15/07 to 15/09 )
+% Quelle für diese Änderung : e-mail Verkehr mit Jérôme Hury, dem Chef der Statec Price Abteilung; und Anne-Sophie.FRAISSE@oecd.org
 
@@
 
-The second of these groups are the Eastern European Countries, the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
+The second of these groups are the Eastern European countries, the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
 In all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
@@
 This trend changed drastically soon after the fixing of the exchange rates leading to higher mean and much higher variances then ever before since the beginning of the HICP series.
+\\
+Italy is quite a nice example to explain the usefullness of the skewness parameter.
+If we focus on the first two periods, with skewness lower than 1 (which indicates that the distribution is skewed to the right) we find that positive shocks are not normally neutralized by negative shocks.
+This changes in the last phase with skewness negative, indicating that the GL is now skewed to the left and that positive shocks will more likely be followed by negative shocks.
+
+% skewness: if delta = 1.1 --> skweness is 0.1, is positive, skewed to the right: linkssteil
+% ie. negative skew (delta <1: rechtssteil, in englisch: skewed to the left 
+% (psigamma(0.5, deriv = 2) - psigamma(1, deriv = 2))/(psigamma(1, deriv = 1) + psigamma(0.5, deriv = 1))^(3/2)
 
@@
 It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.
-We find a structural break in 2008 following the strong contraction of the Irish housing bubble and a beginning deflation.
+We find a structural break in 2008 following the strong contraction triggerd by the bursting of the Irish housing bubble and a beginning deflation.
 
@@
 
-Another interesting group consists of 3 economically very different countries.
+Another interesting group consists of 3 economically rather different countries.
 In all of them we find no evidence of a structural change.
 Whereas the Netherlands and Finland are both on the lower side of the inflation margin, Greece exhibits some very strange behaviour.
-It has by far the highest variance of all series and seems to be completely different from all the others, just like some randomly generated series.
+It has by far the highest variance of all series and seems to be completely different from all the others.
 
@@
 The last group with clear patterns can be seen with the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of the United Kingdom in the EMS}, the United Kingdom and Sweden.
-Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom with the "`Black Wednesday"' in September 1992 and the run on the Swedish currency in 1992.
+Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom cumulating into the "Black Wednesday" in 1992 and the run on the Swedish currency in the same year.
 
@@
 
-No clear message is available for: Austria, Denmark, France, Germany and Portugal.
+No clear interpretation is available for: Austria, Denmark, France, Germany and Portugal.
 
@@
 \subsection{An example--Slovenia}
+\normalsize
 
-Let us go through all the steps of our analysis by means of an example.
-An ideal example for this is the case of Slovenia.
+Let us go through all the steps of our analysis by means of an example, for this we have chosen the case of Slovenia.
 The economic reason why we suppose to find changing dynamics are clear.
@@
 But gradually Slovenia was able to reduce the inflation rate towards much lower values.
-This first phase is not captured in our data since it starts as late as January 1996 for Slovenia.
+This first phase is not captured in our data since it starts as late as January 1996.
 For the period up to today we would expect one or two breaks from an economic point of view.
@@
 
-<<>>=
-i <- 18
-nam <- colnames(HICP)[i]
-fnam <- paste(nam, ".rda", sep = "")
+<<fig=TRUE, height=4, width=8>>=
+x <- results$Slovenia
+plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "")
+lines(x$breakpoints, breaks = x$nbreaks)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
 @
-%\section*{\Sexpr{nam}}
-<<>>=
-if(file.exists(fnam)) {
-load(fnam)
-} else {
-<<analysis>>
-}
+
+<<fig=TRUE, height=4, width=8>>=
+plot(x$efp, functional = supLM(0.1), main = "supLM test", xaxt = "n", xlim = range(time(x$data)))
+Axis(time(x$data), side = 1)
+lines(x$breakpoints, breaks = x$nbreaks)
 @
-\begin{center}
-<<slovplot, fig=TRUE, height=4, width=8>>=
-<<plot1>>
-@
-\end{center}
 
-To be as precise as possible with break points, we also include a graph showing the values of the LWZ (Liu, Wu and Zidek) criteria as developed in \citet{liu}.
+The supLM test graph  suggests one break around 2003 and possibly another towards the end of the sample.
+This second break point is not estimated, since we restricted a regime towards a length of at least 24 month.
+Thus we find one break in August 2003 with a much lower mean value (only a third now of its previous value) and interestingly enough, a higher volatility.\\
+To be as precise as possible with the detection of break points, we also include a graph showing the values of the LWZ (Liu, Wu and Zidek) criteria as developed in \citet{liu}.
 Whereas the supLM test suggest at least one break, the LWZ criterion favors no break.
@@
 
-We find a break in August 2003 d with a much lower mean value (only a third now of its previous value) and interestingly enough, a higher volatility.
-The supLM test graph on the right hand side suggests one break around 2003 and possibly another towards the end of the sample.
-This second break point is not estimated, since we restricted a regime towars a length of at least 24 month.
-\\
 Slovenia did achieve this by two very important measures.
 First it liberalized the financial markets in 2003 and -- most important -- it restricted the year--on--year growth of the monetary base M1 severely starting in 2003.
-Combined these were enough to ensure  Slovenian participation in the EURO--zone.
+Combined, these actions were enough to ensure  Slovenian participation in the EURO--zone.
 
@@
 \clearpage
+\section{Discussion}
+
+% actually too detailed for so small a picture
+
+\begin{center}
+<<greyplot, fig=TRUE, height=8, width=8>>=
+ ## compute GL mean (if no NA)
+  glmean <- function(x) {
+    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
+  }
+  ## returns
+  ## why not include Finland?
+  ## hicp <- diff(log(HICP[, -6]))
+  hicp <- diff(log(HICP[,]))
+
+  ## rolling GL means
+  hicp_glmean <- rollapply(hicp, 12, glmean)
+
+  ## semi-transparent gray and red
+  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+
+  ## plot returns (gray) and rolling means (red)
+  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
+  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+@
+\end{center}
+
+To clarify some issues, lets take a look at the data aggregated in just one graph.
+If more countries have the same inflation rate in a given month, the lines are overlayed and thus get darker.\footnote{We wold seee one black line if all countries would show the same month--on--month increase at any one time.} We do see, that many of these countries follow a rather similar path.
+One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
+However, this was not enough for most of the countries -- with the exception of Ireland -- to lead towards a structural break in inflation rate dynamics.
+\\
+One interesting aspect we would like to focus on is the clear trend in means.
+What we have plotted here actually is a rolling mean, where the mean is calculated using the first moment of a fitted GL distribution.
+It is clearly visible that these rates converged towards some common mean (graphically emphasized by the redder lines) which, from 2005 onwards, comes rather close to the ECB target rate (which would translate to a monthly increase of roughly 2 per mill).
+So concerning the mean of the inflation rates, we see a clear convergence pattern.
+However, this is not the case for variance where we find clear evidence of an increase.
+This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
+\\
+
+% message of these graphs not so clear
+%On this graph we just show the current EURO member countries\footnote{with Estonia included}. The clear trend towards a common rather lower mean is clearly visible.
+% just focus on EURO countries
+%\begin{center}
+%<<greyploteuro, fig=TRUE, height=8, width=8>>=
+%  plot(hicp[, c(1,2,5,6,7,8,9,11,12,13,14,16,17,18,19)], plot.type = "single", lwd = 1.5, col = gray_red[1], ylab="EURO", xlab="Time")
+%  for(i in c(1,2,5,6,7,8,9,11,12,13,14,16,17,18,19)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+%@
+%\end{center}
+%
+%This convergence seems to be missing regarding the countries not part of the ERM~II.
+%
+%
+%% just focus on Non-ERM~II countries
+%\begin{center}
+%<<greyplotnerm2, fig=TRUE, height=8, width=8>>=
+%  plot(hicp[, c(3,10,15,20,21)], plot.type = "single", lwd = 1.5, col = gray_red[1], ylab="Non ERM II", xlab="Time")
+%  for(i in c(3,10,15,20,21)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+%@
+%\end{center}
+
+%\begin{center}
+%<<greyplot, fig=TRUE, height=8, width=8>>=
+%glvar <- function(x) {
+%    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[2]
+%  }
+%  ## returns
+%  ## why not include Finland?
+%  ## hicp <- diff(log(HICP[, -6]))
+%  hicp <- diff(log(HICP[,]))
+%
+%  ## rolling GL means
+%  hicp_glvar <- rollapply(hicp, 12, glvar)
+%  ## since var very small value, increase such as to show it better
+%	 hicp_glvar <- hicp_glvar*100
+%
+%  ## semi-transparent gray and red
+%  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+%
+%  ## plot returns (gray) and rolling variance (red)
+%  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1])
+%  for(i in 1:ncol(hicp)) lines(hicp_glvar[,i], lwd = 1.5, col = gray_red[2])
+%\end{center}
+% idee: concentrate on EURO countries in one graph
+
+\clearpage
 \section{Conlcusion}

### 2010-11-11 pkg/paper/Paper/draft.Rnw
updated Draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@72 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-\graphicspath{{Figures/}}
+% all created dynamically
+%\graphicspath{{Figures/}}
 
@@
   The aim of this paper is to investigate how a countries decision to join one or all of the steps leading towards the common European currency changed its inflation rate dynamics.
-Our attempt is to model the inflation rates via a Generalized Logistic Distribution (GL) and then test for breaks in the parameters of the distribution.
+Our attempt is to model the inflation rates via a Generalized Logistic Distribution (GL) and then test for breaks in its parameters.
 }
@@
 The inflation is modeled by a hybrid Phillips curve  specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
-What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union. \\
-%But monetary policy in a monetary union does not explicitly stabilize the output gap and inflation rate in case of national economic shocks. 
-The effects of joining a monetary union on inflation rate variability depends on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-Due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependend on assumptions.
-\\
+What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
+The effects of joining a monetary union on inflation rate variability depends -- in this model --  on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
+Due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on these assumptions.
+
+\clearpage
 \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
@@
 \\
-In section 2 we will present the data, in section 3 we will present our approach, in section 4 we will present the results, in section 5 we will discuss our findings and try to bring them into context with the overall economic condition, section 6 concludes.
+In section 2 we will present the data, in section 3 we will explain our approach, in section 4 we display the results, in section 5 we will discuss our findings and try to bring them into context with the overall economic condition, section 6 concludes.
 
@@
 The economic reason why this should be so is rather clear cut for more than half of the countries in our sample.
-We would expect changes in the dynamics of all Eastern European countries due to their transition from previously centrally planned economies towards free market economies.
+We would expect changes in the dynamics of all Eastern European countries due to their transition process away from previously centrally planned economies towards free market economies.
 All of these countries experienced some problems with very high inflation rates at the beginning and they all were successfull in eliminating excess inflation afterwards.
@@
 To do this, we assume that our data is a draw from a Generalized Logistic Distribution (GL) and we proceed by estimating the parameters of the GL--distribution via maximum likelihood.
-We then test for structural change and estimate them using dynamic programming.
+We then test for structural changes and -- in case we find some -- estimate them using dynamic programming.
 
@@
 
+\clearpage
+
 An enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a GL distribution is provided with this paper.
@@
 \\
-The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process can be backtracked to \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
+The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process werde developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
 
@@
 
+\clearpage
 \subsection{Test}
@@
 To help with the interpretation, we give a short overview over the history of the EMU.
-It basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2002) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of the exchange rates of the Member States.
+It basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of the exchange rates of the Member States.
 \\
-It should be pointed out however, that these monthly series do still contain all the seasonality\footnote{We would like to mention that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal proquirements.
+It should be pointed out however, that these monthly series do still contain all the seasonality.\footnote{We would like to mention that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal  requirements.
 In addition to Belgium and Luxembourg, there are a number of other states, where the sales periods are regulated by law.
@@
 In Greece there are also two set sales--periods in winter and summer.
-Given the probelems concerning the Greek data, this does not pose any addional difficulty.
+Given the problems concerning the Greek data, this does not pose any additional difficulty.
 In Italy the sales periods are fixed each year by the Chambers of Commerce.
-Portugal has two set sales periods as well, but for two month (January and February as well das Mid--July to Mid--September.
-The source for this information is: $http://www.konsumenteuropa.se/en/Shopping-in-Europe$}.
-If we correct for this by using the lag of order 12 instead of order 1, we find somewhat less pronounced changes with the exception of the latest observations, with the financial crises and the overall lowering in inflation rates, which of course leads to a considerable increase in variance for more than half of our sample.
+Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
+The source for this information is: $http://www.konsumenteuropa.se/en/Shopping-in-Europe$} If we correct for this by using the lag of order 12 instead of order 1, we find somewhat less pronounced changes with the exception of the latest observations, with the financial crises and the overall lowering in inflation rates, which of course leads to a considerable increase in variance for more than half of our sample.
 5 countries change their inflation dynamics after October 2007, coinciding with the ongoing financial crises.
@@
 But in the case of Belgium and Luxembourg, what we actually observe is a change due to the definition of the HICP.
-According to an official of the OECD, these changes can be traced back  to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation.\footnote{for technical details see: http://eur-lex.europa.eu/LexUriServ/LexUriServ.do?uri=OJ:L:2000:300:0016:0017:EN:PDF} This lead to a 6 month seasonality pattern (January and July) afterwards (meaning after the breakpoints induced by sales inclusion), since sales periods in Belgium and Luxembourg are regulated by law and have to occur in the first two weeks of these months and therefor we observe this sharp price decreases in both month.
+According to an official of the OECD, these changes can be traced back  to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation.\footnote{http://eur-lex.europa.eu/LexUriServ/LexUriServ.do?uri=OJ:L:2000:300:0016:0017:EN:PDF} This lead to a 6 month seasonality pattern (January and July) afterwards (meaning after the breakpoints induced by sales inclusion), since sales periods in Belgium and Luxembourg are regulated by law and have to occur in the first two weeks of these months and therefor we observe this sharp price decreases in both month.
 Interestingly enough, both these price decreases are immediately followed by increases in February and August of even higher magnitude.
@@
 
-The second of these groups are the Eastern European countries, the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
-In all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
+The second of these groups are the Eastern European countries: the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
+In almost all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
 Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
@@
 
-Another combination of countries with -- this time -- identical results are the two biggest southern European countries Italy and Spain.
+Another combination of countries with -- this time -- identical results are the two biggest southern European countries, Italy and Spain.
 Both of them experienced high rates of inflation in the early 90ies.
@@
 \\
-Italy is quite a nice example to explain the usefullness of the skewness parameter.
+Italy is quite a nice example to explain the usefulness of the skewness parameter.
 If we focus on the first two periods, with skewness lower than 1 (which indicates that the distribution is skewed to the right) we find that positive shocks are not normally neutralized by negative shocks.
@@
 It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.
-We find a structural break in 2008 following the strong contraction triggerd by the bursting of the Irish housing bubble and a beginning deflation.
+We find a structural break in 2008 following the strong contraction triggered by the bursting of the Irish housing bubble and a beginning deflation.
 
@@
 Another interesting group consists of 3 economically rather different countries.
-In all of them we find no evidence of a structural change.
+In none of them, do we find any evidence of a structural change.
 Whereas the Netherlands and Finland are both on the lower side of the inflation margin, Greece exhibits some very strange behaviour.
@@
 
-The last group with clear patterns can be seen with the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of the United Kingdom in the EMS}, the United Kingdom and Sweden.
+The last group with clear patterns consists of the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of the United Kingdom in the EMS}, the United Kingdom and Sweden.
 Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom cumulating into the "Black Wednesday" in 1992 and the run on the Swedish currency in the same year.
@@
 
-Let us go through all the steps of our analysis by means of an example, for this we have chosen the case of Slovenia.
+Let us go through all the steps of our analysis by means of an example, for this purpose, we have chosen the case of Slovenia.
 The economic reason why we suppose to find changing dynamics are clear.
@@
 Whereas the supLM test suggest at least one break, the LWZ criterion favors no break.
-But the decline in the negative log--likelihood from 0 breaks to 1 is the strongest, we concluded that one break point actually would be the best alternative.
+But, since the decline in the negative log--likelihood from 0 breaks to 1 is the strongest, we concluded that one break point actually would be the best alternative.
 
@@
 
-Slovenia did achieve this by two very important measures.
+Slovenia did achieve its goals by two very important measures.
 First it liberalized the financial markets in 2003 and -- most important -- it restricted the year--on--year growth of the monetary base M1 severely starting in 2003.
@@
 To clarify some issues, lets take a look at the data aggregated in just one graph.
-If more countries have the same inflation rate in a given month, the lines are overlayed and thus get darker.\footnote{We wold seee one black line if all countries would show the same month--on--month increase at any one time.} We do see, that many of these countries follow a rather similar path.
+If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.\footnote{We wold see one black line if all countries would show the same month--on--month increase at any one time.} We do see that many of these countries follow a rather similar path.
 One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
-However, this was not enough for most of the countries -- with the exception of Ireland -- to lead towards a structural break in inflation rate dynamics.
+However, this was not enough for most of the countries -- with the exception of Ireland -- to trigger a structural break in inflation rate dynamics.
 \\
@@
 So concerning the mean of the inflation rates, we see a clear convergence pattern.
-However, this is not the case for variance where we find clear evidence of an increase.
+However, this is not the case for variance, where we find clear evidence of an increase.
 This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
@@
 We were able to outline changes in the dynamics of the inflation rates in various countries across Europe.
-However, we fail to find any clear pattern relating to the EURO introduction, but find clear evidence that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
+However, we fail to find any clear pattern relating to the EURO introduction, but find convincing evidence, that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
 \\
-Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases can be traced back to the strong seasonality in our time series.
+Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases could possibly be traced back to the strong seasonality in our time series.
+The economic reason for this however remains a riddle and should be subject to future research.
 

### 2010-11-19 pkg/paper/Paper/draft.Rnw
updated draft, created analysis3 anew to deal with sales periods, updated bib-file to include links to homepages
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@73 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \usepackage{rotating}
+\usepackage{url}
 %% need no \usepackage{Sweave}
@@
 \Abstract{
-  The aim of this paper is to investigate how a countries decision to join one or all of the steps leading towards the common European currency changed its inflation rate dynamics.
+The aim of this paper is to investigate how a country's decision to join one or all of the steps leading towards the common European currency changed its inflation rate dynamics.
 Our attempt is to model the inflation rates via a Generalized Logistic Distribution (GL) and then test for breaks in its parameters.
@@
 
-\Keywords{inflation rate, structural break, EMU, Generalized Logistic Distribution, EURO}
+\Keywords{inflation rate, structural break, EMU, Generalized Logistic Distribution, euro}
 
@@
 
-The European Central Bank (ECB) defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the Euro area of
-below 2\%''\footnote{http://www.ecb.europa.eu/mopo/strategy/pricestab/html/index.en.html}.
-\citet{emerson} emphasizes the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
+The \citet{ecb} defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the euro area of
+below 2\%''.
+\citet{emerson} emphasize the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
+The impact of lower inflation rates seems to be rather beneficial, as \cite{crespohoe} point out, who suggest that an inflation rate around 1.6\% has favorable effects on economic growth.
+For now, the question of interest centers around the way in which a country's decision to join the Economic and Monetary Union (EMU) changed its inflation rate dynamics.
 \\
-%Therefore a vital aim of every central bank is to achieve a low inflation rate with negligible variation. \\
-The question of interest centers around the way in which a countries decision to join the Economic and Monetary Union (EMU) changed its inflation rate dynamics. 
-%There are a number of reasons why these should have changed indeed.
-Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria were likely to lead to an alternation at least in the mean (since this was required for a number of countries) and possible in the volatility of their respective inflation rates as well.
-This has indeed been the case for a host of EMU countries, like Spain, Italy and Portugal, reflecting their effort to meet the Maastricht Criteria for inflation rates.
+Throughout the literature, there is still a considerable degree of uncertainty as to what extent the introduction of the euro, or monetary unions in general, affect the inflation rate.
+We provide a brief summary of the most important papers in this area, the data used and the conclusion reached:
+
+\begin{longtable}{|p{2cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|}
+\hline 
+Authors & Applied method & Data (region) & Conclusions \\
+\hline
+\citet{busetti} & Unit root tests on inflation differentials & monthly CPI\footnote{Consumer price index} (EMU countries) & Inflation rates diverge after 1998.\\
+\hline
+\citet{cap} & GARCH framework & monthly seasonally adjusted CPI (12 EMU countries) & No consistent pattern can be found for the degree of persistence of inflation. \\
+\hline
+\citet{crespo} & T2 test for significance in changes in price dispersion & Price level data on over 160 products and services (27 European cities) & Price convergence took place at beginning of 1990s.
+Not much evidence that the euro further narrowed price differentials.
 \\
-% no literature overview
-Theory unfortunately does not provide any clear foundation for a proper answer whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
-\citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the alleged risk of deflation.
-It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
-An interesting approach to this question is taken by \citet{holte}. 
-%He creates a two country model for monetary policy analysis along the line of two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}.
-The inflation is modeled by a hybrid Phillips curve  specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
-What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
-The effects of joining a monetary union on inflation rate variability depends -- in this model --  on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-Due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on these assumptions.
+\hline
+\citet{holte} & Dynamic stochastic general equilibrium model; theoretical analysis & no real data (modeled similar to new EU Member States) & Joining a monetary union decreases the variability of the inflation rate.
+\\
+\hline
+\citet{morana} & Markov switching ARFIMA\footnote{Auto regressive fractionally integrated moving average} model & monthly HICP\footnote{Harmonized index of consumer prices} and M3 (aggregated euro Area inflation rate) & Finds 3 regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing core inflation rate.
+\\
+\hline
+\citet{palomba} & Autoregressive metric in ARIMA\footnote{Auto regressive integrated moving average} context to measure similarity & monthly seasonally unadjusted CPI (EU~25 Member States) & The degree of similarity of short-run inflation dynamics across euro area countries is still weak.
+\\
+\hline
+\citet{slacik} & ARDL\footnote{Autoregressive distributed lags} model combined with Bayesian Model Averaging techniques & Various variables (EU~12) & Inflationary jumps from joining a monetary union, in the case of the Czech Republic, may not be completely avoided, but are rather unlikely.
+\\
+\hline
+\end{longtable}
+
+The clear message here is that there is no clear message up to now.
+Most studies fail to find evidence of common behaviour of the inflation time series of the Member States of the EMU.
+Our paper aims to help to fill this gap.
 
 \clearpage
-\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
-%They use a time-varying model with a GARCH specification for the unconditional volatility of inflation and find some instability in the conditional volatility. \\
-In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
-They find that convergence over time of inflation dynamics was only partly observable.
-In a paper studying core inflation and using an aggregated Euro Area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
-\\
-Inquiring into the convergence properties of inflation rates among countries of the EMU, \citet{busetti} find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
-This does not quite go along with our findings, that do not indicate stronger divergence after 1997 by means of the detection of break points.
-\\
-In section 2 we will present the data, in section 3 we will explain our approach, in section 4 we display the results, in section 5 we will discuss our findings and try to bring them into context with the overall economic condition, section 6 concludes.
-
 \section{Data}
 
-Inflation is measured as the monthly change in the HICP rates from 1.1990---3.2010.
-Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	the	United Kingdom.\footnote{Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.} The data is obtained from the OECD Statistics.\footnote{http://www.oecd-ilibrary.org/content/data/data-00285-en} \\
-The countries can be divided into three different groups: the EURO countries (Austria, Belgium, Estonia\footnote{although Estonia has not yet introduced the EURO, but in 2011.}, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain)\footnote{Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.}, EU members not participating in the ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
+The data is obtained from the \citet{oecd}.
+Inflation is measured as the monthly change in the HICP rates from Jan.
+1990---Mar.
+2010.
+Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	the	United Kingdom (Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity).
+\\
+The countries can be divided into three different groups.
+First the euro countries (Austria, Belgium, Estonia (although Estonia has not yet introduced the euro, but in 2011), France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain).
+Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.
+Second, the EU members not participating in the ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
 Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
-% alternative link: http://stats.oecd.org/index.aspx
+\\
+It should be pointed out however, that these monthly series do still contain all the seasonality.\footnote{We would like to mention that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal  requirements.
+In addition to Belgium and Luxembourg, there are a number of other states, where the sales periods are regulated by law.
+France has an individual time span regulated by each department.
+This should pose no problem, since there is no global state level solution.
+In Greece there are also two set sales--periods in winter and summer.
+In Italy the sales periods are fixed each year by the Chambers of Commerce.
+Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
+The source for this information is \citet{konseur}.}
 
-\newpage
 \section{Model}
 
-Our main question of interest is whether or not a countries decision to join one or all of the steps leading towards a common European currency changed its inflation rate dynamics.
+%neuer Titel
+%Strukturbruch
+%slowenisches Beispiel
+
+%allg. Motivation
+Our main question of interest is whether or not a country's decision to join one or all of the steps leading towards a common European currency changed its inflation rate dynamics.
 The economic reason why this should be so is rather clear cut for more than half of the countries in our sample.
@@
 All of these countries experienced some problems with very high inflation rates at the beginning and they all were successfull in eliminating excess inflation afterwards.
-It can be safely assumed that the former constituent republics of Yugoslavia follow the same pattern.
+We might also assume, that the former constituent republics of Yugoslavia follow a rather similar pattern.
 In the case of the Southern European countries, with a history of high inflation and less pronounced fiscal discipline, we would also expect inflation dynamics to change towards a regime of lower inflation.
@@
 \\
-% [although it should be added, that the 90ies are perhaps a bad starting point, since we would expect them to be already rather moderate compared with past values of 60ies and 70ies]
-The approach we use follows one simple idea: i.e.
-we want to track the evolution of mean, variance and skewness of our series over time.
+%explanation
+Our approach  follows one simple idea: we want to track the evolution of mean, variance and skewness of the inflation series over time.
 Thereby we hope to be able to find and estimate possible breakpoints in the data.
-The correlation and the correlation structure of the time series themselves are of no interest and are treated as a nuisance term.\\
-To do this, we assume that our data is a draw from a Generalized Logistic Distribution (GL) and we proceed by estimating the parameters of the GL--distribution via maximum likelihood.
-We then test for structural changes and -- in case we find some -- estimate them using dynamic programming.
+These breakpoints -- where any breakpoint would suggest a change in dynamics -- are then examined for each country individually.
+We focus on countries rather then the whole euro--area due to the still large degree of heterogeneity in the sample.
+If a number of countries would have the same break dates, this would possibly indicate some common factor responsible for changes in the dynamics.
+We hope to be able to trace out similarities and differences in the countries and thereby improve our understanding of what the monetary union actually ment for the inflation rate dynamics.
+\\
 
 \subsection{Generalized Logistic Distribution}
-
-In econometrics, the logistic distribution is often used in income distributions and growth models.
-This is due to  its fatter tails, which fit these kind of data somewhat better.
-The application of the generalized logistic distribution as we will use it in this paper is rather rare.
-\citet{won} use a GL distribution in a regression model with autocorrelated errors.
-They use this type of distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  non symmetric and severely left or right skewed.
-A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they don't use a shape parameter.
-Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context to expected inflation.
-\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
-\\
-We apply the general framework, as developed in \citet{z07}, to a more specific model, in this case by means of a GL distribution.
-Prior to that, we would like to give a short justification for the usage of the GL distribution.
-Regarding the data at hand, it was not possible to utilize the already existing method developed in \citet{z07}, since almost all inflation rates, with the notable exception of Greece, are not normally distributed and clearly exhibit asymmetric properties.
-\\
-Therefore, a somewhat more flexible distribution had to be used.
-We needed a distribution exhibiting rather fatter tails and the property to be both left and right skewed.
-To serve this purpose, we use a generalization of the logistic distribution as defined in \citet{johnson}.
-
-\clearpage
-For our inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we assume a GL distribution given by:
+%Motivation Verteilung
+The first problem to solve was the selection of an appropriate distribution reflecting the properties of the data in use in as good a manner as possible.
+Most of the time series exhibit asymmetric properties (with sometimes rather strong skewness) and -- compared to the normal distribution -- a fatter tail.
+This ruled out all the symmetric distributions.
+In the field of the asymmetric distributions there was of course still no supply shortage, but we have chosen the Generalized logistic distribution (GL) due to its relative low degree of complexity, in comparison -- for example --  with the skewed--t--distribution.
+With its three parameters, we are able to trace mean, variance and skewness of the inflation rates.
+For our inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we therefore assume a GL distribution as defined in \citet{johnson}:
 
@@
 \end{eqnarray}
+% skewness: if delta = 1.1 --> skweness is 0.1, is positive, skewed to the right: linkssteil
+% ie. negative skew (delta <1: rechtssteil, in englisch: skewed to the left 
+% (psigamma(0.5, deriv = 2) - psigamma(1, deriv = 2))/(psigamma(1, deriv = 1) + psigamma(0.5, deriv = 1))^(3/2)
 
 with location ($\theta$), scale ($\sigma$) and shape ($\delta$).
-For $\delta$=1 the distribution simplifies to the logistic distribution, for $\delta$$<$1 it is skewed to the left and for $\delta$$>$1 it is skewed to the right.
+For $\delta$=1 the distribution simplifies to the logistic distribution, for $\delta$$<$1 it is skewed to the left and for $\delta$$>$1 it is skewed to the right.\footnote{more details are provided in section \ref{app}.} \\
+In econometrics, the logistic distribution is often used in income distributions and growth models.
+This is due to  its fatter tail, which fit these kind of data somewhat better.
+The application of the GL distribution, parameterized in a way similar to this paper, is not a very common one.
+\citet{won} use a GL distribution in a regression model with autocorrelated errors.
+They use this type of distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  non symmetric and severely left or right skewed.
+A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they do not use a shape parameter.
+Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context to expected inflation.
+\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
+\\
+
+\clearpage
+\subsection{Estimation}
+%Schätzung
+
+We apply the general framework, as developed in \citet{z07}, to a more specific model, in this case by means of a GL distribution.
+An enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a GL distribution, is provided with this paper.
+The asymptotic testing theory still holds for this generalization.
+The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process werde developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
+\\
+In order to analyze our data, we first estimate the parameters of the GL--distribution by means of maximum likelihood, which is possible, if we assume that our parameter vector $\phi = (\theta, \sigma, \delta)$ is stable over time $t$, i.e.
+we estimate $\hat{\phi} ~=~ \underset{\phi}{argmax} \sum_{t=1}^n \log f(y_t|\phi)$.
+Our problem now simplifies to the following question: Is the assumption valid or do the $\phi_t$ change?
+
+%H0
+\begin{eqnarray}
+& H_0: \phi_t = \phi_0 \quad (t=1,...,n) \nonumber 
+\end{eqnarray}
+
+This can be assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures of model deviation.\footnote{for details see section \ref{app}}  The empirical scores process $\mathit{efp}(\cdot)$ captures deviations from zero over time and is given by:
+
+\begin{eqnarray}
+\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
+\end{eqnarray}
+
+where $\hat{V}^{-1/2}$ is some consistent estimator for the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
+We then employ a Functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ which converges to a 3-dimensional Brownian bridge:
+
+\begin{eqnarray}
+\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot) \nonumber
+\end{eqnarray}
+
+\clearpage
+\subsection{Test}
+
+Actually the $\mathit{efp}(\cdot)$ could be used for aggregated test statistics in various ways.
+In our paper, we employ the Supremum of LM (supLM) test by \citet{andrews}.
+This test statistic is given by:
+
+\begin{eqnarray}
+\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
+\end{eqnarray}
+
+where the appropriate p--values can be computed using:
+
+\begin{eqnarray}
+\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert W^0(t) \Vert_2^2}{t(1-t)} \nonumber
+\end{eqnarray}
+
+The supLM test is the most powerful test for a single abrupt change in unknown time and has the highest power if more than one parameter of the distribution changes over time.
+Throughout the paper we will use the supLM test.
+
+\clearpage
+\section{Breakpoint Estimation}
+
+After a break is detected via the supLM test, we proceed by estimating the timing of the break. We estimate $B$ breakpoints $\tau_1,...,\tau_B$ via maximization of the full segmented likelihood:
+
+\begin{eqnarray}
+\sum_{b=1}^{B+1} \sum_{t=\tau_b + 1}^{\tau_b} \log f(y_t | \phi^{(b)}) \nonumber
+\end{eqnarray}
+
+All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming. More details for this estimation technique are provided in \cite{z10}.
+
+\subsection{An Example}
+
+To illustrate our methods, we will go through all the steps of our analysis by means of an example.
+For this purpose, we have chosen the case of Slovenia.
+The economic reason why we suppose to find changing dynamics can be outlined in the following way: In the first phase of the consolidation after the crackup of Yugoslavia and the short war of 1991, Slovenia experienced very high inflation rates, with yearly growth rates as high as 90\% in August 1990.
+But gradually Slovenia was able to reduce the inflation rate towards much lower values.
+This first phase is not captured in our data, since it starts as late as Jan.
+1996.
+For the period up to today, we would expect one or two breaks from an economic point of view.
+The efforts to reach the Maastricht criteria as soon as possible are likely to lead to a further lowering of the mean and possibly also of the variance of the inflation rate.
+Another break point might possibly be found towards the introduction of the euro coins themselves in 2007.
+\\
+Actually when we look at the data, this is precisely what we find, a breakpoint (represented by the dotted line) at Aug.
+2003, with a clear lowering of the mean (represented by the straight blue lines).
+
+<<fig=TRUE, height=4, width=8>>=
+x <- results$Slovenia
+plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "")
+lines(x$breakpoints, breaks = x$nbreaks)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+@
+
+\begin{figure}[!]
+<<fig=TRUE, height=4, width=8>>=
+plot(x$efp, functional = supLM(0.1), main = "supLM test", xaxt = "n", xlim = range(time(x$data)))
+Axis(time(x$data), side = 1)
+lines(x$breakpoints, breaks = x$nbreaks)
+@
+\end{figure}
+
+The supLM test graph  suggests one break around 2003 and possibly another towards the end of the sample.
+This second break point is not estimated, since we restricted a regime towards a length of at least 24 month.\\
+To be as precise as possible with the detection of break points, we also include a graph showing the values of the LWZ (Liu, Wu and Zidek) criteria as developed in \citet{liu}.
+Whereas the supLM test suggest at least one break, the LWZ criterion favors no break.
+But, since the decline in the negative log--likelihood from 0 breaks to 1 is the strongest, we concluded that one break point actually is the best alternative.
+
+\begin{center}
+<<slovplotbreak, fig=TRUE, height=6, width=8>>=
+<<plot2>>
+@
+\end{center}
+
+Slovenia did achieve its goals by two very important measures.
+First, it liberalized the financial markets in 2003 and -- most important -- it restricted the year--on--year growth of the monetary base M1 severely starting in 2003.
+Combined, these actions were enough to ensure  Slovenian participation in the euro--zone.
+The \cite{imad} explains this somewhat differently by the cooperation of both the government and the Bank of Slovenia, which resulted in a careful handling of administered prices, a counter--cyclical adjustment of excise duties on fuels and moderate wage increases based on social agreements.
+According to them, due to theses measures, the effect of the euro introduction on the Slovenian inflation rate was 0.24 percentage points.
+% alternative explanation \cite{slacik} In Slovenia the government and the Bank of Slovenia helped to curb inflation with coordinated policy measures such as strict implementation of administered prices (increases not above headline in ation), counter-cyclical adjustment of excise duties on fuels and wage moderation based on a social agreement (see IMAD 2007).
+% the total effect of the euro changeover on prices to amount to 0.24 percentage points, being among the smallest recorded in the euro area --> quelle: http://www.umar.gov.si/fileadmin/user_upload/sporocila_za_javnost/euro.pdf
+
+\newpage
+\section{Results}
+
+The results as we show them here are grouped according to patterns, which we think have some common properties.
+We present the breakdates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II or the date of euro introduction.
+To give a guidline for interpretation, we provide a short overview over the history of the EMU.
+It basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of the exchange rates of the Member States.
+\\
+
+%
+%das package colortbl (\usepackege[dvips]{colortbl}) ermöglicht z.B. mit
+%\rowcolor[gray]{0.90} in einer Tabellenzeile diese farbig (bzw. hier grau) zu unterlegen.
+
+\small
+\begin{longtable}{|l|l|r|r|r|l|l|l|}
+\hline
+\multicolumn{8}{|c|}{Belgium and Luxembourg} \\
+\hline 
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(2, 13), format_moments)))
+@
+\multicolumn{8}{|c|}{The Eastern Countries} \\
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(3,5,10,15,17,18), format_moments)))
+@
+\multicolumn{8}{|c|}{Italy and Spain} \\
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(12, 19), format_moments)))
+@
+\multicolumn{8}{|c|}{Ireland} \\
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(11), format_moments)))
+@
+\multicolumn{8}{|c|}{No change countries} \\
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(9,14,6), format_moments)))
+@
+\multicolumn{8}{|c|}{Non participating countries} \\
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(20, 21), format_moments)))
+@
+\multicolumn{8}{|c|}{The other countries} \\
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(1,4,7,8,16), format_moments)))
+@
+
+\end{longtable}
+% results3 hier einbauen
+% warning: the groups then change; Italy and Greece have not yet answered
+% Italian Statistics: Your request has been correctly inserted in the system with the reference number 13078/2010.
+% comment: the greek break is NOT sure, since they have not yet answered about sales periods
+
+The first  distinct group of contries consists of Belgium and Luxembourg.
+This may come as no surprise, since both countries formed a monetary union prior to the start of the ERM.
+In the past, they enjoyed very stable inflation rates with low values of both mean and variance, this changed around the timing of the irrevocable fixing of the exchange rates.
+But in the case of Belgium and Luxembourg, what we actually observe is an alternation due to the definition of the HICP.
+According to their national statistical offices, these changes can be traced back  to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation, details may be found in: \citet{eurlex} This lead to a 6 month seasonality pattern (January and July) afterwards (meaning after the breakpoints induced by sales inclusion), since sales periods in Belgium and Luxembourg are regulated by law and have to occur in the first two weeks of these months and therefor we observe this sharp price decreases in both month.
+Interestingly enough, both these price decreases are immediately followed by increases in February and August of even higher magnitude.
+\\
+The second of these groups are the Eastern European countries: the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
+In almost all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
+Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
+\\
+Another combination of countries with -- this time -- identical results are the two biggest southern European countries, Italy and Spain.
+Both of them experienced high rates of inflation in the early 90ies.
+Afterwards, in the phase leading up to the monetary union, mean and variance were significantly reduced.
+This trend changed drastically soon after the fixing of the exchange rates leading to higher mean and much higher variances then ever before since the beginning of the HICP series.
+\\
+Italy is quite a nice example to explain the usefulness of the skewness parameter.
+If we focus on the first two periods, with skewness lower than 1 (which indicates that the distribution is skewed to the right) we find that positive shocks are not normally neutralized by negative shocks.
+This changes in the last phase with skewness negative, indicating that the GL is now skewed to the left and that positive shocks will more likely be followed by negative shocks.
+\\
+Ireland can be considered a rather special case.
+It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.
+We find a structural break in 2008 following the strong contraction triggered by the bursting of the Irish housing bubble and a beginning deflation.
+\\
+Another interesting group consists of 3 economically rather different countries.
+In none of them do we find any evidence of a structural change.
+Whereas the Netherlands and Finland are both on the lower side of the inflation margin, Greece exhibits some very strange behaviour.
+It has by far the highest variance of all series and seems to be completely different from all the others.
+\\
+The last group with clear patterns consists of the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of the United Kingdom in the EMS.}, the United Kingdom and Sweden.
+Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and the run on the Swedish currency in the same year.
+\\
+No clear interpretation is available for: Austria, Denmark, France, Germany and Portugal. \\
+\cite{busetti} find 3 clusters in their paper, if they include all the 12 countries in question (Austria, Belgium, Finland, France, Germany, Italy, the Netherlands, Luxembourg, Spain, Greece, Portugal, Ireland).
+The first 5 countries form the ``low inflation club'', Italy, the Netherlands and Luxembourg the ``medium inflation club'', whereas the other four countries form a high inflation club.
+We do not find much similarity between our cluster (which focuses on the timing of the break and the change in the parameters) and what they find via aggregation into inflation clubs.
+They also find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
+This does not quite go along with our findings, that do not indicate stronger divergence after 1997 by means of break point detection.
+\\
+
+\clearpage
+\section{Discussion}
+
+% was i zum sega ket hob ba busetti --> vgl. archivierte Datei vom 16.11
+
+% actually too detailed for so small a picture
+
+%\begin{center}
+%<<greyplot, fig=TRUE, height=8, width=8>>=
+% ## compute GL mean (if no NA)
+%  glmean <- function(x) {
+%    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
+%  }
+%  ## returns
+%  ## why not include Finland?
+%  ## hicp <- diff(log(HICP[, -6]))
+%  hicp <- diff(log(HICP[,]))
+%
+%  ## rolling GL means
+%  hicp_glmean <- rollapply(hicp, 12, glmean)
+%
+%  ## semi-transparent gray and red
+%  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+%
+%  ## plot returns (gray) and rolling means (red)
+%  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
+%  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+%@
+%\end{center}
+
+To clarify some issues, lets take a look at the data aggregated in just one graph.
+If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.\footnote{We wold see one black line if all countries would show the same month--on--month increase at any one time.} We do see that many of these countries follow a rather similar path.
+One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
+However, this was not enough for most of the countries -- with the exception of Ireland -- to trigger a structural break in inflation rate dynamics.
+\\
+One interesting aspect we would like to focus on is the clear trend in means.
+What we have plotted here actually is a rolling mean, where the mean is calculated using the first moment of a fitted GL distribution.
+It is clearly visible that these rates converged towards some common mean (graphically emphasized by the redder lines) which, from 2005 onwards, comes rather close to the ECB target rate (which would translate to a monthly increase of roughly 2 per mill).
+So concerning the mean of the inflation rates, we see a clear convergence pattern.
+However, this is not the case for variance, where we find clear evidence of an increase.
+This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
+\\
+
+\clearpage
+\section{Conlcusion}
+
+We were able to outline changes in the dynamics of the inflation rates in various countries across Europe.
+However, we fail to find any clear pattern relating to the euro introduction, but find convincing evidence, that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
+\\
+Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases could possibly be traced back to the strong seasonality in our time series.
+The economic reason for this however remains a riddle and should be subject to future research.
+
+\clearpage
+\section{Appendix}\label{app}
+
+\subsection{GL distribution}
+
 The moments are:
@@
 
-\clearpage
-
-An enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a GL distribution is provided with this paper.
-The asymptotic testing theory still holds for this generalization.
-\\
-The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process werde developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
-
-\subsection{Estimation Framework}
-
-In order to analyze our data, we first estimate the parameters of the GL--distribution by means of maximum likelihood, which is possible, if we assume that $\phi = (\theta, \sigma, \delta)$ is stable over time $t$:
-
-\begin{eqnarray}
-& \hat{\phi} ~=~ \underset{\phi}{argmax} \sum_{t=1}^n \log f(y_t|\phi), \nonumber\\
-& \sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0, \nonumber
-\end{eqnarray}
-
-\medskip
-
-Our problem now simplifies to the following question: Is the assumption valid or do the $\phi_t$ change?
-
-\begin{eqnarray}
-& H_0: \phi_t = \phi_0 \quad (t=1,...,n) \nonumber 
-\end{eqnarray}
-
-This can be assessed using the empirical scores $s(y_t|\hat{\phi})$, as presented above, as measures of model deviation.
-The empirical scores process $\mathit{efp}(\cdot)$ captures deviations from zero over time and is given by:
-
-\begin{eqnarray}
-\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
-\end{eqnarray}
-
-where $\hat{V}^{-1/2}$ is some consistent estimator for the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
-We then employ a Functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ which converges to a 3-dimensional Brownian bridge:
-
-\begin{eqnarray}
-\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot) \nonumber
-\end{eqnarray}
+The scores of the parameters $\phi = (\theta, \sigma, \delta)$ sum up to 0: $\sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0$.
 
 \clearpage
-Lets elaborate on this basic idea. Consider some time series given by the following graph. We would not need any test, since we see a clear break at the middle of the sample.
+\subsection{Graphs}
 
-<<fig=TRUE, height=5, width=8>>=
-x <- rnorm(500,0,.1)
-x[251:500] <- rnorm(250,.5,.1) 
-print(plot(x, type="l", ylab = "series", xlab = "Time"))
-abline(lm(x~1))
+% generate graphs in another fashion (a graph with data of results3) und zwar: results$Belgium$data
+% include fitted means --> still some work to do, not yet figured out due to xyplot and two lines...
+<<bellux, fig=TRUE, height=5, width=8>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(xyplot(cbind(results$Belgium$data, results$Luxembourg$data), xlab="Time", ylab="", main="", screens=list("Belgium", "Luxembourg")), ylim=c(-2,2))
 @
 
-If we  estimate a simple least squares regression we get a straight regression line at the mean of the series.
-What we do now is to sum up the scores (which can be viewed as similar to the residuals) by means of the efp (i.e.
-something like a cumulated sum of residuals).
-This summing up of the residuals would be constructed by first aggregating all the negative residuals down to the lowest point, then  we proceed by accruing the other (positive) residuals.
-The bottom of this cumulated sum will be our estimation of the breakpoint.
-If this peak crosses some test statistics, then it can be viewed as a significant deviation from the null hypothesis.
-
-\clearpage
-\subsection{Test}
-
-Actually the $\mathit{efp}(\cdot)$ could be used for aggregated test statistics in various ways.
-In our paper we employ the Supremum of LM (supLM) test by \citet{andrews}.
-This test statistic is given by:
-
-\begin{eqnarray}
-\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
-\end{eqnarray}
-
-where the appropriate p--values can be computed using:
-
-\begin{eqnarray}
-\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert W^0(t) \Vert_2^2}{t(1-t)} \nonumber
-\end{eqnarray}
-
-The supLM test is the most powerful test for a single abrupt change in unknown time and has the highest power if more than one parameter of the distribution changes over time.
-Throughout the paper we will use the supLM test.
-
-\subsection{Breakpoint Estimation}
-
-After a break is detected via the supLM test, we proceed by estimating the timing of the break. We estimate $B$ breakpoints $\tau_1,...,\tau_B$ via maximization of the full segmented likelihood:
-
-\begin{eqnarray}
-\sum_{b=1}^{B+1} \sum_{t=\tau_b + 1}^{\tau_b} \log f(y_t | \phi^{(b)}) \nonumber
-\end{eqnarray}
-
-All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming. More details for this estimation technique are provided in \cite{z10}.
-
-\newpage
-\section{Results}
-
-To help with the interpretation, we give a short overview over the history of the EMU.
-It basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the Euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of the exchange rates of the Member States.
-\\
-It should be pointed out however, that these monthly series do still contain all the seasonality.\footnote{We would like to mention that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal  requirements.
-In addition to Belgium and Luxembourg, there are a number of other states, where the sales periods are regulated by law.
-France has an individual time span regulated by each department.
-This should pose no problem, since there is no global state level solution.
-In Greece there are also two set sales--periods in winter and summer.
-Given the problems concerning the Greek data, this does not pose any additional difficulty.
-In Italy the sales periods are fixed each year by the Chambers of Commerce.
-Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
-The source for this information is: $http://www.konsumenteuropa.se/en/Shopping-in-Europe$} If we correct for this by using the lag of order 12 instead of order 1, we find somewhat less pronounced changes with the exception of the latest observations, with the financial crises and the overall lowering in inflation rates, which of course leads to a considerable increase in variance for more than half of our sample.
-5 countries change their inflation dynamics after October 2007, coinciding with the ongoing financial crises.
-
-% interesting graph to financial crisis: http://de.wikipedia.org/wiki/Datei:Einlagefazilit%C3%A4t.PNG
-
-%If the EMU had a significant effect upon inflation rate dynamics, we would expect to find a break either in the 90ies, reflecting the various waves of integration or after the introduction of the Euro due to the popular argument that the introduction of Euro paper money led to a considerable price increase.
-\\
-%[some considerations: 1.
-necessary to include all the results?
-seems to me preferable to deal with them as groups of individual countries rather than all on one page; 2.
-show all the groups?
-3.
-show example like in Slovenian case?]
-
-% http://en.wikipedia.org/wiki/European_Exchange_Rate_Mechanism#Replacement_with_the_euro_and_ERM_II
-% http://en.wikipedia.org/wiki/Eurozone
-% automatic longtable provided by R 
-% rotate table, not sure if this actually looks better
-
-%\begin{landscape}
-
-%\small
-%\begin{longtable}{|l|l|l|l|l|l|l|l|}
-%\hline
-%Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-%\hline
-%<<echo=FALSE, results=tex>>=
-%writeLines(unlist(lapply(c(seq(1,21)), format_moments)))
-%@
-%\end{longtable}
-%\end{landscape}
-
-% EMS was never very stable
-% careful: ERM II für nicht EURO Mitglieder, EMS war Vorgänger, auch für EURO Mitglieder
-% http://de.wikipedia.org/wiki/Europ%C3%A4isches_W%C3%A4hrungssystem
-% mit cecchetti Paper im Unterlagen Ordner abgleichen
-
-\subsection{Belgium and Luxembourg}
-
-On the basis of these results, we are able to trace out different groups.
-The first of this groups consists of Belgium and Luxembourg.
-This may come as no surprise since both countries formed a monetary union prior to the start of the Exchange Rate Mechanism (ERM).
-While both countries enjoyed very stable inflation rates with low values of both mean and variance, this changed around the timing of the irrevocable fixing of the exchange rates.
-But in the case of Belgium and Luxembourg, what we actually observe is a change due to the definition of the HICP.
-According to an official of the OECD, these changes can be traced back  to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation.\footnote{http://eur-lex.europa.eu/LexUriServ/LexUriServ.do?uri=OJ:L:2000:300:0016:0017:EN:PDF} This lead to a 6 month seasonality pattern (January and July) afterwards (meaning after the breakpoints induced by sales inclusion), since sales periods in Belgium and Luxembourg are regulated by law and have to occur in the first two weeks of these months and therefor we observe this sharp price decreases in both month.
-Interestingly enough, both these price decreases are immediately followed by increases in February and August of even higher magnitude.
-
-%%%%% Seasonality Problems:
-% for the others: http://www.konsumenteuropa.se/en/Shopping-in-Europe/Luxembourg/
-% we do not find many more regulated sales periods in Europe (given the data above is correct, which I hope :)
-% exemption: France (regulated by each department individually (so changes hopefully leveled out) --> no problem
-% 	         Greece (In Greece there are two set sales-periods. Winter sales: usually during January – February. Summer sales: usually during July – August.)
-%            --> no problem, dont trust series anyway
-%            Italy (There are two set sales periods in Italy, in winter and summer. Dates and duration are fixed by the Chambers of Commerce.)
-%            --> what if these changes???  --> e-mail sent to italian chamber of commerce in austria, awaiting reply
-%            Portugal (There are two set sale periods: Winter sales: 28/12 to 28/02 Summer sales: 15/07 to 15/09 )
-% Quelle für diese Änderung : e-mail Verkehr mit Jérôme Hury, dem Chef der Statec Price Abteilung; und Anne-Sophie.FRAISSE@oecd.org
-
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(2, 13), format_moments)))
-@
-\end{longtable}
-
-% could be looking much better
-\begin{center}
-<<bellux, fig=TRUE, height=5, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Belgium"]))), 100*diff(log(na.omit(HICP[, "Luxembourg"])))), xlab="Time", ylab="", main="", screens=list("Belgium", "Luxembourg")), ylim=c(-2,2))
-@
-\end{center}
-
-\newpage
-\subsection{The Eastern Countries}
-\normalsize
-
-The second of these groups are the Eastern European countries: the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
-In almost all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
-Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
-
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(3,5,10,15,17,18), format_moments)))
-@
-\end{longtable}
-
-\begin{center}
 <<eastern, fig=TRUE, height=7, width=8>>=
 trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(100*diff(log(na.omit(HICP[, "CzechRepublic"]))), 100*diff(log(na.omit(HICP[, "Estonia"]))), 100*diff(log(na.omit(HICP[, "Hungary"]))), 100*diff(log(na.omit(HICP[, "Poland"]))), 100*diff(log(na.omit(HICP[, "Slovakia"]))), 100*diff(log(na.omit(HICP[, "Slovenia"])))), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia"), ylim=c(-2.7,2.7)))
+print(xyplot(cbind(results$CzechRepublic$data, results$Estonia$data, results$Hungary$data, results$Poland$data, results$Slovakia$data, results$Slovenia$data), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia"), ylim=c(-2.7,2.7)))
 @
-\end{center}
 
-%, ylim=c(-2.7,2.7)
-
-\newpage
-\subsection{Italy and Spain}
-\normalsize
-
-Another combination of countries with -- this time -- identical results are the two biggest southern European countries, Italy and Spain.
-Both of them experienced high rates of inflation in the early 90ies.
-Afterwards, in the phase leading up to the monetary union, mean and variance were significantly reduced.
-This trend changed drastically soon after the fixing of the exchange rates leading to higher mean and much higher variances then ever before since the beginning of the HICP series.
-\\
-Italy is quite a nice example to explain the usefulness of the skewness parameter.
-If we focus on the first two periods, with skewness lower than 1 (which indicates that the distribution is skewed to the right) we find that positive shocks are not normally neutralized by negative shocks.
-This changes in the last phase with skewness negative, indicating that the GL is now skewed to the left and that positive shocks will more likely be followed by negative shocks.
-
-% skewness: if delta = 1.1 --> skweness is 0.1, is positive, skewed to the right: linkssteil
-% ie. negative skew (delta <1: rechtssteil, in englisch: skewed to the left 
-% (psigamma(0.5, deriv = 2) - psigamma(1, deriv = 2))/(psigamma(1, deriv = 1) + psigamma(0.5, deriv = 1))^(3/2)
-
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(12, 19), format_moments)))
-@
-\end{longtable}
-
-% could be looking much better
-
-\begin{center}
 <<itaspa, fig=TRUE, height=5, width=8>>=
 trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Italy"]))), 100*diff(log(na.omit(HICP[, "Spain"])))), xlab="Time", ylab="", main="", screens=list("Italy", "Spain"), ylim=c(-2,2)))
+print(xyplot(cbind(results$Spain$data, results$Italy$data), xlab="Time", ylab="", main="", screens=list("Italy", "Spain"), ylim=c(-2,2)))
 @
-\end{center}
 
-\newpage
-\normalsize
-\subsection{Ireland}
+<<ireland, fig=TRUE, height=4, width=8>>=
+plot(results$Ireland$data, main="Ireland", ylab="", xlab ="Time", ylim=c(-2,2))
+@
 
-Ireland can be considered a rather special case.
-It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.
-We find a structural break in 2008 following the strong contraction triggered by the bursting of the Irish housing bubble and a beginning deflation.
-
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(11), format_moments)))
-@
-\end{longtable}
-
-\begin{center}
-<<ireland, fig=TRUE, height=4, width=8>>=
-plot(100*diff(log(na.omit(HICP[, "Ireland"]))), main="Ireland", ylab="", xlab ="Time", ylim=c(-2,2))
-@
-\end{center}
-
-\newpage
-\normalsize
-\subsection{No change countries}
-
-Another interesting group consists of 3 economically rather different countries.
-In none of them, do we find any evidence of a structural change.
-Whereas the Netherlands and Finland are both on the lower side of the inflation margin, Greece exhibits some very strange behaviour.
-It has by far the highest variance of all series and seems to be completely different from all the others.
-
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(9,14,6), format_moments)))
-@
-\end{longtable}
-
-\begin{center}
 <<nochange, fig=TRUE, height=5, width=8>>=
 trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Finland"]))), 100*diff(log(na.omit(HICP[, "Greece"]))), 100*diff(log(na.omit(HICP[, "Netherlands"])))), xlab="Time", ylab="", main="", screens=list("Finland", "Greece", "Netherlands"), ylim=c(-2.7,2.7)))
-@
-\end{center}
-
-\newpage
-\normalsize
-\subsection{The Non--participating Countries}
-
-The last group with clear patterns consists of the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of the United Kingdom in the EMS}, the United Kingdom and Sweden.
-Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom cumulating into the "Black Wednesday" in 1992 and the run on the Swedish currency in the same year.
-
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(20, 21), format_moments)))
-@
-\end{longtable}
-
-% could be looking much better
-\begin{center}
-<<sweuk, fig=TRUE, height=7, width=8>>=
-print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Sweden"]))), 100*diff(log(na.omit(HICP[, "UnitedKingdom"])))), xlab="Time", ylab="", main="", screens=list("Sweden", "United Kingdom")), ylim=c(-2,2))
-@
-\end{center}
-
-\newpage
-\normalsize
-\subsection{The Other Countries}
-
-No clear interpretation is available for: Austria, Denmark, France, Germany and Portugal.
-
-\small
-\begin{longtable}{|l|l|l|l|l|l|l|l|}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & EURO \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(1,4,7,8,16), format_moments)))
-@
-\end{longtable}
-
-\begin{center}
-<<nopattern, fig=TRUE, height=7, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(100*diff(log(na.omit(HICP[, "Austria"]))), 100*diff(log(na.omit(HICP[, "Denmark"]))), 100*diff(log(na.omit(HICP[, "France"]))), 100*diff(log(na.omit(HICP[, "Germany"]))), 100*diff(log(na.omit(HICP[, "Portugal"])))), xlab="Time", ylab="", main="", screens=list("Austria", "Denmark", "France", "Germany", "Portugal"), ylim=c(-2,2)))
-@
-\end{center}
-
-\subsection{An example--Slovenia}
-\normalsize
-
-Let us go through all the steps of our analysis by means of an example, for this purpose, we have chosen the case of Slovenia.
-The economic reason why we suppose to find changing dynamics are clear.
-In the first phase of the consolidation after the break--up of Yugoslavia and after the short war of 1991, Slovenia experienced very high inflation rates,
-with yearly growth rates as high as 90\% in August 1990.
-But gradually Slovenia was able to reduce the inflation rate towards much lower values.
-This first phase is not captured in our data since it starts as late as January 1996.
-For the period up to today we would expect one or two breaks from an economic point of view.
-The efforts to reach the Maastricht criteria as soon as possible are likely to lead to a further lowering of the mean and possibly also of the variance of the inflation rate.
-Another break point might possibly be found towards the introduction of the EURO coins themselves in 2007.
-\\
-Actually when we look at the data, this is precisely what we get:
-
-<<fig=TRUE, height=4, width=8>>=
-x <- results$Slovenia
-plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "")
-lines(x$breakpoints, breaks = x$nbreaks)
-lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+print(xyplot(cbind(results$Finland$data, results$Greece$data, results$Netherlands$data), xlab="Time", ylab="", main="", screens=list("Finland", "Greece", "Netherlands"), ylim=c(-2.7,2.7)))
 @
 
-<<fig=TRUE, height=4, width=8>>=
-plot(x$efp, functional = supLM(0.1), main = "supLM test", xaxt = "n", xlim = range(time(x$data)))
-Axis(time(x$data), side = 1)
-lines(x$breakpoints, breaks = x$nbreaks)
+<<sweuk, fig=TRUE, height=7, width=8>>=
+print(xyplot(cbind(results$Sweden$data, results$UnitedKingdom$data), xlab="Time", ylab="", main="", screens=list("Sweden", "United Kingdom")), ylim=c(-2,2))
 @
 
-The supLM test graph  suggests one break around 2003 and possibly another towards the end of the sample.
-This second break point is not estimated, since we restricted a regime towards a length of at least 24 month.
-Thus we find one break in August 2003 with a much lower mean value (only a third now of its previous value) and interestingly enough, a higher volatility.\\
-To be as precise as possible with the detection of break points, we also include a graph showing the values of the LWZ (Liu, Wu and Zidek) criteria as developed in \citet{liu}.
-Whereas the supLM test suggest at least one break, the LWZ criterion favors no break.
-But, since the decline in the negative log--likelihood from 0 breaks to 1 is the strongest, we concluded that one break point actually would be the best alternative.
-
-\begin{center}
-<<slovplotbreak, fig=TRUE, height=6, width=8>>=
-<<plot2>>
-@
-\end{center}
-
-Slovenia did achieve its goals by two very important measures.
-First it liberalized the financial markets in 2003 and -- most important -- it restricted the year--on--year growth of the monetary base M1 severely starting in 2003.
-Combined, these actions were enough to ensure  Slovenian participation in the EURO--zone.
-
-%% Quelle für diese Behaupunten: http://www.oenb.at/isaweb/dyna2.do?go=initIndikatoren&definitionLaden=false&hierarchieIdSelected=108
-%\newpage
-%\subsection{Alternative Explanations}
-%Another possible explanation pattern could be the growth in the  money supply.
-A drawback is the current lack of data\footnote{hopefully to be solved}.
-But for some countries we can find some plausible findings.
-Regarding the case of Denmark, the M1 money supply started contracting in June 2000 (on a monthly base).
-In Estonia we find the highpoint in M3 expansion ending with March 1998, with much lower M3 growth rates afterwards.
-The same happended to Hungary in April and May 1998.
-In the Polnish case we find a severe contraction of the money supply in May 2001.
-Similar findings hold true for Germany as well.
-\\
-%In Spain we see a stronger increase in M1 in May and June 1996 (nothing for the other break, data scarceity).
-In Portugal we find a modest increase in year on year M1 money supply growth starting in July 1992.
-We find nothing peculiar for Slovakia and Hungary.
-In Estonia the M1 money supply starts to contract severely on a yearly base following March 1998.
-In Slovenia we find a contraction of the Money Supply in July 2003, with the whole year 2003 witnessing a sharp decrease in the year on year growth.
-Poland starts contracting the year on year M1 money supply after June 2000 with very low growth rates till January 2002.
-%
-%%\readme{irgendwo noch inne (nach \cite{duarte}), dass es normal isch, dass die relativen Preise fluktuieren, got uf asymmetrische productivity shocks zruck}
-%
-%
-%% for denmark: see: C:\Dissertation\Daten\DSI\M1\daten - the denmark analysis, alos: Spain, Portugal, Slovakia, Hungary Estonia noch am absatz für M1
-%
-%Concerning monetary unions, it is interesting to notice that Belgium and Luxembourg, that constituted a monetary union prior to the European Monetary Union, did not experience the same timing in their break towards a higher mean and a significantly higher volatility regime.
-%
-%With regard to the alleged benefit of monetary unions, the import of credibility to former high inflation countries (like Italy and Spain), did nothing to decrease their mean or volatility, rather to the contrary.
-Spain and Italy did have a -- rather -- low mean low volatility regime from 1996:6 to 2000:12 but then things changed towards higher volatility.
-It appears as if both countries tried to decrease inflation prior to the fixing of the exchange rates and let go afterwards.
-%
-%Following the reasoning of a study by \cite{duarte}, high volatility of inflation rates within regions of a union may be explained a countries decision to lower the inflation differential by means of fiscal policy.
-This explanation might fit to Spain and Italy, both were above the the average EU inflation.
-%
-
-% Quelle: für Deutschland M3 bis 1995
-%http://www.bundesbank.de/statistik/statistik_zeitreihen.php?lang=de&open=ewu&func=row&tr=TSD304
-%
-%Deutschland: M3 bis 1995
-%dort ersichtlihc, dass Deutschland zu Bruchzeitpunkten jeweils änderungen im m3 aggregat hat
-
-%The AR distances table, measuring the similarity of short--run inflation dynamics as in \citet{palomba} more or less supports this group building, although it is much more accurate.
-The similarity of the structural breaks thus is interesting topic for future research.
-
-%give the 3 most similar in AR distances: in Palomba Paper (table 2)
-%
-%Italy: Greece, France, Portugal
-%Poland: Ireland, Spain, UK
-%Sweden: Denmark, Netherlands, UK
-%Netherlands: Sweden, UK, Portugal
-%France: Italy, Greece, Belgium
-%
-%
-%either support or not support group 1
-%AR distance: the lower the more similar
-%
-
-\clearpage
-\section{Discussion}
-
-% actually too detailed for so small a picture
-
-\begin{center}
-<<greyplot, fig=TRUE, height=8, width=8>>=
- ## compute GL mean (if no NA)
-  glmean <- function(x) {
-    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
-  }
-  ## returns
-  ## why not include Finland?
-  ## hicp <- diff(log(HICP[, -6]))
-  hicp <- diff(log(HICP[,]))
-
-  ## rolling GL means
-  hicp_glmean <- rollapply(hicp, 12, glmean)
-
-  ## semi-transparent gray and red
-  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-
-  ## plot returns (gray) and rolling means (red)
-  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
-  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-@
-\end{center}
-
-To clarify some issues, lets take a look at the data aggregated in just one graph.
-If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.\footnote{We wold see one black line if all countries would show the same month--on--month increase at any one time.} We do see that many of these countries follow a rather similar path.
-One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
-However, this was not enough for most of the countries -- with the exception of Ireland -- to trigger a structural break in inflation rate dynamics.
-\\
-One interesting aspect we would like to focus on is the clear trend in means.
-What we have plotted here actually is a rolling mean, where the mean is calculated using the first moment of a fitted GL distribution.
-It is clearly visible that these rates converged towards some common mean (graphically emphasized by the redder lines) which, from 2005 onwards, comes rather close to the ECB target rate (which would translate to a monthly increase of roughly 2 per mill).
-So concerning the mean of the inflation rates, we see a clear convergence pattern.
-However, this is not the case for variance, where we find clear evidence of an increase.
-This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
-\\
-
-% message of these graphs not so clear
-%On this graph we just show the current EURO member countries\footnote{with Estonia included}. The clear trend towards a common rather lower mean is clearly visible.
-% just focus on EURO countries
-%\begin{center}
-%<<greyploteuro, fig=TRUE, height=8, width=8>>=
-%  plot(hicp[, c(1,2,5,6,7,8,9,11,12,13,14,16,17,18,19)], plot.type = "single", lwd = 1.5, col = gray_red[1], ylab="EURO", xlab="Time")
-%  for(i in c(1,2,5,6,7,8,9,11,12,13,14,16,17,18,19)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+%<<nopattern, fig=TRUE, height=7, width=8>>=
+%trellis.par.set(theme = standard.theme(color = FALSE))
+%print(xyplot(cbind(results$Austria$data, results$Demark$data, results$France$data, results$Germany$data, results$Portugal$data), xlab="Time", ylab="", main="", screens=list("Austria", "Denmark", "France", "Germany", "Portugal"), ylim=c(-2,2)))
 %@
-%\end{center}
-%
-%This convergence seems to be missing regarding the countries not part of the ERM~II.
-%
-%
-%% just focus on Non-ERM~II countries
-%\begin{center}
-%<<greyplotnerm2, fig=TRUE, height=8, width=8>>=
-%  plot(hicp[, c(3,10,15,20,21)], plot.type = "single", lwd = 1.5, col = gray_red[1], ylab="Non ERM II", xlab="Time")
-%  for(i in c(3,10,15,20,21)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-%@
-%\end{center}
-
-%\begin{center}
-%<<greyplot, fig=TRUE, height=8, width=8>>=
-%glvar <- function(x) {
-%    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[2]
-%  }
-%  ## returns
-%  ## why not include Finland?
-%  ## hicp <- diff(log(HICP[, -6]))
-%  hicp <- diff(log(HICP[,]))
-%
-%  ## rolling GL means
-%  hicp_glvar <- rollapply(hicp, 12, glvar)
-%  ## since var very small value, increase such as to show it better
-%	 hicp_glvar <- hicp_glvar*100
-%
-%  ## semi-transparent gray and red
-%  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-%
-%  ## plot returns (gray) and rolling variance (red)
-%  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1])
-%  for(i in 1:ncol(hicp)) lines(hicp_glvar[,i], lwd = 1.5, col = gray_red[2])
-%\end{center}
-% idee: concentrate on EURO countries in one graph
-
-\clearpage
-\section{Conlcusion}
-
-We were able to outline changes in the dynamics of the inflation rates in various countries across Europe.
-However, we fail to find any clear pattern relating to the EURO introduction, but find convincing evidence, that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
-\\
-Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases could possibly be traced back to the strong seasonality in our time series.
-The economic reason for this however remains a riddle and should be subject to future research.
 

### 2010-11-22 pkg/paper/Paper/draft.Rnw
revised version
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@74 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 }
+
+get_data <- function(i) {
+  rval <- do.call("merge",
+    lapply(i, function(j) results[[j]]$data))
+  colnames(rval) <- names(results)[i]
+  rval
+}
+
+get_fitted <- function(i) {
+  rval <- do.call("merge",
+    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks=results[[j]]$nbreaks, type = "mean")))
+  colnames(rval) <- names(results)[i]
+  rval
+}
+
+get_xyplot <- function(i)
+  xyplot(merge(get_data(i), get_fitted(i)),
+    screen = rep(names(results)[i], 2),
+    col = rep(c(1, 4), each = length(i)))
 
@@
 \\
-It should be pointed out however, that these monthly series do still contain all the seasonality.\footnote{We would like to mention that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal  requirements.
+It should be pointed out however, that these monthly series do still contain all the seasonality.\footnote{We would like to meantion that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal  requirements.
 In addition to Belgium and Luxembourg, there are a number of other states, where the sales periods are regulated by law.
@@
 If a number of countries would have the same break dates, this would possibly indicate some common factor responsible for changes in the dynamics.
-We hope to be able to trace out similarities and differences in the countries and thereby improve our understanding of what the monetary union actually ment for the inflation rate dynamics.
+We hope to be able to trace out similarities and differences in the countries and thereby improve our understanding of what the monetary union actually meant for the inflation rate dynamics.
 \\
@@
 The asymptotic testing theory still holds for this generalization.
-The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process werde developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
+The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process were developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
 \\
@@
 x <- results$Slovenia
-plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "")
+plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "Slovenia")
 lines(x$breakpoints, breaks = x$nbreaks)
@@
 We present the breakdates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II or the date of euro introduction.
-To give a guidline for interpretation, we provide a short overview over the history of the EMU.
+To give a guideline for interpretation, we provide a short overview over the history of the EMU.
 It basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of the exchange rates of the Member States.
@@
 
-The first  distinct group of contries consists of Belgium and Luxembourg.
+The first  distinct group of countries consists of Belgium and Luxembourg.
 This may come as no surprise, since both countries formed a monetary union prior to the start of the ERM.
@@
 \clearpage
-\section{Conlcusion}
+\section{Conclusion}
 
@@
 % include fitted means --> still some work to do, not yet figured out due to xyplot and two lines...
-<<bellux, fig=TRUE, height=5, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$Belgium$data, results$Luxembourg$data), xlab="Time", ylab="", main="", screens=list("Belgium", "Luxembourg")), ylim=c(-2,2))
+%<<bellux, fig=TRUE, height=5, width=8>>=
+%trellis.par.set(theme = standard.theme(color = FALSE))
+%print(xyplot(cbind(results$Belgium$data, results$Luxembourg$data), xlab="Time", ylab="", main="", screens=list("Belgium", "Luxembourg")), ylim=c(-2,2))
+%@
+
+Here we present the actual data where the blue lines in the graphs correspond to the estimated mean of the series.
+
+<<bellux, fig=TRUE, height=8, width=8>>=
+print(get_xyplot(c(2, 13)))
 @
 
-<<eastern, fig=TRUE, height=7, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$CzechRepublic$data, results$Estonia$data, results$Hungary$data, results$Poland$data, results$Slovakia$data, results$Slovenia$data), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia"), ylim=c(-2.7,2.7)))
+%<<eastern, fig=TRUE, height=7, width=8>>=
+%trellis.par.set(theme = standard.theme(color = FALSE))
+%print(xyplot(cbind(results$CzechRepublic$data, results$Estonia$data, results$Hungary$data, results$Poland$data, results$Slovakia$data, results$Slovenia$data), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia"), ylim=c(-2.7,2.7)))
+%@
+
+<<eastern, fig=TRUE, height=12, width=8>>=
+print(get_xyplot(c(3, 5,10,17,18)))
 @
 
-<<itaspa, fig=TRUE, height=5, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$Spain$data, results$Italy$data), xlab="Time", ylab="", main="", screens=list("Italy", "Spain"), ylim=c(-2,2)))
+%<<itaspa, fig=TRUE, height=5, width=8>>=
+%trellis.par.set(theme = standard.theme(color = FALSE))
+%print(xyplot(cbind(results$Spain$data, results$Italy$data), xlab="Time", ylab="", main="", screens=list("Italy", "Spain"), ylim=c(-2,2)))
+%@
+
+<<itaspa, fig=TRUE, height=8, width=8>>=
+print(get_xyplot(c(19,12)))
 @
 
-<<ireland, fig=TRUE, height=4, width=8>>=
-plot(results$Ireland$data, main="Ireland", ylab="", xlab ="Time", ylim=c(-2,2))
+%<<ireland, fig=TRUE, height=4, width=8>>=
+%plot(results$Ireland$data, main="Ireland", ylab="", xlab ="Time", ylim=c(-2,2))
+%@
+<<ireland, fig=TRUE, height=8, width=8>>=
+x <- results$Ireland
+plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "Ireland")
+lines(x$breakpoints, breaks = x$nbreaks)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+@
+%<<nochange, fig=TRUE, height=5, width=8>>=
+%trellis.par.set(theme = standard.theme(color = FALSE))
+%print(xyplot(cbind(results$Finland$data, results$Greece$data, results$Netherlands$data), xlab="Time", ylab="", main="", screens=list("Finland", "Greece", "Netherlands"), ylim=c(-2.7,2.7)))
+%@
+
+<<nochange, fig=TRUE, height=12, width=8>>=
+print(get_xyplot(c(6,9,14)))
 @
 
-<<nochange, fig=TRUE, height=5, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$Finland$data, results$Greece$data, results$Netherlands$data), xlab="Time", ylab="", main="", screens=list("Finland", "Greece", "Netherlands"), ylim=c(-2.7,2.7)))
-@
+%<<sweuk, fig=TRUE, height=7, width=8>>=
+%print(xyplot(cbind(results$Sweden$data, results$UnitedKingdom$data), xlab="Time", ylab="", main="", screens=list("Sweden", "United Kingdom")), ylim=c(-2,2))
+%@
 
-<<sweuk, fig=TRUE, height=7, width=8>>=
-print(xyplot(cbind(results$Sweden$data, results$UnitedKingdom$data), xlab="Time", ylab="", main="", screens=list("Sweden", "United Kingdom")), ylim=c(-2,2))
+<<sweuk, fig=TRUE, height=8, width=8>>=
+print(get_xyplot(c(20,21)))
 @
@@
 %trellis.par.set(theme = standard.theme(color = FALSE))
-%print(xyplot(cbind(results$Austria$data, results$Demark$data, results$France$data, results$Germany$data, results$Portugal$data), xlab="Time", ylab="", main="", screens=list("Austria", "Denmark", "France", "Germany", "Portugal"), ylim=c(-2,2)))
+%print(xyplot(cbind(results$Austria$data, results$Denmark$data, results$France$data, results$Germany$data, results$Portugal$data), xlab="Time", ylab="", main="", screens=list("Austria", "Denmark", "France", "Germany", "Portugal"), ylim=c(-2,2)))
 %@
+
+<<nopattern, fig=TRUE, height=12, width=8>>=
+print(get_xyplot(c(1,4,7,8,16)))
+@
 

### 2011-03-10 pkg/paper/Paper/draft.Rnw
updated draft, dataset (now till end of 2010) and results3 (dealing with sales periods inclusion)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@76 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2010-11-19 pkg/paper/Paper/draft.Rnw
@@
 	Sweden = NA ,
-	UnitedKingdom = NA
+	UK = NA
 ))
@@
 
+get_data <- function(i) {
+  rval <- do.call("merge",
+    lapply(i, function(j) results[[j]]$data))
+  colnames(rval) <- names(results)[i]
+  rval
+}
+
+get_fitted <- function(i) {
+  rval <- do.call("merge",
+    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks=results[[j]]$nbreaks, type = "mean")))
+  colnames(rval) <- names(results)[i]
+  rval
+}
+
+get_xyplot <- function(i)
+  xyplot(merge(get_data(i), get_fitted(i)),
+    screen = rep(names(results)[i], 2),
+    col = rep(c(1, 4), each = length(i)))
+
 xyplot.gefp <- function(x, panel = NULL, ylim = NULL, ...)
@@
 HICP <- HICP[, colnames(HICP) != "EU"]
+
+## to change to shorter names if necessary
+names(HICP) <- c("Austria",  "Belgium", "CzechRepublic", "Denmark", "Estonia", 
+"Finland", "France",  "Germany", "Greece", "Hungary", "Ireland",  "Italy", "Luxembourg", "Netherlands",
+"Poland", "Portugal",  "Slovakia", "Slovenia",  "Spain", "Sweden", "UK")
 @
@@
 ## plot "macros" (assuming existence "x", "x_gf", ...)
-<<plot1, eval=FALSE>>=
-par(mfrow = c(1, 2))
-plot(x, main = "Series with Fitted Mean", ylab = "100 * diff(log(HICP))", xlab = "Time")
-lines(x_bp, breaks = x_nbreaks)
-lines(fitted(x_bp, breaks = x_nbreaks, type = "mean"), col = 4)
-plot(x_efp, functional = supLM(0.1), main = "supLM test")
-lines(x_bp, breaks = x_nbreaks)
+
+<<plot3, eval=FALSE>>=
+  glmean <- function(x) {
+    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
+  }
+  ## returns
+  hicp <- diff(log(HICP[,]))
+
+  ## rolling GL means
+  hicp_glmean <- rollapply(hicp, 12, glmean)
+
+  ## semi-transparent gray and red
+  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+
+  ## plot returns (gray) and rolling means (red)
+  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
+  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
 @
 
-<<plot2, eval=FALSE>>=
-par(mfrow = c(1, 1))
-plot(x_bp)
+<<plot4, eval=TRUE>>=
+## plot convergence in mean
+  glmean <- function(x) {
+    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
+  }
+  ## returns
+  hicp <- diff(log(HICP[,]))
+
+  ## rolling GL means
+  ## without Estonia,  Slovakia, Slovenia
+  hicp_glmean <- rollapply(hicp, 12, glmean)
+  hicp_euro <- diff(log(HICP[, c(1,2,6,7,8,9,11,12,13,14,16,19)]))
+	mean_euro <- hicp_glmean[, c(1,2,6,7,8,9,11,12,13,14,16,19)]
+
+	diff <- rep(1,length(mean_euro[,1]))
+	n <- length(mean_euro[,1])
+
+	for (i in 1:n) {
+	bla <- subset(mean_euro[i,], select=colMeans(is.na(mean_euro[i,])) == 0) 
+	diff[i] = max(bla) - min(bla)
+	}
+	diff <- ts(diff , start = c(1990,2), frequency =12)
+
+	plot(diff)
 @
@@
 \Abstract{
-The aim of this paper is to investigate how a country's decision to join one or all of the steps leading towards the common European currency changed its inflation rate dynamics.
-Our attempt is to model the inflation rates via a Generalized Logistic Distribution (GL) and then test for breaks in its parameters.
+(... text when everything else finished, present innovation of paper...)
 }
@@
 
-The \citet{ecb} defines price stability ``as a year-on-year increase in the Harmonized Index of Consumer Prices (HICP) for the euro area of
-below 2\%''.
-\citet{emerson} emphasize the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
-The impact of lower inflation rates seems to be rather beneficial, as \cite{crespohoe} point out, who suggest that an inflation rate around 1.6\% has favorable effects on economic growth.
-For now, the question of interest centers around the way in which a country's decision to join the Economic and Monetary Union (EMU) changed its inflation rate dynamics.
+(... Jesus: journals, EE (empirical economics): mehr EMU, weniger Methode, Journal of Applied Econometrics: genau umgekehrt...)
+(... economic story, big transition ...) \\
+Ever since the beginning of the European Union, the topic of a common currency was a controversial issue.
+Although the Economic and Monetary Union (EMU) is now a fact, the discussion about the economic effects of the Euro is far from being settled.
+The controversial topics range from the question of whether or not the Euro--zone is indeed an optimal currency area (OCA) all the way to the very survival of the Euro in light of the budgetary problems of a number of its member states.
+The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context is in the center of an ongoing debate, for instance the question about short-run and steady-state inflation uncertainty dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries which is the topic of an investigation by \citet{palomba}.
 \\
+(... big transition continued, our contribution...) \\
+The way towards the EMU basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
+Any new countries who want to join the EMU are first obliged to fulfill the Maastricht which, besides setting rules for government debts and interest rates, also requires the participation in the ERM~II for two years, while the exchange rates towards the euro are not allowed to cross the nominal band.
+These criteria acknowledge, that a stable inflation rate is one of the key goals of every good macroeconomic policy.
+\cite{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
+Failing to stabilize inflation leads to severe economic problems.
 Throughout the literature, there is still a considerable degree of uncertainty as to what extent the introduction of the euro, or monetary unions in general, affect the inflation rate.
-We provide a brief summary of the most important papers in this area, the data used and the conclusion reached:
-
-\begin{longtable}{|p{2cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|}
-\hline 
-Authors & Applied method & Data (region) & Conclusions \\
-\hline
-\citet{busetti} & Unit root tests on inflation differentials & monthly CPI\footnote{Consumer price index} (EMU countries) & Inflation rates diverge after 1998.\\
-\hline
-\citet{cap} & GARCH framework & monthly seasonally adjusted CPI (12 EMU countries) & No consistent pattern can be found for the degree of persistence of inflation. \\
-\hline
-\citet{crespo} & T2 test for significance in changes in price dispersion & Price level data on over 160 products and services (27 European cities) & Price convergence took place at beginning of 1990s.
-Not much evidence that the euro further narrowed price differentials.
+There are a number of reasons for suspecting changes.
+Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria were likely to lead to an alteration at least in the mean (since this was required for a number of countries) and possibly in the volatility of their respective inflation rates as well.
+The EMU requires a country to possibly change a number of economic policies, which can only be achieved by big transitionally efforts.
 \\
-\hline
-\citet{holte} & Dynamic stochastic general equilibrium model; theoretical analysis & no real data (modeled similar to new EU Member States) & Joining a monetary union decreases the variability of the inflation rate.
+(...introduction needs to include the value added of this paper, why is it better, how does it compare to others...) \\
+We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates, which focuses on these changes and thereby upon the effect of the EMU on a number of European countries both within and without the Euro--zone.
+The framework developed in this paper allows testing for any significant changes of a countries inflation dynamics and thus helps interpreting the effects of the EMU.
+For this purpose, we developed a new estimation approach building on previous research by \citet{z07}.
 \\
-\hline
-\citet{morana} & Markov switching ARFIMA\footnote{Auto regressive fractionally integrated moving average} model & monthly HICP\footnote{Harmonized index of consumer prices} and M3 (aggregated euro Area inflation rate) & Finds 3 regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing core inflation rate.
+(...effects of monetary unions, literature overview (more literature in discussion of results)...) \\
+Two of the three most important channels that change inflation, the money supply and the degree of openness  are subject to fundamental changes once a country joins a monetary union.
+The original countries participating in the Exchange Rate Mechanism (ERM) had to synchronize money supply growth and interest rates in order to fix exchanges rates or otherwise risk the chance of severe disequilibria.
+Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
 \\
-\hline
-\citet{palomba} & Autoregressive metric in ARIMA\footnote{Auto regressive integrated moving average} context to measure similarity & monthly seasonally unadjusted CPI (EU~25 Member States) & The degree of similarity of short-run inflation dynamics across euro area countries is still weak.
+Openness, i.e.
+the inflow of new goods and services or the abolishment of tariffs and capital restraint and regulations, is more likely to generate deflationary pressure, due to the substitution of cheaper goods for more expensive ones.
+Since deflation is rather unusual, with general price increases being the norm, the logical conclusion is that the most important inflationary channel are changes in money supply.
+%\footnote{Indeed, a general rise in prices is impossible without a simultaneous increase in money supply.}. 
+These changes in turn are influenced by a countries monetary policy and the credit behavior of its banking industry.
+External shocks, like the spillover effects of a hike in oil prices or the financial crisis may be important as well.
 \\
-\hline
-\citet{slacik} & ARDL\footnote{Autoregressive distributed lags} model combined with Bayesian Model Averaging techniques & Various variables (EU~12) & Inflationary jumps from joining a monetary union, in the case of the Czech Republic, may not be completely avoided, but are rather unlikely.
+Economic theory unfortunately does not provide any clear foundation for a proper answer whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
+%\citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the alleged risk of deflation.
+It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
+An interesting approach to this question is taken by \citet{holte}. 
+%He creates a two country model for monetary policy analysis along the line of two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}.
+The inflation is modeled by a hybrid Phillips curve  specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
+What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
+The effects of joining a monetary union on inflation variability in his framework depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
+However, due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on a variety of assumptions.
 \\
-\hline
-\end{longtable}
-
-The clear message here is that there is no clear message up to now.
-Most studies fail to find evidence of common behaviour of the inflation time series of the Member States of the EMU.
-Our paper aims to help to fill this gap.
+The remainder of this paper is structured as follows: section \ref{s2} presents the data, section \ref{s3} presents the model and the estimation techniques used, section \ref{s4} illustrates our approach using Slovenia as an example, section \ref{s5} presents the results and section \ref{s6} concludes.
 
 \clearpage
-\section{Data}
+\section{Data}\label{s2}
 
-The data is obtained from the \citet{oecd}.
-Inflation is measured as the monthly change in the HICP rates from Jan.
-1990---Mar.
-2010.
-Countries included are Austria,	Belgium, the	Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	the Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	the	United Kingdom (Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity).
+(... Jesus would present data later, first describe what we are going to to do...) \\
+HICP series\footnote{seasonally unadjusted} from Jan.1990--Mar.2010 are obtained for 21 countries the \citet{oecd}.
+Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom.
+Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.
+Cypria and Malta are not included since they are very small economies.
 \\
-The countries can be divided into three different groups.
-First the euro countries (Austria, Belgium, Estonia (although Estonia has not yet introduced the euro, but in 2011), France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain).
-Cypria, Malta, Andorra, Monaco, San Marino and the Vatican are left out due their minor importance.
-Second, the EU members not participating in the ERM~II (Exchange Rate Mechanism): the Czech Republic, Hungary, Poland, the United Kingdom and Sweden.
+The countries in this sample can be divided into three different groups.
+First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain).
+Second, the EU members not participating in the ERM~II: Czech Republic, Hungary, Poland, United Kingdom, and Sweden.
 Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
 \\
-It should be pointed out however, that these monthly series do still contain all the seasonality.\footnote{We would like to mention that in the case of Belgium and Luxembourg, the seasonality patterns can be traced back to changes in HICP definition and legal  requirements.
-In addition to Belgium and Luxembourg, there are a number of other states, where the sales periods are regulated by law.
-France has an individual time span regulated by each department.
-This should pose no problem, since there is no global state level solution.
-In Greece there are also two set sales--periods in winter and summer.
-In Italy the sales periods are fixed each year by the Chambers of Commerce.
-Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
-The source for this information is \citet{konseur}.}
 
-\section{Model}
+% greece: -for winter discounts: First 10 days of January-end of February; -for summer discounts: First 10 days of July -end od August; but has always been included
+
+\clearpage
+\section{Model}\label{s3}
+
+% generell strategie: estimate, test parameter stability, breaks model
 
@@
 
-%allg. Motivation
-Our main question of interest is whether or not a country's decision to join one or all of the steps leading towards a common European currency changed its inflation rate dynamics.
-The economic reason why this should be so is rather clear cut for more than half of the countries in our sample.
-We would expect changes in the dynamics of all Eastern European countries due to their transition process away from previously centrally planned economies towards free market economies.
-All of these countries experienced some problems with very high inflation rates at the beginning and they all were successfull in eliminating excess inflation afterwards.
-We might also assume, that the former constituent republics of Yugoslavia follow a rather similar pattern.
-In the case of the Southern European countries, with a history of high inflation and less pronounced fiscal discipline, we would also expect inflation dynamics to change towards a regime of lower inflation.
-The case for the Central European countries of course is different.
-There is no reason why the European Central Bank (ECB) should actually  follow the rather stable monetary policies of the former German central bank in a one--to--one fashion.
+% laut achim noch zuviel bla bla im text --> weniger als anektode, präziser
+(...less anectodic, more precise...) \\
+To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks.
+First, the model is fitted for the whole sample using the generalized logistic distribution.
+This then enables testing for the stability of the parameter estimates.
+If there is evidence of significant changes, one proceeds by computing the number and the timing of the break points.
 \\
-%explanation
-Our approach  follows one simple idea: we want to track the evolution of mean, variance and skewness of the inflation series over time.
-Thereby we hope to be able to find and estimate possible breakpoints in the data.
-These breakpoints -- where any breakpoint would suggest a change in dynamics -- are then examined for each country individually.
-We focus on countries rather then the whole euro--area due to the still large degree of heterogeneity in the sample.
-If a number of countries would have the same break dates, this would possibly indicate some common factor responsible for changes in the dynamics.
-We hope to be able to trace out similarities and differences in the countries and thereby improve our understanding of what the monetary union actually ment for the inflation rate dynamics.
+This framework allows to trace the evolution of mean, variance and skewness of the inflation series and thus interpreting deviations in any of these parameters as possibly due to a change in monetary policy, is made feasible.
+Any huge intervention or change is captured by fluctuations in the parameter stability of the series and can be quantified and related to a distinct time frame.
 \\
+(...discuss correlation...) \\
+The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation. 
+%To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.
 
-\subsection{Generalized Logistic Distribution}
+% korrelation besser herausarbeiten
+% mehr: link to, match with known events, like intervention
+
+\subsection{Generalized Logistic Distribution}\label{s31}
+
+(... formulation problems ...) \\
+(...Jesus: GL not standard, explain why it is worth using it, what makes it better than ARIMA?...) \\
 %Motivation Verteilung
-The first problem to solve was the selection of an appropriate distribution reflecting the properties of the data in use in as good a manner as possible.
-Most of the time series exhibit asymmetric properties (with sometimes rather strong skewness) and -- compared to the normal distribution -- a fatter tail.
-This ruled out all the symmetric distributions.
-In the field of the asymmetric distributions there was of course still no supply shortage, but we have chosen the Generalized logistic distribution (GL) due to its relative low degree of complexity, in comparison -- for example --  with the skewed--t--distribution.
-With its three parameters, we are able to trace mean, variance and skewness of the inflation rates.
-For our inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we therefore assume a GL distribution as defined in \citet{johnson}:
+The data suggested that the use of a normal distribution might cause some estimation problems, since most of the time series exhibit asymmetric properties (with sometimes rather strong skewness) and -- compared to the normal distribution -- a fatter tail.
+To accommodate for this fact, the use of the generalized logistic distribution (GL) with its relative low degree of complexity, seemed appropriate.
+On the one hand, it is a simple distribution and on the other hand, it is flexible enough to capture essential properties of the data.
+For the inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we therefore assume a GL distribution as defined in \citet{johnson}:
 
 \begin{eqnarray}
-f(y | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma} \cdot \exp^{-\frac{y-\theta}{\sigma}}}{(1+\exp^{-\frac{y-\theta}{\sigma}})^{(\delta+1)}} \nonumber
+f(y | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma} \cdot \exp^{-\frac{y-\theta}{\sigma}}}{(1+\exp^{-\frac{y-\theta}{\sigma}})^{(\delta+1)}} 
+\label{eqn1}
 \end{eqnarray}
@@
 
-with location ($\theta$), scale ($\sigma$) and shape ($\delta$).
-For $\delta$=1 the distribution simplifies to the logistic distribution, for $\delta$$<$1 it is skewed to the left and for $\delta$$>$1 it is skewed to the right.\footnote{more details are provided in section \ref{app}.} \\
+with location $\theta$, scale $\sigma$ and shape $\delta$.
+For $\delta=1$ the distribution simplifies to the logistic distribution, for $\delta<1$ it is skewed to the left and for $\delta$$>$1 it is skewed to the right (see appendix \ref{glscore} for more details).
+\\
 In econometrics, the logistic distribution is often used in income distributions and growth models.
 This is due to  its fatter tail, which fit these kind of data somewhat better.
-The application of the GL distribution, parameterized in a way similar to this paper, is not a very common one.
-\citet{won} use a GL distribution in a regression model with autocorrelated errors.
+\citet{won} employ a GL distribution in a regression model with autocorrelated errors.
 They use this type of distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  non symmetric and severely left or right skewed.
@@
 
-\clearpage
 \subsection{Estimation}
+
+% anders erzählen: um ein Modell zu anzupassen, dann den restlichen teil
 %Schätzung
+%obiger teil erst später; den verweis auf R erst am Ende in den computational details; auch das zeug ab an enhancement... in die computatinal details aufnehmen
+%(... better separation between estimation and test, rest to computational details...) \\
 
-We apply the general framework, as developed in \citet{z07}, to a more specific model, in this case by means of a GL distribution.
-An enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a GL distribution, is provided with this paper.
-The asymptotic testing theory still holds for this generalization.
-The first part of the results we are going to present here, i.e., the tests and the graphical illustration of the empirical fluctuation process werde developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
-\\
-In order to analyze our data, we first estimate the parameters of the GL--distribution by means of maximum likelihood, which is possible, if we assume that our parameter vector $\phi = (\theta, \sigma, \delta)$ is stable over time $t$, i.e.
-we estimate $\hat{\phi} ~=~ \underset{\phi}{argmax} \sum_{t=1}^n \log f(y_t|\phi)$.
-Our problem now simplifies to the following question: Is the assumption valid or do the $\phi_t$ change?
+Under the assumption, that $y_t$ is distributed according to the generalized logistic distribution, we test for the hypothesis that the parameter vector $\hat{\phi}$ stays stable over time:
 
-%H0
 \begin{eqnarray}
@@
 
-This can be assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures of model deviation.\footnote{for details see section \ref{app}}  The empirical scores process $\mathit{efp}(\cdot)$ captures deviations from zero over time and is given by:
+against the $H_1$ of changes in at least one of the parameters.
+We first estimate the parameters of the GL--distribution by means of maximum likelihood $\hat{\phi} ~=~ \underset{\phi}{argmax} \sum_{t=1}^n \log f(y_t|\phi)$.
+The parameter stability is assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures of model deviation.\footnote{for details see Appendix \ref{glscore} and \citet{z07}}  The empirical scores process $\mathit{efp}(\cdot)$ captures deviations from a hypothesized zero mean over time and is given by:
 
@@
 \mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
+\label{eqn2}
 \end{eqnarray}
 
-where $\hat{V}^{-1/2}$ is some consistent estimator for the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
-We then employ a Functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ which converges to a 3-dimensional Brownian bridge:
+where $\hat{V}^{-1/2}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
+Then a functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian bridge:
 
@@
 
-\clearpage
 \subsection{Test}
+(.. Jesus would cite a paper by Andrews and Ploberger 1993 that provided the key for creating confidence intervals around the estimates...) \\
 
-Actually the $\mathit{efp}(\cdot)$ could be used for aggregated test statistics in various ways.
-In our paper, we employ the Supremum of LM (supLM) test by \citet{andrews}.
-This test statistic is given by:
+After the empirical fluctuation process is computed, it conveys information about departures from the $H_0$ of parameter stability over time.
+For a precise quantification of the actual deviation from its hypothesized zero--mean path, the main idea of the test statistics presented here is to aggregate the $\mathit{efp}(\cdot)$ in such a way as to be able to construct a test statistics so that critical values and p values can be derived.
+In this paper, we employ the Supremum of LM (supLM) test by \citet{andrews}.
+This test is well suited for single break alternatives and performs better if more than just one of the distribution parameters change.
+It is given by:
 
@@
 
-The supLM test is the most powerful test for a single abrupt change in unknown time and has the highest power if more than one parameter of the distribution changes over time.
-Throughout the paper we will use the supLM test.
+The supLM test computes a test statistics for all possible change points in a given time frame and rejects the null of no change if the maximum is too high.
 
-\clearpage
-\section{Breakpoint Estimation}
+\subsection{Breakpoint Estimation}
 
-After a break is detected via the supLM test, we proceed by estimating the timing of the break. We estimate $B$ breakpoints $\tau_1,...,\tau_B$ via maximization of the full segmented likelihood:
+If a break is detected via the supLM test, the next step is estimating the number and the timing of these break points.
+Breakpoints $\tau_1,...,\tau_B$ are estimated via maximization of the full segmented likelihood:
 
@@
 
-All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming. More details for this estimation technique are provided in \cite{z10}.
+If the test statistics is significant, at least one break is estimated.
+In principle, the optimal segmentation (i.e.
+the concrete dates of a break) can be computed once B (the number of breakpoints) is known.
+However, B needs to be based on the observed data as well.
+The solution is to compute the timing of the breaks for a number of different values for B and then choose B by means of an information criterion, in this case the LWZ (Liu, Wu and Zidek) as developed in \citet{liu}.
+All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming.\footnote{More details for this estimation technique are provided in \cite{z10}.}
 
-\subsection{An Example}
+% noch mehr dazu: finds 0 wenn nicht signifikant, 1 wenn signifikant, dann auswahl über das lwz criterion
 
-To illustrate our methods, we will go through all the steps of our analysis by means of an example.
-For this purpose, we have chosen the case of Slovenia.
-The economic reason why we suppose to find changing dynamics can be outlined in the following way: In the first phase of the consolidation after the crackup of Yugoslavia and the short war of 1991, Slovenia experienced very high inflation rates, with yearly growth rates as high as 90\% in August 1990.
-But gradually Slovenia was able to reduce the inflation rate towards much lower values.
-This first phase is not captured in our data, since it starts as late as Jan.
-1996.
-For the period up to today, we would expect one or two breaks from an economic point of view.
-The efforts to reach the Maastricht criteria as soon as possible are likely to lead to a further lowering of the mean and possibly also of the variance of the inflation rate.
-Another break point might possibly be found towards the introduction of the euro coins themselves in 2007.
-\\
-Actually when we look at the data, this is precisely what we find, a breakpoint (represented by the dotted line) at Aug.
-2003, with a clear lowering of the mean (represented by the straight blue lines).
+\clearpage
+\section{An Example}\label{s4}
 
+% anders: the data encompasses, thus not included....
+%(... shortend, changed formulation...) \\
+For an illustration of the method, consider the case of Slovenia.
+The data collected encompasses the period from Jan 1996 onwards.
+During the beginning of this period, Slovenia experienced very high inflation rates but was successfull in continually decreasing them.
+A glance at the data suggests a break around 2003/04, with an evident change in the mean.
+
+\begin{figure}[h!tp]
 <<fig=TRUE, height=4, width=8>>=
 x <- results$Slovenia
-plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "")
+par(mfrow = c(1, 2))
+plot(x$data, main = "Slovenia", ylab = "100 * diff(log(HICP))", xlab = "Time")
 lines(x$breakpoints, breaks = x$nbreaks)
 lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+plot(x$efp, functional = supLM(0.1), main = "supLM test")
+lines(x$breakpoints, breaks = x$nbreakss)
 @
-
-\begin{figure}[!]
-<<fig=TRUE, height=4, width=8>>=
-plot(x$efp, functional = supLM(0.1), main = "supLM test", xaxt = "n", xlim = range(time(x$data)))
-Axis(time(x$data), side = 1)
-lines(x$breakpoints, breaks = x$nbreaks)
-@
+\caption{Slovenia: breakpoint estimate and supLM test}
 \end{figure}
 
-The supLM test graph  suggests one break around 2003 and possibly another towards the end of the sample.
-This second break point is not estimated, since we restricted a regime towards a length of at least 24 month.\\
-To be as precise as possible with the detection of break points, we also include a graph showing the values of the LWZ (Liu, Wu and Zidek) criteria as developed in \citet{liu}.
-Whereas the supLM test suggest at least one break, the LWZ criterion favors no break.
-But, since the decline in the negative log--likelihood from 0 breaks to 1 is the strongest, we concluded that one break point actually is the best alternative.
+The supLM test confirms this belief by picturing a hike of the $efp(\cdot)$ around 2003. The LWZ criterion suggests one break point, which is estimated at August 2003. \\
+In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth of money supply.
+Combined, these actions were enough to ensure the succesfull participation in the ERM~II a year later.
 
-\begin{center}
-<<slovplotbreak, fig=TRUE, height=6, width=8>>=
-<<plot2>>
-@
-\end{center}
+\section{Power Analysis}
 
-Slovenia did achieve its goals by two very important measures.
-First, it liberalized the financial markets in 2003 and -- most important -- it restricted the year--on--year growth of the monetary base M1 severely starting in 2003.
-Combined, these actions were enough to ensure  Slovenian participation in the euro--zone.
-The \cite{imad} explains this somewhat differently by the cooperation of both the government and the Bank of Slovenia, which resulted in a careful handling of administered prices, a counter--cyclical adjustment of excise duties on fuels and moderate wage increases based on social agreements.
-According to them, due to theses measures, the effect of the euro introduction on the Slovenian inflation rate was 0.24 percentage points.
-% alternative explanation \cite{slacik} In Slovenia the government and the Bank of Slovenia helped to curb inflation with coordinated policy measures such as strict implementation of administered prices (increases not above headline in ation), counter-cyclical adjustment of excise duties on fuels and wage moderation based on a social agreement (see IMAD 2007).
-% the total effect of the euro changeover on prices to amount to 0.24 percentage points, being among the smallest recorded in the euro area --> quelle: http://www.umar.gov.si/fileadmin/user_upload/sporocila_za_javnost/euro.pdf
+(...Jesus would not do this ...) \\
+
+Basis idea: show that with a GL instead of a Normal, the probability of detecting minimal effects in the series (which is GL--distributed) is higher.
+If for example, just the skewness parameter changes, is a test build on the normality assumption able to capture this?
+(...more to come if we decide to include this.
+In principle, we could then argue in section \ref{s31} that a power analysis suggested that we might miss changes in the third moment or so, i.e.
+normal has lower power if series is GL--distributed)...)
 
 \newpage
-\section{Results}
+\section{Results}\label{s5}
 
-The results as we show them here are grouped according to patterns, which we think have some common properties.
-We present the breakdates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II or the date of euro introduction.
-To give a guidline for interpretation, we provide a short overview over the history of the EMU.
-It basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender and started with the Exchange Rate Mechanism~II (ERM~II) for Non--members and the irrevocable fixing of the exchange rates of the Member States.
-\\
+% besser formulieren
 
@@
 
-\small
+\footnotesize
 \begin{longtable}{|l|l|r|r|r|l|l|l|}
 \hline
-\multicolumn{8}{|c|}{Belgium and Luxembourg} \\
-\hline 
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(2, 13), format_moments)))
-@
 \multicolumn{8}{|c|}{The Eastern Countries} \\
@@
 @
+\hline
 \multicolumn{8}{|c|}{Italy and Spain} \\
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
 \hline
@@
 @
+\hline
 \multicolumn{8}{|c|}{Ireland} \\
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
 \hline
@@
 @
+\hline
 \multicolumn{8}{|c|}{No change countries} \\
 \hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(9,14,7,6), format_moments)))
+@
 \hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(9,14,6), format_moments)))
-@
 \multicolumn{8}{|c|}{Non participating countries} \\
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
 \hline
@@
 @
+\hline
 \multicolumn{8}{|c|}{The other countries} \\
 \hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
-\hline
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(1,4,7,8,16), format_moments)))
+writeLines(unlist(lapply(c(1,2,4,8,13,16), format_moments)))
 @
-
+\caption{Break point and parameter estimates and ERM information}
 \end{longtable}
@@
 
-The first  distinct group of contries consists of Belgium and Luxembourg.
-This may come as no surprise, since both countries formed a monetary union prior to the start of the ERM.
-In the past, they enjoyed very stable inflation rates with low values of both mean and variance, this changed around the timing of the irrevocable fixing of the exchange rates.
-But in the case of Belgium and Luxembourg, what we actually observe is an alternation due to the definition of the HICP.
-According to their national statistical offices, these changes can be traced back  to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation, details may be found in: \citet{eurlex} This lead to a 6 month seasonality pattern (January and July) afterwards (meaning after the breakpoints induced by sales inclusion), since sales periods in Belgium and Luxembourg are regulated by law and have to occur in the first two weeks of these months and therefor we observe this sharp price decreases in both month.
-Interestingly enough, both these price decreases are immediately followed by increases in February and August of even higher magnitude.
+% resultate: besser ordnen; mit paragraph oder subsection*{}; text über sales do inne; indicate implications in mean and discuss changes in var and skewness
+
+\clearpage
+\normalsize
+%(... results on one page, more focus on sales effects and parameter changes indicate implications of these changes ...) \\
+The table above presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of euro introduction. \\
+%The first  distinct group of countries consists of Austria and Belgium. This may come as some surprise, since originally Belgium and Luxembourg formed a monetary union. \\
+(...
+perhaps put the data problems into appendix?
+Presently I deal with this in $results3.rnw$ in the following fashion: I do assume that for Belgium, Luxembourg and Portugal (the countries where the exact time and date of inclusion of sales effects is definitely known) the seasonality in these months follows this logic: DECREASE in prices in January, almost offset by an increase in February and another DECREASE in July followed by an increase in August.
+So the series AFTER the inclusion of the sales periods should take into account that the seasonal effects of these month are now HIGHER than before.
+I do this by substracting the difference between the seasonal dummy in January (prior to inclusion) and the seasonal dummy in January (after the inclusion) from the original series.
+If this is the best way to deal with that, I do not know ...) \\
+In a prior inspection, Belgium and Luxembourg were very similar.
+However, this degree of similarity in the original series as provided by the OECD, actually could be traced back -- according to their national statistical offices -- to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation, details may be found in: \citet{eurlex}.\footnote{In addition to Belgium and Luxembourg, there are a number of other countries, where the sales periods are regulated by law.
+France has an individual time span regulated by each department.
+This should pose no problem, since there is no global state level solution.
+In Greece there are also two set sales--periods in winter and summer.
+In Italy the sales periods are fixed each year by the Chambers of Commerce.
+Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
+The source for this information is \citet{konseur}.
+Dummies were estimated for Belgium, Luxembourg and Portugal.
+Italy and Greece were not estimated, since -- as in the case of France -- these periods do not seem to be set for the whole country at once.} This lead to a 6 month seasonality pattern (January and July) after the inclusion of sales periods into the HICP.
+We corrected for this by estimating dummies for these month and subtracting them from the original time series.
+In the case of Belgium, the estimated break was postponed to late 2007, coinciding with the Austrian break.
+The time after the break is characterized by an increase in mean and a severe increase in the variance and can be traced back to two major influences: the beginning of the financial crisis and potentially some spillover effects of higher oil prices.
 \\
-The second of these groups are the Eastern European countries: the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
+The first of these groups are the Eastern European countries: the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
 In almost all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
 Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
+The break during the late 90ies is a result of these countries efforts to curb in inflation by decreasing money supply and indicates the time when these countries actually overcame the biggest transitionary shocks on their way towards free market economies.
+The case of Slovenia was already described.
+Slovakia is very similar, as its inflation dynamics too changed roughly one year prior to ERM~II entry.
 \\
-Another combination of countries with -- this time -- identical results are the two biggest southern European countries, Italy and Spain.
+Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain.
 Both of them experienced high rates of inflation in the early 90ies.
-Afterwards, in the phase leading up to the monetary union, mean and variance were significantly reduced.
-This trend changed drastically soon after the fixing of the exchange rates leading to higher mean and much higher variances then ever before since the beginning of the HICP series.
-\\
-Italy is quite a nice example to explain the usefulness of the skewness parameter.
-If we focus on the first two periods, with skewness lower than 1 (which indicates that the distribution is skewed to the right) we find that positive shocks are not normally neutralized by negative shocks.
-This changes in the last phase with skewness negative, indicating that the GL is now skewed to the left and that positive shocks will more likely be followed by negative shocks.
+Afterwards, in the phase leading up to the monetary union, mean and variance were greatly reduced.
+This trend changed significantly soon after the fixing of the exchange rates leading to higher mean and much higher variances then ever before since the beginning of the HICP series.
 \\
@@
 \\
-Another interesting group consists of 3 economically rather different countries.
-In none of them do we find any evidence of a structural change.
-Whereas the Netherlands and Finland are both on the lower side of the inflation margin, Greece exhibits some very strange behaviour.
-It has by far the highest variance of all series and seems to be completely different from all the others.
-\\
-The last group with clear patterns consists of the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of the United Kingdom in the EMS.}, the United Kingdom and Sweden.
+Another interesting group consists of 4 economically rather different countries.
+In none of them do we find any evidence of a change in inflation dynamics.
+Whereas France, Finland and the Netherlands are on the lower side of the inflation margin, the Greek inflation rate exhibits some very strange behaviour.
+It has by far the highest variance of all series and seems to be completely different from them.\footnote{This suggested some possible data problems, but unfortunately numerous efforts of communication failed with official sources failed.} \\
+The last group with clear patterns consists of the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of both countries in the EMS.}, the United Kingdom and Sweden.
 Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and the run on the Swedish currency in the same year.
 \\
-No clear interpretation is available for: Austria, Denmark, France, Germany and Portugal. \\
-\cite{busetti} find 3 clusters in their paper, if they include all the 12 countries in question (Austria, Belgium, Finland, France, Germany, Italy, the Netherlands, Luxembourg, Spain, Greece, Portugal, Ireland).
-The first 5 countries form the ``low inflation club'', Italy, the Netherlands and Luxembourg the ``medium inflation club'', whereas the other four countries form a high inflation club.
-We do not find much similarity between our cluster (which focuses on the timing of the break and the change in the parameters) and what they find via aggregation into inflation clubs.
-They also find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
-This does not quite go along with our findings, that do not indicate stronger divergence after 1997 by means of break point detection.
+No clear interpretation is available for: Austria, Belgium, Denmark,  Germany, Luxembourg and Portugal.
+For Austria, we observe a change in 2007, possibly explained by the oil price shock, which is the official explanation of the Statistics Austria.
+Belgium changed in 2000 with a clear reduction in mean and an increase in variance.
+In the case of Denmark we observe a increase in variance after 2000 with a very distinct change in skewness.
+Germany saw an increase in its inflation mean and volatility after 2000.
+Luxembourg suffered a severe increase in variance, similar to Italy and Spain.
+Portugal on the other hand was successfull in reducing its very high inflation rate substantially following the participation in the ERM arrangement.
+\\
+A disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
+A very volatile inflation rate is unlikely to contribute to greater macroeconomic stability.
+Although \cite{grier} find no evidence, that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --  \cite{jaro} finds a positive correlation between the level and the standard deviation of inflation.
+This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro--zone countries indicating that in future higher inflation rates are more likely.
+
+% finland  changes in skewness (but only 0.055) not significantly
+% slovenia around 2006
+
+\clearpage
+\subsection{Inflation Differentials}
+
+%(...some problems with graph creation inside .rnw, hope to fix it later ...) \\
+% was i zum sega ket hob ba busetti --> vgl. archivierte Datei vom 16.11
+% actually too detailed for so small a picture
+% des if un die } for am ät dazuatoa
+
+\begin{figure}[h!tp]
+<<greyplot, fig=TRUE, height=8, width=8>>=
+##if(!file.exists(file.path(root_dir, "Draft", "draft-greyplot.pdf"))) {
+<<plot3>>
+##}
+@
+\caption{Inflation rates and mean inflation rates over time}
+\end{figure}
+
+%\begin{figure}[h!tp]
+%\includegraphics[width=150mm, height=100mm]{draft-greyplot}
+%\caption{Inflation rates and mean inflation rates over time}
+%\end{figure}
+
+The disparity of the national inflation rates is still an issue, many researchers, like \citet{hofmann} still find a considerable amount of inflation differentials.
+\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
+In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
+They find that convergence over time of inflation dynamics was only partly observable.
+In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
+\\
+Concerning inflation differentials, we present a graph depicting the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
+If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.\footnote{We wold see one black line if all countries would show the same month--on--month increase at any one time.} We do see that many of these countries follow a rather similar path.
+One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
+However, only a small number of countries exhibits a break around this time (Austria and Ireland).
+
+%\begin{figure}[h!tp]
+%\includegraphics[width=150mm, height=100mm]{diffmean}
+%\caption{Differences in Euro--are mean without Ireland}
+%\end{figure}
+
+\begin{figure}[h!tp]
+<<diffmean, fig=TRUE, height=8, width=8>>=
+##if(!file.exists(file.path(root_dir, "Draft", "diffmean.pdf"))) {
+<<plot4>>
+##}
+@
+\caption{Differences in Euro excluding Estonia, Slovenia and Slovakia}
+\end{figure}
+
+Concerning the mean of the inflation rates, we see a clear convergence pattern.
+This result is reinforced by plotting the difference between the highest and the lowest mean inflation rate in the euro subsample, excluding the new Euro--member states Estonia, Slovakia and Slovenia.\footnote{Including them blurs the clear result somewhat by driving the difference up, since at the beginning, these countries still had very high inflation rates.} After a short hike at the beginning of the convergence phase in 1995, the difference declined clearly to a fraction of its previous value.
+However, the financial crisis clearly contributed to widening of the convergence gap.
+Our finding contradicts \citet{busetti} who find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
+This does not quite go along with the depicted means,  that do not indicate stronger divergence after 1997 but a continuing decrease of inflation differentials until the hikes after 2008.
+\\
+However, this is not the case for variance, where we find clear evidence of an increase.
+Doing the same thing for volatility, we see the second moment almost tripling after 1995.\footnote{Due to a very low volatility of countries like Spain and Luxembourg at the beginning.} This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
 \\
@@
 \clearpage
-\section{Discussion}
+\section{Conclusion}\label{s6}
 
-% was i zum sega ket hob ba busetti --> vgl. archivierte Datei vom 16.11
+(... write when rest is finished, talk about method and empirical findings...)
 
-% actually too detailed for so small a picture
+%We were able to outline changes in the dynamics of the inflation rates in various countries across Europe.
+However, we fail to find any clear pattern relating to the euro introduction, but find convincing evidence, that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
+\\
+%Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases could possibly be traced back to the strong seasonality in our time series.
+The economic reason for this however remains a riddle and should be subject to future research.
 
-%\begin{center}
-%<<greyplot, fig=TRUE, height=8, width=8>>=
-% ## compute GL mean (if no NA)
-%  glmean <- function(x) {
-%    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
-%  }
-%  ## returns
-%  ## why not include Finland?
-%  ## hicp <- diff(log(HICP[, -6]))
-%  hicp <- diff(log(HICP[,]))
-%
-%  ## rolling GL means
-%  hicp_glmean <- rollapply(hicp, 12, glmean)
-%
-%  ## semi-transparent gray and red
-%  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-%
-%  ## plot returns (gray) and rolling means (red)
-%  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
-%  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-%@
-%\end{center}
+% füge noch cpmputatinal details hinzu
 
-To clarify some issues, lets take a look at the data aggregated in just one graph.
-If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.\footnote{We wold see one black line if all countries would show the same month--on--month increase at any one time.} We do see that many of these countries follow a rather similar path.
-One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
-However, this was not enough for most of the countries -- with the exception of Ireland -- to trigger a structural break in inflation rate dynamics.
-\\
-One interesting aspect we would like to focus on is the clear trend in means.
-What we have plotted here actually is a rolling mean, where the mean is calculated using the first moment of a fitted GL distribution.
-It is clearly visible that these rates converged towards some common mean (graphically emphasized by the redder lines) which, from 2005 onwards, comes rather close to the ECB target rate (which would translate to a monthly increase of roughly 2 per mill).
-So concerning the mean of the inflation rates, we see a clear convergence pattern.
-However, this is not the case for variance, where we find clear evidence of an increase.
-This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
+\clearpage
+\section*{Computational Details}
+
+The results shown are computed with a new \proglang{R} package \pkg{glogis}, which is an enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a GL distribution.
+The tests and the graphical illustration of the empirical fluctuation process were developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
 \\
 
-\clearpage
-\section{Conlcusion}
+% appendix HINTER reference
+\begin{appendix}\ref{app}
 
-We were able to outline changes in the dynamics of the inflation rates in various countries across Europe.
-However, we fail to find any clear pattern relating to the euro introduction, but find convincing evidence, that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
-\\
-Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases could possibly be traced back to the strong seasonality in our time series.
-The economic reason for this however remains a riddle and should be subject to future research.
+\section{GL Distribution}\label{glscore}
 
-\clearpage
-\section{Appendix}\label{app}
-
-\subsection{GL distribution}
-
-The moments are:
+The moments of the GL distribution as given in equation \ref{eqn1} are:
 
@@
 
-where $\gamma(\cdot)$ is the digamma function and its derivatives. \\
+where $\gamma(\cdot)'$ and $\gamma(\cdot)''$ are its first and second derivatives, respectively. \\
 The log-likelihood is:
@@
 
-The resulting score function, where the scores are the derivatives of the log--likelihood function, has 3 components $(s_{\theta}, s_{\sigma}, s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$.
-These are used later on for the construction of the test statistics and are a means of measuring the estimation error in a maximum likelihood framework.
+The resulting score function, where the scores are the derivatives of the log--likelihood function, has three components $(s_{\theta}, s_{\sigma}, s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$.
+These are employed in equation \ref{eqn2} for the construction of the test statistics and are a means of measuring the estimation error in a maximum likelihood framework.
 
@@
 
-The scores of the parameters $\phi = (\theta, \sigma, \delta)$ sum up to 0: $\sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0$.
+The scores of the parameters $\phi = (\theta, \sigma, \delta)$ sum up to 0: $\sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0$. The first order condition is given by:
+
+\begin{eqnarray}
+\sum_{i=1}^n s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) = 0
+\end{eqnarray}
+
+\end{appendix}
 
 \clearpage
-\subsection{Graphs}
+\section{Graphs}
 
-% generate graphs in another fashion (a graph with data of results3) und zwar: results$Belgium$data
-% include fitted means --> still some work to do, not yet figured out due to xyplot and two lines...
-<<bellux, fig=TRUE, height=5, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$Belgium$data, results$Luxembourg$data), xlab="Time", ylab="", main="", screens=list("Belgium", "Luxembourg")), ylim=c(-2,2))
+(... less space, more on one page ...)\\
+Here we present the actual data where the blue lines in the graphs correspond to the estimated mean of the series.
+
+\begin{figure}[h!tp]
+<<itaspa, fig=TRUE, height=4, width=8>>=
+print(get_xyplot(c(19,12)))
 @
+\caption{Italy and Spain}
+\end{figure}
 
-<<eastern, fig=TRUE, height=7, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$CzechRepublic$data, results$Estonia$data, results$Hungary$data, results$Poland$data, results$Slovakia$data, results$Slovenia$data), xlab="Time", ylab="", main="", screens=list("Czech Republic", "Estonia", "Hungary", "Poland", "Slovakia", "Slovenia"), ylim=c(-2.7,2.7)))
+\begin{figure}[h!tp]
+<<sweuk, fig=TRUE, height=4, width=8>>=
+print(get_xyplot(c(20,21)))
 @
+\caption{Sweden and United Kingdom}
+\end{figure}
 
-<<itaspa, fig=TRUE, height=5, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$Spain$data, results$Italy$data), xlab="Time", ylab="", main="", screens=list("Italy", "Spain"), ylim=c(-2,2)))
+\begin{figure}[h!tp]
+<<eastern, fig=TRUE, height=10, width=8>>=
+print(get_xyplot(c(3, 5,10,17,18)))
 @
+\caption{The Eastern Countries}
+\end{figure}
 
+\begin{figure}[h!tp]
 <<ireland, fig=TRUE, height=4, width=8>>=
-plot(results$Ireland$data, main="Ireland", ylab="", xlab ="Time", ylim=c(-2,2))
+x <- results$Ireland
+plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "Ireland")
+lines(x$breakpoints, breaks = x$nbreaks)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
 @
+\caption{Ireland}
+\end{figure}
 
-<<nochange, fig=TRUE, height=5, width=8>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(xyplot(cbind(results$Finland$data, results$Greece$data, results$Netherlands$data), xlab="Time", ylab="", main="", screens=list("Finland", "Greece", "Netherlands"), ylim=c(-2.7,2.7)))
+\begin{figure}[h!tp]
+<<nochange, fig=TRUE, height=6, width=8>>=
+print(get_xyplot(c(6,7,9,14)))
 @
+\caption{Nochange Countries}
+\end{figure}
 
-<<sweuk, fig=TRUE, height=7, width=8>>=
-print(xyplot(cbind(results$Sweden$data, results$UnitedKingdom$data), xlab="Time", ylab="", main="", screens=list("Sweden", "United Kingdom")), ylim=c(-2,2))
+\begin{figure}[h!tp]
+<<nopattern, fig=TRUE, height=8, width=8>>=
+print(get_xyplot(c(1,2,4,8,13, 16)))
 @
-
-%<<nopattern, fig=TRUE, height=7, width=8>>=
-%trellis.par.set(theme = standard.theme(color = FALSE))
-%print(xyplot(cbind(results$Austria$data, results$Demark$data, results$France$data, results$Germany$data, results$Portugal$data), xlab="Time", ylab="", main="", screens=list("Austria", "Denmark", "France", "Germany", "Portugal"), ylim=c(-2,2)))
-%@
+\caption{Countries without discernible pattern}
+\end{figure}
 
@@
 
-\clearpage
-
 \end{document}

### 2011-03-11 pkg/paper/Paper/draft.Rnw
updated results, draft and papers.bib
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@77 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 setwd(file.path(root_dir, "Analysis"))
-Sweave("results.Rnw")
+## run results3 (with dummy creation for sales periods)
+Sweave("results3.Rnw")
 setwd(file.path(root_dir, "Draft"))
 }
+
 ## query result file names
@@
 "Poland", "Portugal",  "Slovakia", "Slovenia",  "Spain", "Sweden", "UK")
-@
-%
-%% analysis "macro" (assumes existence of "i", "HICP", and "add_breaks")
-<<analysis, eval=FALSE>>=
-## compute returns
-x <- 100 * diff(log(na.omit(HICP[,i])))
 
-## fit full-sample model
-x_gf <- glogisfit(x)
-
-## assess stability of full-sample model
-x_efp <- gefp(x_gf, fit = NULL)
-
-## estimate breakpoints (with minimal segment size of 2 years)
-x_bp <- breakpoints(x_gf, h = 24)
-
-## (minimal) number of breaks
-x_nbreaks <- if(sctest(x_efp, functional = supLM(0.1))$p.value > 0.05) {
-  0
-} else {
-  max(1, length(breakdates(x_bp))) + add_breaks[i]
-}
-
-## refit segmented model
-x_rf <- refit(x_bp, breaks = x_nbreaks)
-
-## save everything
-save(x, x_gf, x_efp, x_bp, x_nbreaks, x_rf, file = paste(colnames(HICP)[i], ".rda", sep = ""))
-
-## plot "macros" (assuming existence "x", "x_gf", ...)
-
-<<plot3, eval=FALSE>>=
+## include glmean dataset
+if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
   glmean <- function(x) {
@@
   hicp <- diff(log(HICP[,]))
+  ## rolling GL means
 
-  ## rolling GL means
   hicp_glmean <- rollapply(hicp, 12, glmean)
 
-  ## semi-transparent gray and red
-  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-
-  ## plot returns (gray) and rolling means (red)
-  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
-  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-@
-
-<<plot4, eval=TRUE>>=
-## plot convergence in mean
-  glmean <- function(x) {
-    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
-  }
-  ## returns
-  hicp <- diff(log(HICP[,]))
-
-  ## rolling GL means
-  ## without Estonia,  Slovakia, Slovenia
-  hicp_glmean <- rollapply(hicp, 12, glmean)
   hicp_euro <- diff(log(HICP[, c(1,2,6,7,8,9,11,12,13,14,16,19)]))
@@
 
-	plot(diff)
+  setwd(file.path(root_dir, "Analysis"))
+	save(hicp, hicp_glmean, diff, file = "glmean.rda")
+	setwd(file.path(root_dir, "Draft"))
+
+}
+
 @
@@
 
-%(...some problems with graph creation inside .rnw, hope to fix it later ...) \\
-% was i zum sega ket hob ba busetti --> vgl. archivierte Datei vom 16.11
-% actually too detailed for so small a picture
-% des if un die } for am ät dazuatoa
-
 \begin{figure}[h!tp]
 <<greyplot, fig=TRUE, height=8, width=8>>=
-##if(!file.exists(file.path(root_dir, "Draft", "draft-greyplot.pdf"))) {
-<<plot3>>
-##}
+  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+  ## plot returns (gray) and rolling means (red)
+  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
+  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
 @
@@
 \end{figure}
-
-%\begin{figure}[h!tp]
-%\includegraphics[width=150mm, height=100mm]{draft-greyplot}
-%\caption{Inflation rates and mean inflation rates over time}
-%\end{figure}
 
@@
 
-%\begin{figure}[h!tp]
-%\includegraphics[width=150mm, height=100mm]{diffmean}
-%\caption{Differences in Euro--are mean without Ireland}
-%\end{figure}
-
 \begin{figure}[h!tp]
 <<diffmean, fig=TRUE, height=8, width=8>>=
-##if(!file.exists(file.path(root_dir, "Draft", "diffmean.pdf"))) {
-<<plot4>>
-##}
+	plot(diff)
 @

### 2011-03-11 pkg/paper/Paper/draft.Rnw
"Draft" -> "Paper" directory
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@78 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 Sweave("results3.Rnw")
-setwd(file.path(root_dir, "Draft"))
+setwd(file.path(root_dir, "Paper"))
 }
@@
 	save(hicp, hicp_glmean, diff, file = "glmean.rda")
-	setwd(file.path(root_dir, "Draft"))
+	setwd(file.path(root_dir, "Paper"))
 
@@
 % appendix HINTER reference
-\begin{appendix}\ref{app}
+\begin{appendix}\label{app}
 

### 2011-04-02 pkg/paper/Paper/draft.Rnw
commented comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@79 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \Abstract{
-(... text when everything else finished, present innovation of paper...)
+%(... text when everything else finished, present innovation of paper...)
 }
@@
 
-(... Jesus: journals, EE (empirical economics): mehr EMU, weniger Methode, Journal of Applied Econometrics: genau umgekehrt...)
-(... economic story, big transition ...) \\
+%(... Jesus: journals, EE (empirical economics): mehr EMU, weniger Methode, Journal of Applied Econometrics: genau umgekehrt...)
+%(... economic story, big transition ...) \\
 Ever since the beginning of the European Union, the topic of a common currency was a controversial issue.
@@
 \\
-(... big transition continued, our contribution...) \\
+%(... big transition continued, our contribution...) \\
 The way towards the EMU basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
@@
 \\
-(...introduction needs to include the value added of this paper, why is it better, how does it compare to others...) \\
+%(...introduction needs to include the value added of this paper, why is it better, how does it compare to others...) \\
 We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates, which focuses on these changes and thereby upon the effect of the EMU on a number of European countries both within and without the Euro--zone.
@@
 \\
-(...effects of monetary unions, literature overview (more literature in discussion of results)...) \\
+%(...effects of monetary unions, literature overview (more literature in discussion of results)...) \\
 Two of the three most important channels that change inflation, the money supply and the degree of openness  are subject to fundamental changes once a country joins a monetary union.
@@
 
-(... Jesus would present data later, first describe what we are going to to do...) \\
+%(... Jesus would present data later, first describe what we are going to to do...) \\
 HICP series\footnote{seasonally unadjusted} from Jan.1990--Mar.2010 are obtained for 21 countries the \citet{oecd}.
@@
 % laut achim noch zuviel bla bla im text --> weniger als anektode, präziser
-(...less anectodic, more precise...) \\
+%(...less anectodic, more precise...) \\
 To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks.
@@
 \\
-(...discuss correlation...) \\
+%(...discuss correlation...) \\
 The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation. 
@@
 
-(... formulation problems ...) \\
-(...Jesus: GL not standard, explain why it is worth using it, what makes it better than ARIMA?...) \\
+%(... formulation problems ...) \\
+%(...Jesus: GL not standard, explain why it is worth using it, what makes it better than ARIMA?...) \\
 %Motivation Verteilung
@@
 \subsection{Test}
-(.. Jesus would cite a paper by Andrews and Ploberger 1993 that provided the key for creating confidence intervals around the estimates...) \\
+%(.. Jesus would cite a paper by Andrews and Ploberger 1993 that provided the key for creating confidence intervals around the estimates...) \\
 
@@
 
-\section{Power Analysis}
-
-(...Jesus would not do this ...) \\
-
-Basis idea: show that with a GL instead of a Normal, the probability of detecting minimal effects in the series (which is GL--distributed) is higher.
+%\section{Power Analysis}
+%
+%(...Jesus would not do this ...) \\
+%
+%Basis idea: show that with a GL instead of a Normal, the probability of detecting minimal effects in the series (which is GL--distributed) is higher.
 If for example, just the skewness parameter changes, is a test build on the normality assumption able to capture this?
@@
 %The first  distinct group of countries consists of Austria and Belgium. This may come as some surprise, since originally Belgium and Luxembourg formed a monetary union. \\
-(...
+%(...
 perhaps put the data problems into appendix?
@@
 
-(... write when rest is finished, talk about method and empirical findings...)
+%(... write when rest is finished, talk about method and empirical findings...)
 

### 2011-04-27 pkg/paper/Paper/draft.Rnw
updated paper and papers.bib results4 treats with seasonal effects
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@80 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 library("lattice")
+library("moments")
 
 ## additional functions
+momentgen <- function(flag=1,x, start=yearmon(c(1990,1998,2007)), end=yearmon(c(1998+11/12,2007+11/12,2010+11/12))) {
+# 1:mean, 2:variance, 3:skewness
+if(flag==1){
+mean = mat.or.vec(ncol(x),length(start))
+for(jj in 1:length(start)){
+x1 <- window(x, start=start[jj], end=end[jj])
+for(i in 1:ncol(x1)) {
+mean[i,jj] = mean(na.omit(x1[,i]))
+}
+}
+ret1 <- c(quantile(mean[,1], probs=.25), quantile(mean[,1], probs=.50), quantile(mean[,1], probs=.75))
+ret1 <- cbind(ret1)
+ret2 <- c(quantile(mean[,2], probs=.25), quantile(mean[,2], probs=.50), quantile(mean[,2], probs=.75))
+ret2 <- cbind(ret2)
+ret3 <- c(quantile(mean[,3], probs=.25), quantile(mean[,3], probs=.50), quantile(mean[,3], probs=.75))
+ret3 <- cbind(ret3)
+ret <- cbind(ret1,ret2,ret3)
+colnames(ret) <- c(start[1], start[2], start[3])
+rownames(ret) <- c("lower", "medium", "upper")
+return(ret)
+}
+if(flag==2){
+var = mat.or.vec(ncol(x),length(start))
+for(jj in 1:length(start)){
+x1 <- window(x, start=start[jj], end=end[jj])
+for(i in 1:ncol(x1)) {
+var[i,jj] = var(na.omit(x1[,i]))
+}
+}
+ret1 <- c(quantile(var[,1], probs=.25), quantile(var[,1], probs=.50), quantile(var[,1], probs=.75))
+ret1 <- cbind(ret1)
+ret2 <- c(quantile(var[,2], probs=.25), quantile(var[,2], probs=.50), quantile(var[,2], probs=.75))
+ret2 <- cbind(ret2)
+ret3 <- c(quantile(var[,3], probs=.25), quantile(var[,3], probs=.50), quantile(var[,3], probs=.75))
+ret3 <- cbind(ret3)
+ret <- cbind(ret1,ret2,ret3)
+colnames(ret) <- c(start[1], start[2], start[3])
+rownames(ret) <- c("lower", "medium", "upper")
+return(ret)
+}
+if(flag==3){
+skewness= mat.or.vec(ncol(x),length(start))
+for(jj in 1:length(start)){
+x1 <- window(x, start=start[jj], end=end[jj])
+for(i in 1:ncol(x1)) {
+skewness[i,jj] = skewness(na.omit(x1[,i]))
+}
+}
+ret1 <- c(quantile(skewness[,1], probs=.25), quantile(skewness[,1], probs=.50), quantile(skewness[,1], probs=.75))
+ret1 <- cbind(ret1)
+ret2 <- c(quantile(skewness[,2], probs=.25), quantile(skewness[,2], probs=.50), quantile(skewness[,2], probs=.75))
+ret2 <- cbind(ret2)
+ret3 <- c(quantile(skewness[,3], probs=.25), quantile(skewness[,3], probs=.50), quantile(skewness[,3], probs=.75))
+ret3 <- cbind(ret3)
+ret <- cbind(ret1,ret2,ret3)
+colnames(ret) <- c(start[1], start[2], start[3])
+rownames(ret) <- c("lower", "medium", "upper")
+return(ret)
+}
+}
+
 moments <- function(object, ...) UseMethod("moments")
@@
 setwd(file.path(root_dir, "Analysis"))
-## run results3 (with dummy creation for sales periods)
-Sweave("results3.Rnw")
+## run results4 (with dummy creation for sales periods)
+Sweave("results4.Rnw")
 setwd(file.path(root_dir, "Paper"))
@@
 
+table2<- vector(length = length(fnam), mode = "list")
+names(table2) <- sapply(strsplit(sapply(strsplit(fnam, "/", fixed = TRUE), tail, 1), ".", fixed = TRUE), head, 1)
+
+for(qq in seq_along(fnam)) {
+  load(file.path(root_dir, "Analysis", fnam[qq]))
+  table2[[qq]] <- x
+}
+
 ## compute moments
@@
 
+format_momenteuro <- function(momeuro, digits = 6, name="Euro") {
+  rval <- format(round(momeuro, digits = digits), nsmall = digits)
+  rval <- cbind(name,"", matrix(paste("$", rval, "$", sep = ""), ncol = 3))
+  rval[1,1] <- name
+  rval[2,1] <- ""
+  rval[3,1] <- ""
+  rval[1,2] <- "lower"
+  rval[2,2] <- "medium"
+  rval[3,2] <- "upper"
+  rval <- apply(rval, 1, paste, collapse = " & ")
+  rval <- paste(rval, "\\\\")
+  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
+  return(rval)
+}
 get_data <- function(i) {
@@
 
-%(... Jesus: journals, EE (empirical economics): mehr EMU, weniger Methode, Journal of Applied Econometrics: genau umgekehrt...)
-%(... economic story, big transition ...) \\
 Ever since the beginning of the European Union, the topic of a common currency was a controversial issue.
 Although the Economic and Monetary Union (EMU) is now a fact, the discussion about the economic effects of the Euro is far from being settled.
-The controversial topics range from the question of whether or not the Euro--zone is indeed an optimal currency area (OCA) all the way to the very survival of the Euro in light of the budgetary problems of a number of its member states.
-The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context is in the center of an ongoing debate, for instance the question about short-run and steady-state inflation uncertainty dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries which is the topic of an investigation by \citet{palomba}.
+The controversial topics range from the question of whether or not the Eurozone is indeed an optimal currency area (OCA) all the way to the very survival of the Euro in light of the budgetary problems of some of its member states.
+The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context, is in the center of an ongoing debate: this regards the question about short-run and steady-state inflation uncertainty dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries, which is the topic of an investigation by \citet{palomba}.
 \\
-%(... big transition continued, our contribution...) \\
 The way towards the EMU basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
-Any new countries who want to join the EMU are first obliged to fulfill the Maastricht which, besides setting rules for government debts and interest rates, also requires the participation in the ERM~II for two years, while the exchange rates towards the euro are not allowed to cross the nominal band.
-These criteria acknowledge, that a stable inflation rate is one of the key goals of every good macroeconomic policy.
-\cite{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
+A possible fourth stage is the continuing integration of new member states, mostly former communist countries.
+Any new state that wants to join the EMU is first obliged to fulfill the Maastricht criteria, which, besides setting rules for government debts and interest rates, also requires the participation in the ERM~II (Exchange Rate Mechanism~II) for two years, while the exchange rates towards the euro are not allowed to cross the nominal band.\\
+\cite{emerson} put emphasize on the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
 Failing to stabilize inflation leads to severe economic problems.
 Throughout the literature, there is still a considerable degree of uncertainty as to what extent the introduction of the euro, or monetary unions in general, affect the inflation rate.
-There are a number of reasons for suspecting changes.
-Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria were likely to lead to an alteration at least in the mean (since this was required for a number of countries) and possibly in the volatility of their respective inflation rates as well.
-The EMU requires a country to possibly change a number of economic policies, which can only be achieved by big transitionally efforts.
+The EMU requires a country to possibly alter a number of economic policies and institutions, which may demand for significant short term adjustments to take place.
+\clearpage
+Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria are likely to lead to an alteration at least in the mean (since this was one of the requirements) and possibly in the volatility of their respective inflation rates as well.
 \\
-%(...introduction needs to include the value added of this paper, why is it better, how does it compare to others...) \\
-We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates, which focuses on these changes and thereby upon the effect of the EMU on a number of European countries both within and without the Euro--zone.
-The framework developed in this paper allows testing for any significant changes of a countries inflation dynamics and thus helps interpreting the effects of the EMU.
+%(...too short: need: methodological ideas and main questions..., value added: schiefe, schwere ränder)\\
+% http://home.iitk.ac.in/~kundu/paper152.pdf  (gl-beschreibung in unterlagen), p.4 more peaked in center, but heavier tails
+However, a deeper inquiry into this problem is in need of a method able to capture changes in the inflation dynamics.
+We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates, which focuses on these changes and thereby upon the effect of the EMU on a number of European countries both within and without the Eurozone.
+The framework developed in this paper makes it possible to test for any significant changes of a countries inflation dynamics and thus helps interpreting the effects of the EMU.
+We focus on changes in the first three moments of these inflation rates, which we assume to be distributed according to a generalized logistic distribution (GL), as defined in Section \ref{s3}.
+The GL distribution is a suitable one due to its flexibility to capture nonsymmetry and because of its fatter tails.
 For this purpose, we developed a new estimation approach building on previous research by \citet{z07}.
 \\
-%(...effects of monetary unions, literature overview (more literature in discussion of results)...) \\
-Two of the three most important channels that change inflation, the money supply and the degree of openness  are subject to fundamental changes once a country joins a monetary union.
-The original countries participating in the Exchange Rate Mechanism (ERM) had to synchronize money supply growth and interest rates in order to fix exchanges rates or otherwise risk the chance of severe disequilibria.
-Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
-\\
-Openness, i.e.
-the inflow of new goods and services or the abolishment of tariffs and capital restraint and regulations, is more likely to generate deflationary pressure, due to the substitution of cheaper goods for more expensive ones.
-Since deflation is rather unusual, with general price increases being the norm, the logical conclusion is that the most important inflationary channel are changes in money supply.
-%\footnote{Indeed, a general rise in prices is impossible without a simultaneous increase in money supply.}. 
-These changes in turn are influenced by a countries monetary policy and the credit behavior of its banking industry.
-External shocks, like the spillover effects of a hike in oil prices or the financial crisis may be important as well.
-\\
-Economic theory unfortunately does not provide any clear foundation for a proper answer whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate.
-%\citet{cooper} points out, that ``a central bank under a monetary union will internalize the interdependence between countries and optimally choose a lower inflation rate" and he argues that a ``central Bank governing the growth of money supply will optimally choose zero inflation." This is not the case with the ECB which targets at 2\% so as to avoid the alleged risk of deflation.
-It is thus not quite clear how a monetary union will affect the volatility and the level of the inflation rate.\\
-An interesting approach to this question is taken by \citet{holte}. 
-%He creates a two country model for monetary policy analysis along the line of two models by \citet{mc1} and \citet{mc2} as well as \citet{gali}.
-The inflation is modeled by a hybrid Phillips curve  specification and there are a number of different home country interest rate rules, like strict inflation targeting, flexible targeting, pegging to a currency and --  monetary union.
-What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
-The effects of joining a monetary union on inflation variability in his framework depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-However, due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on a variety of assumptions.
-\\
-The remainder of this paper is structured as follows: section \ref{s2} presents the data, section \ref{s3} presents the model and the estimation techniques used, section \ref{s4} illustrates our approach using Slovenia as an example, section \ref{s5} presents the results and section \ref{s6} concludes.
+%(...
+am Schluss so enden, dass nun interessant, wie sich die Momente so genau verändert haben; auf Basis der theorie--> suche nach brüchen, schiefe noch hinein/besser hinein; 3 saha simultan aschoa...)\\
+This framework allows us to observe the evolution of mean, variance and skewness over time. 
+%Considering the prior knowledge available about the timing of a countries participation in any of the aforementioned steps towards monetary union, an exact timing of breakpoints in the moments of the respective inflation series enables us to draw conclusions about the exact effects of the EMU the inflation rates of its members.
+Our estimation technique enables the simultaneous estimation of breaks in the mean (which is important in the context of the Maastricht criteria), the variance (which should be low for the price mechanism to be efficient) and the skewness (where a change in skewness is evidence of a shift of the distribution towards a new ``center of gravity'') of the inflation rates of interest.
+This than can be used to search for answers regarding the question about the effect of the EMU on the inflation dynamics of its member states or more general to investigate the effects of external shocks.\\
+The remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.
 
@@
 
-%(... Jesus would present data later, first describe what we are going to to do...) \\
-HICP series\footnote{seasonally unadjusted} from Jan.1990--Mar.2010 are obtained for 21 countries the \citet{oecd}.
+We employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}.
 Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom.
@@
 The countries in this sample can be divided into three different groups.
-First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain).
+First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).
 Second, the EU members not participating in the ERM~II: Czech Republic, Hungary, Poland, United Kingdom, and Sweden.
@@
 \\
-
+%(...
+schlussverkäufe schon hier erwähnen, einfach cool, ja das gibts, wir haben die aber weggeschätzt; generell viel angelsächsischer/präpotenter schreiben; nur 1 weg der richtige und zwar unserer; nicht erwähnen, dass andere dinge auch noch möglich; so nach motto: mich interessierts, ich machs, andere erst am ende erwähnen...)
 % greece: -for winter discounts: First 10 days of January-end of February; -for summer discounts: First 10 days of July -end od August; but has always been included
+In the case of Belgium and Luxembourg, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium and January 1999 for Luxembourg) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months} and subtracting them from the original time series.
 
@@
 
-% generell strategie: estimate, test parameter stability, breaks model
-
-%neuer Titel
-%Strukturbruch
-%slowenisches Beispiel
-
-% laut achim noch zuviel bla bla im text --> weniger als anektode, präziser
-%(...less anectodic, more precise...) \\
 To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks.
@@
 This then enables testing for the stability of the parameter estimates.
-If there is evidence of significant changes, one proceeds by computing the number and the timing of the break points.
+If there is evidence of significant changes, the number and the timing of the break points are computed.
 \\
@@
 %(...discuss correlation...) \\
-The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation. 
+The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative testes, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.}
 %To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.
@@
 
-%(... formulation problems ...) \\
-%(...Jesus: GL not standard, explain why it is worth using it, what makes it better than ARIMA?...) \\
-%Motivation Verteilung
 The data suggested that the use of a normal distribution might cause some estimation problems, since most of the time series exhibit asymmetric properties (with sometimes rather strong skewness) and -- compared to the normal distribution -- a fatter tail.
@@
 with location $\theta$, scale $\sigma$ and shape $\delta$.
-For $\delta=1$ the distribution simplifies to the logistic distribution, for $\delta<1$ it is skewed to the left and for $\delta$$>$1 it is skewed to the right (see appendix \ref{glscore} for more details).
+For $\delta=1$ the distribution simplifies to the logistic distribution, for $\delta<1$ it is skewed to the left and for $\delta$$>$1 it is skewed to the right (see Appendix \ref{glscore} for more details).
 \\
 In econometrics, the logistic distribution is often used in income distributions and growth models.
-This is due to  its fatter tail, which fit these kind of data somewhat better.
+This is due to  its fatter tails, which fit these kind of data somewhat better.
 \citet{won} employ a GL distribution in a regression model with autocorrelated errors.
@@
 A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they do not use a shape parameter.
-Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context to expected inflation.
-\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates.
+Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context with expected inflation.
 \\
+\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates. \\
 
@@
 
-% anders erzählen: um ein Modell zu anzupassen, dann den restlichen teil
-%Schätzung
-%obiger teil erst später; den verweis auf R erst am Ende in den computational details; auch das zeug ab an enhancement... in die computatinal details aufnehmen
-%(... better separation between estimation and test, rest to computational details...) \\
-
-Under the assumption, that $y_t$ is distributed according to the generalized logistic distribution, we test for the hypothesis that the parameter vector $\hat{\phi}$ stays stable over time:
+Under the assumption that $y_t$ is distributed according to the generalized logistic distribution, we test for the hypothesis that the parameter vector $\hat{\phi}$ stays stable over time:
 
@@
 
-where $\hat{V}^{-1/2}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
+where $\hat{V}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
 Then a functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian bridge:
@@
 \subsection{Test}
-%(.. Jesus would cite a paper by Andrews and Ploberger 1993 that provided the key for creating confidence intervals around the estimates...) \\
 
@@
 This test is well suited for single break alternatives and performs better if more than just one of the distribution parameters change.
+
+\clearpage
 It is given by:
@@
 \subsection{Breakpoint Estimation}
-
+%(...cite bai and perron 2003 paper here, nachlesen, findet achim super, bei slowenien will er noch konfidenzintervalle dazu...)\\
+The dating of structural breaks builds upon the work of \cite{bp03}.
 If a break is detected via the supLM test, the next step is estimating the number and the timing of these break points.
@@
 
-% noch mehr dazu: finds 0 wenn nicht signifikant, 1 wenn signifikant, dann auswahl über das lwz criterion
-
 \clearpage
 \section{An Example}\label{s4}
-
-% anders: the data encompasses, thus not included....
-%(... shortend, changed formulation...) \\
-For an illustration of the method, consider the case of Slovenia.
-The data collected encompasses the period from Jan 1996 onwards.
-During the beginning of this period, Slovenia experienced very high inflation rates but was successfull in continually decreasing them.
-A glance at the data suggests a break around 2003/04, with an evident change in the mean.
+%(...more text: 1. motivation, 2.clear methodological steps, 3.interpretation...) \\
+<<echo=FALSE>>=
+        x <- results$Slovenia
+        slovpval <- sctest(x$efp, functional = supLM(0.1))$p.value
+@
+To illustrate our method, we consider the case of Slovenia.
+After a period of relative instability following the independence from Yugoslavia, Slovenia was successfull in realigning its economy and introduced a number of reforms leading to a stable growth in the past years.
+The good economic performance made it possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce the Euro in January 2007.
+The data collected contains the period from January 1996 onwards.
+In the first few years, Slovenia experienced very high inflation rates, but was successfull in reducing inflation to a much lower level.
+\\
+A glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period.
+The $efp(\cdot)$ starts to significantly deviate as early as 2001 and reaches the peak somewhere between 2003 and 2004.
+The p--value of the supLM test is \Sexpr{round(slovpval, 5)} and clearly indicates at least one break.
+The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.
+In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate.
+Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the Euro two years later.
 
@@
 \caption{Slovenia: breakpoint estimate and supLM test}
+\label{fig1}
 \end{figure}
-
-The supLM test confirms this belief by picturing a hike of the $efp(\cdot)$ around 2003. The LWZ criterion suggests one break point, which is estimated at August 2003. \\
-In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth of money supply.
-Combined, these actions were enough to ensure the succesfull participation in the ERM~II a year later.
-
-%\section{Power Analysis}
-%
-%(...Jesus would not do this ...) \\
-%
-%Basis idea: show that with a GL instead of a Normal, the probability of detecting minimal effects in the series (which is GL--distributed) is higher.
-If for example, just the skewness parameter changes, is a test build on the normality assumption able to capture this?
-(...more to come if we decide to include this.
-In principle, we could then argue in section \ref{s31} that a power analysis suggested that we might miss changes in the third moment or so, i.e.
-normal has lower power if series is GL--distributed)...)
 
@@
 
-% besser formulieren
-
-%
-%das package colortbl (\usepackege[dvips]{colortbl}) ermöglicht z.B. mit
+% das package colortbl (\usepackege[dvips]{colortbl}) ermöglicht z.B. mit
 %\rowcolor[gray]{0.90} in einer Tabellenzeile diese farbig (bzw. hier grau) zu unterlegen.
@@
 \hline
-\multicolumn{8}{|c|}{The Eastern Countries} \\
+\multicolumn{8}{|c|}{Austria and Belgium} \\
 \hline
 Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(1, 2), format_moments)))
+@
+\hline
+\multicolumn{8}{|c|}{The Eastern countries} \\
 \hline
@@
 \hline
-\multicolumn{8}{|c|}{Non participating countries} \\
+\multicolumn{8}{|c|}{The 90s} \\
 \hline
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(20, 21), format_moments)))
+writeLines(unlist(lapply(c(16,20, 21), format_moments)))
 @
 \hline
-\multicolumn{8}{|c|}{The other countries} \\
+\multicolumn{8}{|c|}{Former DM--zone} \\
 \hline
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(1,2,4,8,13,16), format_moments)))
+writeLines(unlist(lapply(c(4,8,13), format_moments)))
 @
 \caption{Break point and parameter estimates and ERM information}
+\label{resulttable}
 \end{longtable}
-% results3 hier einbauen
 % warning: the groups then change; Italy and Greece have not yet answered
@@
 
-% resultate: besser ordnen; mit paragraph oder subsection*{}; text über sales do inne; indicate implications in mean and discuss changes in var and skewness
-
 \clearpage
 \normalsize
-%(... results on one page, more focus on sales effects and parameter changes indicate implications of these changes ...) \\
-The table above presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of euro introduction. \\
-%The first  distinct group of countries consists of Austria and Belgium. This may come as some surprise, since originally Belgium and Luxembourg formed a monetary union. \\
-%(...
-perhaps put the data problems into appendix?
-Presently I deal with this in $results3.rnw$ in the following fashion: I do assume that for Belgium, Luxembourg and Portugal (the countries where the exact time and date of inclusion of sales effects is definitely known) the seasonality in these months follows this logic: DECREASE in prices in January, almost offset by an increase in February and another DECREASE in July followed by an increase in August.
-So the series AFTER the inclusion of the sales periods should take into account that the seasonal effects of these month are now HIGHER than before.
-I do this by substracting the difference between the seasonal dummy in January (prior to inclusion) and the seasonal dummy in January (after the inclusion) from the original series.
-If this is the best way to deal with that, I do not know ...) \\
-In a prior inspection, Belgium and Luxembourg were very similar.
-However, this degree of similarity in the original series as provided by the OECD, actually could be traced back -- according to their national statistical offices -- to the inclusion of sales (winter and summer) in January 1999 for Belgium and January 2000 for Luxembourg which was demanded by a Commission regulation, details may be found in: \citet{eurlex}.\footnote{In addition to Belgium and Luxembourg, there are a number of other countries, where the sales periods are regulated by law.
-France has an individual time span regulated by each department.
-This should pose no problem, since there is no global state level solution.
-In Greece there are also two set sales--periods in winter and summer.
-In Italy the sales periods are fixed each year by the Chambers of Commerce.
-Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
-The source for this information is \citet{konseur}.
-Dummies were estimated for Belgium, Luxembourg and Portugal.
-Italy and Greece were not estimated, since -- as in the case of France -- these periods do not seem to be set for the whole country at once.} This lead to a 6 month seasonality pattern (January and July) after the inclusion of sales periods into the HICP.
-We corrected for this by estimating dummies for these month and subtracting them from the original time series.
-In the case of Belgium, the estimated break was postponed to late 2007, coinciding with the Austrian break.
-The time after the break is characterized by an increase in mean and a severe increase in the variance and can be traced back to two major influences: the beginning of the financial crisis and potentially some spillover effects of higher oil prices.
+
+%(...3 punkte darstellung noch einbauen...) \\
+%(...andere struktur: 1.short introduction, 2. countries in detail (6.1), 3. summary (6.2) ...) \\
+Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction.
+We formed seven groups of countries according to the observed pattern of break points.
+These seven groups consist of Austria and Belgium, the Eastern European countries, the two big southern European countries, Italy and Spain, Ireland, the countries with no change in the inflation dynamics, the countries with breaks in the early 90s and the member countries of the former DM--zone (excluding Belgium).
 \\
-The first of these groups are the Eastern European countries: the Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
-In almost all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90ies.
+
+\subsection{The countries in detail}
+
+Regarding Austria and Belgium, the time after the break is characterized by an increase in the mean and a severe increase in the variance of the inflation rates.
+The skewness is now negative for both series, which indicates that the distribution has shifted to the right, implying that overall inflation is likely to increase.
+This change can be traced back to two major influences: the financial crisis and potentially some spillover effects of higher oil prices in late 2007.
+\\
+The second group consists of the Eastern European countries: Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
+In almost all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90s (Slovenia is an exception with an increase in variance after the break).
 Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
-The break during the late 90ies is a result of these countries efforts to curb in inflation by decreasing money supply and indicates the time when these countries actually overcame the biggest transitionary shocks on their way towards free market economies.
-The case of Slovenia was already described.
-Slovakia is very similar, as its inflation dynamics too changed roughly one year prior to ERM~II entry.
+The break during the late 90s is a result of these countries efforts to curb in inflation by decreasing money supply and indicates the time when these countries actually overcame the biggest transitionally shocks on their way towards free market economies.
+Slovakia is very similar to Slovenia, as its inflation dynamics too changed roughly one year prior to its entry into ERM~II.
 \\
 Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain.
-Both of them experienced high rates of inflation in the early 90ies.
+Both of them experienced high rates of inflation in the early 90s.
 Afterwards, in the phase leading up to the monetary union, mean and variance were greatly reduced.
-This trend changed significantly soon after the fixing of the exchange rates leading to higher mean and much higher variances then ever before since the beginning of the HICP series.
+Soon after the fixing of the exchange rates, this trend changed significantly, leading to a higher mean and a much higher variance.
+A possible reason for this is that the fiscal responsibility that was required for an entry into the Eurozone was softened after this goal was achieved.
+This hypothesis gains some credibility considering the huge government deficits of these countries after 1999.
 \\
+%(... in irland der schock schon persistenter als in anderen ländern, bei der grauen graphik zeigen (greyplot...)\\
 Ireland can be considered a rather special case.
 It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.
-We find a structural break in 2008 following the strong contraction triggered by the bursting of the Irish housing bubble and a beginning deflation.
+In Ireland, compared to other countries, the deflationary effect of the financial crises was more persistent.
+This deflationary trend, which for a time affected all countries, can be seen in Figure \ref{greyplot}, which depicts a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
+There is evidence of a decline in mean inflation, which returns to its previous path roughly a year later.
+The break in 2008 can be attributed to the strong contraction triggered by the bursting of the Irish housing bubble and a beginning deflation in concert with the massive costs of the bailout of the countries banking system.
+Surprisingly, this deflationary pressure is not visible in Spain, which was affected by a housing bubble of roughly the same size as Ireland.
 \\
-Another interesting group consists of 4 economically rather different countries.
+%(...change formulation for greece...) \\
+\clearpage
+Another interesting group consists of four economically rather different countries.
 In none of them do we find any evidence of a change in inflation dynamics.
-Whereas France, Finland and the Netherlands are on the lower side of the inflation margin, the Greek inflation rate exhibits some very strange behaviour.
-It has by far the highest variance of all series and seems to be completely different from them.\footnote{This suggested some possible data problems, but unfortunately numerous efforts of communication failed with official sources failed.} \\
-The last group with clear patterns consists of the two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of both countries in the EMS.}, the United Kingdom and Sweden.
-Both have a clear break in the early 90ies, and both of these breaks can be traced back to crisis, the currency crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and the run on the Swedish currency in the same year.
+Whereas France, Finland and Netherlands are on the lower side of the inflation margin, the Greek inflation rate is among the highest.
 \\
-No clear interpretation is available for: Austria, Belgium, Denmark,  Germany, Luxembourg and Portugal.
-For Austria, we observe a change in 2007, possibly explained by the oil price shock, which is the official explanation of the Statistics Austria.
-Belgium changed in 2000 with a clear reduction in mean and an increase in variance.
-In the case of Denmark we observe a increase in variance after 2000 with a very distinct change in skewness.
-Germany saw an increase in its inflation mean and volatility after 2000.
+The next group consists of two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of both countries in the EMS.}, the United Kingdom and Sweden.
+Portugal is included due to the same break point behavior.
+All three show a clear break in the early 90s.
+Two of them can be traced back to economic crisis, the currency crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and the run on the Swedish currency in the same year.
+In the case of Portugal, the inflation rate after 1992 was much more stable, showing that the inflationary problems of Portugal were less severe thereafter.\\
+The last group consists of the former DM--zone countries Denmark, Germany and Luxembourg.
+In the case of Denmark we observe an increase in variance after 2000 with a very distinct change in skewness.
+In Germany, the inflation rate increased both on average and in its variability.
 Luxembourg suffered a severe increase in variance, similar to Italy and Spain.
-Portugal on the other hand was successfull in reducing its very high inflation rate substantially following the participation in the ERM arrangement.
+All of these countries, which profited from the stability of the previous German currency, still have a very low inflation rate.
+
+\subsection{Overview over the general trend}
+
+The disparity of the national inflation rates is still an issue, many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials.
+\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
+In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
+They find that convergence over time of inflation dynamics was only partly observable.
+In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
 \\
-A disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
-A very volatile inflation rate is unlikely to contribute to greater macroeconomic stability.
-Although \cite{grier} find no evidence, that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --  \cite{jaro} finds a positive correlation between the level and the standard deviation of inflation.
-This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro--zone countries indicating that in future higher inflation rates are more likely.
-
-% finland  changes in skewness (but only 0.055) not significantly
-% slovenia around 2006
-
-\clearpage
-\subsection{Inflation Differentials}
+Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.
+We wold see one black line if all countries would show the same month--on--month increase at any one time.
+The red lines depict the means.} We do see that many of these countries follow a rather similar path.
+One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
+Over the time, there is clear evidence, that the inflation rates converged towards some common -- albeit broad -- mean level.
 
@@
 @
-\caption{Inflation rates and mean inflation rates over time}
+\caption{Inflation rates and estimated GL mean over time}
+\label{greyplot}
 \end{figure}
 
-The disparity of the national inflation rates is still an issue, many researchers, like \citet{hofmann} still find a considerable amount of inflation differentials.
-\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
-In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
-They find that convergence over time of inflation dynamics was only partly observable.
-In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
+\clearpage
+
+Regarding the general trend, Table \ref{resulttable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series during the convergence phase (January 1990 until December 1998), the Euro phase (January 1999 until August 2008) and the time after the financial crisis (since September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.} until now), for the countries in the ``core'' of the Eurozone (Austria, Belgium, Finland, France, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal and Spain) and all the others.
 \\
-Concerning inflation differentials, we present a graph depicting the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
-If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.\footnote{We wold see one black line if all countries would show the same month--on--month increase at any one time.} We do see that many of these countries follow a rather similar path.
-One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
-However, only a small number of countries exhibits a break around this time (Austria and Ireland).
+Regarding the mean, we see a clear evolution towards a lower mean in both groups, where the core countries show much lower values.
+Almost the opposite is true concerning the variance of the inflation rates of the core countries, that has increased to a much higher value than at the start of the sample, with the upper percentile during the 90s now almost equaling the lower percentile for the time after the financial crisis; the variance for the other countries has declined significantly.
+Concerning skewness, we see a clear trend in the core countries for skewness to become negative, indicating that the bulk of the values now lies right to the mean.
+The same is true -- although to a lesser extent -- for the other countries.
+\\
 
-\begin{figure}[h!tp]
-<<diffmean, fig=TRUE, height=8, width=8>>=
-	plot(diff)
+\begin{longtable}{|l|l|l|l|l|}
+\hline
+\multicolumn{5}{|c|}{Mean} \\
+\hline
+Zone & Percentile & 1990(1)--1998(12) & 1999(1)--2008(8) & 2008(9)--2010(12)\\
+\hline
+<<echo=FALSE, results=tex>>=
+euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[9]],table2[[11]],table2[[12]],table2[[13]],table2[[14]],table2[[16]],table2[[19]])
+noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
+writeLines(unlist(format_momenteuro(momentgen(1,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
 @
-\caption{Differences in Euro excluding Estonia, Slovenia and Slovakia}
-\end{figure}
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(format_momenteuro(momentgen(1,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+@
+\hline
+\multicolumn{5}{|c|}{Variance} \\
+\hline
+Zone & Quantil & 1990(1)--1998(12) & 1999(1)--2007(12) & 2008(1)--2010(12)\\
+\hline
+<<echo=FALSE, results=tex>>=
+euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[9]],table2[[11]],table2[[12]],table2[[13]],table2[[14]],table2[[16]],table2[[19]])
+noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
+writeLines(unlist(format_momenteuro(momentgen(2,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+@
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(format_momenteuro(momentgen(2,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+@
+\hline
+\multicolumn{5}{|c|}{Skewness} \\
+\hline
+Zone & Quantil & 1990(1)--1998(12) & 1999(1)--2007(12) & 2008(1)--2010(12)\\
+\hline
+<<echo=FALSE, results=tex>>=
+euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[9]],table2[[11]],table2[[12]],table2[[13]],table2[[14]],table2[[16]],table2[[19]])
+noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
+writeLines(unlist(format_momenteuro(momentgen(3,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+@
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(format_momenteuro(momentgen(3,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+@
+\hline
+\caption{Development of first three moments}
+\label{resulttable}
+\end{longtable}
 
-Concerning the mean of the inflation rates, we see a clear convergence pattern.
+\clearpage
+\section{Conclusion}\label{s6}
+
+Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
+\\
+An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
+What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
+The effects of joining a monetary union on inflation variability in his framework depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
+However, due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on a variety of assumptions.
+\\
+Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
+This is due to the fact that two of the most important inflationary channels, the money supply and the degree of openness, had to be arranged according to the criteria necessary for EMU entry.
+To achieve convergence in inflation rates, the growth in money supply had to be in concert with the M3 growth rate as decided by the European Central Bank (ECB).
+The opening of the economies made cheaper goods available and thus contributed to lower prices.
+Like the original countries participating in the ERM, the Eastern European countries had to synchronize money supply growth and interest rates in order to fix exchanges rates or otherwise risk the chance of severe disequilibria.
+\\
+%Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
+For the countries in the core of the Eurozone, the most disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
+A very volatile inflation rate will likely  contribute to greater macroeconomic instability.
+Although \cite{grier} find no evidence that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --  \cite{jaro} finds a positive correlation between the level and the standard deviation of inflation.
+This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro--zone countries indicating that in future higher inflation rates are more likely.
+This will put pressure on the ECB to achieve its inflation target of 2\%.
+
+%\begin{figure}[h!tp]
+%<<diffmean, fig=TRUE, height=8, width=8>>=
+%	plot(diff)
+%@
+%\caption{Differences in Euro excluding Estonia, Slovenia and Slovakia}
+%\end{figure}
+%
+%Concerning the mean of the inflation rates, we see a clear convergence pattern.
 This result is reinforced by plotting the difference between the highest and the lowest mean inflation rate in the euro subsample, excluding the new Euro--member states Estonia, Slovakia and Slovenia.\footnote{Including them blurs the clear result somewhat by driving the difference up, since at the beginning, these countries still had very high inflation rates.} After a short hike at the beginning of the convergence phase in 1995, the difference declined clearly to a fraction of its previous value.
@@
 \\
-However, this is not the case for variance, where we find clear evidence of an increase.
+%However, this is not the case for variance, where we find clear evidence of an increase.
 Doing the same thing for volatility, we see the second moment almost tripling after 1995.\footnote{Due to a very low volatility of countries like Spain and Luxembourg at the beginning.} This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
 \\
-
-\clearpage
-\section{Conclusion}\label{s6}
-
-%(... write when rest is finished, talk about method and empirical findings...)
-
-%We were able to outline changes in the dynamics of the inflation rates in various countries across Europe.
-However, we fail to find any clear pattern relating to the euro introduction, but find convincing evidence, that the zeal of the Eastern European countries to join the European currency at the earliest possible date, clearly helped in reducing mean and variance of their inflation rates.
-\\
-%Considering the severe increase in overall volatility of the European inflation rates, it should be pointed out however, that most of these increases could possibly be traced back to the strong seasonality in our time series.
-The economic reason for this however remains a riddle and should be subject to future research.
-
-% füge noch cpmputatinal details hinzu
 
@@
 
-where $\gamma(\cdot)'$ and $\gamma(\cdot)''$ are its first and second derivatives, respectively. \\
+where $\gamma'(\cdot)$ and $\gamma''(\cdot)$ are the first and the second derivatives of the GL distribution, respectively. \\
 The log-likelihood is:
@@
 \begin{eqnarray}
-\log f(y|\theta,\sigma, \delta) & = & \log(\delta) - \log(\sigma)   \\
-& - &  \frac{1}{\sigma} (y-\theta) - (\delta+1) \nonumber \\
+\log f(y|\theta,\sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{1}{\sigma} (y-\theta) - (\delta+1) \nonumber \\
 & \times & \log (1+\exp^{-\frac{y-\theta}{\sigma}}) \nonumber 
@@
 
+\section{Data and Sales Periods}\label{dataprobs}
+
+In addition to Belgium and Luxembourg, there are a number of other countries, where the sales periods are regulated by law.
+France has an individual time span regulated by each department.
+This should pose no problem, since there is no global state level solution.
+In Greece there are also two set sales--periods in winter and summer.
+In Italy the sales periods are fixed each year by the Chambers of Commerce.
+Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
+The source for this information is \citet{konseur}.
+According to information by their national statistical offices, the official dates of the inclusion of these sales periods are: January 2000 for Belgium and January 1999 for Luxembourg.
+Italy, Greece and Portugal were not estimated, since -- as in the case of France -- these periods either do not seem to be set for the whole country at once, or information about the treatment of sales periods in the HICP was not provided by the national statistical offices upon  request.
+
 \end{appendix}
@@
 
-(... less space, more on one page ...)\\
 Here we present the actual data where the blue lines in the graphs correspond to the estimated mean of the series.

### 2011-05-10 pkg/paper/Paper/draft.Rnw
corrected typo in appendix
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@81 b679b2f7-6539-45d0-a605-5ddcb976cde9
new draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@84 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-\Keywords{inflation rate, structural break, EMU, Generalized Logistic Distribution, euro}
+\Keywords{inflation rate, structural break, EMU, Generalized Logistic Distribution 
+}
 
@@
 Ever since the beginning of the European Union, the topic of a common currency was a controversial issue.
-Although the Economic and Monetary Union (EMU) is now a fact, the discussion about the economic effects of the Euro is far from being settled.
-The controversial topics range from the question of whether or not the Eurozone is indeed an optimal currency area (OCA) all the way to the very survival of the Euro in light of the budgetary problems of some of its member states.
-The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context, is in the center of an ongoing debate: this regards the question about short-run and steady-state inflation uncertainty dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries, which is the topic of an investigation by \citet{palomba}.
+Although the Economic and Monetary Union (EMU) is now a fact, the discussion about the economic effects of the euro is far from being settled.
+The controversial topics range from the question of whether or not the eurozone is indeed an optimal currency area (OCA) all the way to the very survival of the euro in light of the budgetary problems of some of its member states.
+The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context, is in the center of an ongoing debate: this regards the question about short-run and steady-state inflation uncertainty (in context to inflation expectations) dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries, which is the topic of an investigation by \citet{palomba}.
 \\
+%(... will jesus früher anetua ...) \\
 The way towards the EMU basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
@@
 \\
-%(...too short: need: methodological ideas and main questions..., value added: schiefe, schwere ränder)\\
-% http://home.iitk.ac.in/~kundu/paper152.pdf  (gl-beschreibung in unterlagen), p.4 more peaked in center, but heavier tails
-However, a deeper inquiry into this problem is in need of a method able to capture changes in the inflation dynamics.
-We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates, which focuses on these changes and thereby upon the effect of the EMU on a number of European countries both within and without the Eurozone.
+However, a deeper inquiry into this problem is in need of a method able to capture changes in inflation dynamics.
+We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates and thereby upon the effect of the EMU on a number of European countries both within and without the eurozone.
 The framework developed in this paper makes it possible to test for any significant changes of a countries inflation dynamics and thus helps interpreting the effects of the EMU.
-We focus on changes in the first three moments of these inflation rates, which we assume to be distributed according to a generalized logistic distribution (GL), as defined in Section \ref{s3}.
+We focus on changes in the first three moments of these inflation rates, which we assume to be distributed according to a generalized logistic distribution (GL).
 The GL distribution is a suitable one due to its flexibility to capture nonsymmetry and because of its fatter tails.
-For this purpose, we developed a new estimation approach building on previous research by \citet{z07}.
+For this purpose, we develop a new estimation approach building on previous research by \citet{z07}.
 \\
@@
 am Schluss so enden, dass nun interessant, wie sich die Momente so genau verändert haben; auf Basis der theorie--> suche nach brüchen, schiefe noch hinein/besser hinein; 3 saha simultan aschoa...)\\
-This framework allows us to observe the evolution of mean, variance and skewness over time. 
+This framework allows us to assess the evolution of mean, variance and skewness over time.
+The skewness parameter turns out to be quite important to measure changes in the long run inflation dynamics, as it changes considerably over the time of the analysis.
+This trend would be all but invisible using a normal distribution.
 %Considering the prior knowledge available about the timing of a countries participation in any of the aforementioned steps towards monetary union, an exact timing of breakpoints in the moments of the respective inflation series enables us to draw conclusions about the exact effects of the EMU the inflation rates of its members.
@@
 This than can be used to search for answers regarding the question about the effect of the EMU on the inflation dynamics of its member states or more general to investigate the effects of external shocks.\\
+%(... jesus: maybe comment on results here ...) \\
 The remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.
@@
 \section{Data}\label{s2}
-
+%(... first method, than data...) \\
 We employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}.
@@
 % greece: -for winter discounts: First 10 days of January-end of February; -for summer discounts: First 10 days of July -end od August; but has always been included
-In the case of Belgium and Luxembourg, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium and January 1999 for Luxembourg) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months} and subtracting them from the original time series.
+As shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Belgium, Italy, Luxembourg and Spain); failing to take this into account may lead to considerable distortions in the estimation.
+In these cases, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium, January 2001 for Italy, January 1999 for Luxembourg and January 2001 for Spain) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months (February and August).} and subtracting them from the original time series.
 
@@
 This then enables testing for the stability of the parameter estimates.
-If there is evidence of significant changes, the number and the timing of the break points are computed.
+If there is evidence of significant changes, the number and the timing of the break points are estimated.
 \\
@@
 %(...discuss correlation...) \\
-The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative testes, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.}
+The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative tests, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.}
 %To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.
@@
 This is due to  its fatter tails, which fit these kind of data somewhat better.
-\citet{won} employ a GL distribution in a regression model with autocorrelated errors.
-They use this type of distribution rather than a Student's t-distribution, as to model the fact that these are oftentimes  non symmetric and severely left or right skewed.
+\citet{won} develop a maximum likelihood estimator if the underlying distribution is a GL distribution.
 A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they do not use a shape parameter.
@@
 We first estimate the parameters of the GL--distribution by means of maximum likelihood $\hat{\phi} ~=~ \underset{\phi}{argmax} \sum_{t=1}^n \log f(y_t|\phi)$.
-The parameter stability is assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures of model deviation.\footnote{for details see Appendix \ref{glscore} and \citet{z07}}  The empirical scores process $\mathit{efp}(\cdot)$ captures deviations from a hypothesized zero mean over time and is given by:
+The parameter stability is assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures of model deviation.\footnote{for details see Appendix \ref{glscore} and \citet{z07}}  The empirical fluctuation process $\mathit{efp}(\cdot)$ captures deviations from a hypothesized zero mean over time and is given by:
 
@@
 \end{eqnarray}
-
+%(... jesus: thinks this is incomplete ...) \\
 where $\hat{V}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
@@
 
-After the empirical fluctuation process is computed, it conveys information about departures from the $H_0$ of parameter stability over time.
+After the $\mathit{efp}(\cdot)$ is computed, it conveys information about departures from the $H_0$ of parameter stability over time.
 For a precise quantification of the actual deviation from its hypothesized zero--mean path, the main idea of the test statistics presented here is to aggregate the $\mathit{efp}(\cdot)$ in such a way as to be able to construct a test statistics so that critical values and p values can be derived.
@@
 \subsection{Breakpoint Estimation}
-%(...cite bai and perron 2003 paper here, nachlesen, findet achim super, bei slowenien will er noch konfidenzintervalle dazu...)\\
+
 The dating of structural breaks builds upon the work of \cite{bp03}.
@@
 
-If the test statistics is significant, at least one break is estimated.
+If the test statistics is significant (in this case the supLM test), at least one break is estimated.
 In principle, the optimal segmentation (i.e.
@@
 \clearpage
-\section{An Example}\label{s4}
-%(...more text: 1. motivation, 2.clear methodological steps, 3.interpretation...) \\
+\subsection{An Example}\label{s4}
+
 <<echo=FALSE>>=
@@
 After a period of relative instability following the independence from Yugoslavia, Slovenia was successfull in realigning its economy and introduced a number of reforms leading to a stable growth in the past years.
-The good economic performance made it possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce the Euro in January 2007.
+The good economic performance made it possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce the euro in January 2007.
 The data collected contains the period from January 1996 onwards.
@@
 A glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period.
-The $efp(\cdot)$ starts to significantly deviate as early as 2001 and reaches the peak somewhere between 2003 and 2004.
+The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches the peak somewhere between 2003 and 2004.
 The p--value of the supLM test is \Sexpr{round(slovpval, 5)} and clearly indicates at least one break.
-The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.
-In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate.
-Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the Euro two years later.
+The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.\footnote{For the changes in the parameters, see Table \ref{resulttable}.} In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate.
+Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the euro two years later.
 
@@
 \section{Results}\label{s5}
-
+%(... jesus: show test statistic: no room...) \\
 % das package colortbl (\usepackege[dvips]{colortbl}) ermöglicht z.B. mit
@@
 @
-\caption{Break point and parameter estimates and ERM information}
+\caption{Break point, parameter estimates and ERM information}
 \label{resulttable}
@@
 %(...andere struktur: 1.short introduction, 2. countries in detail (6.1), 3. summary (6.2) ...) \\
+%(... jesus: opening confusing and obscure...)\\
 Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction.
 We formed seven groups of countries according to the observed pattern of break points.
-These seven groups consist of Austria and Belgium, the Eastern European countries, the two big southern European countries, Italy and Spain, Ireland, the countries with no change in the inflation dynamics, the countries with breaks in the early 90s and the member countries of the former DM--zone (excluding Belgium).
+These seven groups consist of Austria and Belgium, the Eastern European countries, the two big southern European countries, Italy and Spain, Ireland, the countries with no change in the inflation dynamics, the countries with breaks in the early 90s and the member countries of the former DM--zone (excluding Austria and Belgium).
 \\
@@
 Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
-The break during the late 90s is a result of these countries efforts to curb in inflation by decreasing money supply and indicates the time when these countries actually overcame the biggest transitionally shocks on their way towards free market economies.
+The break during the late 90s is a result of these countries efforts to curb in inflation by decreasing money supply and indicates the time when these countrie's actually overcame the biggest transitionally shocks on their way towards free market economies.
 Slovakia is very similar to Slovenia, as its inflation dynamics too changed roughly one year prior to its entry into ERM~II.
 \\
+%(... jesus: maybe some reference on MP (what is MP???)...) \\
+%(... thinks this is just data artefact, however even after desaisonalization the same effect...) \\
 Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain.
@@
 Soon after the fixing of the exchange rates, this trend changed significantly, leading to a higher mean and a much higher variance.
-A possible reason for this is that the fiscal responsibility that was required for an entry into the Eurozone was softened after this goal was achieved.
-This hypothesis gains some credibility considering the huge government deficits of these countries after 1999.
+A possible reason for this is that the fiscal responsibility that was required for an entry into the eurozone was softened after this goal was achieved.
 \\
@@
 Ireland can be considered a rather special case. 
-It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.
+%It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.  
 In Ireland, compared to other countries, the deflationary effect of the financial crises was more persistent.
@@
 The break in 2008 can be attributed to the strong contraction triggered by the bursting of the Irish housing bubble and a beginning deflation in concert with the massive costs of the bailout of the countries banking system.
-Surprisingly, this deflationary pressure is not visible in Spain, which was affected by a housing bubble of roughly the same size as Ireland.
+Surprisingly, this deflationary pressure is not visible in Spain, which was affected by a housing bubble of roughly the same size as Ireland, however in Spain housing prices did not adjust so abruptly and in such a degree.
 \\
@@
 
-The disparity of the national inflation rates is still an issue, many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials.
+<<echo=FALSE>>=
+data("HICP", package = "glogis")
+euhi <- HICP[, colnames(HICP) == "EU"]
+euhi <- 100 * diff(log(na.omit(euhi)))
+euhim1 <- mean(window(euhi, start=yearmon(1990), end= yearmon(1998+11/12)))
+euhim2 <- mean(window(euhi, start=yearmon(1998), end=yearmon(2008+7/12)))
+euhim3 <- mean(window(euhi, start=yearmon(2008+8/10), end=yearmon(2010+11/12)))
+euhiv1 <- var(window(euhi, start=yearmon(1990), end= yearmon(1998+11/12)))
+euhiv2 <- var(window(euhi, start=yearmon(1998), end=yearmon(2008+7/12)))
+euhiv3 <- var(window(euhi, start=yearmon(2008+8/10), end=yearmon(2010+11/12)))
+euhis1 <- skewness(window(euhi, start=yearmon(1990), end= yearmon(1998+11/12)))
+euhis2 <- skewness(window(euhi, start=yearmon(1998), end=yearmon(2008+7/12)))
+euhis3 <- skewness(window(euhi, start=yearmon(2008+8/10), end=yearmon(2010+11/12)))
+@
+
+The disparity of the national inflation rates is still an issue.
+Many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials.
 \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
-In a paper examining structural convergence of the inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
+In a paper examining structural convergence of inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
 They find that convergence over time of inflation dynamics was only partly observable.
@@
 One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
-Over the time, there is clear evidence, that the inflation rates converged towards some common -- albeit broad -- mean level.
+Over time, there are hints, that the inflation rates converged towards some common -- albeit broad -- mean level.
+For the inflation series of the whole EMU, as published by the \citet{oecd}, we see an increase in mean from \Sexpr{round(euhim1, 5)} (1990(1)--1998(12)) to \Sexpr{round(euhim2, 5)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 5)} (2008(9)--2010(12)) following the financial crises.
+Whereas variance has increased over time from    \Sexpr{round(euhiv1, 5)} to \Sexpr{round(euhim3, 5)}.
+Regarding skewness, we see a clear change towards a negative skewness from \Sexpr{round(euhis1, 5)} to \Sexpr{round(euhis3, 5)}.
+
+%(... compare this with the emu inflation rate...)\\
 
@@
 
-Regarding the general trend, Table \ref{resulttable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series during the convergence phase (January 1990 until December 1998), the Euro phase (January 1999 until August 2008) and the time after the financial crisis (since September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.} until now), for the countries in the ``core'' of the Eurozone (Austria, Belgium, Finland, France, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal and Spain) and all the others.
+Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series during the convergence phase (January 1990 until December 1998), the euro phase (January 1999 until August 2008) and the time after the financial crisis (since September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.} until now), for the countries in the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
 \\
-Regarding the mean, we see a clear evolution towards a lower mean in both groups, where the core countries show much lower values.
-Almost the opposite is true concerning the variance of the inflation rates of the core countries, that has increased to a much higher value than at the start of the sample, with the upper percentile during the 90s now almost equaling the lower percentile for the time after the financial crisis; the variance for the other countries has declined significantly.
-Concerning skewness, we see a clear trend in the core countries for skewness to become negative, indicating that the bulk of the values now lies right to the mean.
-The same is true -- although to a lesser extent -- for the other countries.
-\\
+%(... jesus would further differentiate between core and southeren core (or PIIGS)...)\\
 
@@
 <<echo=FALSE, results=tex>>=
-euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[9]],table2[[11]],table2[[12]],table2[[13]],table2[[14]],table2[[16]],table2[[19]])
+euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[13]],table2[[14]])
 noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
+piigs <- cbind(table2[[9]], table2[[11]],table2[[12]], table2[[16]],table2[[19]])
 writeLines(unlist(format_momenteuro(momentgen(1,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+@
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(format_momenteuro(momentgen(1,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
 @
@@
 @
-\hline
 \multicolumn{5}{|c|}{Variance} \\
@@
 <<echo=FALSE, results=tex>>=
-euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[9]],table2[[11]],table2[[12]],table2[[13]],table2[[14]],table2[[16]],table2[[19]])
-noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
 writeLines(unlist(format_momenteuro(momentgen(2,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+@
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(format_momenteuro(momentgen(2,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
 @
@@
 <<echo=FALSE, results=tex>>=
-euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[9]],table2[[11]],table2[[12]],table2[[13]],table2[[14]],table2[[16]],table2[[19]])
-noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
 writeLines(unlist(format_momenteuro(momentgen(3,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+@
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(format_momenteuro(momentgen(3,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
 @
@@
 \caption{Development of first three moments}
-\label{resulttable}
+\label{grouptable}
 \end{longtable}
+
+Regarding the mean, we see a clear evolution towards a lower mean in all groups, where the core countries show much lower values at the median.
+Almost the opposite is true concerning the variance of the inflation rates of the core countries, that has increased to a much higher value than at the start of the sample, with the upper percentile during the 90s now even smaller than the lower percentile for the time after the financial crisis.
+The same is true for the PIIGS countries, although the core country median variance is still lower; the variance for the other countries has declined significantly.
+Concerning skewness, we see a clear trend in the core countries for skewness to become negative, indicating that the bulk of the values now lies right to the mean.
+The same is true -- although to a lesser extent -- for the PIIGS and the other countries.
+\\
 
@@
 
-Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
+%(... jesus does not like the commented part very much...)\\
+%Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
 \\
-An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
+%An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
 What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
@@
 \\
-Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
+%Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
 This is due to the fact that two of the most important inflationary channels, the money supply and the degree of openness, had to be arranged according to the criteria necessary for EMU entry.
@@
 %Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
-For the countries in the core of the Eurozone, the most disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
+For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
 A very volatile inflation rate will likely  contribute to greater macroeconomic instability.
@@
 The moments of the GL distribution as given in equation \ref{eqn1} are:
+%Quelle: "C:\Dissertation\Paper-1-Z\Unterlagen\GL-ml-estimation.pdf"
 
@@
 
-where $\gamma'(\cdot)$ and $\gamma''(\cdot)$ are the first and the second derivatives of the GL distribution, respectively. \\
+where $\gamma'(\cdot)$ and $\gamma''(\cdot)$ are the first and the second derivatives of the digamma function, respectively. \\
 The log-likelihood is:
@@
 
-In addition to Belgium and Luxembourg, there are a number of other countries, where the sales periods are regulated by law.
+In addition to Belgium, Italy, Luxembourg and Spain, there are a number of other countries, where the sales periods are regulated by law.
 France has an individual time span regulated by each department.
@@
 In Greece there are also two set sales--periods in winter and summer.
-In Italy the sales periods are fixed each year by the Chambers of Commerce.
-Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September.
+%In Italy the sales periods are fixed each year by the Chambers of Commerce.
+Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September).
 The source for this information is \citet{konseur}.
-According to information by their national statistical offices, the official dates of the inclusion of these sales periods are: January 2000 for Belgium and January 1999 for Luxembourg.
-Italy, Greece and Portugal were not estimated, since -- as in the case of France -- these periods either do not seem to be set for the whole country at once, or information about the treatment of sales periods in the HICP was not provided by the national statistical offices upon  request.
+According to information by their national statistical offices, the official dates of the inclusion of these sales periods are: January 2000 for Belgium and January 1999 for Luxembourg, while Italy and Spain introduced them in January 2001.
+Greece and Portugal were not estimated, since -- as in the case of France -- these periods either do not seem to be set for the whole country at once, or there does not seem to be additional seasonality as a result of sales periods.
 

### 2011-05-10 pkg/paper/Paper/draft.Rnw
small style touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@85 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 colnames(ret) <- c(start[1], start[2], start[3])
-rownames(ret) <- c("lower", "medium", "upper")
+rownames(ret) <- c("lower", "median", "upper")
 return(ret)
@@
   rval[1,2] <- "lower"
-  rval[2,2] <- "medium"
+  rval[2,2] <- "median"
   rval[3,2] <- "upper"
@@
 
-where the appropriate p--values can be computed using:
+where the appropriate $p$~values can be computed using:
 
@@
 The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches the peak somewhere between 2003 and 2004.
-The p--value of the supLM test is \Sexpr{round(slovpval, 5)} and clearly indicates at least one break.
+The $p$~value of the supLM test is \Sexpr{round(slovpval, 5)} and clearly indicates at least one break.
 The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.\footnote{For the changes in the parameters, see Table \ref{resulttable}.} In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate.
@@
 They find that convergence over time of inflation dynamics was only partly observable.
-In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
+In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980--1984, 1984--1993 and 1993--2000) governing the core inflation rate.
 \\
-Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.
-We wold see one black line if all countries would show the same month--on--month increase at any one time.
-The red lines depict the means.} We do see that many of these countries follow a rather similar path.
-One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
-Over time, there are hints, that the inflation rates converged towards some common -- albeit broad -- mean level.
-For the inflation series of the whole EMU, as published by the \citet{oecd}, we see an increase in mean from \Sexpr{round(euhim1, 5)} (1990(1)--1998(12)) to \Sexpr{round(euhim2, 5)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 5)} (2008(9)--2010(12)) following the financial crises.
-Whereas variance has increased over time from    \Sexpr{round(euhiv1, 5)} to \Sexpr{round(euhim3, 5)}.
-Regarding skewness, we see a clear change towards a negative skewness from \Sexpr{round(euhis1, 5)} to \Sexpr{round(euhis3, 5)}.
+Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the
+first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get
+darker. We wold see one black line if all countries would show the same month--on--month increase at any one time. The red lines depict the means.} We do
+see that many of these countries follow a rather similar path. One event clearly visible is the financial crises and the short deflationary phase starting
+somewhere around the autumn of 2007 and lasting a few months. Over time, there are hints, that the inflation rates converged towards some common -- albeit
+broad -- mean level. For the inflation series of the whole EMU, as published by the \citet{oecd}, we see an increase in mean from \Sexpr{round(euhim1, 3)}
+(1990(1)--1998(12)) to \Sexpr{round(euhim2, 3)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 3)} (2008(9)--2010(12)) following the
+financial crises. Whereas variance has increased over time from    \Sexpr{round(euhiv1, 3)} to \Sexpr{round(euhim3, 3)}. Regarding skewness, we see a clear
+change towards a negative skewness from \Sexpr{round(euhis1, 3)} to $\Sexpr{round(euhis3, 3)}$.
 
@@
 
-Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series during the convergence phase (January 1990 until December 1998), the euro phase (January 1999 until August 2008) and the time after the financial crisis (since September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.} until now), for the countries in the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
+Regarding the general trend, Table \ref{grouptable} presents lower, median and upper (25\%, 50\%, 75\%) percentiles of the first three moments of the inflation series during the convergence phase (January 1990 until December 1998), the euro phase (January 1999 until August 2008) and the time after the financial crisis (since September 2008\footnote{After the insolvency of Lehman Brothers in September 2008.} until now), for the countries in the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
 \\
@@
 
-\begin{longtable}{|l|l|l|l|l|}
+\begin{longtable}{|l|l|r|r|r|}
 \hline
@@
 piigs <- cbind(table2[[9]], table2[[11]],table2[[12]], table2[[16]],table2[[19]])
-writeLines(unlist(format_momenteuro(momentgen(1,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+writeLines(unlist(format_momenteuro(momentgen(1,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="Core euro")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(1,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
+writeLines(unlist(format_momenteuro(momentgen(1,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="PIIGS")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(1,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+writeLines(unlist(format_momenteuro(momentgen(1,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="Others")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(2,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+writeLines(unlist(format_momenteuro(momentgen(2,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="Core euro")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(2,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
+writeLines(unlist(format_momenteuro(momentgen(2,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="PIIGS")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(2,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+writeLines(unlist(format_momenteuro(momentgen(2,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="Others")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(3,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+writeLines(unlist(format_momenteuro(momentgen(3,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="Core euro")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(3,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
+writeLines(unlist(format_momenteuro(momentgen(3,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="PIIGS")))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(3,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+writeLines(unlist(format_momenteuro(momentgen(3,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=3, name="Others")))
 @
@@
 However, the financial crisis clearly contributed to widening of the convergence gap.
-Our finding contradicts \citet{busetti} who find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
+Our finding contradicts \citet{busetti} who find that from 1980--1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
 This does not quite go along with the depicted means,  that do not indicate stronger divergence after 1997 but a continuing decrease of inflation differentials until the hikes after 2008.

### 2011-05-12 pkg/paper/Paper/draft.Rnw
new draft with fixmes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@86 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2011-05-10 pkg/paper/Paper/draft.Rnw
@@
 ## additional functions
-momentgen <- function(flag=1,x, start=yearmon(c(1990,1998,2007)), end=yearmon(c(1998+11/12,2007+11/12,2010+11/12))) {
-# 1:mean, 2:variance, 3:skewness
-if(flag==1){
-mean = mat.or.vec(ncol(x),length(start))
-for(jj in 1:length(start)){
-x1 <- window(x, start=start[jj], end=end[jj])
-for(i in 1:ncol(x1)) {
-mean[i,jj] = mean(na.omit(x1[,i]))
+cross_country_moments <- function(countries, date,
+  probs = c(0.25, 0.5, 0.75))
+{
+  rval <- sapply(countries, function(country) {
+    nb <- results[[country]]$nbreaks
+    m <- t(sapply(results[[country]]$model, moments))
+    bd <- if(nb < 1) {
+      end(results[[country]]$data)
+    } else {
+      breakdates(results[[country]]$breakpoints, breaks = nb)
     }
-}
-ret1 <- c(quantile(mean[,1], probs=.25), quantile(mean[,1], probs=.50), quantile(mean[,1], probs=.75))
-ret1 <- cbind(ret1)
-ret2 <- c(quantile(mean[,2], probs=.25), quantile(mean[,2], probs=.50), quantile(mean[,2], probs=.75))
-ret2 <- cbind(ret2)
-ret3 <- c(quantile(mean[,3], probs=.25), quantile(mean[,3], probs=.50), quantile(mean[,3], probs=.75))
-ret3 <- cbind(ret3)
-ret <- cbind(ret1,ret2,ret3)
-colnames(ret) <- c(start[1], start[2], start[3])
-rownames(ret) <- c("lower", "medium", "upper")
-return(ret)
-}
-if(flag==2){
-var = mat.or.vec(ncol(x),length(start))
-for(jj in 1:length(start)){
-x1 <- window(x, start=start[jj], end=end[jj])
-for(i in 1:ncol(x1)) {
-var[i,jj] = var(na.omit(x1[,i]))
-}
-}
-ret1 <- c(quantile(var[,1], probs=.25), quantile(var[,1], probs=.50), quantile(var[,1], probs=.75))
-ret1 <- cbind(ret1)
-ret2 <- c(quantile(var[,2], probs=.25), quantile(var[,2], probs=.50), quantile(var[,2], probs=.75))
-ret2 <- cbind(ret2)
-ret3 <- c(quantile(var[,3], probs=.25), quantile(var[,3], probs=.50), quantile(var[,3], probs=.75))
-ret3 <- cbind(ret3)
-ret <- cbind(ret1,ret2,ret3)
-colnames(ret) <- c(start[1], start[2], start[3])
-rownames(ret) <- c("lower", "medium", "upper")
-return(ret)
-}
-if(flag==3){
-skewness= mat.or.vec(ncol(x),length(start))
-for(jj in 1:length(start)){
-x1 <- window(x, start=start[jj], end=end[jj])
-for(i in 1:ncol(x1)) {
-skewness[i,jj] = skewness(na.omit(x1[,i]))
-}
-}
-ret1 <- c(quantile(skewness[,1], probs=.25), quantile(skewness[,1], probs=.50), quantile(skewness[,1], probs=.75))
-ret1 <- cbind(ret1)
-ret2 <- c(quantile(skewness[,2], probs=.25), quantile(skewness[,2], probs=.50), quantile(skewness[,2], probs=.75))
-ret2 <- cbind(ret2)
-ret3 <- c(quantile(skewness[,3], probs=.25), quantile(skewness[,3], probs=.50), quantile(skewness[,3], probs=.75))
-ret3 <- cbind(ret3)
-ret <- cbind(ret1,ret2,ret3)
-colnames(ret) <- c(start[1], start[2], start[3])
-rownames(ret) <- c("lower", "medium", "upper")
-return(ret)
-}
+    m[sum(as.yearmon(date) > bd) + 1, ]
+  })
+  apply(rval, 1, quantile, probs = probs)
 }
@@
 
-format_momenteuro <- function(momeuro, digits = 6, name="Euro") {
-  rval <- format(round(momeuro, digits = digits), nsmall = digits)
+format_cross_country <- function(crossmom, digits = 6, name="Euro") {
+  rval <- format(round(crossmom, digits = digits), nsmall = digits)
   rval <- cbind(name,"", matrix(paste("$", rval, "$", sep = ""), ncol = 3))
@@
 }
+
 get_data <- function(i) {
@@
 am Schluss so enden, dass nun interessant, wie sich die Momente so genau verändert haben; auf Basis der theorie--> suche nach brüchen, schiefe noch hinein/besser hinein; 3 saha simultan aschoa...)\\
-This framework allows us to assess the evolution of mean, variance and skewness over time.
+This \fixme{Jesus: sell method better, why this one and not another} framework allows us to assess the evolution of mean, variance and skewness over time.
 The skewness parameter turns out to be quite important to measure changes in the long run inflation dynamics, as it changes considerably over the time of the analysis.
@@
 %(... jesus: maybe comment on results here ...) \\
-The remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.
+The \fixme{Jesus: maybe comment on results here} remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.
 
@@
 %(... first method, than data...) \\
-We employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}.
+We \fixme{Jesus: yearly inflation rates better; don't start with data in paper} employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}.
 Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom.
@@
 % greece: -for winter discounts: First 10 days of January-end of February; -for summer discounts: First 10 days of July -end od August; but has always been included
-As shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Belgium, Italy, Luxembourg and Spain); failing to take this into account may lead to considerable distortions in the estimation.
+As \readme{Jesus: Italy and Spain are problematic, fixed with same method as Belgium and Luxembourg} shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Belgium, Italy, Luxembourg and Spain); failing to take this into account may lead to considerable distortions in the estimation.
 In these cases, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium, January 2001 for Italy, January 1999 for Luxembourg and January 2001 for Spain) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months (February and August).} and subtracting them from the original time series.
@@
 %(...discuss correlation...) \\
-The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative tests, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.}
+The correlation \fixme{Jesus: yes, but a DGP needs to be assumed, it would need to take care of any autororrelation, wouldn't it?} dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative tests, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.}
 %To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.
@@
 This is due to  its fatter tails, which fit these kind of data somewhat better.
-\citet{won} develop a maximum likelihood estimator if the underlying distribution is a GL distribution.
+\citet{won} develop a maximum likelihood \fixme{Jesus: is this not better than what you do here: answer: no, they use completely other framework} estimator if the underlying distribution is a GL distribution.
 A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they do not use a shape parameter.
@@
 %(... jesus: thinks this is incomplete ...) \\
-where $\hat{V}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
+\fixme{Jesus: this seems a bit incomplete} where $\hat{V}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
 Then a functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian bridge:
@@
 @
+
 To illustrate our method, we consider the case of Slovenia.
@@
 \\
-A glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period.
+A \fixme{Jesus: confidence interval for break} glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period.
 The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches the peak somewhere between 2003 and 2004.
@@
 Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the euro two years later.
+\readme{Jesus: include information on how the parameters change: I did give a footnote here}
 
@@
 %(... jesus: opening confusing and obscure...)\\
-Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction.
+\readme{Jesus: include p-values in table; no place left; Jesus: opening seems a bit confusing and obscure} Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction.
 We formed seven groups of countries according to the observed pattern of break points.
@@
 %(... thinks this is just data artefact, however even after desaisonalization the same effect...) \\
-Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain.
+\fixme{Jesus: this is just a data construct: but after desaisonalization still almost the same} Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain.
 Both of them experienced high rates of inflation in the early 90s.
@@
 Over time, there are hints, that the inflation rates converged towards some common -- albeit broad -- mean level.
-For the inflation series of the whole EMU, as published by the \citet{oecd}, we see an increase in mean from \Sexpr{round(euhim1, 5)} (1990(1)--1998(12)) to \Sexpr{round(euhim2, 5)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 5)} (2008(9)--2010(12)) following the financial crises.
+For the inflation series of the whole EMU, as published by the \citet{oecd}, we see \readme{Jesus: why not do this for EMU inflation (i.e.
+the one time series of the OECD} an increase in mean from \Sexpr{round(euhim1, 5)} (1990(1)--1998(12)) to \Sexpr{round(euhim2, 5)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 5)} (2008(9)--2010(12)) following the financial crises.
 Whereas variance has increased over time from    \Sexpr{round(euhiv1, 5)} to \Sexpr{round(euhim3, 5)}.
@@
 
-Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series during the convergence phase (January 1990 until December 1998), the euro phase (January 1999 until August 2008) and the time after the financial crisis (since September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.} until now), for the countries in the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
+Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series  at the beginning of the convergence phase (January 1995), the beginning of the financial crisis (in September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.}) and for the last observation in the sample.
+The countries are grouped into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
 \\
-%(... jesus would further differentiate between core and southeren core (or PIIGS)...)\\
 
@@
 \hline
-\multicolumn{5}{|c|}{Mean} \\
+\multicolumn{5}{|c|}{1995} \\
 \hline
-Zone & Percentile & 1990(1)--1998(12) & 1999(1)--2008(8) & 2008(9)--2010(12)\\
+Zone & Percentile & Mean & Variance & Skewness\\
 \hline
 <<echo=FALSE, results=tex>>=
-euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[13]],table2[[14]])
-noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
-piigs <- cbind(table2[[9]], table2[[11]],table2[[12]], table2[[16]],table2[[19]])
-writeLines(unlist(format_momenteuro(momentgen(1,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+euro <- c(1, 2, 6, 7, 8, 13, 14)
+noneuro <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
+piigs <- c(9, 11, 12, 16, 19)
+writeLines(format_cross_country(cross_country_moments(euro, 1995), name ="Euro"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(1,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
+writeLines(format_cross_country(cross_country_moments(piigs, 1995), name ="PIIGS"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(1,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+writeLines(format_cross_country(cross_country_moments(noneuro, 1995), name ="Noneuro"))
 @
-\multicolumn{5}{|c|}{Variance} \\
 \hline
-Zone & Quantil & 1990(1)--1998(12) & 1999(1)--2007(12) & 2008(1)--2010(12)\\
+\multicolumn{5}{|c|}{2008(9)} \\
 \hline
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(2,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+writeLines(format_cross_country(cross_country_moments(euro, 2008+8/12), name ="Euro"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(2,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
+writeLines(format_cross_country(cross_country_moments(piigs, 2008+8/12), name ="PIIGS"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(2,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+writeLines(format_cross_country(cross_country_moments(noneuro, 2008+8/12), name ="Noneuro"))
 @
 \hline
-\multicolumn{5}{|c|}{Skewness} \\
-\hline
-Zone & Quantil & 1990(1)--1998(12) & 1999(1)--2007(12) & 2008(1)--2010(12)\\
+\multicolumn{5}{|c|}{2010(12)} \\
 \hline
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(3,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
+writeLines(format_cross_country(cross_country_moments(euro, 2010+11/12), name ="Euro"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(3,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
+writeLines(format_cross_country(cross_country_moments(piigs, 2010+11/12), name ="PIIGS"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(format_momenteuro(momentgen(3,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
+writeLines(format_cross_country(cross_country_moments(noneuro, 2010+11/12), name ="Noneuro"))
 @
@@
 
-Regarding the mean, we see a clear evolution towards a lower mean in all groups, where the core countries show much lower values at the median.
-Almost the opposite is true concerning the variance of the inflation rates of the core countries, that has increased to a much higher value than at the start of the sample, with the upper percentile during the 90s now even smaller than the lower percentile for the time after the financial crisis.
-The same is true for the PIIGS countries, although the core country median variance is still lower; the variance for the other countries has declined significantly.
-Concerning skewness, we see a clear trend in the core countries for skewness to become negative, indicating that the bulk of the values now lies right to the mean.
-The same is true -- although to a lesser extent -- for the PIIGS and the other countries.
-\\
+Regarding the mean, there is a slight increase in median inflation in the euro countries, with a bigger increase in the PIIGS countries, whereas the other countries witness a strong decrease in inflation (as a result of their efforts to reach the qualifications required for eurozone entry).
+The variance for the euro countries almost triples over the past 15 years, whereas the variance of the PIIGS countries decreases slightly.
+Regarding skewness, in the euro countries and the PIIGS, there is a tendency of decreasing skewness, indicating a shift of the distribution to the right, whereas the  countries outside the euro do not show much change.\readme{this way, there is a completely different story, not the big changes that we get if we take whole subperiods in the data.}
 
@@
 %(... jesus does not like the commented part very much...)\\
-%Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
+Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
 \\
-%An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
+\fixme{Jesus: this belongs to the intro section; answer: but where??} An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
 What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
@@
 \\
-%Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
+\fixme{Jesus: maybe, but you don't show that in your analysis} Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
 This is due to the fact that two of the most important inflationary channels, the money supply and the degree of openness, had to be arranged according to the criteria necessary for EMU entry.
@@
 \\
-%Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
+Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
 For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.

### 2011-05-13 pkg/R/breakpoints.glogisfit.R
Added confidence intervals for breakpoint estimates in glogisfit.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@88 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@ -41,3 +41,113 @@ fitted.breakpoints.glogisfit <- function(object,
   return(rval)
 }
 
+index.breakpoints.glogisfit <- function(x, ...) x$index
+
+confint.breakpoints.glogisfit <- function(object, parm = NULL, level = 0.95, breaks = NULL,  meat. = NULL, ...)
+{
+  ## parameters: level, breaks
+  a2 <- (1 - level)/2
+  if(!is.null(parm) & !is.null(breaks))
+    warning("`parm' and `breaks' are both specified: `breaks' is used")
+  else
+    if(!is.null(parm)) breaks <- parm
+
+  ## extract estimates
+  bp <- breakpoints(object, breaks = breaks)$breakpoints
+  if(any(is.na(bp))) stop("cannot compute confidence interval when `breaks = 0'")
+  
+  ## set up intervals
+  nbp <- length(bp)
+  upper <- rep(0, nbp)
+  lower <- rep(0, nbp)
+  bp <- c(0, bp, object$nobs)
+
+  ## fitted models
+  rf <- refit(object, breaks = breaks)
+  cf <- lapply(rf, coef)
+  Q <- lapply(rf, function(x) solve(bread(x)))
+  Omega <- if(is.null(meat.)) Q else lapply(rf, meat.)
+
+  ## utility functions
+  myfun <- function(x, level = 0.975, xi = 1, phi1 = 1, phi2 = 1)
+    (pargmaxV(x, xi = xi, phi1 = phi1, phi2 = phi2) - level)
+  myprod <- function(delta, mat) as.vector(crossprod(delta, mat) %*% delta)
+
+  ## loop over breaks
+  for(i in 1:nbp)
+  {
+    delta <- cf[[i+1]] - cf[[i]]            
+    Oprod1 <- myprod(delta, Omega[[i]])
+    Oprod2 <- myprod(delta, Omega[[i+1]])
+    Qprod1 <- myprod(delta, Q[[i]])
+    Qprod2 <- myprod(delta, Q[[i+1]])
+
+    xi <- Qprod2/Qprod1
+    phi1 <- sqrt(Oprod1/Qprod1)
+    phi2 <- sqrt(Oprod2/Qprod2)
+ 
+    p0 <- pargmaxV(0, phi1 = phi1, phi2 = phi2, xi = xi)
+    if(is.nan(p0) || p0 < a2 || p0 > (1-a2)) {
+      warning(paste("Confidence interval", as.integer(i),
[…]

### 2011-05-15 pkg/paper/Paper/draft.Rnw
updated greyplot graphics
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@92 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-## include glmean dataset
 if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
@@
   }
-  ## returns
-  hicp <- diff(log(HICP[,]))
+
+  hicps <- hicps[, colnames(hicps) != "EU"]
+
   ## rolling GL means
-
-  hicp_glmean <- rollapply(hicp, 12, glmean)
-
-  hicp_euro <- diff(log(HICP[, c(1,2,6,7,8,9,11,12,13,14,16,19)]))
-	mean_euro <- hicp_glmean[, c(1,2,6,7,8,9,11,12,13,14,16,19)]
-
-	diff <- rep(1,length(mean_euro[,1]))
-	n <- length(mean_euro[,1])
-
-	for (i in 1:n) {
-	bla <- subset(mean_euro[i,], select=colMeans(is.na(mean_euro[i,])) == 0) 
-	diff[i] = max(bla) - min(bla)
-	}
-	diff <- ts(diff , start = c(1990,2), frequency =12)
+  hicp_glmean <- rollapply(hicps, 12, glmean)
 
   setwd(file.path(root_dir, "Analysis"))
-	save(hicp, hicp_glmean, diff, file = "glmean.rda")
+	save(hicps, hicp_glmean, file = "glmean.rda")
 	setwd(file.path(root_dir, "Paper"))
@@
 %(... first method, than data...) \\
+
+% wichtig: http://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=122.ICP.M.U2.N.000000.4.ANR
+% sieht man die daten von sales prices: 1999-1 portugal,  2000-1 lux, 2001-1 bel, 2002 ita-spa (allerdings in anderem paper ein jahr früher) hm
+
 We \fixme{Jesus: yearly inflation rates better; don't start with data in paper} employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}.
@@
 In these cases, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium, January 2001 for Italy, January 1999 for Luxembourg and January 2001 for Spain) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months (February and August).} and subtracting them from the original time series.
+\fixme{update for using of X-12-ARIMA method, argue with seasonal effects in \cite{pltm}: In Germany and Portugal, sales prices are included from 1998 onwards, in Ireland and Luxembourg from 1999 onwards and in Belgium from 2000.
+In January 2001, Italy and Spain were the last two EU15 countries to include sales into their HICP statistics (The official date may be 12 months later.
+However, the indices of the previous 12 months had been revised in order to avoid distortions in the inflation rates.)}
 
@@
 \\
-Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.
+Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation (with a window size of 12 months), where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.
 We wold see one black line if all countries would show the same month--on--month increase at any one time.
@@
 
-%(... compare this with the emu inflation rate...)\\
+% if(!file.exists(file.path(root_dir, "Paper", "draft-greyplot.pdf")))
+
+%\begin{figure}[h!tp]
+%<<greyplot, fig=TRUE, height=8, width=8>>=
+%
+% ## compute GL mean (if no NA)
+%  glmean <- function(x) {
+%    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
+%  }
+%  hicps <- hicps[, colnames(hicps) != "EU"]
+%
+%  ## rolling GL means
+%  hicp_glmean <- rollapply(hicps, 12, glmean)
+%
+%  ## semi-transparent gray and red
+%  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+%
+%  ## plot returns (gray) and rolling means (red)
+%  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
+%  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+%@
+%\caption{Inflation rates and estimated GL mean over time}
+%\label{greyplot}
+%\end{figure}
 
@@
   ## plot returns (gray) and rolling means (red)
-  plot(hicp, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
-  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+  hicps <- hicps[, colnames(hicps) != "EU"]
+  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
+  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
 @
@@
 
-Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series  at the beginning of the convergence phase (January 1995), the beginning of the financial crisis (in September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.}) and for the last observation in the sample.
+Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series  at the beginning of the convergence phase (January 1995), in December 2007 as a point in time before the financial crisis  and for the last observation in the sample.
 The countries are grouped into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
@@
 \hline
-\multicolumn{5}{|c|}{2008(9)} \\
+\multicolumn{5}{|c|}{2007(12)} \\
 \hline
 <<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(euro, 2008+8/12), name ="Euro"))
+writeLines(format_cross_country(cross_country_moments(euro, 2007+11/12), name ="Euro"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(piigs, 2008+8/12), name ="PIIGS"))
+writeLines(format_cross_country(cross_country_moments(piigs, 2007+11/12), name ="PIIGS"))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(noneuro, 2008+8/12), name ="Noneuro"))
+writeLines(format_cross_country(cross_country_moments(noneuro, 2007+11/12), name ="Noneuro"))
 @

### 2011-05-16 pkg/R/glogisfit.R
Improvements to glogis package:   - added seasonally adjusted hicps data   - modified visualization code
Reflected changes also in results1s.Rnw.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@93 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@ -125,6 +125,7 @@ glogisfit.default <- function(x, weights = NULL,
     loglik = -opt$value,
     df = length(cf),
     n = length(x),
+    nobs = sum(w > 0),
     weights = if(isTRUE(all.equal(as.vector(w), rep.int(1L, length(x))))) NULL else w,
     optim = opt,
     method = method,
@@ -158,7 +159,7 @@ vcov.glogisfit <- function(object, log = TRUE, ...) {
   }
   return(vc)
 }
-logLik.glogisfit <- function(object, ...) structure(object$loglik, df = object$df, nobs = object$n, class = "logLik")
+logLik.glogisfit <- function(object, ...) structure(object$loglik, df = object$df, nobs = object$nobs, class = "logLik")
 estfun.glogisfit <- function(x, ...) {
   rval <- t(t(sglogis(x$x, x$parameters[1], x$parameters[2], x$parameters[3])) * c(1, x$parameters[2], x$parameters[3]))[, is.na(x$fixed)]
   colnames(rval) <- names(coef(x))
@@ -167,7 +168,7 @@ estfun.glogisfit <- function(x, ...) {
   return(rval)
 }
 bread.glogisfit <- function(x, log = TRUE, ...) {
-  vcov(x, log = log) * x$n
+  vcov(x, log = log) * x$nobs
 }
 residuals.glogisfit <- function(object, ...) object$x - object$moments[1]
 
@@ -247,39 +248,50 @@ print.summary.glogisfit <- function(x, digits = max(3, getOption("digits") - 3),
 
 ## visualization
 plot.glogisfit <- function(x, main = "", xlab = NULL, fill = "lightgray",
-  col = "blue", lwd = 1, lty = 1, ylim = NULL, legend = "topright", ...)
+  col = "blue", lwd = 1, lty = 1, xlim = NULL, ylim = NULL, legend = "topright", moments = FALSE, ...)
 {
   if(is.null(x$x)) stop("Data not stored in 'glogisfit' object.")
   if(is.null(ylim)) {
     aux1 <- seq(min(x$x) - 3 * x$parameters[2], max(x$x) + 3 * x$parameters[2], length = 100)
-    aux2 <- hist(x$x, plot = FALSE)$density
+    aux2 <- hist(x$x, plot = FALSE, ...)$density
     ylim <- range(c(dglogis(aux1, x$parameters[1], x$parameters[2], x$parameters[3]), aux2))
[…]

### 2011-05-19 pkg/R/glogisfit.R
changed just one word in hope to get r-forge to recompile new version
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@94 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@ -73,7 +73,7 @@ glogisfit.default <- function(x, weights = NULL,
     -colSums(w * t(t(sglogis(x, location = p1, scale = exp(p2), shape = exp(p3))) * c(1, exp(p2), exp(p3)))[, is.na(fixed)])
   }
 
-  ## choose starting values
+  ## choose the starting values
   if(is.null(start)) start <- c(0, 0, 0)[is.na(fixed)]
   
   ## call optim

### 2011-05-19 pkg/paper/Paper/draft.Rnw
updated draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@95 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 ## additional functions
-cross_country_moments <- function(countries, date,
+cross_country_moments <- function(countries, date=yearmon(c(1990,1995,2000,2010)),
   probs = c(0.25, 0.5, 0.75))
 {
+ret <- mat.or.vec(9,length(date)) 
+for(jj in 1:length(date)) {
   rval <- sapply(countries, function(country) {
@@
     }
-    m[sum(as.yearmon(date) > bd) + 1, ]
+    m[sum(as.yearmon(date[jj]) > bd) + 1, ] 
   })
-  apply(rval, 1, quantile, probs = probs)
+  v1 <- apply(rval, 1, quantile, probs = probs)
+  ret[1:3,jj] <- v1[,1] 
+  ret[4:6,jj] <- v1[,2] 
+  ret[7:9,jj] <- v1[,3] 
+}
+  return(ret)
 }
@@
 ## run results4 (with dummy creation for sales periods)
-Sweave("results4.Rnw")
+Sweave("results1s.Rnw")
 setwd(file.path(root_dir, "Paper"))
@@
 
-format_cross_country <- function(crossmom, digits = 6, name="Euro") {
+format_cross_country <- function(crossmom, digits = 3, moment="mean", quantil="lower") {
   rval <- format(round(crossmom, digits = digits), nsmall = digits)
-  rval <- cbind(name,"", matrix(paste("$", rval, "$", sep = ""), ncol = 3))
-  rval[1,1] <- name
-  rval[2,1] <- ""
-  rval[3,1] <- ""
+  rval <- cbind(moment, quantil, matrix(paste("$", rval, "$", sep=""), ncol = 4))
+  rval[1,1] <- "mean"
+  rval[2:3,1] <- ""
   rval[1,2] <- "lower"
@@
   rval[3,2] <- "upper"
+  rval[4,1] <- "variance"
+  rval[5:6,1] <- ""
+  rval[4,2] <- "lower"
+  rval[5,2] <- "medium"
+  rval[6,2] <- "upper"
+  rval[7,1] <- "skewness"
+  rval[8:9,1] <- ""
+  rval[7,2] <- "lower"
+  rval[8,2] <- "medium"
+  rval[9,2] <- "upper"
   rval <- apply(rval, 1, paste, collapse = " & ")
@@
 \\
-%(... will jesus früher anetua ...) \\
 The way towards the EMU basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
@@
 \\
-%(...
-am Schluss so enden, dass nun interessant, wie sich die Momente so genau verändert haben; auf Basis der theorie--> suche nach brüchen, schiefe noch hinein/besser hinein; 3 saha simultan aschoa...)\\
 This \fixme{Jesus: sell method better, why this one and not another} framework allows us to assess the evolution of mean, variance and skewness over time.
@@
 This trend would be all but invisible using a normal distribution.
-%Considering the prior knowledge available about the timing of a countries participation in any of the aforementioned steps towards monetary union, an exact timing of breakpoints in the moments of the respective inflation series enables us to draw conclusions about the exact effects of the EMU the inflation rates of its members.
 Our estimation technique enables the simultaneous estimation of breaks in the mean (which is important in the context of the Maastricht criteria), the variance (which should be low for the price mechanism to be efficient) and the skewness (where a change in skewness is evidence of a shift of the distribution towards a new ``center of gravity'') of the inflation rates of interest.
 This than can be used to search for answers regarding the question about the effect of the EMU on the inflation dynamics of its member states or more general to investigate the effects of external shocks.\\
-%(... jesus: maybe comment on results here ...) \\
 The \fixme{Jesus: maybe comment on results here} remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.
@@
 \section{Data}\label{s2}
-%(... first method, than data...) \\
-
-% wichtig: http://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=122.ICP.M.U2.N.000000.4.ANR
-% sieht man die daten von sales prices: 1999-1 portugal,  2000-1 lux, 2001-1 bel, 2002 ita-spa (allerdings in anderem paper ein jahr früher) hm
 
@@
 \\
-%(...
-schlussverkäufe schon hier erwähnen, einfach cool, ja das gibts, wir haben die aber weggeschätzt; generell viel angelsächsischer/präpotenter schreiben; nur 1 weg der richtige und zwar unserer; nicht erwähnen, dass andere dinge auch noch möglich; so nach motto: mich interessierts, ich machs, andere erst am ende erwähnen...)
-% greece: -for winter discounts: First 10 days of January-end of February; -for summer discounts: First 10 days of July -end od August; but has always been included
-As \readme{Jesus: Italy and Spain are problematic, fixed with same method as Belgium and Luxembourg} shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Belgium, Italy, Luxembourg and Spain); failing to take this into account may lead to considerable distortions in the estimation.
-In these cases, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium, January 2001 for Italy, January 1999 for Luxembourg and January 2001 for Spain) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months (February and August).} and subtracting them from the original time series.
-\fixme{update for using of X-12-ARIMA method, argue with seasonal effects in \cite{pltm}: In Germany and Portugal, sales prices are included from 1998 onwards, in Ireland and Luxembourg from 1999 onwards and in Belgium from 2000.
-In January 2001, Italy and Spain were the last two EU15 countries to include sales into their HICP statistics (The official date may be 12 months later.
-However, the indices of the previous 12 months had been revised in order to avoid distortions in the inflation rates.)}
+As  shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Germany, Portugal, Ireland, Luxembourg, Belgium, Italy and Spain); failing to take this into account may lead to considerable distortions in the estimation.
+We dealt with the inclusion of sales periods\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} and all the other seasonality in the data by using the X--12--ARIMA method for desaisonalization, as described in \cite{X-12-ARIMA} and \cite{Findley+Monsell+Bell:1998}.
 
@@
 \\
-%(...discuss correlation...) \\
 The correlation \fixme{Jesus: yes, but a DGP needs to be assumed, it would need to take care of any autororrelation, wouldn't it?} dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative tests, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.}
-%To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.
-
-% korrelation besser herausarbeiten
-% mehr: link to, match with known events, like intervention
 
@@
 \end{eqnarray}
-% skewness: if delta = 1.1 --> skweness is 0.1, is positive, skewed to the right: linkssteil
-% ie. negative skew (delta <1: rechtssteil, in englisch: skewed to the left 
-% (psigamma(0.5, deriv = 2) - psigamma(1, deriv = 2))/(psigamma(1, deriv = 1) + psigamma(0.5, deriv = 1))^(3/2)
 
@@
         x <- results$Slovenia
-        slovpval <- sctest(x$efp, functional = supLM(0.1))$p.value
+        spval <- sctest(x$efp, functional = supLM(0.1))$p.value
 @
@@
 \\
-A \fixme{Jesus: confidence interval for break} glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period.
+A glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period.
 The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches the peak somewhere between 2003 and 2004.
-The p--value of the supLM test is \Sexpr{round(slovpval, 5)} and clearly indicates at least one break.
-The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.\footnote{For the changes in the parameters, see Table \ref{resulttable}.} In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate.
+The p--value of the supLM test is \Sexpr{round(spval, 8)} and clearly indicates at least one break.
+The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.
+The goodness of fit test depicted in Figure \ref{figtest} confirms the choice of the GL distribution.
+The mean inflation rate was cut to only a third of its previous value, whereas variance increased slightly and the distribution slightly skewed to the left, but peaking at a lower value than previously.
+\\
+In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate.
 Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the euro two years later.
-\readme{Jesus: include information on how the parameters change: I did give a footnote here}
 
@@
 lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
 plot(x$efp, functional = supLM(0.1), main = "supLM test")
@@
 
+\begin{figure}[h!tp]
+<<fig=TRUE, height=4, width=8>>=
+x <- results$Slovenia
+par(mfrow=c(1,2))
+for(i in 1:2)
+plot(x$model[[i]], moments = TRUE,
+  xlim = c(-1.5, 2), ylim = c(0, 1.3),
+  breaks = seq(-1.5, 2, by = 0.25),
+  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
+    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
+    sep = ""))
+@
+\caption{Slovenia: Goodness of fit test}
+\label{figtest}
+\end{figure}
+
+\clearpage
 \newpage
 \section{Results}\label{s5}
-%(... jesus: show test statistic: no room...) \\
-% das package colortbl (\usepackege[dvips]{colortbl}) ermöglicht z.B. mit
-%\rowcolor[gray]{0.90} in einer Tabellenzeile diese farbig (bzw. hier grau) zu unterlegen.
 
@@
 \hline
-\multicolumn{8}{|c|}{Austria and Belgium} \\
+\multicolumn{8}{|c|}{Financial Crisis} \\
 \hline
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(1, 2), format_moments)))
+writeLines(unlist(lapply(c(11, 19), format_moments)))
 @
@@
 \hline
-\multicolumn{8}{|c|}{Italy and Spain} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(12, 19), format_moments)))
-@
-\hline
-\multicolumn{8}{|c|}{Ireland} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(11), format_moments)))
-@
-\hline
 \multicolumn{8}{|c|}{No change countries} \\
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(9,14,7,6), format_moments)))
+writeLines(unlist(lapply(c(2,4,8,14), format_moments)))
 @
@@
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(16,20, 21), format_moments)))
+writeLines(unlist(lapply(c(20, 21), format_moments)))
 @
 \hline
-\multicolumn{8}{|c|}{Former DM--zone} \\
+\multicolumn{8}{|c|}{Phase 1 convergence} \\
 \hline
 <<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(4,8,13), format_moments)))
+writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
+@
+\hline
+\multicolumn{8}{|c|}{Phase 2 convergence} \\
+\hline
+<<echo=FALSE, results=tex>>=
+writeLines(unlist(lapply(c(9,12,13,16), format_moments)))
 @
@@
 \end{longtable}
-% warning: the groups then change; Italy and Greece have not yet answered
-% Italian Statistics: Your request has been correctly inserted in the system with the reference number 13078/2010.
-% comment: the greek break is NOT sure, since they have not yet answered about sales periods
 
@@
 
-%(...3 punkte darstellung noch einbauen...) \\
-%(...andere struktur: 1.short introduction, 2. countries in detail (6.1), 3. summary (6.2) ...) \\
-%(... jesus: opening confusing and obscure...)\\
 \readme{Jesus: include p-values in table; no place left; Jesus: opening seems a bit confusing and obscure} Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction.
-We formed seven groups of countries according to the observed pattern of break points.
-These seven groups consist of Austria and Belgium, the Eastern European countries, the two big southern European countries, Italy and Spain, Ireland, the countries with no change in the inflation dynamics, the countries with breaks in the early 90s and the member countries of the former DM--zone (excluding Austria and Belgium).
-\\
 
@@
 
-Regarding Austria and Belgium, the time after the break is characterized by an increase in the mean and a severe increase in the variance of the inflation rates.
-The skewness is now negative for both series, which indicates that the distribution has shifted to the right, implying that overall inflation is likely to increase.
-This change can be traced back to two major influences: the financial crisis and potentially some spillover effects of higher oil prices in late 2007.
+The first group that draws particular attention is a subgroup of the PIIGS (Portugal, Ireland, Italy, Greece and Spain) countries consisting of Ireland and Spain.
+Both show a break in the summer of 2008 following the financial crisis and the severe problems in both countries concerning the real estate situation.
+In both countries this had a strong deflationary impact, with Ireland actually showing negative inflation rates for an extended period of time; in Spain, which was affected by a housing bubble of roughly the same size as Ireland, the housing prices did not adjust so abruptly and in such a degree.
+This deflationary trend, which for a time affected all countries, can be seen in Figure \ref{greyplot}, which depicts a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
+There is evidence of a decline in mean inflation, which returns to its previous path roughly a year later.
 \\
@@
 \\
-%(... jesus: maybe some reference on MP (what is MP???)...) \\
-%(... thinks this is just data artefact, however even after desaisonalization the same effect...) \\
-\fixme{Jesus: this is just a data construct: but after desaisonalization still almost the same} Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain.
-Both of them experienced high rates of inflation in the early 90s.
-Afterwards, in the phase leading up to the monetary union, mean and variance were greatly reduced.
-Soon after the fixing of the exchange rates, this trend changed significantly, leading to a higher mean and a much higher variance.
-A possible reason for this is that the fiscal responsibility that was required for an entry into the eurozone was softened after this goal was achieved.
-\\
-%(... in irland der schock schon persistenter als in anderen ländern, bei der grauen graphik zeigen (greyplot...)\\
-Ireland can be considered a rather special case. 
-%It is the only economy where the effects of the ongoing financial crisis of 2008 are visible.  
-In Ireland, compared to other countries, the deflationary effect of the financial crises was more persistent.
-This deflationary trend, which for a time affected all countries, can be seen in Figure \ref{greyplot}, which depicts a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
-There is evidence of a decline in mean inflation, which returns to its previous path roughly a year later.
-The break in 2008 can be attributed to the strong contraction triggered by the bursting of the Irish housing bubble and a beginning deflation in concert with the massive costs of the bailout of the countries banking system.
-Surprisingly, this deflationary pressure is not visible in Spain, which was affected by a housing bubble of roughly the same size as Ireland, however in Spain housing prices did not adjust so abruptly and in such a degree.
-\\
-%(...change formulation for greece...) \\
-\clearpage
-Another interesting group consists of four economically rather different countries.
-In none of them do we find any evidence of a change in inflation dynamics.
-Whereas France, Finland and Netherlands are on the lower side of the inflation margin, the Greek inflation rate is among the highest.
+Another interesting group consists -- with the exception of Austria and Luxembourg -- of the countries of the former DM--zone.
+All these countries have a very low inflation rate in both mean and variance and their inflation dynamics did not change.
 \\
 The next group consists of two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of both countries in the EMS.}, the United Kingdom and Sweden.
-Portugal is included due to the same break point behavior.
-All three show a clear break in the early 90s.
-Two of them can be traced back to economic crisis, the currency crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and the run on the Swedish currency in the same year.
-In the case of Portugal, the inflation rate after 1992 was much more stable, showing that the inflationary problems of Portugal were less severe thereafter.\\
-The last group consists of the former DM--zone countries Denmark, Germany and Luxembourg.
-In the case of Denmark we observe an increase in variance after 2000 with a very distinct change in skewness.
-In Germany, the inflation rate increased both on average and in its variability.
-Luxembourg suffered a severe increase in variance, similar to Italy and Spain.
-All of these countries, which profited from the stability of the previous German currency, still have a very low inflation rate.
+Both show a clear break in the early 90s.
+They can be traced back to economic crisis, the currency crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and the run on the Swedish currency in the same year.
+\\
+Austria, Finland and France adjusted to the requirements of the EMU during the first phase of convergence.
+These are also the countries with the lowest inflation rates in both mean and variance after the adjustment.
+\\
+The last group consists of three PIIGS countries and Luxembourg.
+Greece and Portugal adjusted to the ERM shortly before or afterwards its introduction.
+Italy was successfull in reducing its inflation rate, whereas Luxembourg experienced a doubling of its former -- very low -- inflation rate with the variance rising by a factor of 14 with a strong shift in skewness towards the left.
 
@@
 <<echo=FALSE>>=
-data("HICP", package = "glogis")
-euhi <- HICP[, colnames(HICP) == "EU"]
-euhi <- 100 * diff(log(na.omit(euhi)))
+data("hicps", package = "glogis")
+euhi <- hicps[, colnames(hicps) == "EU"]
+euhi <- na.omit(euhi)
 euhim1 <- mean(window(euhi, start=yearmon(1990), end= yearmon(1998+11/12)))
@@
 
-% if(!file.exists(file.path(root_dir, "Paper", "draft-greyplot.pdf")))
-
-%\begin{figure}[h!tp]
-%<<greyplot, fig=TRUE, height=8, width=8>>=
-%
-% ## compute GL mean (if no NA)
-%  glmean <- function(x) {
-%    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
-%  }
-%  hicps <- hicps[, colnames(hicps) != "EU"]
-%
-%  ## rolling GL means
-%  hicp_glmean <- rollapply(hicps, 12, glmean)
-%
-%  ## semi-transparent gray and red
-%  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-%
-%  ## plot returns (gray) and rolling means (red)
-%  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
-%  for(i in 1:ncol(hicp)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-%@
-%\caption{Inflation rates and estimated GL mean over time}
-%\label{greyplot}
-%\end{figure}
-
 \begin{figure}[h!tp]
@@
 
-Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series  at the beginning of the convergence phase (January 1995), in December 2007 as a point in time before the financial crisis  and for the last observation in the sample.
+Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series in January 1994, 1998, 2007 and 2010.
 The countries are grouped into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
+For this, we took account of possible breaks in the moments.
 \\
 
-\begin{longtable}{|l|l|l|l|l|}
+\begin{longtable}{|l|l|r|r|r|r|}
 \hline
-\multicolumn{5}{|c|}{1995} \\
+\multicolumn{6}{|c|}{Core--euro} \\
 \hline
-Zone & Percentile & Mean & Variance & Skewness\\
+Moment & Quantil & 1994 & 1998 & 2007 & 2010\\
 \hline
@@
 euro <- c(1, 2, 6, 7, 8, 13, 14)
-noneuro <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
+others <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
 piigs <- c(9, 11, 12, 16, 19)
-writeLines(format_cross_country(cross_country_moments(euro, 1995), name ="Euro"))
+writeLines(format_cross_country(cross_country_moments(euro, date=yearmon(c(1994,1998,2007,2010)))))
 @
 \hline
+\multicolumn{6}{|c|}{Others} \\
+\hline
 <<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(piigs, 1995), name ="PIIGS"))
+writeLines(format_cross_country(cross_country_moments(others, date=yearmon(c(1994,1998,2007,2010)))))
 @
 \hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(noneuro, 1995), name ="Noneuro"))
-@
-\hline
-\multicolumn{5}{|c|}{2007(12)} \\
+\multicolumn{6}{|c|}{PIIGS} \\
 \hline
 <<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(euro, 2007+11/12), name ="Euro"))
-@
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(piigs, 2007+11/12), name ="PIIGS"))
-@
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(noneuro, 2007+11/12), name ="Noneuro"))
-@
-\hline
-\multicolumn{5}{|c|}{2010(12)} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(euro, 2010+11/12), name ="Euro"))
-@
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(piigs, 2010+11/12), name ="PIIGS"))
-@
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(noneuro, 2010+11/12), name ="Noneuro"))
+writeLines(format_cross_country(cross_country_moments(piigs, date=yearmon(c(1994,1998,2007,2010)))))
 @
@@
 
-Regarding the mean, there is a slight increase in median inflation in the euro countries, with a bigger increase in the PIIGS countries, whereas the other countries witness a strong decrease in inflation (as a result of their efforts to reach the qualifications required for eurozone entry).
-The variance for the euro countries almost triples over the past 15 years, whereas the variance of the PIIGS countries decreases slightly.
-Regarding skewness, in the euro countries and the PIIGS, there is a tendency of decreasing skewness, indicating a shift of the distribution to the right, whereas the  countries outside the euro do not show much change.\readme{this way, there is a completely different story, not the big changes that we get if we take whole subperiods in the data.}
+In the core--zone, we see a decrease in the lower quartil from 1994 to 2010, with no change in the medium inflation rate, which is due to the fact, that there is no break after 1998 for these countries; however, the interquartile range decreased by almost 20\%.
+The variance increased slightly (but with a 15\% decrease in interquartile range), whereas skewness stayed symmetric in the medium, with almost no change in the interquartile range.
+\\
+The other countries were successfull in decreasing all three moments, but they are still higher than the euro--zone countries.
+Most importantly, the interquartile range for all moments decreased considerably showing a clear convergence towards the core--zone countries.
+The PIIGS countries decreased inflation rates as well with a strong change of skewness towards more negative values.
+The interquartile range of their inflation variance increased considerably, which is mostly due to Spain and Ireland and their strong reaction to the financial crisis.
 
@@
 \section{Conclusion}\label{s6}
-
-%(... jesus does not like the commented part very much...)\\
-Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
+%
+%Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
 \\
-\fixme{Jesus: this belongs to the intro section; answer: but where??} An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
+\fixme{Jesus: this belongs to the intro section} An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
 What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
@@
 \\
-\fixme{Jesus: maybe, but you don't show that in your analysis} Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
+%\fixme{Jesus: maybe, but you don't show that in your analysis} Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
 This is due to the fact that two of the most important inflationary channels, the money supply and the degree of openness, had to be arranged according to the criteria necessary for EMU entry.
@@
 Although \cite{grier} find no evidence that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --  \cite{jaro} finds a positive correlation between the level and the standard deviation of inflation.
-This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro--zone countries indicating that in future higher inflation rates are more likely.
+This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro--zone countries, and especially the PIIGS, indicating that in future higher inflation rates are more likely.
 This will put pressure on the ECB to achieve its inflation target of 2\%.
-
-%\begin{figure}[h!tp]
-%<<diffmean, fig=TRUE, height=8, width=8>>=
-%	plot(diff)
-%@
-%\caption{Differences in Euro excluding Estonia, Slovenia and Slovakia}
-%\end{figure}
-%
-%Concerning the mean of the inflation rates, we see a clear convergence pattern.
-This result is reinforced by plotting the difference between the highest and the lowest mean inflation rate in the euro subsample, excluding the new Euro--member states Estonia, Slovakia and Slovenia.\footnote{Including them blurs the clear result somewhat by driving the difference up, since at the beginning, these countries still had very high inflation rates.} After a short hike at the beginning of the convergence phase in 1995, the difference declined clearly to a fraction of its previous value.
-However, the financial crisis clearly contributed to widening of the convergence gap.
-Our finding contradicts \citet{busetti} who find that from 1980---1997 there was convergence of inflation rates, but afterwards there is some diverging behavior.
-This does not quite go along with the depicted means,  that do not indicate stronger divergence after 1997 but a continuing decrease of inflation differentials until the hikes after 2008.
-\\
-%However, this is not the case for variance, where we find clear evidence of an increase.
-Doing the same thing for volatility, we see the second moment almost tripling after 1995.\footnote{Due to a very low volatility of countries like Spain and Luxembourg at the beginning.} This of course does make some economic sense, since the convergence criteria and the ECB focus on mean inflation only and do not take volatility into account.
-\\
 
@@
 
-The results shown are computed with a new \proglang{R} package \pkg{glogis}, which is an enhancement to the already existing \proglang{R} package \pkg{strucchange}, which currently does not support a GL distribution.
+The results shown are computed with a new \readme{add reference} \proglang{R} package \pkg{glogis}, which is an enhancement to the already existing \proglang{R} package \readme{add reference} \pkg{strucchange}, which currently does not support a GL distribution.
 The tests and the graphical illustration of the empirical fluctuation process were developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
 \\
-
-% appendix HINTER reference
 \begin{appendix}\label{app}
@@
 The moments of the GL distribution as given in equation \ref{eqn1} are:
-%Quelle: "C:\Dissertation\Paper-1-Z\Unterlagen\GL-ml-estimation.pdf"
 
@@
 
-% scores given in dp02_15 p.14
-% bemerkungen zu den herleitungen: siehe herleitung-scores-gen-logistic-neu; wobei vor allem das "`-"' beim term der scale ableitung in mathematica mit seltsamerweise mit "`+"' berechnet wird; beim händischen nachrechnen stellt sich allerdings das - als richtig heraus
-
 \begin{eqnarray}
@@
 
-In addition to Belgium, Italy, Luxembourg and Spain, there are a number of other countries, where the sales periods are regulated by law.
+In addition to Germany, Portugal, Belgium, Italy, Luxembourg and Spain, there are a number of other countries, where the sales periods are regulated by law.
 France has an individual time span regulated by each department.
@@
 In Greece there are also two set sales--periods in winter and summer.
-%In Italy the sales periods are fixed each year by the Chambers of Commerce.
+In Italy the sales periods are fixed each year by the Chambers of Commerce.
 Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September).
 The source for this information is \citet{konseur}.
-According to information by their national statistical offices, the official dates of the inclusion of these sales periods are: January 2000 for Belgium and January 1999 for Luxembourg, while Italy and Spain introduced them in January 2001.
-Greece and Portugal were not estimated, since -- as in the case of France -- these periods either do not seem to be set for the whole country at once, or there does not seem to be additional seasonality as a result of sales periods.
+According to \cite{pltm}, the official dates of the inclusion of these sales periods are: January 1998 for Germany and Portugal, January 2000 for Belgium and January 1999 for Luxembourg and Ireland, while Italy and Spain introduced them in January 2001.
+\\
+Since X--12--ARIMA is not implemented in \proglang{R}, we used the gretl software, as described in \cite{Smith+Mixon:2006}, to desaisonalize the data without correction for outliers.
+For the results, we used the seasonally adjusted series provided by gretl using X--12--ARIMA
 

### 2011-05-23 pkg/paper/Paper/draft.Rnw
Tried to finalize data analysis. Lots of style improvements.
Data analysis: R code simplified. Code chunks moved to the relevant
sections throughout the paper (and not clustered at beginning). The
only remaining major point is how to present the results for the EU
series. (Previoulsy, this used slightly wrong data windows and only
empirical moments as opposed to model-based estimates.)

Style:
Enhanced graphics by improving aspect ratio, size (as included in the
paper), and detailed captions.

Tables have been more unified using less lines (and omitting all vertical
lines as many journals don'

@@
 \documentclass[nojss]{jss}
-\usepackage[latin1]{inputenc}
-\usepackage{longtable}
-\usepackage{lscape}
-\usepackage{multirow}
 \usepackage{amsfonts,amstext,amsmath}
-\usepackage{rotating}
-\usepackage{url}
+\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
+
 %% need no \usepackage{Sweave}
-
 \setkeys{Gin}{width=\textwidth}
@@
 
+%% for internal use
+\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
+\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
+
+\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}
+
+\author{Thomas Windberger\\Universit\"at Innsbruck \And
+        Achim Zeileis\\Universit\"at Innsbruck}
+\Plainauthor{Thomas Windberger, Achim Zeileis}
+
+\Abstract{
+  text when everything else finished, present innovation of paper
+}
+
+\Keywords{inflation rate, structural break, EMU, generalized logistic distribution}
+
+\Address{
+  Thomas Windberger, Achim Zeileis\\
+  Department of Statistics\\
+  Universit\"at Innsbruck\\
+  Universit\"atsstra{\ss}e 15\\
+  A-6020 Innsbruck, Austria\\
+  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
+}
+
+\begin{document}
+
 <<preliminaries>>=
+## locale settings
+if(Sys.getlocale() != "C") Sys.setlocale("LC_TIME", "English")
+
 ## packages
-if(Sys.getlocale() != "C") Sys.setlocale("LC_TIME", "English")
 library("glogis")
 library("fxregime")
-library("lattice")
-library("moments")
 
-## additional functions
-cross_country_moments <- function(countries, date=yearmon(c(1990,1995,2000,2010)),
-  probs = c(0.25, 0.5, 0.75))
-{
-ret <- mat.or.vec(9,length(date)) 
-for(jj in 1:length(date)) {
-  rval <- sapply(countries, function(country) {
-    nb <- results[[country]]$nbreaks 
-    m <- t(sapply(results[[country]]$model, moments)) 
-    bd <- if(nb < 1) { 
-      end(results[[country]]$data) 
-    } else {
-      breakdates(results[[country]]$breakpoints, breaks = nb) 
-    }
-    m[sum(as.yearmon(date[jj]) > bd) + 1, ] 
-  })
-  v1 <- apply(rval, 1, quantile, probs = probs)
-  ret[1:3,jj] <- v1[,1] 
-  ret[4:6,jj] <- v1[,2] 
-  ret[7:9,jj] <- v1[,3] 
-}
-  return(ret)
-}
-
+## additional moment method
 moments <- function(object, ...) UseMethod("moments")
@@
   t(sapply(refit(object, breaks = breaks), "[[", "moments"))
+@
 
-## For the creation of the tables
-## <<tabelle, eval=FALSE>>=
+\section{Introduction}
+
+Ever since the beginning of the European Union, the topic of a common currency was a controversial issue.
+Although the Economic and Monetary Union (EMU) is now a fact, the discussion about the economic effects of the euro is far from being settled.
+The controversial topics range from the question of whether or not the eurozone is indeed an optimal currency area (OCA) all the way to the very survival of the euro in light of the budgetary problems of some of its member states.
+The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context, is in the center of an ongoing debate: this regards the question about short-run and steady-state inflation uncertainty (in context to inflation expectations) dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries, which is the topic of an investigation by \citet{palomba}.
+
+The way towards the EMU basically consisted of 3 stages: stage I (1990--1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
+A possible fourth stage is the continuing integration of new member states, mostly former communist countries.
+Any new state that wants to join the EMU is first obliged to fulfill the Maastricht criteria, which, besides setting rules for government debts and interest rates, also requires the participation in the ERM~II (Exchange Rate Mechanism~II) for two years, while the exchange rates towards the euro are not allowed to cross the nominal band.
+
+\cite{emerson} put emphasize on the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
+Failing to stabilize inflation leads to severe economic problems.
+Throughout the literature, there is still a considerable degree of uncertainty as to what extent the introduction of the euro, or monetary unions in general, affect the inflation rate.
+The EMU requires a country to possibly alter a number of economic policies and institutions, which may demand for significant short term adjustments to take place.
+
+Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria are likely to lead to an alteration at least in the mean (since this was one of the requirements) and possibly in the volatility of their respective inflation rates as well.
+
+However, a deeper inquiry into this problem is in need of a method able to capture changes in inflation dynamics.
+We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates and thereby upon the effect of the EMU on a number of European countries both within and without the eurozone.
+The framework developed in this paper makes it possible to test for any significant changes of a countries inflation dynamics and thus helps interpreting the effects of the EMU.
+We focus on changes in the first three moments of these inflation rates, which we assume to be distributed according to a generalized logistic distribution (GL).
+The GL distribution is a suitable one due to its flexibility to capture nonsymmetry and because of its fatter tails.
+For this purpose, we develop a new estimation approach building on previous research by \citet{z07}.
+
+This \fixme{Jesus: sell method better, why this one and not another} framework allows us to assess the evolution of mean, variance and skewness over time.
+The skewness parameter turns out to be quite important to measure changes in the long run inflation dynamics, as it changes considerably over the time of the analysis.
+This trend would be all but invisible using a normal distribution.
+Our estimation technique enables the simultaneous estimation of breaks in the mean (which is important in the context of the Maastricht criteria), the variance (which should be low for the price mechanism to be efficient) and the skewness (where a change in skewness is evidence of a shift of the distribution towards a new ``center of gravity'') of the inflation rates of interest.
+This than can be used to search for answers regarding the question about the effect of the EMU on the inflation dynamics of its member states or more general to investigate the effects of external shocks.
+
+The \fixme{Jesus: maybe comment on results here} remainder of this paper is structured as follows: Section~\ref{sec:data} presents the data, Section~\ref{sec:model} presents the model and the estimation techniques used, Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an example, Section~\ref{sec:results} presents the results and Section~\ref{sec:conclusion} concludes.
+
+\clearpage
+\section{Data}\label{sec:data}
+
+<<data>>=
+data("hicps", package = "glogis")
+eu <- na.omit(hicps[, "EU"])
+hicps <- hicps[, colnames(hicps) != "EU"]
+@
+
+We \fixme{Jesus: yearly inflation rates better; don't start with data in paper}
+employ seasonally unadjusted HICP series of 21 countries ranging from January
+1990 to December 2010 that are obtained from the \citet{oecd}. Countries
+included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,
+Finland,	France,	Germany,	Greece,	Hungary,	Ireland,
+Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia,
+Slovenia,  Spain,	Sweden, and	United Kingdom. Latvia and Lithuania as
+well as Bulgaria and Romania are excluded due to data scarcity. Cypria and Malta
+are not included since they are very small economies.
+
+The countries in this sample can be divided into three different groups. First
+the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece,
+Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and
+Spain).  Second, the EU members not participating in the ERM~II: Czech Republic,
+Hungary, Poland, United Kingdom, and Sweden. Denmark stands on its own as a
+member of the EU and the ERM~II, but not yet a member of the EMU.
+
+As  shown in a paper by \cite{pltm}, the harmonization of the treatment of sales
+prices led to considerable changes in some countries (Germany, Portugal,
+Ireland, Luxembourg, Belgium, Italy and Spain); failing to take this into
+account may lead to considerable distortions in the estimation. We dealt with
+the inclusion of sales periods\footnote{This was demanded by a Commission
+regulation, details may be found in: \citet{eurlex} and in the
+Appendix~\ref{sec:sales}.} and all the other seasonality in the data by using
+the X-12-ARIMA method for desaisonalization, as described in \cite{X-12-ARIMA}
+and \cite{Findley+Monsell+Bell:1998}.
+
+\clearpage
+\section{Model}\label{sec:model}
+
+To monitor changes in the dynamics of the inflation rates, we follow the usual
+strategy employed when testing for structural breaks. First, the model is fitted
+for the whole sample using the generalized logistic distribution. This then
+enables testing for the stability of the parameter estimates. If there is
+evidence of significant changes, the number and the timing of the break points
+are estimated.
+
+This framework allows to trace the evolution of mean, variance and skewness of
+the inflation series and thus interpreting deviations in any of these parameters
+as possibly due to a change in monetary policy, is made feasible. Any huge
+intervention or change is captured by fluctuations in the parameter stability of
+the series and can be quantified and related to a distinct time frame.
+
+The correlation \fixme{Jesus: yes, but a DGP needs to be assumed, it would need
+to take care of any autororrelation, wouldn't it?} dynamics of the series is of
+less importance to this framework, which concentrates on changes in mean,
+variance and skewness and does not focus on autocorrelation, the use of
+alternative tests, as described in \cite{andreou}, is also hindered by the low
+number of observations, due to the monthly frequencies.\footnote{In some years
+the Billion Prices Project of the \cite{bpp}, which provides a much higher data
+frequency, could solve this problem.}
+
+\subsection{Generalized logistic distribution}\label{sec:gld}
+
+The data suggested that the use of a normal distribution might cause some
+estimation problems, since most of the time series exhibit asymmetric properties
+(with sometimes rather strong skewness) and -- compared to the normal
+distribution -- a fatter tail. To accommodate for this fact, the use of the
+generalized logistic distribution (GL) with its relative low degree of
+complexity, seemed appropriate. On the one hand, it is a simple distribution and
+on the other hand, it is flexible enough to capture essential properties of the
+data. For the inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we
+therefore assume a GL distribution as defined in \citet{johnson}:
+
+\begin{eqnarray}
+f(y | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma} \cdot \exp^{-\frac{y-\theta}{\sigma}}}{(1+\exp^{-\frac{y-\theta}{\sigma}})^{(\delta+1)}} 
+\label{eq:dglogis}
+\end{eqnarray}
+
+with location $\theta$, scale $\sigma$ and shape $\delta$. For $\delta=1$ the
+distribution simplifies to the logistic distribution, for $\delta<1$ it is
+skewed to the left and for $\delta$$>$1 it is skewed to the right (see
+Appendix~\ref{sec:gld2} for more details).
+
+In econometrics, the logistic distribution is often used in income distributions
+and growth models. This is due to  its fatter tails, which fit these kind of
+data somewhat better. \citet{won} develop a maximum likelihood \fixme{Jesus: is
+this not better than what you do here: answer: no, they use completely other
+framework} estimator if the underlying distribution is a GL distribution. A
+similar GL distribution is also used in \citet{tolikas} who analyze extreme risk
+and value-at-risk in the German stock market, although they do not use a shape
+parameter. Regarding inflation rates, the GL distribution is -- to our best
+knowledge -- only used in context with expected inflation.
+
+\citet{batchelor} use a logistic distribution (not its generalization) to model
+the distribution of mean expected inflation rates.
+
+\subsection{Estimation}
+
+Under the assumption that $y_t$ is distributed according to the generalized
+logistic distribution, we test for the hypothesis that the parameter vector
+$\hat{\phi}$ stays stable over time:
+
+\begin{eqnarray}
+& H_0: \phi_t = \phi_0 \quad (t=1,...,n) \nonumber 
+\end{eqnarray}
+
+against the $H_1$ of changes in at least one of the parameters. We first
+estimate the parameters of the GL-distribution by means of maximum likelihood
+$\hat{\phi} ~=~ \argmax_\phi \sum_{t=1}^n \log f(y_t|\phi)$. The parameter
+stability is assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures
+of model deviation.\footnote{for details see Appendix~\ref{sec:gld2} and
+\citet{z07}}  The empirical fluctuation process $\mathit{efp}(\cdot)$ captures
+deviations from a hypothesized zero mean over time and is given by
+%
+\begin{eqnarray}
+\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
+\label{eq:efp}
+\end{eqnarray}
+%
+where $\hat{V}$ is some consistent estimator of the variance of the scores,
+which is used to decorrelate the series as to avoid problems due to serial
+correlation. Then a functional central limit theorem (FCLT) for
+$\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian
+bridge: $\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot)$.
+
+\subsection{Test}
+
+After the $\mathit{efp}(\cdot)$ is computed, it conveys information about
+departures from the $H_0$ of parameter stability over time. For a precise
+quantification of the actual deviation from its hypothesized zero-mean path, the
+main idea of the test statistics presented here is to aggregate the
+$\mathit{efp}(\cdot)$ in such a way as to be able to construct a test statistics
+so that critical values and p values can be derived. In this paper, we employ
+the sup$\mathit{LM}$ test of \citet{andrews}. This test is well suited for
+single break alternatives and performs better if more than just one of the
+distribution parameters change.
+
+It is given by:
+%
+\begin{eqnarray}
+\sup_{t \in [0.1, 0.9]} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
+\end{eqnarray}
+%
+where the appropriate $p$~values can be computed using from the corresponding
+limiting distribution -- with $\mathit{efp}(t)$ replaced by  $W^0(t)$.
+The sup$\mathit{LM}$ test computes a test statistics for all possible change points in a
+given time frame and rejects the null of no change if the maximum is too high.
+
+\subsection{Breakpoint estimation}
+
+The dating of structural breaks builds upon the work of \cite{bp03}. If a break
+is detected via the sup$\mathit{LM}$ test, the next step is estimating the number and the
+timing of these break points. Breakpoints $\tau_1, \dots,\tau_B$ are estimated
+via maximization of the full segmented likelihood
+%
+\begin{eqnarray}
+\sum_{b=1}^{B+1} \sum_{t=\tau_b + 1}^{\tau_b} \log f(y_t | \phi^{(b)}) \nonumber
+\end{eqnarray}
+%
+If the test statistics is significant (in this case the sup$\mathit{LM}$ test), at least
+one break is estimated. In principle, the optimal segmentation (i.e. the
+concrete dates of a break) can be computed once B (the number of breakpoints) is
+known. However, B needs to be based on the observed data as well. The solution
+is to compute the timing of the breaks for a number of different values for B
+and then choose B by means of an information criterion, in this case the LWZ
+(Liu, Wu and Zidek) as developed in \citet{liu}. All parameters $\tau_1, \dots,
+\tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic
+programming.\footnote{More details for this estimation technique are provided in
+\cite{z10}.}
+
+<<analysis-results>>=
+## run analysis (for seasonally adjusted data)
 root_dir <- do.call("file.path", as.list(head(strsplit(getwd(), .Platform$file.sep)[[1]], -1)))
@@
 setwd(file.path(root_dir, "Analysis"))
-## run results4 (with dummy creation for sales periods)
 Sweave("results1s.Rnw")
@@
 fnam <- Sys.glob(file.path("..", "Analysis", "*.rda"))
-fnam2 <- grep("2", fnam, fixed = TRUE)
-if(length(fnam2) > 0) fnam <- fnam[-fnam2]
+fnam_omit <- c(grep("2", fnam, fixed = TRUE), grep("glmean", fnam, fixed = TRUE))
+if(length(fnam_omit) > 0) fnam <- fnam[-fnam_omit]
 
@@
 
-## erm <- yearmon(c(
-##   Austria = 1995,
-##   ...,
-##   France = 1995 + 1/12,
-##   ...))
-
-## create erm (Exchange rate mechanism bzw. exchange rate mechanism 2
-## ERM~II, ERM~I (part of european monetary system expired in 1.1.1999 gave way to ERM II
+## Exchange Rate Mechanism information
+## ERM~II, ERM~I (part of european monetary system expired in 1999-01-01 gave way to ERM II
 ## if never entered ""right?
 ## http://web.archive.org/web/20080101022356/http://lexikon.meyers.de/meyers/Europ%C3%A4isches_W%C3%A4hrungssystem
-
 erm <- yearmon(c(
@@
 
-## sequentially load data
-## old: load(file.path("..", "Analysis", fnam[i]))
-
 for(i in seq_along(fnam)) {
@@
 }
+@
 
-table2<- vector(length = length(fnam), mode = "list")
-names(table2) <- sapply(strsplit(sapply(strsplit(fnam, "/", fixed = TRUE), tail, 1), ".", fixed = TRUE), head, 1)
+\subsection{An example: Slovenia}\label{sec:slovenia}
 
-for(qq in seq_along(fnam)) {
-  load(file.path(root_dir, "Analysis", fnam[qq]))
-  table2[[qq]] <- x
-}
+<<slovenia>>=
+x <- results$Slovenia
+spval <- sctest(x$efp, functional = supLM(0.1))$p.value
+@
 
+To illustrate our method, we consider the case of Slovenia. After a period of
+relative instability following the independence from Yugoslavia, Slovenia was
+successfull in realigning its economy and introduced a number of reforms leading
+to a stable growth in the past years. The good economic performance made it
+possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce
+the euro in January 2007. The data collected contains the period from January
+1996 onwards. In the first few years, Slovenia experienced very high inflation
+rates, but was successfull in reducing inflation to a much lower level.
+
+A glance at the time series depicted in Figure~\ref{fig:slovenia1} of the monthly inflation
+rates reveals a decline in the mean around the middle of the observed time
+period. The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches
+the peak somewhere between 2003 and 2004. The $p$~value of the sup$\mathit{LM}$ test is
+\Sexpr{round(spval, 8)} and clearly indicates at least one break.  The number of
+breaks is calculated by means of the LWZ criteria, which chooses one break in
+August 2003. The goodness-of-fit test depicted in Figure~\ref{fig:slovenia2} confirms
+the choice of the GL distribution. The mean inflation rate was cut to only a
+third of its previous value, whereas variance increased slightly and the
+distribution slightly skewed to the left, but peaking at a lower value than
+previously.
+
+In this year, Slovenia introduced a number of financial reforms
+and severely decreased the year on year growth in the M3 monetary aggregate.
+Combined, these actions were enough to ensure the successfull participation in
+the ERM~II a year later and the introduction of the euro two years later.
+
+\begin{figure}[t!]
+<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
+par(mfrow = c(1, 2))
+plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
+lines(x$breakpoints, breaks = x$nbreaks)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
+plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
+lines(x$breakpoints, breaks = x$nbreakss)
+@
+\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
+with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
+from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
+with sequence of $\mathit{LM}$~statistics,
+critical value at 5\% level (vertical dashed line), and maximum (vertical dashed line).}
+
+<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
+par(mfrow = c(1, 2))
+for(i in 1:2) plot(x$model[[i]], moments = TRUE,
+  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
+  breaks = seq(-1.5, 2, by = 0.25),
+  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
+    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
+    sep = ""))
+@
+\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
+Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
+along with fitted generalized logistic probability density function, associated
+moments, and $\chi^2$ goodness-of-fit test.}
+\end{figure}
+
+\clearpage
+\section{Results}\label{sec:results}
+
+\readme{Jesus: include p-values in table; no place left; Jesus: opening seems a
+bit confusing and obscure}
+
+Table~\ref{tab:results1} presents the break dates and the parameter values of
+the segments as well as the dates of entry into the ERM/ERM~II and the date of
+the euro introduction.
+
+<<results-moments>>=
 ## compute moments
 mom <- lapply(results, function(obj) t(sapply(obj$model, moments)))
+names(mom)[names(mom) == "CzechRepublic"] <- "Czech Rep."
 
@@
 }
+@
 
-format_cross_country <- function(crossmom, digits = 3, moment="mean", quantil="lower") {
-  rval <- format(round(crossmom, digits = digits), nsmall = digits)
-  rval <- cbind(moment, quantil, matrix(paste("$", rval, "$", sep=""), ncol = 4))
-  rval[1,1] <- "mean"
-  rval[2:3,1] <- ""
-  rval[1,2] <- "lower"
-  rval[2,2] <- "medium"
-  rval[3,2] <- "upper"
-  rval[4,1] <- "variance"
-  rval[5:6,1] <- ""
-  rval[4,2] <- "lower"
-  rval[5,2] <- "medium"
-  rval[6,2] <- "upper"
-  rval[7,1] <- "skewness"
-  rval[8:9,1] <- ""
-  rval[7,2] <- "lower"
-  rval[8,2] <- "medium"
-  rval[9,2] <- "upper"
+\begin{table}[t!]
+\begin{center}
+\footnotesize
+\begin{tabular}{llrrrlll}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{No change}} \\
+\hline
+<<moments-table1, results=tex>>=
+writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
+\hline
+<<moments-table2, results=tex>>=
+writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
+\hline
+<<moments-table3, results=tex>>=
+writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Financial crisis}} \\
+\hline
+<<moments-table4, results=tex>>=
+writeLines(unlist(lapply(c(11, 19), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Non-euro}} \\
+\hline
+<<moments-table5, results=tex>>=
+writeLines(unlist(lapply(c(20, 21), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Eastern countries}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
+@
+
+\end{tabular}
+
+\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
+distribution per segment, and ERM information for all countries under investigation.}
+\end{center}
+\end{table}
+
+\subsection{The countries in detail}
+
+The first group that draws particular attention is a subgroup of the PIIGS
+(Portugal, Ireland, Italy, Greece and Spain) countries consisting of Ireland and
+Spain. Both show a break in the summer of 2008 following the financial crisis
+and the severe problems in both countries concerning the real estate situation.
+In both countries this had a strong deflationary impact, with Ireland actually
+showing negative inflation rates for an extended period of time; in Spain, which
+was affected by a housing bubble of roughly the same size as Ireland, the
+housing prices did not adjust so abruptly and in such a degree. This
+deflationary trend, which for a time affected all countries, can be seen in
+Figure~\ref{fig:glmean}, which depicts a rolling mean estimation, where the mean
+is calculated using the first moment of a fitted GL distribution. There is
+evidence of a decline in mean inflation, which returns to its previous path
+roughly a year later.
+
+The second group consists of the Eastern European countries: Czech Republic,
+Estonia, Hungary, Poland, Slovakia and Slovenia. In almost all of these
+countries, the mean and the variance of their inflation rates declined in the
+later part of the 90s (Slovenia is an exception with an increase in variance
+after the break). Most of them experienced a break in 1997/1998, with Poland,
+Slovakia and Slovenia somewhat later. The break during the late 90s is a result
+of these countries efforts to curb in inflation by decreasing money supply and
+indicates the time when these countrie's actually overcame the biggest
+transitionally shocks on their way towards free market economies. Slovakia is
+very similar to Slovenia, as its inflation dynamics too changed roughly one year
+prior to its entry into ERM~II.
+
+Another interesting group consists -- with the exception of Austria and
+Luxembourg -- of the countries of the former DM-zone. All these countries have a
+very low inflation rate in both mean and variance and their inflation dynamics
+did not change.
+
+The next group consists of two countries that decided not to be part of any
+European Monetary System\footnote{If we ignore the short participating time of
+both countries in the EMS.}, the United Kingdom and Sweden. Both show a clear
+break in the early 90s. They can be traced back to economic crisis, the currency
+crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and
+the run on the Swedish currency in the same year.
+
+Austria, Finland and France adjusted to the requirements of the EMU during the
+first phase of convergence. These are also the countries with the lowest
+inflation rates in both mean and variance after the adjustment.
+
+The last group consists of three PIIGS countries and Luxembourg. Greece and
+Portugal adjusted to the ERM shortly before or afterwards its introduction.
+Italy was successfull in reducing its inflation rate, whereas Luxembourg
+experienced a doubling of its former -- very low -- inflation rate with the
+variance rising by a factor of 14 with a strong shift in skewness towards the
+left.
+
+\subsection{Overview over the general trend}
+
+<<eu, echo=FALSE>>=
+## impose breaks "by hand"
+eu_breaks <- yearmon(c(1998 + 11/12, 2008 + 7/12))
+eu_seg <- 1 + (time(eu) > eu_breaks[1]) + (time(eu) > eu_breaks[2])
+library("e1071")
+eu_mom_old1 <- aggregate(eu, eu_seg, function(x) c(mean(x), var(x), skewness(x)))
+
+## better: use model-implied moments
+eu_mom_old2 <- aggregate(eu, eu_seg, function(x) moments(glogisfit(x)))
+
+## or even better: estimate breaks?
+## eu_bp <- breakpoints(glogisfit(eu), h = 24)
+## moments(eu_bp, breaks = 2)
+## replicate results "by hand"
+eu_breaks <- yearmon(c(1999 + 5/12, 2007 + 7/12))
+eu_seg <- 1 + (time(eu) > eu_breaks[1]) + (time(eu) > eu_breaks[2])
+eu_mom <- aggregate(eu, eu_seg, function(x) moments(glogisfit(x)))
+@
+
+The disparity of the national inflation rates is still an issue. Many researchers, like \citet{hofmann}  find a considerable amount of inflation
+differentials. \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity
+across EMU countries in terms of average inflation and its degree of persistence. In a paper examining structural convergence of inflation rates in EU
+countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar. They find that
+convergence over time of inflation dynamics was only partly observable.  In a paper studying core inflation and using an aggregated euro area inflation
+rate, \citet{morana} finds three regimes (roughly 1980--1984, 1984--1993 and 1993--2000) governing the core inflation rate.
+
+Concerning inflation differentials, Figure~\ref{fig:glmean} depicts the inflation rates and a rolling mean estimation (with a window size of 12 months),
+where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month,
+the lines are overlapped and thus get darker. We wold see one black line if all countries would show the same month-on-month increase at any one time. The
+red lines depict the means.} We do see that many of these countries follow a rather similar path. One event clearly visible is the financial crises and the
+short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months. Over time, there are hints, that the inflation rates
+converged towards some common -- albeit broad -- mean level. For the inflation series of the whole EMU, as published by the \citet{oecd}, we see
+an increase in mean from
+$\Sexpr{round(eu_mom[1, 1], digits = 3)}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
+$\Sexpr{round(eu_mom[2, 1], digits = 3)}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a strong decrease to
+$\Sexpr{round(eu_mom[3, 1], digits = 3)}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) following the financial crises.
+Whereas variance has increased over time from
+$\Sexpr{round(eu_mom[1, 2], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 2], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 2], digits = 3)}$.
+Regarding skewness, we see a clear change towards a negative skewness from
+$\Sexpr{round(eu_mom[1, 3], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 3], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 3], digits = 3)}$.
+\fixme{Z: interpretation needs to be adapted to modified analysis}
+
+<<glmean>>=
+if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
+   glmean <- function(x) {
+    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
+  }  
+  hicp_glmean <- rollapply(hicps, 12, glmean)
+  setwd(file.path(root_dir, "Analysis"))
+  save(hicp_glmean, file = "glmean.rda")
+  setwd(file.path(root_dir, "Paper"))
+} else {
+  load(file.path(root_dir, "Analysis", "glmean.rda"))
+}
+@
+
+\setkeys{Gin}{width=0.8\textwidth}
+\begin{figure}[t!]
+\begin{center}
+<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
+  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
+  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
+    xlab = "Time", ylab = "Monthly inflation rates")
+  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+@
+\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from rolling GL fits (red)
+over time for all countries.}
+\end{center}
+\end{figure}
+\setkeys{Gin}{width=\textwidth}
+
+<<cross-country, echo=FALSE, results=hide>>=
+cross_country_moments <- function(countries, date,
+  probs = c(0.25, 0.5, 0.75))
+{
+  ret <- mat.or.vec(9, length(date)) 
+  for(jj in 1:length(date)) {
+    rval <- sapply(countries, function(country) {
+      nb <- results[[country]]$nbreaks 
+      m <- t(sapply(results[[country]]$model, moments)) 
+      bd <- if(nb < 1) { 
+        end(results[[country]]$data) 
+      } else {
+        breakdates(results[[country]]$breakpoints, breaks = nb) 
+      }
+      m[sum(as.yearmon(date[jj]) > bd) + 1, ] 
+    })
+    v1 <- apply(rval, 1, quantile, probs = probs)
+    ret[1:3, jj] <- v1[,1] 
+    ret[4:6, jj] <- v1[,2] 
+    ret[7:9, jj] <- v1[,3]  
+  }
+  return(ret)
+}
+
+format_cross_country <- function(x, digits = 3, columns = 1:NCOL(x)) {
+  rval <- format(round(x, digits = digits), nsmall = digits)
+  for(i in which(!(1:NCOL(rval) %in% columns))) rval[, i] <- c(rep("", 4), "\\cdots", rep("", 4))
+  rval <- cbind("", "", matrix(paste("$", rval, "$", sep=""), ncol = 4))
+  rval[, 1] <- c("Mean", "", "", "Variance", "", "", "Skewness", "", "")
+  rval[, 2] <- rep(c("Lower", "Median", "Upper"), 3)
   rval <- apply(rval, 1, paste, collapse = " & ")
@@
 
+yrs <- c(1994, 1998, 2007, 2010)
+euro <- c(1, 2, 6, 7, 8, 13, 14)
+others <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
+piigs <- c(9, 11, 12, 16, 19)
+@
+
+Regarding the general trend, Table~\ref{tab:results2} presents lower quartile,
+median, and upper quartile (25\%, 50\%, 75\% quantiles) of the first three moments of the
+inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}. The countries are grouped
+into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany,
+Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland,
+Greece and Spain) and all the others. For this, we took account of possible
+breaks in the moments.
+
+\begin{table}[t!]
+\begin{center}
+\begin{tabular}{llrrrr}
+\hline
+Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
+\hline
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Core euro}} \\
+\hline
+<<cross-country-table1, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{PIIGS}} \\
+\hline
+<<cross-country-table2, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Other}} \\
+\hline
+<<cross-country-table3, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
+@
+
+\end{tabular}
+\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
+January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
+changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
+\end{center}
+\end{table}
+
+In the core-zone, we see a decrease in the lower quartil from \Sexpr{yrs[1]} to \Sexpr{yrs[4]}, with no change in the median inflation rate, which is due to the fact, that there is no break after \Sexpr{yrs[2]} for these countries; however, the interquartile range decreased by almost 20\%.
+The variance increased slightly (but with a 15\% decrease in interquartile range), whereas skewness stayed symmetric in the median, with almost no change in the interquartile range.
+
+The other countries were successfull in decreasing all three moments, but they are still higher than the euro-zone countries.
+Most importantly, the interquartile range for all moments decreased considerably showing a clear convergence towards the core-zone countries.
+The PIIGS countries decreased inflation rates as well with a strong change of skewness towards more negative values.
+The interquartile range of their inflation variance increased considerably, which is mostly due to Spain and Ireland and their strong reaction to the financial crisis.
+
+\clearpage
+\section{Conclusion}\label{sec:conclusion}
+
+%Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
+
+\fixme{Jesus: this belongs to the intro section} An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
+What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
+The effects of joining a monetary union on inflation variability in his framework depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
+However, due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on a variety of assumptions.
+
+%\fixme{Jesus: maybe, but you don't show that in your analysis} Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
+This is due to the fact that two of the most important inflationary channels, the money supply and the degree of openness, had to be arranged according to the criteria necessary for EMU entry.
+To achieve convergence in inflation rates, the growth in money supply had to be in concert with the M3 growth rate as decided by the European Central Bank (ECB).
+The opening of the economies made cheaper goods available and thus contributed to lower prices.
+Like the original countries participating in the ERM, the Eastern European countries had to synchronize money supply growth and interest rates in order to fix exchanges rates or otherwise risk the chance of severe disequilibria.
+
+Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
+For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
+A very volatile inflation rate will likely  contribute to greater macroeconomic instability.
+Although \cite{grier} find no evidence that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --  \cite{jaro} finds a positive correlation between the level and the standard deviation of inflation.
+This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro-zone countries, and especially the PIIGS, indicating that in future higher inflation rates are more likely.
+This will put pressure on the ECB to achieve its inflation target of 2\%.
+
+\clearpage
+\section*{Computational details}
+
+Our results were obtained using 
+\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} \citep{R} with the packages
+\pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]},
+\pkg{strucchange}~\Sexpr{packageDescription("strucchange")["Version"]} \citep{Zeileis+Leisch+Hornik:2002}, and
+\pkg{fxregime}~\Sexpr{packageDescription("fxregime")["Version"]} \citep{z10},
+all of which are available under the General Public License (GPL) from the
+Comprehensive \proglang{R} Archive Network (\url{http://CRAN.R-project.org/}). Package \pkg{strucchange} provides
+various techniques for inference in structural-change settings which are complemented
+with additional maximum-likelihood-based methods by \pkg{fxregime}. Inference for the
+generalized logistic model (that can be combined with \pkg{strucchange}/\pkg{fxregime})
+is implemented in the new package \pkg{glogis} that also contains the data used, both
+the raw monthly price indexes \citep[as provided by][]{oecd} and the seasonally adjusted
+monthly inflation rates. The seasonal adjustment was performed in \pkg{X-12-ARIMA}~0.3
+(\citealp{X-12-ARIMA}, see also \citealp{Findley+Monsell+Bell:1998}) through the interface
+provided by \proglang{gretl}~1.9.5 (\citealp{gretl}, see also \citealp{Smith+Mixon:2006} for a review).
+
+\bibliography{papers}
+
+\clearpage
+
+\begin{appendix}
+
+\section{GL distribution}\label{sec:gld2}
+
+The moments of the GL distribution as given in Equation~\ref{eq:dglogis} are
+%
+\begin{eqnarray}
+\mathrm{E}(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
+\mathrm{Var}(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
+\mathrm{Skew}(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
+\end{eqnarray}
+%
+where $\gamma'(\cdot)$ and $\gamma''(\cdot)$ are the first and the second
+derivatives of the digamma function, respectively. The log-likelihood is
+%
+\begin{eqnarray}
+\log f(y|\theta,\sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{1}{\sigma} (y-\theta) - (\delta+1) \nonumber \\
+& & \times \log (1+\exp^{-\frac{y-\theta}{\sigma}}) \nonumber 
+\end{eqnarray}
+%
+The resulting score function, where the scores are the derivatives of the
+log-likelihood function, has three components $(s_{\theta}, s_{\sigma},
+s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$. These are
+employed in Equation~\ref{eq:efp} for the construction of the test statistics
+and are a means of measuring the estimation error in a maximum likelihood
+framework.
+
+\begin{eqnarray}
+s_{\theta}(y|\theta, \sigma, \delta) & =  &\frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \theta}   \\
+&=&  \frac{1}{\sigma} - (\delta+1)\cdot\frac{\frac{1}{\sigma}\tilde{y}}{(1+\tilde{y})}  \nonumber \\
+s_{\sigma}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \sigma}   \\
+& =&  -\frac{1}{\sigma} + \frac{1}{\sigma^2}(y-\theta) - (\delta+1) \times \frac{\frac{1}{\sigma^2}(y-\theta)\tilde{y}}{(1+\tilde{y})} \nonumber \\
+s_{\delta}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \delta}  \\
+& =&   \frac{1}{\delta} - \log(1+\tilde{y}) \nonumber 
+\end{eqnarray}
+%
+The scores of the parameters $\phi = (\theta, \sigma, \delta)$ sum up to 0: $\sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0$. The first order condition is given by:
+%
+\begin{eqnarray}
+\sum_{i=1}^n s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) = 0
+\end{eqnarray}
+
+\clearpage
+\section{Data and sales periods}\label{sec:sales}
+
+In addition to Germany, Portugal, Belgium, Italy, Luxembourg and Spain, there
+are a number of other countries, where the sales periods are regulated by law.
+France has an individual time span regulated by each department. This should
+pose no problem, since there is no global state level solution. In Greece there
+are also two set sales-periods in winter and summer. In Italy the sales periods
+are fixed each year by the Chambers of Commerce.  Portugal has two set sales
+periods as well, but for two month (January and February as well as Mid-July to
+Mid-September). The source for this information is \citet{konseur}. According to
+\cite{pltm}, the official dates of the inclusion of these sales periods are:
+January 1998 for Germany and Portugal, January 2000 for Belgium and January 1999
+for Luxembourg and Ireland, while Italy and Spain introduced them in January
+2001.
+
+\clearpage
+\section{Graphs}
+
+Figures~\ref{fig:nochange}--\ref{fig:eastern} depict the data underlying the results
+in Section~\ref{sec:results} along with the GL-based means from Table~\ref{tab:results1}.
+
+<<graphs-code>>=
 get_data <- function(i) {
@@
   rval <- do.call("merge",
-    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks=results[[j]]$nbreaks, type = "mean")))
+    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks = results[[j]]$nbreaks, type = "mean")))
   colnames(rval) <- names(results)[i]
@@
 
-get_xyplot <- function(i)
+library("lattice")
+get_xyplot <- function(i) {
+  nam <- names(results)[i]
+  nam[nam == "CzechRepublic"] <- "Czech Republic"
   xyplot(merge(get_data(i), get_fitted(i)),
-    screen = rep(names(results)[i], 2),
+    screen = rep(nam, 2),
     col = rep(c(1, 4), each = length(i))) 
-
-xyplot.gefp <- function(x, panel = NULL, ylim = NULL, ...)
-{
-  cval <- maxBB$computeCritval(0.05, ncol(x$process))
-  if(is.null(ylim)) ylim <- range(c(range(x$process), 1.1 * cval, -1.1 * cval))
-
-  gefp_plot <- function(x, y, subscripts, groups, panel = panel.xyplot,
-    col = 1, type = "p", pch = 20, lty = 1, lwd = 1, ...)
-  {
-    col <- rep(as.list(col), length = nlevels(groups))
-    type <- rep(as.list(type), length = nlevels(groups))
-    pch <- rep(as.list(pch), length = nlevels(groups))
-    lty <- rep(as.list(lty), length = nlevels(groups))
-    lwd <- rep(as.list(lwd), length = nlevels(groups))
-
-    for(g in 1:nlevels(groups)) {
-      idx <- levels(groups)[g] == groups[subscripts]
-      if (any(idx)) panel(x[idx], y[idx], ...,
-        col = col[[g]], type = type[[g]], pch = pch[[g]],
-        lty = lty[[g]], lwd = lwd[[g]])
-      grid::grid.lines(y = grid::unit(0, "native"))
-      grid::grid.lines(y = grid::unit(cval, "native"), gp = grid::gpar(col = 2))
-      grid::grid.lines(y = grid::unit(-cval, "native"), gp = grid::gpar(col = 2))
 }
-  }
-  if(is.null(panel)) panel <- gefp_plot
-
-  xyplot(x$process, panel = panel, ylim = ylim, ...)
-}
-
-## data
-data("HICP", package = "glogis")
-HICP <- HICP[, colnames(HICP) != "EU"]
-
-## to change to shorter names if necessary
-names(HICP) <- c("Austria",  "Belgium", "CzechRepublic", "Denmark", "Estonia", 
-"Finland", "France",  "Germany", "Greece", "Hungary", "Ireland",  "Italy", "Luxembourg", "Netherlands",
-"Poland", "Portugal",  "Slovakia", "Slovenia",  "Spain", "Sweden", "UK")
-
-if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
-   glmean <- function(x) {
-    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
-  }
-
-  hicps <- hicps[, colnames(hicps) != "EU"]
-
-  ## rolling GL means
-  hicp_glmean <- rollapply(hicps, 12, glmean)
-
-  setwd(file.path(root_dir, "Analysis"))
-	save(hicps, hicp_glmean, file = "glmean.rda")
-	setwd(file.path(root_dir, "Paper"))
-
-}
-
 @
 
-% all created dynamically
-%\graphicspath{{Figures/}}
+\setkeys{Gin}{width=0.8\textwidth}
+\begin{figure}[p!]
+\begin{center}
+<<graphs-nochange, fig=TRUE, height=6.3, width=7.5>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(get_xyplot(c(2, 4, 8, 14)))
+@
+\caption{\label{fig:nochange} No change.}
 
-\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}
-
-%% for internal use
-\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
-\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
-
-\author{Thomas Windberger\\Universit\"at Innsbruck \And
-        Achim Zeileis\\Universit\"at Innsbruck
-}
-\Plainauthor{Thomas Windberger, Achim Zeileis}
-
-\Abstract{
-%(... text when everything else finished, present innovation of paper...)
-}
-
-\Keywords{inflation rate, structural break, EMU, Generalized Logistic Distribution 
-}
-
-\Address{
-  Thomas Windberger, Achim Zeileis\\
-  Department of Statistics\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstraße 15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
-}
-
-\begin{document}
-
-\section{Introduction}
-
-Ever since the beginning of the European Union, the topic of a common currency was a controversial issue.
-Although the Economic and Monetary Union (EMU) is now a fact, the discussion about the economic effects of the euro is far from being settled.
-The controversial topics range from the question of whether or not the eurozone is indeed an optimal currency area (OCA) all the way to the very survival of the euro in light of the budgetary problems of some of its member states.
-The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context, is in the center of an ongoing debate: this regards the question about short-run and steady-state inflation uncertainty (in context to inflation expectations) dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries, which is the topic of an investigation by \citet{palomba}.
-\\
-The way towards the EMU basically consisted of 3 stages: stage I (1990-1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
-A possible fourth stage is the continuing integration of new member states, mostly former communist countries.
-Any new state that wants to join the EMU is first obliged to fulfill the Maastricht criteria, which, besides setting rules for government debts and interest rates, also requires the participation in the ERM~II (Exchange Rate Mechanism~II) for two years, while the exchange rates towards the euro are not allowed to cross the nominal band.\\
-\cite{emerson} put emphasize on the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
-Failing to stabilize inflation leads to severe economic problems.
-Throughout the literature, there is still a considerable degree of uncertainty as to what extent the introduction of the euro, or monetary unions in general, affect the inflation rate.
-The EMU requires a country to possibly alter a number of economic policies and institutions, which may demand for significant short term adjustments to take place.
-\clearpage
-Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria are likely to lead to an alteration at least in the mean (since this was one of the requirements) and possibly in the volatility of their respective inflation rates as well.
-\\
-However, a deeper inquiry into this problem is in need of a method able to capture changes in inflation dynamics.
-We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates and thereby upon the effect of the EMU on a number of European countries both within and without the eurozone.
-The framework developed in this paper makes it possible to test for any significant changes of a countries inflation dynamics and thus helps interpreting the effects of the EMU.
-We focus on changes in the first three moments of these inflation rates, which we assume to be distributed according to a generalized logistic distribution (GL).
-The GL distribution is a suitable one due to its flexibility to capture nonsymmetry and because of its fatter tails.
-For this purpose, we develop a new estimation approach building on previous research by \citet{z07}.
-\\
-This \fixme{Jesus: sell method better, why this one and not another} framework allows us to assess the evolution of mean, variance and skewness over time.
-The skewness parameter turns out to be quite important to measure changes in the long run inflation dynamics, as it changes considerably over the time of the analysis.
-This trend would be all but invisible using a normal distribution.
-Our estimation technique enables the simultaneous estimation of breaks in the mean (which is important in the context of the Maastricht criteria), the variance (which should be low for the price mechanism to be efficient) and the skewness (where a change in skewness is evidence of a shift of the distribution towards a new ``center of gravity'') of the inflation rates of interest.
-This than can be used to search for answers regarding the question about the effect of the EMU on the inflation dynamics of its member states or more general to investigate the effects of external shocks.\\
-The \fixme{Jesus: maybe comment on results here} remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.
-
-\clearpage
-\section{Data}\label{s2}
-
-We \fixme{Jesus: yearly inflation rates better; don't start with data in paper} employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}.
-Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom.
-Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity.
-Cypria and Malta are not included since they are very small economies.
-\\
-The countries in this sample can be divided into three different groups.
-First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).
-Second, the EU members not participating in the ERM~II: Czech Republic, Hungary, Poland, United Kingdom, and Sweden.
-Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.
-\\
-As  shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Germany, Portugal, Ireland, Luxembourg, Belgium, Italy and Spain); failing to take this into account may lead to considerable distortions in the estimation.
-We dealt with the inclusion of sales periods\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} and all the other seasonality in the data by using the X--12--ARIMA method for desaisonalization, as described in \cite{X-12-ARIMA} and \cite{Findley+Monsell+Bell:1998}.
-
-\clearpage
-\section{Model}\label{s3}
-
-To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks.
-First, the model is fitted for the whole sample using the generalized logistic distribution.
-This then enables testing for the stability of the parameter estimates.
-If there is evidence of significant changes, the number and the timing of the break points are estimated.
-\\
-This framework allows to trace the evolution of mean, variance and skewness of the inflation series and thus interpreting deviations in any of these parameters as possibly due to a change in monetary policy, is made feasible.
-Any huge intervention or change is captured by fluctuations in the parameter stability of the series and can be quantified and related to a distinct time frame.
-\\
-The correlation \fixme{Jesus: yes, but a DGP needs to be assumed, it would need to take care of any autororrelation, wouldn't it?} dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative tests, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.}
-
-\subsection{Generalized Logistic Distribution}\label{s31}
-
-The data suggested that the use of a normal distribution might cause some estimation problems, since most of the time series exhibit asymmetric properties (with sometimes rather strong skewness) and -- compared to the normal distribution -- a fatter tail.
-To accommodate for this fact, the use of the generalized logistic distribution (GL) with its relative low degree of complexity, seemed appropriate.
-On the one hand, it is a simple distribution and on the other hand, it is flexible enough to capture essential properties of the data.
-For the inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we therefore assume a GL distribution as defined in \citet{johnson}:
-
-\begin{eqnarray}
-f(y | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma} \cdot \exp^{-\frac{y-\theta}{\sigma}}}{(1+\exp^{-\frac{y-\theta}{\sigma}})^{(\delta+1)}} 
-\label{eqn1}
-\end{eqnarray}
-
-with location $\theta$, scale $\sigma$ and shape $\delta$.
-For $\delta=1$ the distribution simplifies to the logistic distribution, for $\delta<1$ it is skewed to the left and for $\delta$$>$1 it is skewed to the right (see Appendix \ref{glscore} for more details).
-\\
-In econometrics, the logistic distribution is often used in income distributions and growth models.
-This is due to  its fatter tails, which fit these kind of data somewhat better.
-\citet{won} develop a maximum likelihood \fixme{Jesus: is this not better than what you do here: answer: no, they use completely other framework} estimator if the underlying distribution is a GL distribution.
-A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they do not use a shape parameter.
-Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context with expected inflation.
-\\
-\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates. \\
-
-\subsection{Estimation}
-
-Under the assumption that $y_t$ is distributed according to the generalized logistic distribution, we test for the hypothesis that the parameter vector $\hat{\phi}$ stays stable over time:
-
-\begin{eqnarray}
-& H_0: \phi_t = \phi_0 \quad (t=1,...,n) \nonumber 
-\end{eqnarray}
-
-against the $H_1$ of changes in at least one of the parameters.
-We first estimate the parameters of the GL--distribution by means of maximum likelihood $\hat{\phi} ~=~ \underset{\phi}{argmax} \sum_{t=1}^n \log f(y_t|\phi)$.
-The parameter stability is assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures of model deviation.\footnote{for details see Appendix \ref{glscore} and \citet{z07}}  The empirical fluctuation process $\mathit{efp}(\cdot)$ captures deviations from a hypothesized zero mean over time and is given by:
-
-\begin{eqnarray}
-\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
-\label{eqn2}
-\end{eqnarray}
-%(... jesus: thinks this is incomplete ...) \\
-\fixme{Jesus: this seems a bit incomplete} where $\hat{V}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation.
-Then a functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian bridge:
-
-\begin{eqnarray}
-\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot) \nonumber
-\end{eqnarray}
-
-\subsection{Test}
-
-After the $\mathit{efp}(\cdot)$ is computed, it conveys information about departures from the $H_0$ of parameter stability over time.
-For a precise quantification of the actual deviation from its hypothesized zero--mean path, the main idea of the test statistics presented here is to aggregate the $\mathit{efp}(\cdot)$ in such a way as to be able to construct a test statistics so that critical values and p values can be derived.
-In this paper, we employ the Supremum of LM (supLM) test by \citet{andrews}.
-This test is well suited for single break alternatives and performs better if more than just one of the distribution parameters change.
-
-\clearpage
-It is given by:
-
-\begin{eqnarray}
-\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
-\end{eqnarray}
-
-where the appropriate p--values can be computed using:
-
-\begin{eqnarray}
-\underset{t \in [0.1, 0.9]}{sup} \frac{\Vert W^0(t) \Vert_2^2}{t(1-t)} \nonumber
-\end{eqnarray}
-
-The supLM test computes a test statistics for all possible change points in a given time frame and rejects the null of no change if the maximum is too high.
-
-\subsection{Breakpoint Estimation}
-
-The dating of structural breaks builds upon the work of \cite{bp03}.
-If a break is detected via the supLM test, the next step is estimating the number and the timing of these break points.
-Breakpoints $\tau_1,...,\tau_B$ are estimated via maximization of the full segmented likelihood:
-
-\begin{eqnarray}
-\sum_{b=1}^{B+1} \sum_{t=\tau_b + 1}^{\tau_b} \log f(y_t | \phi^{(b)}) \nonumber
-\end{eqnarray}
-
-If the test statistics is significant (in this case the supLM test), at least one break is estimated.
-In principle, the optimal segmentation (i.e.
-the concrete dates of a break) can be computed once B (the number of breakpoints) is known.
-However, B needs to be based on the observed data as well.
-The solution is to compute the timing of the breaks for a number of different values for B and then choose B by means of an information criterion, in this case the LWZ (Liu, Wu and Zidek) as developed in \citet{liu}.
-All parameters $\tau_1, \dots, \tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic programming.\footnote{More details for this estimation technique are provided in \cite{z10}.}
-
-\clearpage
-\subsection{An Example}\label{s4}
-
-<<echo=FALSE>>=
-        x <- results$Slovenia
-        spval <- sctest(x$efp, functional = supLM(0.1))$p.value
+<<graphs-phase1, fig=TRUE, height=4.9, width=7.5>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(get_xyplot(c(1, 6, 7)))
 @
-
-To illustrate our method, we consider the case of Slovenia.
-After a period of relative instability following the independence from Yugoslavia, Slovenia was successfull in realigning its economy and introduced a number of reforms leading to a stable growth in the past years.
-The good economic performance made it possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce the euro in January 2007.
-The data collected contains the period from January 1996 onwards.
-In the first few years, Slovenia experienced very high inflation rates, but was successfull in reducing inflation to a much lower level.
-\\
-A glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period.
-The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches the peak somewhere between 2003 and 2004.
-The p--value of the supLM test is \Sexpr{round(spval, 8)} and clearly indicates at least one break.
-The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.
-The goodness of fit test depicted in Figure \ref{figtest} confirms the choice of the GL distribution.
-The mean inflation rate was cut to only a third of its previous value, whereas variance increased slightly and the distribution slightly skewed to the left, but peaking at a lower value than previously.
-\\
-In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate.
-Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the euro two years later.
-
-\begin{figure}[h!tp]
-<<fig=TRUE, height=4, width=8>>=
-x <- results$Slovenia
-par(mfrow = c(1, 2))
-plot(x$data, main = "Slovenia", ylab = "100 * diff(log(HICP))", xlab = "Time")
-lines(x$breakpoints, breaks = x$nbreaks)
-lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
-if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
-plot(x$efp, functional = supLM(0.1), main = "supLM test")
-lines(x$breakpoints, breaks = x$nbreakss)
-@
-\caption{Slovenia: breakpoint estimate and supLM test}
-\label{fig1}
+\caption{\label{fig:phase1} Phase~1 convergence.}
+\end{center}
 \end{figure}
 
-\begin{figure}[h!tp]
-<<fig=TRUE, height=4, width=8>>=
-x <- results$Slovenia
-par(mfrow=c(1,2))
-for(i in 1:2)
-plot(x$model[[i]], moments = TRUE,
-  xlim = c(-1.5, 2), ylim = c(0, 1.3),
-  breaks = seq(-1.5, 2, by = 0.25),
-  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
-    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
-    sep = ""))
+\begin{figure}[p!]
+\begin{center}
+<<graphs-phase2, fig=TRUE, height=6.3, width=7.5>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(get_xyplot(c(9, 12, 13, 16)))
 @
-\caption{Slovenia: Goodness of fit test}
-\label{figtest}
+\caption{\label{fig:phase2} Phase~2 convergence.}
+
+<<graphs-crisis, fig=TRUE, height=3.5, width=7.5>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(get_xyplot(c(11, 19)))
+@
+\caption{\label{fig:crisis} Financial crisis.}
+\end{center}
 \end{figure}
 
-\clearpage
-\newpage
-\section{Results}\label{s5}
+\begin{figure}[p!]
+\begin{center}
+<<graphs-noneuro, fig=TRUE, height=3.5, width=7.5>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(get_xyplot(c(20, 21)))
+@
+\caption{\label{fig:noneuro} Non-euro.}
 
-\footnotesize
-\begin{longtable}{|l|l|r|r|r|l|l|l|}
-\hline
-\multicolumn{8}{|c|}{Financial Crisis} \\
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(11, 19), format_moments)))
+<<graphs-eastern, fig=TRUE, height=9.1, width=7.5>>=
+trellis.par.set(theme = standard.theme(color = FALSE))
+print(get_xyplot(c(3, 5, 10, 15, 17, 18)))
 @
-\hline
-\multicolumn{8}{|c|}{The Eastern countries} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(3,5,10,15,17,18), format_moments)))
-@
-\hline
-\multicolumn{8}{|c|}{No change countries} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(2,4,8,14), format_moments)))
-@
-\hline
-\multicolumn{8}{|c|}{The 90s} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(20, 21), format_moments)))
-@
-\hline
-\multicolumn{8}{|c|}{Phase 1 convergence} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
-@
-\hline
-\multicolumn{8}{|c|}{Phase 2 convergence} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(unlist(lapply(c(9,12,13,16), format_moments)))
-@
-\caption{Break point, parameter estimates and ERM information}
-\label{resulttable}
-\end{longtable}
-
-\clearpage
-\normalsize
-
-\readme{Jesus: include p-values in table; no place left; Jesus: opening seems a bit confusing and obscure} Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction.
-
-\subsection{The countries in detail}
-
-The first group that draws particular attention is a subgroup of the PIIGS (Portugal, Ireland, Italy, Greece and Spain) countries consisting of Ireland and Spain.
-Both show a break in the summer of 2008 following the financial crisis and the severe problems in both countries concerning the real estate situation.
-In both countries this had a strong deflationary impact, with Ireland actually showing negative inflation rates for an extended period of time; in Spain, which was affected by a housing bubble of roughly the same size as Ireland, the housing prices did not adjust so abruptly and in such a degree.
-This deflationary trend, which for a time affected all countries, can be seen in Figure \ref{greyplot}, which depicts a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
-There is evidence of a decline in mean inflation, which returns to its previous path roughly a year later.
-\\
-The second group consists of the Eastern European countries: Czech Republic, Estonia, Hungary, Poland, Slovakia and Slovenia.
-In almost all of these countries, the mean and the variance of their inflation rates declined in the later part of the 90s (Slovenia is an exception with an increase in variance after the break).
-Most of them experienced a break in 1997/1998, with Poland, Slovakia and Slovenia somewhat later.
-The break during the late 90s is a result of these countries efforts to curb in inflation by decreasing money supply and indicates the time when these countrie's actually overcame the biggest transitionally shocks on their way towards free market economies.
-Slovakia is very similar to Slovenia, as its inflation dynamics too changed roughly one year prior to its entry into ERM~II.
-\\
-Another interesting group consists -- with the exception of Austria and Luxembourg -- of the countries of the former DM--zone.
-All these countries have a very low inflation rate in both mean and variance and their inflation dynamics did not change.
-\\
-The next group consists of two countries that decided not to be part of any European Monetary System\footnote{If we ignore the short participating time of both countries in the EMS.}, the United Kingdom and Sweden.
-Both show a clear break in the early 90s.
-They can be traced back to economic crisis, the currency crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and the run on the Swedish currency in the same year.
-\\
-Austria, Finland and France adjusted to the requirements of the EMU during the first phase of convergence.
-These are also the countries with the lowest inflation rates in both mean and variance after the adjustment.
-\\
-The last group consists of three PIIGS countries and Luxembourg.
-Greece and Portugal adjusted to the ERM shortly before or afterwards its introduction.
-Italy was successfull in reducing its inflation rate, whereas Luxembourg experienced a doubling of its former -- very low -- inflation rate with the variance rising by a factor of 14 with a strong shift in skewness towards the left.
-
-\subsection{Overview over the general trend}
-
-<<echo=FALSE>>=
-data("hicps", package = "glogis")
-euhi <- hicps[, colnames(hicps) == "EU"]
-euhi <- na.omit(euhi)
-euhim1 <- mean(window(euhi, start=yearmon(1990), end= yearmon(1998+11/12)))
-euhim2 <- mean(window(euhi, start=yearmon(1998), end=yearmon(2008+7/12)))
-euhim3 <- mean(window(euhi, start=yearmon(2008+8/10), end=yearmon(2010+11/12)))
-euhiv1 <- var(window(euhi, start=yearmon(1990), end= yearmon(1998+11/12)))
-euhiv2 <- var(window(euhi, start=yearmon(1998), end=yearmon(2008+7/12)))
-euhiv3 <- var(window(euhi, start=yearmon(2008+8/10), end=yearmon(2010+11/12)))
-euhis1 <- skewness(window(euhi, start=yearmon(1990), end= yearmon(1998+11/12)))
-euhis2 <- skewness(window(euhi, start=yearmon(1998), end=yearmon(2008+7/12)))
-euhis3 <- skewness(window(euhi, start=yearmon(2008+8/10), end=yearmon(2010+11/12)))
-@
-
-The disparity of the national inflation rates is still an issue.
-Many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials.
-\citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence.
-In a paper examining structural convergence of inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar.
-They find that convergence over time of inflation dynamics was only partly observable.
-In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.
-\\
-Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation (with a window size of 12 months), where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker.
-We wold see one black line if all countries would show the same month--on--month increase at any one time.
-The red lines depict the means.} We do see that many of these countries follow a rather similar path.
-One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months.
-Over time, there are hints, that the inflation rates converged towards some common -- albeit broad -- mean level.
-For the inflation series of the whole EMU, as published by the \citet{oecd}, we see \readme{Jesus: why not do this for EMU inflation (i.e.
-the one time series of the OECD} an increase in mean from \Sexpr{round(euhim1, 5)} (1990(1)--1998(12)) to \Sexpr{round(euhim2, 5)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 5)} (2008(9)--2010(12)) following the financial crises.
-Whereas variance has increased over time from    \Sexpr{round(euhiv1, 5)} to \Sexpr{round(euhim3, 5)}.
-Regarding skewness, we see a clear change towards a negative skewness from \Sexpr{round(euhis1, 5)} to \Sexpr{round(euhis3, 5)}.
-
-\begin{figure}[h!tp]
-<<greyplot, fig=TRUE, height=8, width=8>>=
-  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-  ## plot returns (gray) and rolling means (red)
-  hicps <- hicps[, colnames(hicps) != "EU"]
-  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1], xlab="Time", ylab="Series")
-  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-@
-\caption{Inflation rates and estimated GL mean over time}
-\label{greyplot}
+\caption{\label{fig:eastern} Eastern countries.}
+\end{center}
 \end{figure}
-
-\clearpage
-
-Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series in January 1994, 1998, 2007 and 2010.
-The countries are grouped into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.
-For this, we took account of possible breaks in the moments.
-\\
-
-\begin{longtable}{|l|l|r|r|r|r|}
-\hline
-\multicolumn{6}{|c|}{Core--euro} \\
-\hline
-Moment & Quantil & 1994 & 1998 & 2007 & 2010\\
-\hline
-<<echo=FALSE, results=tex>>=
-euro <- c(1, 2, 6, 7, 8, 13, 14)
-others <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
-piigs <- c(9, 11, 12, 16, 19)
-writeLines(format_cross_country(cross_country_moments(euro, date=yearmon(c(1994,1998,2007,2010)))))
-@
-\hline
-\multicolumn{6}{|c|}{Others} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(others, date=yearmon(c(1994,1998,2007,2010)))))
-@
-\hline
-\multicolumn{6}{|c|}{PIIGS} \\
-\hline
-<<echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(piigs, date=yearmon(c(1994,1998,2007,2010)))))
-@
-\hline
-\caption{Development of first three moments}
-\label{grouptable}
-\end{longtable}
-
-In the core--zone, we see a decrease in the lower quartil from 1994 to 2010, with no change in the medium inflation rate, which is due to the fact, that there is no break after 1998 for these countries; however, the interquartile range decreased by almost 20\%.
-The variance increased slightly (but with a 15\% decrease in interquartile range), whereas skewness stayed symmetric in the medium, with almost no change in the interquartile range.
-\\
-The other countries were successfull in decreasing all three moments, but they are still higher than the euro--zone countries.
-Most importantly, the interquartile range for all moments decreased considerably showing a clear convergence towards the core--zone countries.
-The PIIGS countries decreased inflation rates as well with a strong change of skewness towards more negative values.
-The interquartile range of their inflation variance increased considerably, which is mostly due to Spain and Ireland and their strong reaction to the financial crisis.
-
-\clearpage
-\section{Conclusion}\label{s6}
-%
-%Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
-\\
-\fixme{Jesus: this belongs to the intro section} An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
-What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
-The effects of joining a monetary union on inflation variability in his framework depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-However, due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on a variety of assumptions.
-\\
-%\fixme{Jesus: maybe, but you don't show that in your analysis} Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
-This is due to the fact that two of the most important inflationary channels, the money supply and the degree of openness, had to be arranged according to the criteria necessary for EMU entry.
-To achieve convergence in inflation rates, the growth in money supply had to be in concert with the M3 growth rate as decided by the European Central Bank (ECB).
-The opening of the economies made cheaper goods available and thus contributed to lower prices.
-Like the original countries participating in the ERM, the Eastern European countries had to synchronize money supply growth and interest rates in order to fix exchanges rates or otherwise risk the chance of severe disequilibria.
-\\
-Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
-For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
-A very volatile inflation rate will likely  contribute to greater macroeconomic instability.
-Although \cite{grier} find no evidence that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --  \cite{jaro} finds a positive correlation between the level and the standard deviation of inflation.
-This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro--zone countries, and especially the PIIGS, indicating that in future higher inflation rates are more likely.
-This will put pressure on the ECB to achieve its inflation target of 2\%.
-
-\clearpage
-\section*{Computational Details}
-
-The results shown are computed with a new \readme{add reference} \proglang{R} package \pkg{glogis}, which is an enhancement to the already existing \proglang{R} package \readme{add reference} \pkg{strucchange}, which currently does not support a GL distribution.
-The tests and the graphical illustration of the empirical fluctuation process were developed in \citet{z07}, the second part of the results - the dating procedure and the illustration of the densities fitted for the subsamples (divided by the breaks) - can be found in \citet{z10}.
-\\
-\begin{appendix}\label{app}
-
-\section{GL Distribution}\label{glscore}
-
-The moments of the GL distribution as given in equation \ref{eqn1} are:
-
-\begin{eqnarray}
-E(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
-Var(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
-Skew(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
-\end{eqnarray}
-
-where $\gamma'(\cdot)$ and $\gamma''(\cdot)$ are the first and the second derivatives of the digamma function, respectively. \\
-The log-likelihood is:
-
-\begin{eqnarray}
-\log f(y|\theta,\sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{1}{\sigma} (y-\theta) - (\delta+1) \nonumber \\
-& \times & \log (1+\exp^{-\frac{y-\theta}{\sigma}}) \nonumber 
-\end{eqnarray}
-
-The resulting score function, where the scores are the derivatives of the log--likelihood function, has three components $(s_{\theta}, s_{\sigma}, s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$.
-These are employed in equation \ref{eqn2} for the construction of the test statistics and are a means of measuring the estimation error in a maximum likelihood framework.
-
-\begin{eqnarray}
-s_{\theta}(y|\theta, \sigma, \delta) & =  &\frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \theta}   \\
-&=&  \frac{1}{\sigma} - (\delta+1)\cdot\frac{\frac{1}{\sigma}\tilde{y}}{(1+\tilde{y})}  \nonumber \\
-s_{\sigma}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \sigma}   \\
-& =&  -\frac{1}{\sigma} + \frac{1}{\sigma^2}(y-\theta) - (\delta+1) \times \frac{\frac{1}{\sigma^2}(y-\theta)\tilde{y}}{(1+\tilde{y})} \nonumber \\
-s_{\delta}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \delta}  \\
-& =&   \frac{1}{\delta} - \log(1+\tilde{y}) \nonumber 
-\end{eqnarray}
-
-The scores of the parameters $\phi = (\theta, \sigma, \delta)$ sum up to 0: $\sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0$. The first order condition is given by:
-
-\begin{eqnarray}
-\sum_{i=1}^n s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) = 0
-\end{eqnarray}
-
-\section{Data and Sales Periods}\label{dataprobs}
-
-In addition to Germany, Portugal, Belgium, Italy, Luxembourg and Spain, there are a number of other countries, where the sales periods are regulated by law.
-France has an individual time span regulated by each department.
-This should pose no problem, since there is no global state level solution.
-In Greece there are also two set sales--periods in winter and summer.
-In Italy the sales periods are fixed each year by the Chambers of Commerce.
-Portugal has two set sales periods as well, but for two month (January and February as well as Mid--July to Mid--September).
-The source for this information is \citet{konseur}.
-According to \cite{pltm}, the official dates of the inclusion of these sales periods are: January 1998 for Germany and Portugal, January 2000 for Belgium and January 1999 for Luxembourg and Ireland, while Italy and Spain introduced them in January 2001.
-\\
-Since X--12--ARIMA is not implemented in \proglang{R}, we used the gretl software, as described in \cite{Smith+Mixon:2006}, to desaisonalize the data without correction for outliers.
-For the results, we used the seasonally adjusted series provided by gretl using X--12--ARIMA
 
@@
 
-\clearpage
-\section{Graphs}
-
-Here we present the actual data where the blue lines in the graphs correspond to the estimated mean of the series.
-
-\begin{figure}[h!tp]
-<<itaspa, fig=TRUE, height=4, width=8>>=
-print(get_xyplot(c(19,12)))
-@
-\caption{Italy and Spain}
-\end{figure}
-
-\begin{figure}[h!tp]
-<<sweuk, fig=TRUE, height=4, width=8>>=
-print(get_xyplot(c(20,21)))
-@
-\caption{Sweden and United Kingdom}
-\end{figure}
-
-\begin{figure}[h!tp]
-<<eastern, fig=TRUE, height=10, width=8>>=
-print(get_xyplot(c(3, 5,10,17,18)))
-@
-\caption{The Eastern Countries}
-\end{figure}
-
-\begin{figure}[h!tp]
-<<ireland, fig=TRUE, height=4, width=8>>=
-x <- results$Ireland
-plot(x$data, xlab = "Time", ylab = "100 * diff(log(HICP))", main = "Ireland")
-lines(x$breakpoints, breaks = x$nbreaks)
-lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
-@
-\caption{Ireland}
-\end{figure}
-
-\begin{figure}[h!tp]
-<<nochange, fig=TRUE, height=6, width=8>>=
-print(get_xyplot(c(6,7,9,14)))
-@
-\caption{Nochange Countries}
-\end{figure}
-
-\begin{figure}[h!tp]
-<<nopattern, fig=TRUE, height=8, width=8>>=
-print(get_xyplot(c(1,2,4,8,13, 16)))
-@
-\caption{Countries without discernible pattern}
-\end{figure}
-
-\clearpage
-\bibliography{papers}
-
 \end{document}

### 2011-05-24 pkg/paper/Paper/draft.Rnw
Rewrote "Data" section. Seasonal adjustment is now presented as more of a completely natural step (as opposed to a fix of problems in the data). The appendix can probably also be omitted.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@98 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-Ever since the beginning of the European Union, the topic of a common currency was a controversial issue.
-Although the Economic and Monetary Union (EMU) is now a fact, the discussion about the economic effects of the euro is far from being settled.
-The controversial topics range from the question of whether or not the eurozone is indeed an optimal currency area (OCA) all the way to the very survival of the euro in light of the budgetary problems of some of its member states.
-The effects of monetary unions on a number of macroeconomic indicators, with inflation being the most important one in this context, is in the center of an ongoing debate: this regards the question about short-run and steady-state inflation uncertainty (in context to inflation expectations) dealt with in \citet{cap}, or the structural convergence of the inflation rates in EU countries, which is the topic of an investigation by \citet{palomba}.
+Ever since the beginning of the European Union, the topic of a common currency
+was a controversial issue. Although the Economic and Monetary Union (EMU) is now
+a fact, the discussion about the economic effects of the euro is far from being
+settled. The controversial topics range from the question of whether or not the
+eurozone is indeed an optimal currency area (OCA) \fixme{Z: include a citation for
+OCA or is this completely common knowledge?} all the way to the very
+survival of the euro in light of the budgetary problems of some of its member
+states. The effects of monetary unions on a number of macroeconomic indicators,
+with inflation being the most important one in this context, is in the center of
+an ongoing debate: this regards the question about short-run and steady-state
+inflation uncertainty (in context to inflation expectations) dealt with in
+\citet{cap}, or the structural convergence of the inflation rates in EU
+countries, which is the topic of an investigation by \citet{palomba}.
 
-The way towards the EMU basically consisted of 3 stages: stage I (1990--1994) as a phase of liberalization, stage II (1995-1998) a phase of convergence and stage III (1999-2001) the transition period, which ended with the introduction of the euro as legal tender.
-A possible fourth stage is the continuing integration of new member states, mostly former communist countries.
-Any new state that wants to join the EMU is first obliged to fulfill the Maastricht criteria, which, besides setting rules for government debts and interest rates, also requires the participation in the ERM~II (Exchange Rate Mechanism~II) for two years, while the exchange rates towards the euro are not allowed to cross the nominal band.
+The way towards the EMU basically consisted of three stages: stage~I
+(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
+convergence and stage~III (1999--2001) the transition period, which ended with
+the introduction of the euro as legal tender. A possible fourth stage is the
+continuing integration of new member states, mostly former communist countries.
+Any new state that wants to join the EMU is first obliged to fulfill the
+Maastricht criteria, which, besides setting rules for government debts and
+interest rates, also requires the participation in the ERM~II (Exchange Rate
+Mechanism~II) for two years, while the exchange rates towards the euro are not
+allowed to cross the nominal band.
 
-\cite{emerson} put emphasize on the fact that a high inflation rate is also more variable and uncertain and thus causes more relative price variability, leading to a less efficient price mechanism.
-Failing to stabilize inflation leads to severe economic problems.
-Throughout the literature, there is still a considerable degree of uncertainty as to what extent the introduction of the euro, or monetary unions in general, affect the inflation rate.
-The EMU requires a country to possibly alter a number of economic policies and institutions, which may demand for significant short term adjustments to take place.
+\cite{emerson} put emphasize on the fact that a high inflation rate is also more
+variable and uncertain and thus causes more relative price variability, leading
+to a less efficient price mechanism.  Failing to stabilize inflation leads to
+severe economic problems. Throughout the literature, there is still a
+considerable degree of uncertainty as to what extent the introduction of the
+euro, or monetary unions in general, affect the inflation rate. The EMU requires
+a country to possibly alter a number of economic policies and institutions,
+which may demand for significant short term adjustments to take place.
 
-Given that a country experienced quite volatile inflation rates, its efforts to meet the convergence criteria are likely to lead to an alteration at least in the mean (since this was one of the requirements) and possibly in the volatility of their respective inflation rates as well.
+Given that a country experienced quite volatile inflation rates, its efforts to
+meet the convergence criteria are likely to lead to an alteration at least in
+the mean (since this was one of the requirements) and possibly in the volatility
+of their respective inflation rates as well.
 
-However, a deeper inquiry into this problem is in need of a method able to capture changes in inflation dynamics.
-We contribute to this debate by developing a new method for testing for structural breaks in the dynamics of the inflation rates and thereby upon the effect of the EMU on a number of European countries both within and without the eurozone.
-The framework developed in this paper makes it possible to test for any significant changes of a countries inflation dynamics and thus helps interpreting the effects of the EMU.
-We focus on changes in the first three moments of these inflation rates, which we assume to be distributed according to a generalized logistic distribution (GL).
-The GL distribution is a suitable one due to its flexibility to capture nonsymmetry and because of its fatter tails.
-For this purpose, we develop a new estimation approach building on previous research by \citet{z07}.
+However, a deeper inquiry into this problem is in need of a method able to
+capture changes in inflation dynamics. We contribute to this debate by
+developing a new method for testing for structural breaks in the dynamics of the
+inflation rates and thereby upon the effect of the EMU on a number of European
+countries both within and outside the eurozone. The framework developed in this
+paper makes it possible to test for any significant changes of a countries
+inflation dynamics and thus helps interpreting the effects of the EMU. We focus
+on changes in the first three moments of these inflation rates, which we assume
+to be distributed according to a generalized logistic distribution (GL). The GL
+distribution is a suitable one due to its flexibility to capture nonsymmetry and
+because of its fatter tails. For this purpose, we develop a new estimation
+approach building on previous research by \citet{z07}.
 
-This \fixme{Jesus: sell method better, why this one and not another} framework allows us to assess the evolution of mean, variance and skewness over time.
-The skewness parameter turns out to be quite important to measure changes in the long run inflation dynamics, as it changes considerably over the time of the analysis.
-This trend would be all but invisible using a normal distribution.
-Our estimation technique enables the simultaneous estimation of breaks in the mean (which is important in the context of the Maastricht criteria), the variance (which should be low for the price mechanism to be efficient) and the skewness (where a change in skewness is evidence of a shift of the distribution towards a new ``center of gravity'') of the inflation rates of interest.
-This than can be used to search for answers regarding the question about the effect of the EMU on the inflation dynamics of its member states or more general to investigate the effects of external shocks.
+This \fixme{Jesus: sell method better, why this one and not another} framework
+allows us to assess the evolution of mean, variance and skewness over time. The
+skewness parameter turns out to be quite important to measure changes in the
+long run inflation dynamics, as it changes considerably over the time of the
+analysis. This trend would be all but invisible using a normal distribution.  
+Our estimation technique enables the simultaneous estimation of breaks in the
+mean (which is important in the context of the Maastricht criteria), the
+variance (which should be low for the price mechanism to be efficient) and the
+skewness (where a change in skewness is evidence of a shift of the distribution
+towards a new ``center of gravity'') of the inflation rates of interest. This
+than can be used to search for answers regarding the question about the effect
+of the EMU on the inflation dynamics of its member states or more general to
+investigate the effects of external shocks.
 
-The \fixme{Jesus: maybe comment on results here} remainder of this paper is structured as follows: Section~\ref{sec:data} presents the data, Section~\ref{sec:model} presents the model and the estimation techniques used, Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an example, Section~\ref{sec:results} presents the results and Section~\ref{sec:conclusion} concludes.
+The \fixme{Jesus: maybe comment on results here} remainder of this paper is
+structured as follows: Section~\ref{sec:data} presents the data,
+Section~\ref{sec:model} presents the model and the estimation techniques used,
+Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an
+example, Section~\ref{sec:results} presents the results and
+Section~\ref{sec:conclusion} concludes.
 
@@
 
-We \fixme{Jesus: yearly inflation rates better; don't start with data in paper}
-employ seasonally unadjusted HICP series of 21 countries ranging from January
-1990 to December 2010 that are obtained from the \citet{oecd}. Countries
-included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,
-Finland,	France,	Germany,	Greece,	Hungary,	Ireland,
-Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia,
-Slovenia,  Spain,	Sweden, and	United Kingdom. Latvia and Lithuania as
-well as Bulgaria and Romania are excluded due to data scarcity. Cypria and Malta
-are not included since they are very small economies.
+All empirical analyses are based on seasonally adjusted inflation rates (in percent)
+for \Sexpr{ncol(hicps)}~countries
+(\Sexpr{gsub("UK", "United Kingdom", gsub("CzechRep", "Czech Rep", paste(colnames(hicps), collapse = ", "), fixed = TRUE), fixed = TRUE)})
+in the time period from the early or mid-1990s to the end of 2010.
+The original data source are seasonally unadjusted HICP series provided by the
+\citet{oecd} from January~1990 (if available) to December~2010, that were transformed
+to inflation rates using log-returns and subsequent seasonal adjustment
+via X-12-ARIMA \citep{Findley+Monsell+Bell:1998}.
 
-The countries in this sample can be divided into three different groups. First
-the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece,
-Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and
-Spain).  Second, the EU members not participating in the ERM~II: Czech Republic,
-Hungary, Poland, United Kingdom, and Sweden. Denmark stands on its own as a
-member of the EU and the ERM~II, but not yet a member of the EMU.
+The countries in this sample can be divided into three different groups: (1)~euro
+countries -- Austria, Belgium, Estonia, France, Finland, Germany, Greece,
+Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia,
+Spain; (2)~EU members not participating in the ERM~II -- Czech Republic,
+Hungary, Poland, United Kingdom, Sweden; (3)~Denmark which stands on its own as a
+member of the EU and the ERM~II, but not yet of the EMU. 
+Latvia, Lithuania, Bulgaria, Romania are excluded due to data scarcity.
+Cyprus and Malta are not included since they are very small economies.
 
-As  shown in a paper by \cite{pltm}, the harmonization of the treatment of sales
-prices led to considerable changes in some countries (Germany, Portugal,
-Ireland, Luxembourg, Belgium, Italy and Spain); failing to take this into
-account may lead to considerable distortions in the estimation. We dealt with
-the inclusion of sales periods\footnote{This was demanded by a Commission
-regulation, details may be found in: \citet{eurlex} and in the
-Appendix~\ref{sec:sales}.} and all the other seasonality in the data by using
-the X-12-ARIMA method for desaisonalization, as described in \cite{X-12-ARIMA}
-and \cite{Findley+Monsell+Bell:1998}.
+Seasonal adjustment is necessary due to the harmonization of the treatment
+of sales prices\footnote{Inclusion of sales prices in the HICP was demanded by
+a Commission regulation \citep[see][for details]{eurlex}.} in the HICP, resulting in
+changes of the seasonal inflation patters over time within several countries
+and also between the countries. See \cite{pltm} for a detailed discussion.
 

### 2011-05-24 pkg/paper/Paper/draft.Rnw
rewrote intro to modeling section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@99 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-To monitor changes in the dynamics of the inflation rates, we follow the usual
-strategy employed when testing for structural breaks. First, the model is fitted
-for the whole sample using the generalized logistic distribution. This then
-enables testing for the stability of the parameter estimates. If there is
-evidence of significant changes, the number and the timing of the break points
-are estimated.
+The goal of the modeling strategy, that is subsequently introduced, is the
+following: detection of structural changes in the distributional properties of
+the inflation rates over time, e.g., as potentially caused by interventions.
+However, unlike many standard least squares approaches \citep[such as][]{bp03},
+not only changes in the mean level should be addressed but also changes in
+variance and potential skewness.
 
-This framework allows to trace the evolution of mean, variance and skewness of
-the inflation series and thus interpreting deviations in any of these parameters
-as possibly due to a change in monetary policy, is made feasible. Any huge
-intervention or change is captured by fluctuations in the parameter stability of
-the series and can be quantified and related to a distinct time frame.
+Hence, a likelihood-based model is adopted that can also incorporate stylized
+facts for return data, such as heavy tails and skewness. Here, the seasonally
+adjusted returns $y_t$ ($t = 1, \dots, n$) are assumed to come from a
+generalized logistic (GL) distribution \citep[see][]{johnson} with
+three-dimensional parameter $\phi_t = (\theta_t, \sigma_t, \delta_t)^\top$ (for
+location~$\theta_t$, scale~$\sigma_t$, and shape~$\delta_t$ at time $t$). Structural change
+techniques are then employed to check whether the parameter vector $\phi_t$
+remains constant over time and, if that is not the case, when and how the
+parameters change. This framework allows to trace breaks in the
+evolution of mean, variance and skewness of the inflation series which may
+possibly be linked to underlying regime changes, e.g., in monetary policy.
 
-The correlation \fixme{Jesus: yes, but a DGP needs to be assumed, it would need
-to take care of any autororrelation, wouldn't it?} dynamics of the series is of
-less importance to this framework, which concentrates on changes in mean,
-variance and skewness and does not focus on autocorrelation, the use of
-alternative tests, as described in \cite{andreou}, is also hindered by the low
-number of observations, due to the monthly frequencies.\footnote{In some years
-the Billion Prices Project of the \cite{bpp}, which provides a much higher data
-frequency, could solve this problem.}
+For estimation of the parameters, a (quasi-)maximum likelihood framework is
+adopted and the observations are assumed to be (approximately) independent,
+i.e., potential autocorrelation is treated as a nuisance parameter and not
+explicitly incorporated in the model. The motivation for this is to focus on the
+shifts in mean, variance, and skewness -- and it turns out to be a useful model
+for the data under investigation because only negligible amounts of residual
+autocorrelation remain after incorporation of the structural breaks. If the
+focus of the analysis were the dynamics in the autocorrelation, higher frequency
+data would be required, e.g., in order to apply techniques such as those
+expounded by \cite{andreou}.\footnote{For example, it would be conceivable to
+obtain high-frequency data for certain years from the Billion Prices Project of
+the \cite{bpp}; however, this is not pursued here.}
+
+\fixme{Z: potentially say something about the subsequent subsections}
 

### 2011-05-25 pkg/paper/Paper/draft.Rnw
Rewrote modeling section, merging it with the GL appendix.
The intro to the results section has also been enhanced.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@100 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 because of its fatter tails. For this purpose, we develop a new estimation
-approach building on previous research by \citet{z07}.
+approach building on previous research by \citet{Zeileis+Hornik:2007}.
 
@@
 facts for return data, such as heavy tails and skewness. Here, the seasonally
-adjusted returns $y_t$ ($t = 1, \dots, n$) are assumed to come from a
+adjusted returns $y_i$ ($i = 1, \dots, n$) are assumed to come from a
 generalized logistic (GL) distribution \citep[see][]{johnson} with
-three-dimensional parameter $\phi_t = (\theta_t, \sigma_t, \delta_t)^\top$ (for
-location~$\theta_t$, scale~$\sigma_t$, and shape~$\delta_t$ at time $t$). Structural change
-techniques are then employed to check whether the parameter vector $\phi_t$
+three-dimensional parameter $\phi_i = (\theta_i, \sigma_i, \delta_i)^\top$ for
+location~$\theta_i$, scale~$\sigma_i$, and shape~$\delta_i$ at time $i$. Structural change
+techniques are then employed to check whether the parameter vector $\phi_i$
 remains constant over time and, if that is not the case, when and how the
@@
 
-\fixme{Z: potentially say something about the subsequent subsections}
+In the following, details are provided about (1)~the GL distribution and its properties,
+(2)~how it can be estimated (under parameter stability), (3)~how parameter stability
+can be assesed, and (4)~how breakpoints can be estimated in the presence of parameter
+instability.
 
@@
 
-The data suggested that the use of a normal distribution might cause some
-estimation problems, since most of the time series exhibit asymmetric properties
-(with sometimes rather strong skewness) and -- compared to the normal
-distribution -- a fatter tail. To accommodate for this fact, the use of the
-generalized logistic distribution (GL) with its relative low degree of
-complexity, seemed appropriate. On the one hand, it is a simple distribution and
-on the other hand, it is flexible enough to capture essential properties of the
-data. For the inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we
-therefore assume a GL distribution as defined in \citet{johnson}:
+The logistic distribution is often used in econometrics in the
+context of income and growth models due to its fatter tails compared to the 
+normal distribution -- it has also been applied to the modeling of
+inflation rates by \citet{batchelor}. Its generalization -- which additionally
+allows for asymmetries -- has been applied to analyze risk \citep[e.g., by][]{tolikas}
+but has not yet been applied -- to the best of our knowledge -- to inflation rates.
+However, it will prove to be a rather simple model that fits HICP inflation rates
+quite well.
 
+The type~I GL distribution (also known as skew-logistic distribution), as defined in \citet{johnson},
+has the log-density function
+%
 \begin{eqnarray}
-f(y | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma} \cdot \exp^{-\frac{y-\theta}{\sigma}}}{(1+\exp^{-\frac{y-\theta}{\sigma}})^{(\delta+1)}} 
+%% f(y ~|~ \theta, \sigma, \delta) & = &
+%% \frac{\frac{\delta}{\sigma} \cdot \exp\{-\frac{y - \theta}{\sigma}\}}{(1+\exp\{-\frac{y - \theta}{\sigma}\})^{(\delta+1)}} 
+\ell(y ~|~ \theta, \sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{y - \theta}{\sigma} -
+                                         (\delta + 1) \cdot \log\left(1 + \exp\left\{-\frac{y - \theta}{\sigma}\right\}\right)
 \label{eq:dglogis}
 \end{eqnarray}
+%
+with location~$\theta$, scale~$\sigma > 0$ and shape~$\delta > 0$. For $\delta = 1$ the
+distribution simplifies to the logistic distribution, for $\delta < 1$ or $> 1$ it is
+left- or right-skewed, respectively. The corresponding first three moments are
+%
+\begin{eqnarray}
+\mathrm{E}(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
+\mathrm{Var}(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
+\mathrm{Skew}(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
+\end{eqnarray}
+%
+where $\gamma(\cdot)$, $\gamma'(\cdot)$, and $\gamma''(\cdot)$ are the digamma function and its
+first and second derivative, respectively.
 
-with location $\theta$, scale $\sigma$ and shape $\delta$. For $\delta=1$ the
-distribution simplifies to the logistic distribution, for $\delta<1$ it is
-skewed to the left and for $\delta$$>$1 it is skewed to the right (see
-Appendix~\ref{sec:gld2} for more details).
-
-In econometrics, the logistic distribution is often used in income distributions
-and growth models. This is due to its fatter tails, which fit these kind of
-data somewhat better. \citet{won} develop a maximum likelihood \fixme{Jesus: is
-this not better than what you do here: answer: no, they use completely other
-framework} estimator if the underlying distribution is a GL distribution. A
-similar GL distribution is also used in \citet{tolikas} who analyze extreme risk
-and value-at-risk in the German stock market, although they do not use a shape
-parameter. Regarding inflation rates, the GL distribution is -- to our best
-knowledge -- only used in context with expected inflation.
-
-\citet{batchelor} use a logistic distribution (not its generalization) to model
-the distribution of mean expected inflation rates.
+The corresponding score function, i.e., the derivative of the log-density with respect
+to the parameter vector, is given by
+%
+\begin{equation}
+s(y ~|~ \phi) ~=~ \frac{\partial \ell(y ~|~ \phi)}{\partial \phi}  ~=~ \left(
+  \begin{gathered}
+    \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \\
+    \left\{ \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \right\} \cdot \frac{y - \theta}{\sigma} -\frac{1}{\sigma} \\
+    \frac{1}{\delta} - \log(1 + \tilde{y})
+  \end{gathered} \right)
+\end{equation}
+%
+where $\phi = (\theta, \sigma, \delta)^\top$ and
+$\tilde{y} = \exp\{- (y - \theta)/\sigma\}$.
 
@@
 
-Under the assumption that $y_t$ is distributed according to the generalized
-logistic distribution, we test for the hypothesis that the parameter vector
-$\hat{\phi}$ stays stable over time:
+Under the assumption that $y_i \sim \mathit{GL}(\theta, \sigma, \delta)$
+for $i = 1, \dots, n$ independently -- i.e., are independent realizations from
+a GL distribution with parameter vector $\phi = (\theta, \sigma, \delta)^\top$ --
+the parameters can be estimated as usual by maximum likelihood (ML):
+$\hat \phi ~=~ \argmax_\phi \sum_{i = 1}^n \ell(y_i | \phi)$. The corresponding
+first order condition is $\sum_{i = 1}^n s(y_i | \hat \phi) = 0$.
 
-\begin{eqnarray}
-& H_0: \phi_t = \phi_0 \quad (t=1,...,n) \nonumber 
-\end{eqnarray}
+To guard the inference against potential misspecification, e.g., autocorrelation or
+misspecification of higher moments of the distribution, one can treat $\hat \phi$
+as the quasi-maximum-likelihood (QML) estimator and adjust the inference by using
+heteroskedasticity and autocorrelation consistent (HAC) covariance matrix
+estimators \citep[see e.g.,][]{Andrews:1991}.
 
-against the $H_1$ of changes in at least one of the parameters. We first
-estimate the parameters of the GL-distribution by means of maximum likelihood
-$\hat{\phi} ~=~ \argmax_\phi \sum_{t=1}^n \log f(y_t|\phi)$. The parameter
-stability is assessed using the empirical scores $s(y_t|\hat{\phi})$ as measures
-of model deviation.\footnote{for details see Appendix~\ref{sec:gld2} and
-\citet{z07}}  The empirical fluctuation process $\mathit{efp}(\cdot)$ captures
-deviations from a hypothesized zero mean over time and is given by
+If the parameters are potentially varying over time --
+$y_i \sim \mathit{GL}(\theta_i, \sigma_i, \delta_i)$ -- then $\hat \phi$ is the (Q)ML
+estimator under the null hypothesis of parameter stability
 %
-\begin{eqnarray}
-\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
-\label{eq:efp}
-\end{eqnarray}
+\begin{equation} \label{eq:h0}
+  H_0:~ \phi_i ~=~ \phi_0 \quad (i = 1, \dots, n)
+\end{equation}
 %
-where $\hat{V}$ is some consistent estimator of the variance of the scores,
-which is used to decorrelate the series as to avoid problems due to serial
-correlation. Then a functional central limit theorem (FCLT) for
-$\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian
-bridge: $\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot)$.
+which should be tested against the alternative $H_1$ that at least one of the
+parameters changes over time.
 
-\subsection{Test}
+\subsection{Test for structural change}
 
-After the $\mathit{efp}(\cdot)$ is computed, it conveys information about
-departures from the $H_0$ of parameter stability over time. For a precise
-quantification of the actual deviation from its hypothesized zero-mean path, the
-main idea of the test statistics presented here is to aggregate the
-$\mathit{efp}(\cdot)$ in such a way as to be able to construct a test statistics
-so that critical values and p values can be derived. In this paper, we employ
-the sup$\mathit{LM}$ test of \citet{andrews}. This test is well suited for
-single break alternatives and performs better if more than just one of the
-distribution parameters change.
+The null hypothesis of parameter stability (\ref{eq:h0}) can be assessed using the empirical
+scores $s(y_i | \hat \phi)$ as measures of model deviation \citep{Zeileis+Hornik:2007} from the
+model fit under $H_0$. Systematic deviations over time from the full sample estimates $\hat \phi$
+can then be captured by cumulative sums of the empirical scores in an empirical fluctuation
+process $\mathit{efp}(t)$:
+%
+\begin{equation} \label{eq:efp}
+\mathit{efp}(t) ~=~ \hat V^{-1/2} n^{-1/2} \sum_{i = 1}^{\lfloor nt \rfloor}
+  s(y_i | \hat \phi) \quad (0 \leq t \leq 1),
+\end{equation}
+%
+where $\hat V$ is some consistent estimator of the variance of the scores. Below,
+we use the outer product of gradients $s(y_i | \hat \phi)$ to estimate $\hat V$.
+Alternatively, HAC estimators could be used leading to the same qualitative
+results for the data under investigation.
 
-It is given by:
+As usual in structural change analysis, a functional central limit theorem holds for
+$\mathit{efp}(\cdot)$, which converges to a 3-dimensional Brownian
+bridge: $\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot)$. Based on
+this various types of test statistics can be computed
+(see \citealp{Zeileis:2005} for details). In the following, the 
+sup$\mathit{LM}$ test of \cite{Andrews:1993} is employed, which performs particularly
+well for single shift alternatives and if several of the three distribution
+parameters change simultaneously. The test statistic with 10\% trimming is given by
 %
@@
 %
-where the appropriate $p$~values can be computed using from the corresponding
+and the appropriate $p$~values can be computed using from the corresponding
 limiting distribution -- with $\mathit{efp}(t)$ replaced by $W^0(t)$.
-The sup$\mathit{LM}$ test computes a test statistics for all possible change points in a
-given time frame and rejects the null of no change if the maximum is too high.
 
@@
 
-The dating of structural breaks builds upon the work of \cite{bp03}. If a break
-is detected via the sup$\mathit{LM}$ test, the next step is estimating the number and the
-timing of these break points. Breakpoints $\tau_1, \dots,\tau_B$ are estimated
-via maximization of the full segmented likelihood
+If the null hypothesis of parameter stability (\ref{eq:h0}) is rejected, a natural
+strategy is to assume that there are $B$ breakpoints with stable parameters within
+each of the resulting segments. \cite{bp03} have established a rigorous inference
+framework in this situation for least squares estimation which has been extended by
+\cite{Zeileis+Shah+Patnaik:2010} to (Q)ML estimation. Here, we follow the same ideas
+and maximize the full segmented log-likelihood
 %
-\begin{eqnarray}
-\sum_{b=1}^{B+1} \sum_{t=\tau_b + 1}^{\tau_b} \log f(y_t | \phi^{(b)}) \nonumber
-\end{eqnarray}
+\begin{equation*}
+  \sum_{b = 1}^{B+1} \sum_{i = \tau_b + 1}^{\tau_b} \ell(y_i ~|~ \phi^{(b)})
+\end{equation*}
 %
-If the test statistics is significant (in this case the sup$\mathit{LM}$ test), at least
-one break is estimated. In principle, the optimal segmentation (i.e. the
-concrete dates of a break) can be computed once B (the number of breakpoints) is
-known. However, B needs to be based on the observed data as well. The solution
-is to compute the timing of the breaks for a number of different values for B
-and then choose B by means of an information criterion, in this case the LWZ
-(Liu, Wu and Zidek) as developed in \citet{liu}. All parameters $\tau_1, \dots,
-\tau_B, \phi^{(1)}, \dots, \phi^{(B+1)}$ can be estimated jointly using dynamic
-programming.\footnote{More details for this estimation technique are provided in
-\cite{z10}.}
+for joint estimation of the breakpoints $\tau_1, \dots, \tau_B$ and the segment-specific
+GL parameters $\phi^{(b)}$ ($b = 1, \dots, B + 1$). Following the recommendations
+of \cite{bp03}, a modified Bayes information criterion \citep[LWZ, proposed by][]{liu}
+is employed selecting the number of breakpoints $B$.
 
@@
 x <- results$Slovenia
-spval <- sctest(x$efp, functional = supLM(0.1))$p.value
+x$efp2 <- gefp(x$model0, fit = NULL, vcov = kernHAC)
+spval <- c(sctest(x$efp, functional = supLM(0.1))$p.value,
+  sctest(x$efp2, functional = supLM(0.1))$p.value)
+spval <- ifelse(spval < 0.001, "$< 0.001$", format(round(spval, digits = 4), nsmall = 4))
 @
 
-To illustrate our method, we consider the case of Slovenia. After a period of
+To illustrate the proposed strategy, the case of Slovenia is considered. After a period of
 relative instability following the independence from Yugoslavia, Slovenia was
 successfull in realigning its economy and introduced a number of reforms leading
-to a stable growth in the past years. The good economic performance made it
-possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce
-the euro in January 2007. The data collected contains the period from January
-1996 onwards. In the first few years, Slovenia experienced very high inflation
-rates, but was successfull in reducing inflation to a much lower level.
-
-A glance at the time series depicted in Figure~\ref{fig:slovenia1} of the monthly inflation
-rates reveals a decline in the mean around the middle of the observed time
-period. The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches
-the peak somewhere between 2003 and 2004. The $p$~value of the sup$\mathit{LM}$ test is
-\Sexpr{round(spval, 8)} and clearly indicates at least one break.  The number of
-breaks is calculated by means of the LWZ criteria, which chooses one break in
-August 2003. The goodness-of-fit test depicted in Figure~\ref{fig:slovenia2} confirms
-the choice of the GL distribution. The mean inflation rate was cut to only a
-third of its previous value, whereas variance increased slightly and the
-distribution slightly skewed to the left, but peaking at a lower value than
-previously.
-
-In this year, Slovenia introduced a number of financial reforms
-and severely decreased the year on year growth in the M3 monetary aggregate.
-Combined, these actions were enough to ensure the successfull participation in
-the ERM~II a year later and the introduction of the euro two years later.
+to a stable growth in recent years. The good economic performance made it
+possible for Slovenia to enter the ERM~II in June~2004 and later on to introduce
+the euro in January~2007. The seasonally adjusted inflation series for Slovenia encompasses
+the time from \Sexpr{start(x$data)} to \Sexpr{end(x$data)} and is depicted in the left
+panel of Figure~\ref{fig:slovenia1}.
 
@@
 
+Assuming a single stable set of parmaters (\ref{eq:h0}) is clearly not valid, as the
+sequence of $\mathit{LM}$~statistics (see right panel of Figure~\ref{fig:slovenia1}) clearly
+exceeds its 5\% critical value (red horizontal line), leading to a highly significant
+$p$-value \Sexpr{spval[1]}. If a quadratic spectral kernal HAC estimator \citep{Andrews:1991}
+would have been used to adjust (\ref{eq:efp}), the resulting $p$-value would be somewhat
+larger with \Sexpr{spval[2]} but still clearly significant.
+
+As there is evidence for at least one break, the LWZ criterion is employed for $B = 1, 2, \dots$
+breakpoints with a minimal segment size of two years per segment. It assumes its minimum for $B = 1$
+associated with the breakpoint \Sexpr{breakdates(x$breakpoints, breaks = x$nbreaks)}, depcited
+by a vertical dashed line in both panels of Figure~\ref{fig:slovenia1}. The resulting segmented
+fitted mean (horizontal blue lines in the left panel Figure~\ref{fig:slovenia1}) shows that
+in before the break Slovenia experienced very high inflation rates but was successful in reducing
+inflation to a much lower level afterwards.
+
+This decrease in inflation is also conveyed by Figure~\ref{fig:slovenia2} that shows the
+observed histograms of inflation rates before and after the breakpoint along with the
+fitted GL distribution. This highlights that along with the decrease in mean level, the variance increases
+somewhat and -- probably more interestingly -- the distribution changes from being right-skewed to
+slightly left-skewed. This confirms that very high inflation rates that occurred in several
+months before the breakpoint could be avoided afterwards. Furthermore, Figure~\ref{fig:slovenia2}
+shows that the GL distribution fits the observed data very well, both before and after the
+breakpoint, with $p$~values from $\chi^2$ goodness-of-fit tests of
+\Sexpr{format(round(summary(x$model[[1]])$chisq.test$p.value, digits = 3), nsmall = 3)} and
+\Sexpr{format(round(summary(x$model[[2]])$chisq.test$p.value, digits = 3), nsmall = 3)}, respectively.
+
+The estimated breakpoint can be matched very well with the timing of financial reforms
+in Slovenia that severely decreased the annual growth in the M3 monetary aggregate.
+These success of these actions ensured the Slovenian participation in
+the ERM~II only a year later and the introduction of the euro two years later.
+
 \clearpage
@@
 
-\readme{Jesus: include p-values in table; no place left; Jesus: opening seems a
-bit confusing and obscure}
+The modeling strategy proposed in Section~\ref{sec:model} is applied to each of the
+countries included in the data set introduced in Section~\ref{sec:data}.
+Table~\ref{tab:results1} reports the estimated breakpoints along with the segment-specific
+moments (mean, variance, skewness) implied by the corresponding parameter
+estimates. Furthermore, entry dates for ERM/ERM~II and the euro are included as
+additional information. For ease of interpretation, countries with similar patterns of
+change are grouped together in Table~\ref{tab:results1} (and ordered alphabetically
+within each group).
 
-Table~\ref{tab:results1} presents the break dates and the parameter values of
-the segments as well as the dates of entry into the ERM/ERM~II and the date of
-the euro introduction.
+If there is only a single segment, the sup$\mathit{LM}$ test was non-significant at 5\%
+level. In significant cases, the number of breakpoints was selected via the LWZ criterion
+and subsequently all parameters (breakpoints and segment-specific GL parameters) were
+estimated by ML. No HAC~correction was employed as there was only negligible autocorrelation
+in the empirical scores after inclusion of the breakpoints (if any).
 
@@
 
-\begin{table}[t!]
+\begin{table}[p!]
 \begin{center}
@@
 
-The first group that draws particular attention is a subgroup of the PIIGS
-(Portugal, Ireland, Italy, Greece and Spain) countries consisting of Ireland and
+\fixme{Z: I have reordered the groups in Table~1 which tries to be roughly in the order
+of joining EU/ERM/EMU. Probably the following interpretations should start with a
+very short overview of the groupings.}
+
+The first group that draws particular attention is a subgroup of the PIIGS countries
+(Portugal, Ireland, Italy, Greece, Spain) consisting of Ireland and
 Spain. Both show a break in the summer of 2008 following the financial crisis
@@
 Another interesting group consists -- with the exception of Austria and
-Luxembourg -- of the countries of the former DM-zone. All these countries have a
+Luxembourg -- of the countries of the former DEM-zone. All these countries have a
 very low inflation rate in both mean and variance and their inflation dynamics
@@
 
-In the core-zone, we see a decrease in the lower quartil from \Sexpr{yrs[1]} to \Sexpr{yrs[4]}, with no change in the median inflation rate, which is due to the fact, that there is no break after \Sexpr{yrs[2]} for these countries; however, the interquartile range decreased by almost 20\%.
-The variance increased slightly (but with a 15\% decrease in interquartile range), whereas skewness stayed symmetric in the median, with almost no change in the interquartile range.
+In the core-zone, we see a decrease in the lower quartile from \Sexpr{yrs[1]} to
+\Sexpr{yrs[4]}, with no change in the median inflation rate, which is due to the
+fact, that there is no break after \Sexpr{yrs[2]} for these countries; however,
+the interquartile range decreased by almost 20\%. The variance increased
+slightly (but with a 15\% decrease in interquartile range), whereas skewness
+stayed symmetric in the median, with almost no change in the interquartile
+range.
 
-The other countries were successfull in decreasing all three moments, but they are still higher than the euro-zone countries.
-Most importantly, the interquartile range for all moments decreased considerably showing a clear convergence towards the core-zone countries.
-The PIIGS countries decreased inflation rates as well with a strong change of skewness towards more negative values.
-The interquartile range of their inflation variance increased considerably, which is mostly due to Spain and Ireland and their strong reaction to the financial crisis.
+The other countries were successfull in decreasing all three moments, but they
+are still higher than the euro-zone countries. Most importantly, the
+interquartile range for all moments decreased considerably showing a clear
+convergence towards the core-zone countries.  The PIIGS countries decreased
+inflation rates as well with a strong change of skewness towards more negative
+values. The interquartile range of their inflation variance increased
+considerably, which is mostly due to Spain and Ireland and their strong reaction
+to the financial crisis.
 
@@
 \pkg{strucchange}~\Sexpr{packageDescription("strucchange")["Version"]} \citep{Zeileis+Leisch+Hornik:2002}, and
-\pkg{fxregime}~\Sexpr{packageDescription("fxregime")["Version"]} \citep{z10},
+\pkg{fxregime}~\Sexpr{packageDescription("fxregime")["Version"]} \citep{Zeileis+Shah+Patnaik:2010},
 all of which are available under the General Public License (GPL) from the
@@
 \begin{appendix}
-
-\section{GL distribution}\label{sec:gld2}
-
-The moments of the GL distribution as given in Equation~\ref{eq:dglogis} are
-%
-\begin{eqnarray}
-\mathrm{E}(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
-\mathrm{Var}(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
-\mathrm{Skew}(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
-\end{eqnarray}
-%
-where $\gamma'(\cdot)$ and $\gamma''(\cdot)$ are the first and the second
-derivatives of the digamma function, respectively. The log-likelihood is
-%
-\begin{eqnarray}
-\log f(y|\theta,\sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{1}{\sigma} (y-\theta) - (\delta+1) \nonumber \\
-& & \times \log (1+\exp^{-\frac{y-\theta}{\sigma}}) \nonumber 
-\end{eqnarray}
-%
-The resulting score function, where the scores are the derivatives of the
-log-likelihood function, has three components $(s_{\theta}, s_{\sigma},
-s_{\delta})$, with $\tilde{y} = \exp^{-\frac{y-\theta}{\sigma}}$. These are
-employed in Equation~\ref{eq:efp} for the construction of the test statistics
-and are a means of measuring the estimation error in a maximum likelihood
-framework.
-
-\begin{eqnarray}
-s_{\theta}(y|\theta, \sigma, \delta) & =  &\frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \theta}   \\
-&=&  \frac{1}{\sigma} - (\delta+1)\cdot\frac{\frac{1}{\sigma}\tilde{y}}{(1+\tilde{y})}  \nonumber \\
-s_{\sigma}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \sigma}   \\
-& =&  -\frac{1}{\sigma} + \frac{1}{\sigma^2}(y-\theta) - (\delta+1) \times \frac{\frac{1}{\sigma^2}(y-\theta)\tilde{y}}{(1+\tilde{y})} \nonumber \\
-s_{\delta}(y|\theta, \sigma, \delta) & = & \frac{\delta \log f(y|\theta,\sigma, \delta)}{\delta \delta}  \\
-& =&   \frac{1}{\delta} - \log(1+\tilde{y}) \nonumber 
-\end{eqnarray}
-%
-The scores of the parameters $\phi = (\theta, \sigma, \delta)$ sum up to 0: $\sum_{t=1}^n s(y_t|\hat{\phi}) ~=~ 0$. The first order condition is given by:
-%
-\begin{eqnarray}
-\sum_{i=1}^n s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) = 0
-\end{eqnarray}
 

### 2011-05-25 pkg/paper/Paper/draft.Rnw
fixup in caption of Figure 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@102 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 with sequence of $\mathit{LM}$~statistics,
-critical value at 5\% level (vertical dashed line), and maximum (vertical dashed line).}
+critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}
 

### 2011-05-26 pkg/paper/Paper/draft.Rnw
updated paper and draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@103 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \Abstract{
-  text when everything else finished, present innovation of paper
+  In this paper, we test for structural breaks in the inflation dynamics of 21 European countries. To capture
+  changes in the first three moments of the inflation rates, we assumed them to be distributed according to 
+  a generalized logistic distribution, which we implemented in a new \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} 
+  \citep{R} package called \pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]}. We find breaks in all but 4 countries.
+  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a change towards negative skewness.
 }
@@
 settled. The controversial topics range from the question of whether or not the
-eurozone is indeed an optimal currency area (OCA) \fixme{Z: include a citation for
-OCA or is this completely common knowledge?} all the way to the very
-survival of the euro in light of the budgetary problems of some of its member
+eurozone is indeed an optimal currency area (as developed in: \citet{Mundell:1961})\footnote{
+For an overview, the interested reader is referred to \citet{Dellas:2009}}, 
+all the way to the very survival of the euro in light of the budgetary problems of some of its member
 states. The effects of monetary unions on a number of macroeconomic indicators,
@@
 
-The way towards the EMU basically consisted of three stages: stage~I
+\readme{perhaps skip this paragraph?} The way towards the EMU basically consisted of three stages: stage~I
 (1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
@@
 
+\clearpage
 \cite{emerson} put emphasize on the fact that a high inflation rate is also more
@@
 a country to possibly alter a number of economic policies and institutions,
-which may demand for significant short term adjustments to take place.
+which may require significant short term adjustments.
 
@@
 However, a deeper inquiry into this problem is in need of a method able to
-capture changes in inflation dynamics. We contribute to this debate by
+capture changes in inflation dynamics. We contribute to this  by
 developing a new method for testing for structural breaks in the dynamics of the
 inflation rates and thereby upon the effect of the EMU on a number of European
-countries both within and outside the eurozone. The framework developed in this
-paper makes it possible to test for any significant changes of a countries
-inflation dynamics and thus helps interpreting the effects of the EMU. We focus
-on changes in the first three moments of these inflation rates, which we assume
-to be distributed according to a generalized logistic distribution (GL). The GL
-distribution is a suitable one due to its flexibility to capture nonsymmetry and
-because of its fatter tails. For this purpose, we develop a new estimation
-approach building on previous research by \citet{Zeileis+Hornik:2007}.
+countries. The framework developed in this paper makes it possible to test for any 
+significant changes of a countries inflation dynamics and thus helps interpreting 
+the effects of the EMU. We focus on changes in the first three moments of these 
+inflation rates, which we assume to be distributed according to a generalized logistic 
+distribution (GL). The GL distribution is a suitable one due to its flexibility to 
+capture nonsymmetry and because of its fatter tails. For this purpose, we develop a 
+new estimation approach building on previous research by \citet{Zeileis+Hornik:2007}.
 
@@
 allows us to assess the evolution of mean, variance and skewness over time. The
-skewness parameter turns out to be quite important to measure changes in the
+skewness parameter turns out to be quite important to measure alterations in the
 long run inflation dynamics, as it changes considerably over the time of the
@@
 
-The type~I GL distribution (also known as skew-logistic distribution), as defined in \citet{johnson},
-has the log-density function
+The GL distribution in the parameterization used here is defined in \citet{Abberger:2002}
+% The type~I GL distribution (also known as skew-logistic distribution), as defined in \citet{Abberger:2002}
+%\citet{johnson},
+and has the log-density function\footnote{Which is similar to the type~1 GL distribution as defined in \citet{johnson}.}
 %
@@
 
-Assuming a single stable set of parmaters (\ref{eq:h0}) is clearly not valid, as the
+Assuming a single stable set of parmaters (\ref{eq:h0}) is  not valid, as the
 sequence of $\mathit{LM}$~statistics (see right panel of Figure~\ref{fig:slovenia1}) clearly
@@
 
+\clearpage
 This decrease in inflation is also conveyed by Figure~\ref{fig:slovenia2} that shows the
@@
 
-If there is only a single segment, the sup$\mathit{LM}$ test was non-significant at 5\%
+If there is only a single segment, the sup$\mathit{LM}$ test was non-significant at the 5\%
 level. In significant cases, the number of breakpoints was selected via the LWZ criterion
@@
 
-\fixme{Z: I have reordered the groups in Table~1 which tries to be roughly in the order
-of joining EU/ERM/EMU. Probably the following interpretations should start with a
-very short overview of the groupings.}
+For Table~\ref{tab:results1}, we grouped the countries according to either their
+similar timing of the changes in inflation dynamics or according to economic similarities, 
+as in the case of the former communist countries in Eastern Europe.
 
-The first group that draws particular attention is a subgroup of the PIIGS countries
-(Portugal, Ireland, Italy, Greece, Spain) consisting of Ireland and
-Spain. Both show a break in the summer of 2008 following the financial crisis
-and the severe problems in both countries concerning the real estate situation.
-In both countries this had a strong deflationary impact, with Ireland actually
-showing negative inflation rates for an extended period of time; in Spain, which
-was affected by a housing bubble of roughly the same size as Ireland, the
-housing prices did not adjust so abruptly and in such a degree. This
-deflationary trend, which for a time affected all countries, can be seen in
-Figure~\ref{fig:glmean}, which depicts a rolling mean estimation, where the mean
-is calculated using the first moment of a fitted GL distribution. There is
-evidence of a decline in mean inflation, which returns to its previous path
-roughly a year later.
+The first group that draws particular attention consists -- with the exception of Austria and
+Luxembourg -- of the countries of the former DEM-zone. All these countries had and still have 
+a very low inflation rate in terms both of mean and variance and their inflation dynamics
+did not change. This does make economic sense since the ECB is build after the tradition of
+the Deutsche Bundesbank and the ECB did not -- until recently -- much diverge from the 
+monetary policy of the former.
 
-The second group consists of the Eastern European countries: Czech Republic,
+Austria, Finland and France adjusted to the requirements of the EMU during the
+first phase of convergence. These are also the countries with the lowest
+inflation rates in both mean and variance after the adjustment. In these countries,
+the result of the adjustment process was a lower inflation rate and an overall
+decrease in the variance.
+
+The next group consists of three of the PIIGS (Portugal, Ireland, Italy, 
+Greece and Spain) countries and Luxembourg. Greece and
+Portugal adjusted to the ERM shortly before or after its introduction.
+Italy was successfull in reducing its inflation rate, whereas Luxembourg
+experienced a doubling of its former -- very low -- inflation rate with the
+variance rising by a factor of 14 with a strong shift in skewness towards the
+left. It is possible that in the case of Luxembourg, this still reflects the
+adjustment due to the treatment of sales periods, even after seasonal adjustment
+of the original series.
+
+The forth subgroup consists of two of the PIIGS  countries, namely Ireland and Spain.  
+Both show a break in the summer of  2008  following the financial crisis and the severe 
+problems in both countries concerning   the real estate situation. In both countries this  
+had a strong deflationary impact,  with  Ireland actually showing negative inflation rates 
+for an extended period of time;   in Spain, which was affected by a housing bubble of  
+roughly the same size as Ireland,  the housing prices did not adjust so abruptly and  
+in such a degree. This deflationary  trend, which for a time affected all countries,  
+can be seen in Figure~\ref{fig:glmean},  which depicts a rolling mean estimation,  
+where the mean is calculated using the first  moment of a fitted GL distribution.  
+There is evidence of a decline in mean inflation,  which returns to its previous 
+path roughly a year later.
+
+\clearpage
+The next group consists of two countries that decided not to be part of the
+European Monetary System\footnote{If we ignore the short participating time of
+both countries in the EMS.}, the United Kingdom and Sweden. Both show a clear
+break in the early 90s. These can be traced back to economic crisis, the currency
+crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and
+the banking crisis in Sweden during the early 90s.
+
+The last group consists of the Eastern European countries: Czech Republic,
 Estonia, Hungary, Poland, Slovakia and Slovenia. In almost all of these
@@
 Slovakia and Slovenia somewhat later. The break during the late 90s is a result
-of these countries efforts to curb in inflation by decreasing money supply and
-indicates the time when these countrie's actually overcame the biggest
+of these countries efforts to curb in inflation by decreasing the growth of the money 
+supply and indicates the time when these countries actually overcame the biggest
 transitionally shocks on their way towards free market economies. Slovakia is
@@
 prior to its entry into ERM~II.
-
-Another interesting group consists -- with the exception of Austria and
-Luxembourg -- of the countries of the former DEM-zone. All these countries have a
-very low inflation rate in both mean and variance and their inflation dynamics
-did not change.
-
-The next group consists of two countries that decided not to be part of any
-European Monetary System\footnote{If we ignore the short participating time of
-both countries in the EMS.}, the United Kingdom and Sweden. Both show a clear
-break in the early 90s. They can be traced back to economic crisis, the currency
-crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and
-the run on the Swedish currency in the same year.
-
-Austria, Finland and France adjusted to the requirements of the EMU during the
-first phase of convergence. These are also the countries with the lowest
-inflation rates in both mean and variance after the adjustment.
-
-The last group consists of three PIIGS countries and Luxembourg. Greece and
-Portugal adjusted to the ERM shortly before or afterwards its introduction.
-Italy was successfull in reducing its inflation rate, whereas Luxembourg
-experienced a doubling of its former -- very low -- inflation rate with the
-variance rising by a factor of 14 with a strong shift in skewness towards the
-left.
 
@@
 
-The disparity of the national inflation rates is still an issue. Many researchers, like \citet{hofmann}  find a considerable amount of inflation
+The disparity of the national inflation rates within the EMU is still an issue. Many researchers, like \citet{hofmann}  find a considerable amount of inflation
 differentials. \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity
@@
 red lines depict the means.} We do see that many of these countries follow a rather similar path. One event clearly visible is the financial crises and the
-short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months. Over time, there are hints, that the inflation rates
+short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months. The general trend also suggests, that the inflation rates
 converged towards some common -- albeit broad -- mean level. For the inflation series of the whole EMU, as published by the \citet{oecd}, we see
@@
 $\Sexpr{round(eu_mom[1, 1], digits = 3)}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
-$\Sexpr{round(eu_mom[2, 1], digits = 3)}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a strong decrease to
+$\Sexpr{round(eu_mom[2, 1], digits = 3)}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
 $\Sexpr{round(eu_mom[3, 1], digits = 3)}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) following the financial crises.
-Whereas variance has increased over time from
+Whereas variance has increased considerably over time from
 $\Sexpr{round(eu_mom[1, 2], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 2], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 2], digits = 3)}$.
-Regarding skewness, we see a clear change towards a negative skewness from
+Regarding skewness, we witness a movement towards negative skewness from
 $\Sexpr{round(eu_mom[1, 3], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 3], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 3], digits = 3)}$.
-\fixme{Z: interpretation needs to be adapted to modified analysis}
 
@@
 Regarding the general trend, Table~\ref{tab:results2} presents lower quartile,
-median, and upper quartile (25\%, 50\%, 75\% quantiles) of the first three moments of the
+median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
 inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}. The countries are grouped
 into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany,
-Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland,
-Greece and Spain) and all the others. For this, we took account of possible
+Luxembourg and Netherlands), the PIIGS countries  and all the others. For this, we took account of possible
 breaks in the moments.
@@
 \Sexpr{yrs[4]}, with no change in the median inflation rate, which is due to the
-fact, that there is no break after \Sexpr{yrs[2]} for these countries; however,
+fact that there is no break after \Sexpr{yrs[2]} for these countries; however,
 the interquartile range decreased by almost 20\%. The variance increased
@@
 
+The PIIGS countries decreased inflation rates as well with a strong change of skewness 
+towards more negative values. The interquartile range of their inflation variance increased
+considerably, which is mostly due to Spain and Ireland and their strong reaction
+to the financial crisis.
+
 The other countries were successfull in decreasing all three moments, but they
@@
 interquartile range for all moments decreased considerably showing a clear
-convergence towards the core-zone countries.  The PIIGS countries decreased
-inflation rates as well with a strong change of skewness towards more negative
-values. The interquartile range of their inflation variance increased
-considerably, which is mostly due to Spain and Ireland and their strong reaction
-to the financial crisis.
+convergence towards the core-zone countries.
 
@@
 
-%Contrary to economic theory, which unfortunately does not provide any clear foundation for a proper answer to the question of whether or not the creation of a monetary union between two or more states is likely to reduce or increase the variability or even the level of the inflation rate, we were able to circumvent these problems by looking directly at the data and testing for breaks in the inflation rates dynamics.
+In contrast to \citet{holte}, whose simulations suggest that the standard deviation of the home 
+CPI inflation rate can be substantially reduced by joining a monetary union, we fail to find a 
+decrease in variance for the whole samples. However, this may be partly due to the financial crisis. 
+For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in 
+inflation volatility in almost half of the euro countries. A very volatile inflation rate will 
+likely  contribute to greater macroeconomic instability. Although \cite{grier} find no evidence 
+that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --
+\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. 
+This finding might be still more problematic if we take into account that on average the skewness 
+parameter has decreased substantially in most of the euro-zone countries, and especially the PIIGS, 
+indicating that in future higher inflation rates might be more likely. This will put pressure on the 
+ECB to achieve its inflation target of 2\%. The Eastern European countries were greatly 
+successfull in decreasing inflation in both mean and variance.
 
-\fixme{Jesus: this belongs to the intro section} An interesting approach to this question -- using a theoretical framework -- is taken by \citet{holte}.
-What he finds out via simulations of  different interest rate rules is that the standard deviation of the home CPI inflation rate can be substantially reduced by joining a monetary union.
-The effects of joining a monetary union on inflation variability in his framework depend on structural parameters like risk aversion, price flexibility, export demand elasticity, openness and shock correlations.
-However, due to the fact that not all of these parameters are known and that their interaction as well has to be estimated, the whole model is very dependent on a variety of assumptions.
+The results suggest that a GL distribution is appropriate for the modeling of the seasonally 
+adjusted inflation rate series and that failing to account for a change in the skewness of the 
+distribution might lead to wrong conclusions concerning the dynamics.
 
-%\fixme{Jesus: maybe, but you don't show that in your analysis} Regarding our results, it is safe to say, that the Eastern European countries were greatly successfull in decreasing inflation in both mean and variance.
-This is due to the fact that two of the most important inflationary channels, the money supply and the degree of openness, had to be arranged according to the criteria necessary for EMU entry.
-To achieve convergence in inflation rates, the growth in money supply had to be in concert with the M3 growth rate as decided by the European Central Bank (ECB).
-The opening of the economies made cheaper goods available and thus contributed to lower prices.
-Like the original countries participating in the ERM, the Eastern European countries had to synchronize money supply growth and interest rates in order to fix exchanges rates or otherwise risk the chance of severe disequilibria.
-
-Any new country willing to join must first participate in the second exchange rate mechanism (ERM~II) with their currencies floating in a rather narrow band against the euro, which basically requires the same synchronization tools.
-For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in inflation volatility in almost half of the euro countries.
-A very volatile inflation rate will likely  contribute to greater macroeconomic instability.
-Although \cite{grier} find no evidence that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --  \cite{jaro} finds a positive correlation between the level and the standard deviation of inflation.
-This finding might be still more problematic if we take into account that on average the skewness parameter has decreased substantially in most of the euro-zone countries, and especially the PIIGS, indicating that in future higher inflation rates are more likely.
-This will put pressure on the ECB to achieve its inflation target of 2\%.
+Another area for possible improvement is the data. It seems rather disturbing that adjustments 
+in the computation of the HICP should lead to such considerable changes.
 
@@
 In addition to Germany, Portugal, Belgium, Italy, Luxembourg and Spain, there
-are a number of other countries, where the sales periods are regulated by law.
-France has an individual time span regulated by each department. This should
-pose no problem, since there is no global state level solution. In Greece there
+are a number of other countries where the sales periods are regulated by law.
+In France the sale periods time span is regulated by each department. In Greece there
 are also two set sales-periods in winter and summer. In Italy the sales periods
@@
 \cite{pltm}, the official dates of the inclusion of these sales periods are:
-January 1998 for Germany and Portugal, January 2000 for Belgium and January 1999
-for Luxembourg and Ireland, while Italy and Spain introduced them in January
-2001.
+January 1998 for Germany and Portugal,  January 1999 for Luxembourg and Ireland and 
+January 2000 for Belgium,  while Italy and Spain introduced them in January 2001.
 

### 2011-05-30 pkg/paper/Paper/draft.Rnw
updated shao citation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@104 b679b2f7-6539-45d0-a605-5ddcb976cde9
small touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@105 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-The GL distribution in the parameterization used here is defined in \citet{Abberger:2002}
+The GL distribution in the parameterization employed here is also used in \citet{Shao:2002}
 % The type~I GL distribution (also known as skew-logistic distribution), as defined in \citet{Abberger:2002}
@@
 \mathit{efp}(t) ~=~ \hat V^{-1/2} n^{-1/2} \sum_{i = 1}^{\lfloor nt \rfloor}
-  s(y_i | \hat \phi) \quad (0 \leq t \leq 1),
+  s(y_i ~|~ \hat \phi) \quad (0 \leq t \leq 1),
 \end{equation}
@@
 
+Replication scripts for all analyses are provided within the \pkg{glogis} package: An overview
+is provided in \code{demo(package = "glogis")} and the scripts can be launched easily, e.g.,
+via \code{demo("Austria", package = "glogis")} etc.
+
 \bibliography{papers}

### 2011-05-30 pkg/paper/Paper/draft.Rnw
Included computations for EU in results-adjusted.Rnw and modified draft.Rnw accordingly.
Note: To work properly, the latest version of glogis needs to be installed
because the columns in HICP and hicps were reordered.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@109 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
   t(sapply(refit(object, breaks = breaks), "[[", "moments"))
+
+## convenience formatting function
+round3 <- function(x) format(round(x, digits = 3), nsmall = 3)
 @
@@
 <<analysis-results>>=
+## country names
+nam <- c(colnames(hicps), "EU")
+
 ## run analysis (for seasonally adjusted data)
 root_dir <- do.call("file.path", as.list(head(strsplit(getwd(), .Platform$file.sep)[[1]], -1)))
-if(!file.exists(file.path(root_dir, "Analysis", "Austria.rda"))) {
+if(!all(file.exists(file.path(root_dir, "Analysis", paste(nam, "rda", sep = "."))))) {
 setwd(file.path(root_dir, "Analysis"))
-Sweave("results1s.Rnw")
+Sweave("results-adjusted.Rnw")
 setwd(file.path(root_dir, "Paper"))
@@
 
-## query result file names
-fnam <- Sys.glob(file.path("..", "Analysis", "*.rda"))
-fnam_omit <- c(grep("2", fnam, fixed = TRUE), grep("glmean", fnam, fixed = TRUE))
-if(length(fnam_omit) > 0) fnam <- fnam[-fnam_omit]
-
 ## set up named results list
-results <- vector(length = length(fnam), mode = "list")
-names(results) <- sapply(strsplit(sapply(strsplit(fnam, "/", fixed = TRUE), tail, 1), ".", fixed = TRUE), head, 1)
+results <- vector(length = length(nam), mode = "list")
+names(results) <- nam
 
@@
   Sweden = NA ,
-  UK = NA
+  UK = NA,
+  EU = NA
 ))
@@
   Sweden = NA ,
-  UK = NA
+  UK = NA,
+  EU = NA
 ))
 
-for(i in seq_along(fnam)) {
-  load(file.path(root_dir, "Analysis", fnam[i]))
+for(i in seq_along(nam)) {
+  load(file.path(root_dir, "Analysis", paste(nam[i], "rda", sep = ".")))
   results[[i]] <- list(
@@
 breakpoint, with $p$~values from $\chi^2$ goodness-of-fit tests of
-\Sexpr{format(round(summary(x$model[[1]])$chisq.test$p.value, digits = 3), nsmall = 3)} and
-\Sexpr{format(round(summary(x$model[[2]])$chisq.test$p.value, digits = 3), nsmall = 3)}, respectively.
+\Sexpr{round3(summary(x$model[[1]])$chisq.test$p.value)} and
+\Sexpr{round3(summary(x$model[[2]])$chisq.test$p.value)}, respectively.
 
@@
 
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{European Union}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(22, format_moments)))
+@
+
 \end{tabular}
@@
 <<eu, echo=FALSE>>=
-## impose breaks "by hand"
-eu_breaks <- yearmon(c(1998 + 11/12, 2008 + 7/12))
-eu_seg <- 1 + (time(eu) > eu_breaks[1]) + (time(eu) > eu_breaks[2])
-library("e1071")
-eu_mom_old1 <- aggregate(eu, eu_seg, function(x) c(mean(x), var(x), skewness(x)))
-
-## better: use model-implied moments
-eu_mom_old2 <- aggregate(eu, eu_seg, function(x) moments(glogisfit(x)))
-
-## or even better: estimate breaks?
-## eu_bp <- breakpoints(glogisfit(eu), h = 24)
-## moments(eu_bp, breaks = 2)
-## replicate results "by hand"
-eu_breaks <- yearmon(c(1999 + 5/12, 2007 + 7/12))
-eu_seg <- 1 + (time(eu) > eu_breaks[1]) + (time(eu) > eu_breaks[2])
-eu_mom <- aggregate(eu, eu_seg, function(x) moments(glogisfit(x)))
+eu_breaks <- breakdates(results$EU$breakpoints, breaks = results$EU$nbreaks)
+eu_mom <- moments(results$EU$breakpoints, breaks = results$EU$nbreaks)
 @
@@
 an increase in mean from
-$\Sexpr{round(eu_mom[1, 1], digits = 3)}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
-$\Sexpr{round(eu_mom[2, 1], digits = 3)}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
-$\Sexpr{round(eu_mom[3, 1], digits = 3)}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) following the financial crises.
+$\Sexpr{round3(eu_mom[1, 1])}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
+$\Sexpr{round3(eu_mom[2, 1])}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
+$\Sexpr{round3(eu_mom[3, 1])}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) following the financial crises.
 Whereas variance has increased considerably over time from
-$\Sexpr{round(eu_mom[1, 2], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 2], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 2], digits = 3)}$.
+$\Sexpr{round3(eu_mom[1, 2])}$ over $\Sexpr{round3(eu_mom[2, 2])}$ to $\Sexpr{round3(eu_mom[3, 2])}$.
 Regarding skewness, we witness a movement towards negative skewness from
-$\Sexpr{round(eu_mom[1, 3], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 3], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 3], digits = 3)}$.
+$\Sexpr{round3(eu_mom[1, 3])}$ over $\Sexpr{round3(eu_mom[2, 3])}$ to $\Sexpr{round3(eu_mom[3, 3])}$.
 
 <<glmean>>=
-if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
 glmean <- function(x) {
@@
 }
+
+if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
   hicp_glmean <- rollapply(hicps, 12, glmean)
@@
 @
+
+\fixme{need to note somewhere} Rapid changes that persist only for very short amounts of time
+are not captured well. Specifically, short-lasting changes in the mean, may simply be attributed
+to increased variance. For example, the mean inflation in the 
+year before the crisis (\Sexpr{yearmon(2007 + 7/12)}--\Sexpr{yearmon(2008 + 6/12)}) was
+$\Sexpr{round3(glmean(window(eu, start = yearmon(2007 + 7/12), end = yearmon(2008 + 6/12))))}$
+and after the crisis (\Sexpr{yearmon(2008 + 7/12)}--\Sexpr{yearmon(2009 + 6/12)}) it dropped to
+$\Sexpr{round3(glmean(window(eu, start = yearmon(2008 + 7/12), end = yearmon(2009 + 6/12))))}$
+before rising again in the year after that (\Sexpr{yearmon(2009 + 7/12)}--\Sexpr{yearmon(2010 + 6/12)}):
+$\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
 

### 2011-06-02 pkg/paper/Paper/draft.Rnw
updated
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@110 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
   changes in the first three moments of the inflation rates, we assumed them to be distributed according to 
-  a generalized logistic distribution, which we implemented in a new \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} 
-  \citep{R} package called \pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]}. We find breaks in all but 4 countries.
-  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a change towards negative skewness.
+  a generalized logistic distribution. We find breaks in all but 4 countries.
+  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a change
+  towards negative skewness.
+
 }
@@
 
-Ever since the beginning of the European Union, the topic of a common currency
+\readme{perhaps skip this paragraph?} The way towards the EMU basically consisted of three stages: stage~I
+(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
+convergence, and stage~III (1999--2001) the transition period, which ended with
+the introduction of the euro as legal tender. A possible fourth stage is the
+continuing integration of new member states, mostly former communist countries.
+
+Since the beginning of the European Union, the topic of a common currency
 was a controversial issue. Although the Economic and Monetary Union (EMU) is now
@@
 settled. The controversial topics range from the question of whether or not the
-eurozone is indeed an optimal currency area (as developed in: \citet{Mundell:1961})\footnote{
-For an overview, the interested reader is referred to \citet{Dellas:2009}}, 
+eurozone is indeed an optimal currency area \citep[as developed in:][]{Mundell:1961}, 
 all the way to the very survival of the euro in light of the budgetary problems of some of its member
@@
 with inflation being the most important one in this context, is in the center of
-an ongoing debate: this regards the question about short-run and steady-state
-inflation uncertainty (in context to inflation expectations) dealt with in
-\citet{cap}, or the structural convergence of the inflation rates in EU
+an ongoing debate. This concerns the issue of short--run and steady--state
+inflation uncertainty -- in context to inflation expectations -- dealt with in
+\citet{cap}, or the degree of similarity of short--run dynamic properties of the inflation rates in EU
 countries, which is the topic of an investigation by \citet{palomba}.
 
-\readme{perhaps skip this paragraph?} The way towards the EMU basically consisted of three stages: stage~I
-(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
-convergence and stage~III (1999--2001) the transition period, which ended with
-the introduction of the euro as legal tender. A possible fourth stage is the
-continuing integration of new member states, mostly former communist countries.
-Any new state that wants to join the EMU is first obliged to fulfill the
-Maastricht criteria, which, besides setting rules for government debts and
-interest rates, also requires the participation in the ERM~II (Exchange Rate
-Mechanism~II) for two years, while the exchange rates towards the euro are not
-allowed to cross the nominal band.
-
-\clearpage
-\cite{emerson} put emphasize on the fact that a high inflation rate is also more
-variable and uncertain and thus causes more relative price variability, leading
-to a less efficient price mechanism.  Failing to stabilize inflation leads to
-severe economic problems. Throughout the literature, there is still a
+%Any new state that wants to join the EMU is first obliged to fulfill the
+%Maastricht criteria, which, besides setting rules for government debts and
+%interest rates, also requires the participation in the ERM~II (Exchange Rate
+%Mechanism~II) for two years, while the exchange rates towards the euro are not
+%allowed to cross the nominal band.
+
+Throughout the literature, there is still a
 considerable degree of uncertainty as to what extent the introduction of the
-euro, or monetary unions in general, affect the inflation rate. The EMU requires
-a country to possibly alter a number of economic policies and institutions,
-which may require significant short term adjustments.
-
-Given that a country experienced quite volatile inflation rates, its efforts to
-meet the convergence criteria are likely to lead to an alteration at least in
-the mean (since this was one of the requirements) and possibly in the volatility
-of their respective inflation rates as well.
-
-However, a deeper inquiry into this problem is in need of a method able to
-capture changes in inflation dynamics. We contribute to this  by
+euro, or monetary unions in general, affect the inflation rate.
+
+However, a deeper inquiry into the effects of monetary unions on the inflation rates of its member
+countries is in need of a method able to
+capture changes in inflation dynamics. We contribute to this field of research by
 developing a new method for testing for structural breaks in the dynamics of the
-inflation rates and thereby upon the effect of the EMU on a number of European
-countries. The framework developed in this paper makes it possible to test for any 
-significant changes of a countries inflation dynamics and thus helps interpreting 
+inflation rates. The framework utilized in this paper allows testing for any 
+possible alternations  and thus helps interpreting 
 the effects of the EMU. We focus on changes in the first three moments of these 
 inflation rates, which we assume to be distributed according to a generalized logistic 
-distribution (GL). The GL distribution is a suitable one due to its flexibility to 
-capture nonsymmetry and because of its fatter tails. For this purpose, we develop a 
-new estimation approach building on previous research by \citet{Zeileis+Hornik:2007}.
-
-This \fixme{Jesus: sell method better, why this one and not another} framework
-allows us to assess the evolution of mean, variance and skewness over time. The
-skewness parameter turns out to be quite important to measure alterations in the
+distribution (GL). For this purpose, we build on previous research by \citet{Zeileis+Hornik:2007}.
+
+This \fixme{describe method} framework allows us to assess the evolution of mean, variance, and skewness over time.
+Our estimation technique enables the simultaneous estimation of breaks in these moments. 
+The skewness parameter turns out to be quite important to measure alterations in the
 long run inflation dynamics, as it changes considerably over the time of the
 analysis. This trend would be all but invisible using a normal distribution.
-Our estimation technique enables the simultaneous estimation of breaks in the
-mean (which is important in the context of the Maastricht criteria), the
-variance (which should be low for the price mechanism to be efficient) and the
-skewness (where a change in skewness is evidence of a shift of the distribution
-towards a new ``center of gravity'') of the inflation rates of interest. This
-than can be used to search for answers regarding the question about the effect
-of the EMU on the inflation dynamics of its member states or more general to
-investigate the effects of external shocks.
-
-The \fixme{Jesus: maybe comment on results here} remainder of this paper is
+
+\fixme{comment on results here} Overall, we find four periods during which structural breaks
+happened. Two of them in the 90s (early and late), another one in the early 2000s, and the last period covers the financial crisis. There is also a 
+tendency towards a lower mean inflation rate, combined with a decreasing difference between
+individual inflation rates.
+
+The  remainder of this paper is
 structured as follows: Section~\ref{sec:data} presents the data,
@@
 Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an
-example, Section~\ref{sec:results} presents the results and
+example, Section~\ref{sec:results} presents the results, and
 Section~\ref{sec:conclusion} concludes.
@@
 eu <- na.omit(hicps[, "EU"])
+## do not include EU for figures
 hicps <- hicps[, colnames(hicps) != "EU"]
@@
 Latvia, Lithuania, Bulgaria, Romania are excluded due to data scarcity.
-Cyprus and Malta are not included since they are very small economies.
+Cyprus, and Malta are not included since they are very small economies.
 
@@
 changes of the seasonal inflation patters over time within several countries
-and also between the countries. See \cite{pltm} for a detailed discussion.
+and also between the countries. See \citet{pltm} for a detailed discussion.
 
@@
 parameters change. This framework allows to trace breaks in the
-evolution of mean, variance and skewness of the inflation series which may
+evolution of mean, variance, and skewness of the inflation series which may
 possibly be linked to underlying regime changes, e.g., in monetary policy.
@@
 data would be required, e.g., in order to apply techniques such as those
-expounded by \cite{andreou}.\footnote{For example, it would be conceivable to
+expounded by \citet{andreou}.\footnote{For example, it would be conceivable to
 obtain high-frequency data for certain years from the Billion Prices Project of
-the \cite{bpp}; however, this is not pursued here.}
+the \citet{bpp}; however, this is not pursued here.}
 
@@
 context of income and growth models due to its fatter tails compared to the 
-normal distribution -- it has also been applied to the modeling of
+normal distribution -- it has also been applied to the modeling of expected
 inflation rates by \citet{batchelor}. Its generalization -- which additionally
-allows for asymmetries -- has been applied to analyze risk \citep[e.g., by][]{tolikas}
+allows for asymmetries -- has been applied to analyze extreme risks in the context of stock markets \citep[e.g., by][]{tolikas}
 but has not yet been applied -- to the best of our knowledge -- to inflation rates.
@@
 %
-with location~$\theta$, scale~$\sigma > 0$ and shape~$\delta > 0$. For $\delta = 1$ the
+with location~$\theta$, scale~$\sigma > 0$, and shape~$\delta > 0$. For $\delta = 1$ the
 distribution simplifies to the logistic distribution, for $\delta < 1$ or $> 1$ it is
@@
 
-The modeling strategy proposed in Section~\ref{sec:model} is applied to each of the
-countries included in the data set introduced in Section~\ref{sec:data}.
-Table~\ref{tab:results1} reports the estimated breakpoints along with the segment-specific
-moments (mean, variance, skewness) implied by the corresponding parameter
-estimates. Furthermore, entry dates for ERM/ERM~II and the euro are included as
-additional information. For ease of interpretation, countries with similar patterns of
-change are grouped together in Table~\ref{tab:results1} (and ordered alphabetically
-within each group).
-
-If there is only a single segment, the sup$\mathit{LM}$ test was non-significant at the 5\%
-level. In significant cases, the number of breakpoints was selected via the LWZ criterion
-and subsequently all parameters (breakpoints and segment-specific GL parameters) were
-estimated by ML. No HAC~correction was employed as there was only negligible autocorrelation
-in the empirical scores after inclusion of the breakpoints (if any).
-
-<<results-moments>>=
-## compute moments
-mom <- lapply(results, function(obj) t(sapply(obj$model, moments)))
-names(mom)[names(mom) == "CzechRepublic"] <- "Czech Rep."
-
-## LaTeX formatting
-format_moments <- function(i, digits = 3) {
-  rval <- format(round(mom[[i]], digits = digits), nsmall = digits)
-  rval <- cbind("", rownames(rval), matrix(paste("$", rval, "$", sep = ""), ncol = 3), "", "", "")
-  rval[1,1] <- if(is.character(i)) i else names(mom)[i]
-  rval[1,6] <- if(is.na(results[[i]]$erm)) "--" else format(results[[i]]$erm)
-  rval[1,7] <- if(is.na(results[[i]]$erm2)) "--" else format(results[[i]]$erm2)
-  rval[1,8] <- if(is.na(results[[i]]$eur)) "--" else format(results[[i]]$eur)
-  rval <- apply(rval, 1, paste, collapse = " & ")
-  rval <- paste(rval, "\\\\")
-  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
-  return(rval)
-}
-@
-
-\begin{table}[p!]
-\begin{center}
-\footnotesize
-\begin{tabular}{llrrrlll}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
-\hline
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{No change}} \\
-\hline
-<<moments-table1, results=tex>>=
-writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
-\hline
-<<moments-table2, results=tex>>=
-writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
-\hline
-<<moments-table3, results=tex>>=
-writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Financial crisis}} \\
-\hline
-<<moments-table4, results=tex>>=
-writeLines(unlist(lapply(c(11, 19), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Non-euro}} \\
-\hline
-<<moments-table5, results=tex>>=
-writeLines(unlist(lapply(c(20, 21), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Eastern countries}} \\
-\hline
-<<moments-table6, results=tex>>=
-writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{European Union}} \\
-\hline
-<<moments-table6, results=tex>>=
-writeLines(unlist(lapply(22, format_moments)))
-@
-
-\end{tabular}
-
-\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
-distribution per segment, and ERM information for all countries under investigation.}
-\end{center}
-\end{table}
-
-\subsection{The countries in detail}
-
-For Table~\ref{tab:results1}, we grouped the countries according to either their
-similar timing of the changes in inflation dynamics or according to economic similarities, 
-as in the case of the former communist countries in Eastern Europe.
-
-The first group that draws particular attention consists -- with the exception of Austria and
-Luxembourg -- of the countries of the former DEM-zone. All these countries had and still have 
-a very low inflation rate in terms both of mean and variance and their inflation dynamics
-did not change. This does make economic sense since the ECB is build after the tradition of
-the Deutsche Bundesbank and the ECB did not -- until recently -- much diverge from the 
-monetary policy of the former.
-
-Austria, Finland and France adjusted to the requirements of the EMU during the
-first phase of convergence. These are also the countries with the lowest
-inflation rates in both mean and variance after the adjustment. In these countries,
-the result of the adjustment process was a lower inflation rate and an overall
-decrease in the variance.
-
-The next group consists of three of the PIIGS (Portugal, Ireland, Italy, 
-Greece and Spain) countries and Luxembourg. Greece and
-Portugal adjusted to the ERM shortly before or after its introduction.
-Italy was successfull in reducing its inflation rate, whereas Luxembourg
-experienced a doubling of its former -- very low -- inflation rate with the
-variance rising by a factor of 14 with a strong shift in skewness towards the
-left. It is possible that in the case of Luxembourg, this still reflects the
-adjustment due to the treatment of sales periods, even after seasonal adjustment
-of the original series.
-
-The forth subgroup consists of two of the PIIGS  countries, namely Ireland and Spain.  
-Both show a break in the summer of  2008  following the financial crisis and the severe 
-problems in both countries concerning   the real estate situation. In both countries this  
-had a strong deflationary impact,  with  Ireland actually showing negative inflation rates 
-for an extended period of time;   in Spain, which was affected by a housing bubble of  
-roughly the same size as Ireland,  the housing prices did not adjust so abruptly and  
-in such a degree. This deflationary  trend, which for a time affected all countries,  
-can be seen in Figure~\ref{fig:glmean},  which depicts a rolling mean estimation,  
-where the mean is calculated using the first  moment of a fitted GL distribution.  
-There is evidence of a decline in mean inflation,  which returns to its previous 
-path roughly a year later.
-
-\clearpage
-The next group consists of two countries that decided not to be part of the
-European Monetary System\footnote{If we ignore the short participating time of
-both countries in the EMS.}, the United Kingdom and Sweden. Both show a clear
-break in the early 90s. These can be traced back to economic crisis, the currency
-crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and
-the banking crisis in Sweden during the early 90s.
-
-The last group consists of the Eastern European countries: Czech Republic,
-Estonia, Hungary, Poland, Slovakia and Slovenia. In almost all of these
-countries, the mean and the variance of their inflation rates declined in the
-later part of the 90s (Slovenia is an exception with an increase in variance
-after the break). Most of them experienced a break in 1997/1998, with Poland,
-Slovakia and Slovenia somewhat later. The break during the late 90s is a result
-of these countries efforts to curb in inflation by decreasing the growth of the money 
-supply and indicates the time when these countries actually overcame the biggest
-transitionally shocks on their way towards free market economies. Slovakia is
-very similar to Slovenia, as its inflation dynamics too changed roughly one year
-prior to its entry into ERM~II.
-
-\subsection{Overview over the general trend}
-
-<<eu, echo=FALSE>>=
-eu_breaks <- breakdates(results$EU$breakpoints, breaks = results$EU$nbreaks)
-eu_mom <- moments(results$EU$breakpoints, breaks = results$EU$nbreaks)
-@
-
-The disparity of the national inflation rates within the EMU is still an issue. Many researchers, like \citet{hofmann}  find a considerable amount of inflation
-differentials. \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity
-across EMU countries in terms of average inflation and its degree of persistence. In a paper examining structural convergence of inflation rates in EU
-countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar. They find that
-convergence over time of inflation dynamics was only partly observable.  In a paper studying core inflation and using an aggregated euro area inflation
-rate, \citet{morana} finds three regimes (roughly 1980--1984, 1984--1993 and 1993--2000) governing the core inflation rate.
-
-Concerning inflation differentials, Figure~\ref{fig:glmean} depicts the inflation rates and a rolling mean estimation (with a window size of 12 months),
-where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month,
-the lines are overlapped and thus get darker. We wold see one black line if all countries would show the same month-on-month increase at any one time. The
-red lines depict the means.} We do see that many of these countries follow a rather similar path. One event clearly visible is the financial crises and the
-short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months. The general trend also suggests, that the inflation rates
-converged towards some common -- albeit broad -- mean level. For the inflation series of the whole EMU, as published by the \citet{oecd}, we see
-an increase in mean from
-$\Sexpr{round3(eu_mom[1, 1])}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
-$\Sexpr{round3(eu_mom[2, 1])}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
-$\Sexpr{round3(eu_mom[3, 1])}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) following the financial crises.
-Whereas variance has increased considerably over time from
-$\Sexpr{round3(eu_mom[1, 2])}$ over $\Sexpr{round3(eu_mom[2, 2])}$ to $\Sexpr{round3(eu_mom[3, 2])}$.
-Regarding skewness, we witness a movement towards negative skewness from
-$\Sexpr{round3(eu_mom[1, 3])}$ over $\Sexpr{round3(eu_mom[2, 3])}$ to $\Sexpr{round3(eu_mom[3, 3])}$.
-
-<<glmean>>=
-glmean <- function(x) {
-  if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
-}
-
-if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
-  hicp_glmean <- rollapply(hicps, 12, glmean)
-  setwd(file.path(root_dir, "Analysis"))
-  save(hicp_glmean, file = "glmean.rda")
-  setwd(file.path(root_dir, "Paper"))
-} else {
-  load(file.path(root_dir, "Analysis", "glmean.rda"))
-}
-@
-
-\fixme{need to note somewhere} Rapid changes that persist only for very short amounts of time
-are not captured well. Specifically, short-lasting changes in the mean, may simply be attributed
-to increased variance. For example, the mean inflation in the 
-year before the crisis (\Sexpr{yearmon(2007 + 7/12)}--\Sexpr{yearmon(2008 + 6/12)}) was
-$\Sexpr{round3(glmean(window(eu, start = yearmon(2007 + 7/12), end = yearmon(2008 + 6/12))))}$
-and after the crisis (\Sexpr{yearmon(2008 + 7/12)}--\Sexpr{yearmon(2009 + 6/12)}) it dropped to
-$\Sexpr{round3(glmean(window(eu, start = yearmon(2008 + 7/12), end = yearmon(2009 + 6/12))))}$
-before rising again in the year after that (\Sexpr{yearmon(2009 + 7/12)}--\Sexpr{yearmon(2010 + 6/12)}):
-$\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
-
-\setkeys{Gin}{width=0.8\textwidth}
-\begin{figure}[t!]
-\begin{center}
-<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
-  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
-  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
-    xlab = "Time", ylab = "Monthly inflation rates")
-  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-@
-\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from rolling GL fits (red)
-over time for all countries.}
-\end{center}
-\end{figure}
-\setkeys{Gin}{width=\textwidth}
-
 <<cross-country, echo=FALSE, results=hide>>=
@@
 
-Regarding the general trend, Table~\ref{tab:results2} presents lower quartile,
-median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
-inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}. The countries are grouped
-into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany,
-Luxembourg and Netherlands), the PIIGS countries  and all the others. For this, we took account of possible
-breaks in the moments.
+\readme{questions and challenges} The aim of this paper is to test for structural breaks in the dynamics of the inflation rates.
+Making use of the method developed in Section~\ref{sec:model},  the effect of the EMU upon national inflation rates can be investigated.
+Apart from the countries of the euro--zone, the inflation dynamics of the other European countries are also of interest.
+This issue is closely related with the question about the convergence of the national inflation rates of the EMU members towards some common mean.
+Many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials in parts of the sample.\readme{perhaps skip this one??} \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity in terms of average inflation and its degree of persistence.
+
+\readme{methods, approach} To explore these issues, we make use of three different approaches.
+First, in Section~\ref{sec:bpresults}, the estimation technique proposed in Section~\ref{sec:model} is applied to each of the countries included in the data set introduced in Section~\ref{sec:data}.
+Table~\ref{tab:results1} reports the estimated breakpoints along with the segment-specific moments (mean, variance, skewness) implied by the corresponding parameter
+estimates. Furthermore, entry dates for ERM/ERM~II and the euro are included as additional information. For ease of interpretation, countries with similar patterns of
+change are grouped together and ordered alphabetically within each group.
+If there is only a single segment, the sup$\mathit{LM}$ test was non-significant at the 5\% level.
+In significant cases, the number of breakpoints was selected via the LWZ criterion and subsequently all parameters (breakpoints and segment-specific GL parameters) were estimated by ML.
+No HAC~correction was employed as there was only negligible autocorrelation in the empirical scores after inclusion of the breakpoints (if any).
+
+Section~\ref{sec:aggregate} deals with the convergence issue.
+Table~\ref{tab:results2} gives an overview over the data aggregated into three different country groups: the core--euro countries (Austria, Belgium, Finland, France, Germany, Luxembourg, and Netherlands), the PIIGS countries (Portugal, Ireland, Italy, Greece, and Spain), and the other countries.
+This is done by showing mean, variance, and skewness over time and taking possible break points in the data into account.
+
+Finally, Figure~\ref{fig:glmean} in Section~\ref{sec:fcrisis}, depicts the inflation rates and a rolling mean estimation.
+This is another way of investigating into a possible convergence pattern of national inflation rates and is especially useful to depict the influence of the financial crisis.
+
+<<results-moments>>=
+## compute moments
+mom <- lapply(results, function(obj) t(sapply(obj$model, moments)))
+names(mom)[names(mom) == "CzechRepublic"] <- "Czech Rep."
+
+## LaTeX formatting
+format_moments <- function(i, digits = 3) {
+  rval <- format(round(mom[[i]], digits = digits), nsmall = digits)
+  rval <- cbind("", rownames(rval), matrix(paste("$", rval, "$", sep = ""), ncol = 3), "", "", "")
+  rval[1,1] <- if(is.character(i)) i else names(mom)[i]
+  rval[1,6] <- if(is.na(results[[i]]$erm)) "--" else format(results[[i]]$erm)
+  rval[1,7] <- if(is.na(results[[i]]$erm2)) "--" else format(results[[i]]$erm2)
+  rval[1,8] <- if(is.na(results[[i]]$eur)) "--" else format(results[[i]]$eur)
+  rval <- apply(rval, 1, paste, collapse = " & ")
+  rval <- paste(rval, "\\\\")
+  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
+  return(rval)
+}
+@
+
+\begin{table}[p!]
+\begin{center}
+\footnotesize
+\begin{tabular}{llrrrlll}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{No change}} \\
+\hline
+<<moments-table1, results=tex>>=
+writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
+\hline
+<<moments-table2, results=tex>>=
+writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
+\hline
+<<moments-table3, results=tex>>=
+writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Financial crisis}} \\
+\hline
+<<moments-table4, results=tex>>=
+writeLines(unlist(lapply(c(11, 19), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Non-euro}} \\
+\hline
+<<moments-table5, results=tex>>=
+writeLines(unlist(lapply(c(20, 21), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Eastern countries}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Euro--area}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(22, format_moments)))
+@
+
+\end{tabular}
+
+\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
+distribution per segment, and ERM information for all countries under investigation.}
+\end{center}
+\end{table}
+
+\subsection{Breakpoint estimation results}\label{sec:bpresults}
+
+\readme{changed to: 1) individual countries (table 1), 2) EU (table 1), 3) groups of countries (table 2), 4) crisis (fig.3)}
+For Table~\ref{tab:results1}, we grouped the countries according to either a
+similar timing of the changes in inflation dynamics or according to economic similarities, 
+as in the case of the former communist countries in Eastern Europe.
+
+The first group that draws particular attention consists -- with the exception of Austria and
+Luxembourg -- of the countries of the former DEM-zone. All these countries had and still have 
+a very low inflation rate in terms of both mean and variance and their inflation dynamics
+did not change. This does make economic sense since the ECB is modeled after 
+the Deutsche Bundesbank and the ECB did not -- until recently -- much diverge from the 
+monetary policy of the former. Thus the EMU did have no effect since the policies in these countries were
+already much in line with Germany and thus the ECB.
+
+Austria, Finland, and France adjusted to the requirements of the EMU during the
+first phase of convergence. In these countries,
+the result of the adjustment process was a lower inflation rate and an overall
+decrease in the variance.
+
+The next group consists of three of the PIIGS  countries and Luxembourg. Greece and
+Portugal adjusted to the ERM shortly before or after its introduction.
+Italy was successfull in reducing its inflation rate, whereas Luxembourg
+experienced a doubling of its former -- very low -- inflation rate with the
+variance rising by a factor of almost 14 with a strong shift in skewness towards the
+left. Apart from Luxembourg, the countries that had a structural break during the second 
+phase of convergence were also successfull in reducing mean inflation rates. In this group, as well
+as in the previous, the EMU adjustment led to a lower mean inflation rate.
+
+The forth subgroup consists of two of the PIIGS  countries, namely Ireland and Spain.  
+There is a break in both countries in the summer of  2008  following the financial crisis and the severe 
+problems concerning the real estate situation. In Ireland as well as in Spain, this  
+had a strong deflationary impact, with  deflation in Ireland over an extended period of time. 
+In Spain, which was affected by a housing bubble of  
+roughly the same size as Ireland,  the housing prices did not adjust so abruptly and  
+in such a degree.
+
+%This deflationary  trend, which for a time affected all countries,  
+%can be seen in Figure~\ref{fig:glmean},  which depicts a rolling mean estimation,  
+%where the mean is calculated using the first  moment of a fitted GL distribution.  
+%There is evidence of a decline in mean inflation,  which returns to its previous 
+%path roughly a year later.
+
+The next group consists of two countries that, after a very short participation in the
+ERM, decided not to be part of it and in that way also not to participate in the EMU, i.e. the United Kingdom and Sweden. Both show a clear
+break in the early 90s. These can be traced back to economic crisis, the currency
+crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and
+the banking crisis in Sweden during the early 90s.
+
+The last group consists of the Eastern European countries: Czech Republic,
+Estonia, Hungary, Poland, Slovakia, and Slovenia. In almost all of these
+countries, the mean and the variance of their inflation rates declined in the
+later part of the 90s. Czech Republic, Estonia and Hungary experienced a break in 1997/1998. This is a result
+of these countries efforts to curb in inflation by decreasing the growth of the money 
+supply and indicates the time when they actually overcame the biggest
+transitionally shocks on their way towards free market economies. In Poland, this took two more
+years. Slovakia and Slovenia are special cases: for once, they started with lower mean
+inflation values and on the other hand, their inflation dynamics  both changed roughly one year
+prior to their entry into the ERM~II.
+
+<<glmean>>=
+glmean <- function(x) {
+  if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
+}
+
+if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
+  hicp_glmean <- rollapply(hicps, 12, glmean)
+  setwd(file.path(root_dir, "Analysis"))
+  save(hicp_glmean, file = "glmean.rda")
+  setwd(file.path(root_dir, "Paper"))
+} else {
+  load(file.path(root_dir, "Analysis", "glmean.rda"))
+}
+@
+
+<<eu, echo=FALSE>>=
+eu_breaks <- breakdates(results$EU$breakpoints, breaks = results$EU$nbreaks)
+eu_mom <- moments(results$EU$breakpoints, breaks = results$EU$nbreaks)
+@
+
+In the case of the inflation rate of the euro--area, we witness three different regimes. One until 
+1999 with both a low mean and variance, the next shortly after the beginning of the third and final stage of
+monetary integration, accompanied with a significant increase in the mean. The overheating of the economy that
+culminated in the financial crisis shows itself in a big increase in variance and a lower mean inflation, reflecting
+the negative inflation rates is some euro--area countries.
+
+%There is an increase in mean from
+%$\Sexpr{round3(eu_mom[1, 1])}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
+%$\Sexpr{round3(eu_mom[2, 1])}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
+%$\Sexpr{round3(eu_mom[3, 1])}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) shortly before the financial crises.
+%
+
+\clearpage
+\subsection{Convergence on an aggregate level}\label{sec:aggregate}
 
@@
 
-In the core-zone, we see a decrease in the lower quartile from \Sexpr{yrs[1]} to
-\Sexpr{yrs[4]}, with no change in the median inflation rate, which is due to the
-fact that there is no break after \Sexpr{yrs[2]} for these countries; however,
-the interquartile range decreased by almost 20\%. The variance increased
-slightly (but with a 15\% decrease in interquartile range), whereas skewness
-stayed symmetric in the median, with almost no change in the interquartile
-range.
-
-The PIIGS countries decreased inflation rates as well with a strong change of skewness 
-towards more negative values. The interquartile range of their inflation variance increased
-considerably, which is mostly due to Spain and Ireland and their strong reaction
-to the financial crisis.
-
-The other countries were successfull in decreasing all three moments, but they
-are still higher than the euro-zone countries. Most importantly, the
-interquartile range for all moments decreased considerably showing a clear
-convergence towards the core-zone countries.
+Table~\ref{tab:results2} presents lower quartile,
+median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
+inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}. The countries are grouped
+into the ``core'' of the eurozone, the PIIGS countries  and all the others. For this, we took account of possible
+breaks in the moments.
+
+The time periods are chosen according to four fundamental change periods. Namely, the early 90s, with breaks in
+Austria, Finland, France, Sweden, and UK. The late 90s, with breaks in Greece, Italy, Luxembourg, Portugal, Czech Republic,
+Estonia, and Hungary. And also the period after the second phase of convergence, with breaks in the early 2000s (Poland, Slovakia, and Slovenia)
+and the two countries with the most visible effect of the financial crisis (Ireland and Spain).
+
+In the core-zone, there is a decrease in the lower quartile from \Sexpr{yrs[1]} to
+\Sexpr{yrs[2]}. However, the interquartile range decreased by almost 20\%. The variance increased
+slightly, but also with a 15\% decrease in interquartile range. The skewness
+stayed symmetric in the median, with almost an 8\% increase in the interquartile range.
+
+The PIIGS countries significantly decreased the median inflation rates from \Sexpr{yrs[1]} to \Sexpr{yrs[4]}. 
+Whereas the interquartile range was very tight in \Sexpr{yrs[1]}, it increased in \Sexpr{yrs[2]} and again in \Sexpr{yrs[4]}.
+However, the median values are much smaller. The variance increased from \Sexpr{yrs[1]} to \Sexpr{yrs[2]} but did not change much 
+in \Sexpr{yrs[4]}. The interquartile range did not change much too. Median skewness changed considerable 
+from highly positive to negative values. This is mostly due to the changes associated with the financial crisis, that was
+responsible for deflationary shocks. The interquartile range of skewness decreased considerably.
+
+The other countries were successfull in decreasing all three moments from \Sexpr{yrs[1]} to \Sexpr{yrs[3]}. Still, their median
+inflation rate is higher than those of the core countries. Also, the interquartile range of the mean decreased considerably. 
+Contrary to the euro--zone and the PIIGS countries, variance decreased also, both in terms of the median value and the
+interquartile range. The skewness decreased as well, as did the interquartile range of the skewness parameter.
+
+In context of the convergence issue, we find that, with the interquartile range of the mean of all the aggregated zones declining and getting tighter, there is  
+evidence of a convergence towards some common mean inflation rate in the countries of our sample, both within the subgroups and the three groups put together.
+This is also visible in the median values, that  diverge much less than at the beginning in \Sexpr{yrs[1]}.
+
+\clearpage
+\subsection{The influence of the financial crisis}\label{sec:fcrisis}
+
+\setkeys{Gin}{width=0.8\textwidth}
+\begin{figure}[t!]
+\begin{center}
+<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
+  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
+  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
+    xlab = "Time", ylab = "Monthly inflation rates")
+  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+@
+\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from rolling GL fits (red)
+over time for all countries, with a window size of 12 months}
+\end{center}
+\end{figure}
+\setkeys{Gin}{width=\textwidth}
+
+Another way to address the convergence behavior of the inflation rates and the influence of the financial crisis is a graphical representation, as in Figure~\ref{fig:glmean}.
+It depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
+The financial crises and the short deflationary phase starting  around the autumn of 2007 and lasting a few months is clearly visible.
+It is outlined by a strong increase in mean inflation during the boom years of 2007/2008 accompanied by an equally strong deflation following the bust.
+However, the level of mean inflation did not change much, as the inflation rates seem to have returned to their prior path.
+The inflation series of the euro--area conveys the same message.
+With regard to the financial crisis, this shows that for low frequency data, like inflation, rapid changes
+that persist only for a very short period of time
+are not captured well. Specifically, short-lasting changes in the mean may simply be attributed
+to increased variance. For example, the mean inflation of the euro--area in the 
+year before the crisis (\Sexpr{yearmon(2007 + 7/12)}--\Sexpr{yearmon(2008 + 6/12)}) was
+$\Sexpr{round3(glmean(window(eu, start = yearmon(2007 + 7/12), end = yearmon(2008 + 6/12))))}$
+and after the crisis (\Sexpr{yearmon(2008 + 7/12)}--\Sexpr{yearmon(2009 + 6/12)}) it dropped to
+$\Sexpr{round3(glmean(window(eu, start = yearmon(2008 + 7/12), end = yearmon(2009 + 6/12))))}$
+before rising again in the following year  (\Sexpr{yearmon(2009 + 7/12)}--\Sexpr{yearmon(2010 + 6/12)}):
+$\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
+The general trend also suggests	 that the inflation rates converged towards some common -- albeit broad -- mean level. 
+There is still some visible difference in the inflation rates, but the mean estimates are moving closer, albeit in a rather broad band.
+So contrary to \citet{palomba}, who find that the degree of similarity in short--run inflation dynamics is still weak, we find evidence of convergence within the countries in the sample.
 
@@
 
-In contrast to \citet{holte}, whose simulations suggest that the standard deviation of the home 
-CPI inflation rate can be substantially reduced by joining a monetary union, we fail to find a 
-decrease in variance for the whole samples. However, this may be partly due to the financial crisis. 
-For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in 
-inflation volatility in almost half of the euro countries. A very volatile inflation rate will 
-likely  contribute to greater macroeconomic instability. Although \cite{grier} find no evidence 
-that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --
-\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. 
-This finding might be still more problematic if we take into account that on average the skewness 
-parameter has decreased substantially in most of the euro-zone countries, and especially the PIIGS, 
-indicating that in future higher inflation rates might be more likely. This will put pressure on the 
-ECB to achieve its inflation target of 2\%. The Eastern European countries were greatly 
-successfull in decreasing inflation in both mean and variance.
+\readme{start with 1) short recap of what we did, 2) conclusions}
+
+To assess changes in the dynamics of the inflation rates of a number of European countries both within
+and without the EMU, we developed a new method for estimating structural breaks resting on the 
+assumption that the inflation rates are GL distributed.
 
@@
 
-Another area for possible improvement is the data. It seems rather disturbing that adjustments 
-in the computation of the HICP should lead to such considerable changes.
+The results of \citet{holte}, whose simulations indicated that the standard deviation of the home 
+CPI inflation rate can be substantially reduced by joining a monetary union, can not be empirically validated.
+
+For the countries in the core of the eurozone, one clearly visible trend is the rise in 
+inflation volatility in almost half of the euro countries. \citet{emerson} emphasize  the fact that a high inflation rate is also more
+variable and uncertain and thus causes more relative price variability, leading
+to a less efficient price mechanism. From an economic point of view, a very volatile inflation rate will 
+likely  contribute to greater macroeconomic instability. In the literature, there is no consensus about the effects of higher 
+inflation volatility on the inflation rate. \citet{grier} find no evidence 
+that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA. However,
+\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. Naturally, an
+efficient price mechanism will benefit from inflation rates that do not diverge too much from the expectations.
+
+The good news is that mean inflation declined considerably, most notably in the Eastern European countries that clearly 
+benefited from the monetary discipline that is demanded for those countries who want to join the EMU.
 
@@
 
-In addition to Germany, Portugal, Belgium, Italy, Luxembourg and Spain, there
+In addition to Germany, Portugal, Belgium, Italy, Luxembourg, and Spain, there
 are a number of other countries where the sales periods are regulated by law.

### 2011-06-06 pkg/paper/Paper/draft.Rnw
rewritten 4.1 and 4.3
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@111 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
   towards negative skewness.
-
+  %% avoid "we"-constructs
+  %% "adopt" better than "assume"
 }
@@
 \fixme{comment on results here} Overall, we find four periods during which structural breaks
-happened. Two of them in the 90s (early and late), another one in the early 2000s, and the last period covers the financial crisis. There is also a 
+happened. Two of them in the 1990s (early and late), another one in the early 2000s, and the last period covers the financial crisis. There is also a 
 tendency towards a lower mean inflation rate, combined with a decreasing difference between
@@
 relative instability following the independence from Yugoslavia, Slovenia was
-successfull in realigning its economy and introduced a number of reforms leading
+successful in realigning its economy and introduced a number of reforms leading
 to a stable growth in recent years. The good economic performance made it
@@
 
-\clearpage
 This decrease in inflation is also conveyed by Figure~\ref{fig:slovenia2} that shows the
@@
 
-\readme{questions and challenges} The aim of this paper is to test for structural breaks in the dynamics of the inflation rates.
-Making use of the method developed in Section~\ref{sec:model},  the effect of the EMU upon national inflation rates can be investigated.
-Apart from the countries of the euro--zone, the inflation dynamics of the other European countries are also of interest.
-This issue is closely related with the question about the convergence of the national inflation rates of the EMU members towards some common mean.
-Many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials in parts of the sample.\readme{perhaps skip this one??} \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity in terms of average inflation and its degree of persistence.
-
-\readme{methods, approach} To explore these issues, we make use of three different approaches.
-First, in Section~\ref{sec:bpresults}, the estimation technique proposed in Section~\ref{sec:model} is applied to each of the countries included in the data set introduced in Section~\ref{sec:data}.
-Table~\ref{tab:results1} reports the estimated breakpoints along with the segment-specific moments (mean, variance, skewness) implied by the corresponding parameter
-estimates. Furthermore, entry dates for ERM/ERM~II and the euro are included as additional information. For ease of interpretation, countries with similar patterns of
-change are grouped together and ordered alphabetically within each group.
-If there is only a single segment, the sup$\mathit{LM}$ test was non-significant at the 5\% level.
-In significant cases, the number of breakpoints was selected via the LWZ criterion and subsequently all parameters (breakpoints and segment-specific GL parameters) were estimated by ML.
-No HAC~correction was employed as there was only negligible autocorrelation in the empirical scores after inclusion of the breakpoints (if any).
-
-Section~\ref{sec:aggregate} deals with the convergence issue.
-Table~\ref{tab:results2} gives an overview over the data aggregated into three different country groups: the core--euro countries (Austria, Belgium, Finland, France, Germany, Luxembourg, and Netherlands), the PIIGS countries (Portugal, Ireland, Italy, Greece, and Spain), and the other countries.
-This is done by showing mean, variance, and skewness over time and taking possible break points in the data into account.
-
-Finally, Figure~\ref{fig:glmean} in Section~\ref{sec:fcrisis}, depicts the inflation rates and a rolling mean estimation.
-This is another way of investigating into a possible convergence pattern of national inflation rates and is especially useful to depict the influence of the financial crisis.
+To investigate the effects of the EMU on national inflation rates, the
+techniques expounded in Section~\ref{sec:model} are employed to
+uncover changes in the dynamics of European inflation rates. The inflation
+dynamics within countries from the euro-zone (but also other European
+countries) is closely related with the question about the convergence
+of the national inflation rates of the EMU members towards some common mean.
+Many researchers provide evidence against convergence among European countries, e.g.,
+\citet{hofmann} find considerable amounts of inflation differentials and
+\citet{cap} describe rather strong heterogeneity among 12~EMU countries, both
+in terms of average inflation and its degree of persistence.
+
+To explore the question of convergence, we proceed in two steps:
+First, in Section~\ref{sec:bpresults}, the estimation technique proposed
+in Section~\ref{sec:model} is applied to each of the countries included in
+the data set (see Section~\ref{sec:data}). Second, Section~\ref{sec:aggregate}
+aggregates the results from each individual country to reveal different
+potential convergence patterns over time within three groups of countries:
+core-euro (\Sexpr{paste(colnames(hicps)[euro], collapse = ", ")}),
+PIIGS  (\Sexpr{paste(colnames(hicps)[piigs], collapse = ", ")}), and the
+rest.
+
+Finally, Section~\ref{sec:fcrisis} focuses on the question of changes
+in the inflation dynamics during the 2008/2009 financial crisis.
 
@@
 \noalign{\medskip}
-\multicolumn{8}{l}{\emph{Euro--area}} \\
+\multicolumn{8}{l}{\emph{Euro area}} \\
 \hline
@@
 
-\subsection{Breakpoint estimation results}\label{sec:bpresults}
-
-\readme{changed to: 1) individual countries (table 1), 2) EU (table 1), 3) groups of countries (table 2), 4) crisis (fig.3)}
-For Table~\ref{tab:results1}, we grouped the countries according to either a
+\subsection{Inflation dynamics on a national level}\label{sec:bpresults}
+
+As a first step towards the investigation of inflation convergence, the
+structural breaks in the inflation dynamics within each countries are
+investigated using the methods introduced in Section~\ref{sec:model}.
+Table~\ref{tab:results1} reports the estimated breakpoints along with the
+segment-specific moments (mean, variance, skewness) implied by the corresponding
+parameter estimates. Furthermore, entry dates for ERM/ERM~II and the euro are
+included as additional information. For ease of interpretation, countries with
+similar patterns of change are grouped together and ordered alphabetically
+within each group. If there is only a single segment, the sup$\mathit{LM}$ test
+was non-significant at the 5\% level. In significant cases, the number of
+breakpoints was selected via the LWZ criterion and subsequently all parameters
+(breakpoints and segment-specific GL parameters) were estimated by ML. No
+HAC~correction was employed as there was only negligible autocorrelation in the
+empirical scores after inclusion of the breakpoints (if any).
+In Table~\ref{tab:results1} the countries are grouped according to either a
 similar timing of the changes in inflation dynamics or according to economic similarities, 
@@
 
-The first group that draws particular attention consists -- with the exception of Austria and
-Luxembourg -- of the countries of the former DEM-zone. All these countries had and still have 
-a very low inflation rate in terms of both mean and variance and their inflation dynamics
-did not change. This does make economic sense since the ECB is modeled after 
-the Deutsche Bundesbank and the ECB did not -- until recently -- much diverge from the 
-monetary policy of the former. Thus the EMU did have no effect since the policies in these countries were
-already much in line with Germany and thus the ECB.
+The first group that draws particular attention consists of the former
+DEM (Deutsche Mark) zone countries, with the exception of Austria and
+Luxembourg. All these countries had and still have a very low inflation rate,
+both in terms of mean and variance, and their inflation dynamics
+did not change significantly. This is plausible from an economic point of view
+since the European central bank (ECB) is modeled after the Deutsche Bundesbank and the ECB did not
+-- until recently -- diverge much from the monetary policy of the former. Thus,
+the EMU did have no effect on inflation dynamics since the policies in
+these countries were already much in line with Germany and hence the ECB.
 
@@
 Portugal adjusted to the ERM shortly before or after its introduction.
-Italy was successfull in reducing its inflation rate, whereas Luxembourg
-experienced a doubling of its former -- very low -- inflation rate with the
-variance rising by a factor of almost 14 with a strong shift in skewness towards the
+Italy was successful in reducing its inflation rate, whereas Luxembourg
+experienced a strong increase in its formerly very low inflation rate, both
+in terms of mean and variances, also associated with a shift in skewness towards the
 left. Apart from Luxembourg, the countries that had a structural break during the second 
-phase of convergence were also successfull in reducing mean inflation rates. In this group, as well
-as in the previous, the EMU adjustment led to a lower mean inflation rate.
-
-The forth subgroup consists of two of the PIIGS  countries, namely Ireland and Spain.  
-There is a break in both countries in the summer of  2008  following the financial crisis and the severe 
-problems concerning the real estate situation. In Ireland as well as in Spain, this  
-had a strong deflationary impact, with  deflation in Ireland over an extended period of time. 
-In Spain, which was affected by a housing bubble of  
+phase of convergence were also successful in reducing mean inflation rates.
+After the changes in this group, all countries have rather similar mean inflation
+rates, closer to the levels of the previous two groups.
+
+The fourth subgroup consists of two of the PIIGS countries, namely Ireland and Spain.  
+There is a break in both countries in the summer of 2008 that captures the effects of
+the financial crisis and the severe problems concerning the real estate situation.
+In both countries, this had a strong deflationary impact, with deflation in Ireland
+over an extended period of time. In Spain, which was affected by a housing bubble of  
 roughly the same size as Ireland, the housing prices did not adjust so abruptly and  
-in such a degree.
-
-%This deflationary  trend, which for a time affected all countries,  
-%can be seen in Figure~\ref{fig:glmean},  which depicts a rolling mean estimation,  
-%where the mean is calculated using the first  moment of a fitted GL distribution.  
-%There is evidence of a decline in mean inflation,  which returns to its previous 
-%path roughly a year later.
+so strongly.
 
 The next group consists of two countries that, after a very short participation in the
-ERM, decided not to be part of it and in that way also not to participate in the EMU, i.e. the United Kingdom and Sweden. Both show a clear
-break in the early 90s. These can be traced back to economic crisis, the currency
-crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and
-the banking crisis in Sweden during the early 90s.
+ERM, decided not to be part of it and hence also not to participate in the EMU:
+the United Kingdom and Sweden. Both show clear breaks in the early 1990s which can be
+traced back to economic crises: the currency crisis of the United Kingdom cumulating
+into the ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
+1990s.
 
@@
 countries, the mean and the variance of their inflation rates declined in the
-later part of the 90s. Czech Republic, Estonia and Hungary experienced a break in 1997/1998. This is a result
-of these countries efforts to curb in inflation by decreasing the growth of the money 
-supply and indicates the time when they actually overcame the biggest
-transitionally shocks on their way towards free market economies. In Poland, this took two more
-years. Slovakia and Slovenia are special cases: for once, they started with lower mean
-inflation values and on the other hand, their inflation dynamics  both changed roughly one year
-prior to their entry into the ERM~II.
+late 1990s or early 2000s. Czech Republic, Estonia and Hungary experienced a break in 1997/1998
+resulting from efforts to curb in inflation by decreasing the growth of the money 
+supply. The timing of the break thus indicates the timing of the biggest
+transitionally shocks on the way towards free market economies. In Poland, this
+transition took two more years. Slovakia and Slovenia are somewhat different: On one hand,
+they started with lower mean inflation rates and on the other hand, their inflation dynamics
+both changed roughly one year prior to their entry into the ERM~II.
 
@@
 
-In the case of the inflation rate of the euro--area, we witness three different regimes. One until 
-1999 with both a low mean and variance, the next shortly after the beginning of the third and final stage of
-monetary integration, accompanied with a significant increase in the mean. The overheating of the economy that
-culminated in the financial crisis shows itself in a big increase in variance and a lower mean inflation, reflecting
-the negative inflation rates is some euro--area countries.
+Finally, the inflation rates for the whole euro area \fixme{Is EU euro area or European union?},
+as provided by the OECD, are considered and three different regimes are found: (1)~until 
+1999 with both a low mean and variance, (2)~the third and final stage of monetary integration,
+accompanied with a significant increase in the mean, (3)~a period encompassing
+the overheating of the economy that culminated in the financial crisis, showing a big increase
+in variance and a lower mean inflation.
+
+\fixme{some concluding remarks about the EU level bridging the current gap to the next subsection}
 
@@
 
-\clearpage
 \subsection{Convergence on an aggregate level}\label{sec:aggregate}
@@
 
+\fixme{intro needed}
+
+\fixme{also ``zoom out'' more, a big picture needs to be established}
+
 Table~\ref{tab:results2} presents lower quartile,
@@
 
-The time periods are chosen according to four fundamental change periods. Namely, the early 90s, with breaks in
-Austria, Finland, France, Sweden, and UK. The late 90s, with breaks in Greece, Italy, Luxembourg, Portugal, Czech Republic,
+The time periods are chosen according to four fundamental change periods. Namely, the early 1990s, with breaks in
+Austria, Finland, France, Sweden, and UK. The late 1990s, with breaks in Greece, Italy, Luxembourg, Portugal, Czech Republic,
 Estonia, and Hungary. And also the period after the second phase of convergence, with breaks in the early 2000s (Poland, Slovakia, and Slovenia)
@@
 
-The other countries were successfull in decreasing all three moments from \Sexpr{yrs[1]} to \Sexpr{yrs[3]}. Still, their median
+The other countries were successful in decreasing all three moments from \Sexpr{yrs[1]} to \Sexpr{yrs[3]}. Still, their median
 inflation rate is higher than those of the core countries. Also, the interquartile range of the mean decreased considerably. 
@@
 
-\clearpage
 \subsection{The influence of the financial crisis}\label{sec:fcrisis}
@@
 @
-\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from rolling GL fits (red)
-over time for all countries, with a window size of 12 months}
+\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
+12-month rolling GL fits (red) over time for all countries.}
 \end{center}
@@
 
-Another way to address the convergence behavior of the inflation rates and the influence of the financial crisis is a graphical representation, as in Figure~\ref{fig:glmean}.
-It depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.
-The financial crises and the short deflationary phase starting  around the autumn of 2007 and lasting a few months is clearly visible.
-It is outlined by a strong increase in mean inflation during the boom years of 2007/2008 accompanied by an equally strong deflation following the bust.
-However, the level of mean inflation did not change much, as the inflation rates seem to have returned to their prior path.
-The inflation series of the euro--area conveys the same message.
-With regard to the financial crisis, this shows that for low frequency data, like inflation, rapid changes
-that persist only for a very short period of time
-are not captured well. Specifically, short-lasting changes in the mean may simply be attributed
-to increased variance. For example, the mean inflation of the euro--area in the 
+The results from the previous sections convey that the financial
+crisis of 2008/9 did not have a significant effect on the inflation dynamics
+in most of the European countries (except Ireland and Spain).
+This may be somewhat surprising and is not enitrely correct. In fact,
+there is a rather clear effect but as it is very short-lived and not
+persistent it is not picked up by the structural break techniques.
+
+More specifically, the mean inflation of the euro area \fixme{see above} from a GL fit in the 
 year before the crisis (\Sexpr{yearmon(2007 + 7/12)}--\Sexpr{yearmon(2008 + 6/12)}) was
@@
 $\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
-The general trend also suggests	 that the inflation rates converged towards some common -- albeit broad -- mean level. 
-There is still some visible difference in the inflation rates, but the mean estimates are moving closer, albeit in a rather broad band.
-So contrary to \citet{palomba}, who find that the degree of similarity in short--run inflation dynamics is still weak, we find evidence of convergence within the countries in the sample.
+Another way to bring out this pattern in a graphical way is
+Figure~\ref{fig:glmean}. It depicts the inflation rates (in gray) and a 12-month rolling mean (in red),
+calculated again as the first moment of a fitted GL distribution.
+The strong increase in mean inflation during the boom years of 2007/2008
+accompanied by an equally strong deflation following the burst of the bubble
+is clearly visible for most countries. However, the level of mean inflation
+did not change persistently, as the inflation rates seem to have returned to
+their prior path, conveying the same message as the inflation series of the
+euro area.
+
+As for the structural break methods adopted in the previous sections,
+this means that such rapid and non-persistent effects in the mean level
+are not picked up from low frequency data, like inflation rates. Instead,
+they may simply be attributed to increased variance.
+
+Apart from bringing out more details about the short-term crisis effects,
+Figure~\ref{fig:glmean} also provides an alternative view on the
+long-range convergence behavior of the inflation rates studied in the
+previous sections. Over the 1990s and early 2000s, there is a clear
+general trend towards some common -- albeit broad -- mean level. 
+There is still some visible difference in the inflation rates, but the mean
+estimates are moving closer in a rather broad band. So contrary to
+\citet{palomba}, who find that the degree of similarity in short-run inflation
+dynamics is still weak, we find evidence of convergence within the countries in
+the sample.
 
@@
 
-\clearpage
 \section{Data and sales periods}\label{sec:sales}

### 2011-06-06 pkg/paper/Paper/draft.Rnw
small typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@112 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 The results from the previous sections convey that the financial
-crisis of 2008/9 did not have a significant effect on the inflation dynamics
+crisis of 2008/2009 did not have a significant effect on the inflation dynamics
 in most of the European countries (except Ireland and Spain).

### 2011-06-09 pkg/paper/Paper/draft.Rnw
draft changed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@115 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
+\renewcommand{\abstractname}{Summary}
 \Abstract{
-  In this paper, we test for structural breaks in the inflation dynamics of 21 European countries. To capture
-  changes in the first three moments of the inflation rates, we assumed them to be distributed according to 
-  a generalized logistic distribution. We find breaks in all but 4 countries.
-  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a change
-  towards negative skewness.
+	In an attempt to assess the effects of the EMU on inflation rate dynamics of its member states, this paper tests for structural breaks in the inflation dynamics of 21 European   countries.
+To capture   changes in the first three moments of the inflation rates, a model is adopted that makes use of the generalized logistic distribution.
+	There are   breaks in all but 4 countries.  Also, the differences between the inflation dynamics and its changes  are considerable. 
+  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a decrease in skewness.
   %% avoid "we"-constructs
@@
   E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
+  Telephone: +43/ 512/ 507 - 7103 \\
+  Fax: 43/ 512/ 507 - 2851 \\
 }
@@
 
-\readme{perhaps skip this paragraph?} The way towards the EMU basically consisted of three stages: stage~I
-(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
-convergence, and stage~III (1999--2001) the transition period, which ended with
-the introduction of the euro as legal tender. A possible fourth stage is the
-continuing integration of new member states, mostly former communist countries.
-
+\fixme{affiliation should appear on title page}
 Since the beginning of the European Union, the topic of a common currency
@@
 
+Basically, the way towards the EMU basically consisted of three stages: stage~I
+(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
+convergence, and stage~III (1999--2001) the transition period, which ended with
+the introduction of the euro as legal tender. A possible fourth stage is the
+continuing integration of new member states, mostly former communist countries.
+\readme{problem of breaks addressed} The possible influence of these different stages on the inflation rate dynamics
+suggested itself for a more thorough investigation.
+
+To inquire into possible changes in the dynamics of the inflation rate following
+these stages, which may have led to considerable alterations, seems a promising task.
+
 However, a deeper inquiry into the effects of monetary unions on the inflation rates of its member
@@
 
-\fixme{comment on results here} Overall, we find four periods during which structural breaks
-happened. Two of them in the 1990s (early and late), another one in the early 2000s, and the last period covers the financial crisis. There is also a 
-tendency towards a lower mean inflation rate, combined with a decreasing difference between
-individual inflation rates.
+%Overall, we find four periods during which structural breaks
+%happened. Two of them in the 1990s (early and late), another one in the early 2000s, and the last period covers the financial crisis. There is also a 
+%tendency towards a lower mean inflation rate, combined with a decreasing difference between
+%individual inflation rates.
+
+\readme{more about aggregation...} Overall, we find that the countries of the core euro--area (as defined in Section~\ref{sec:results}) did not change after the second stage and ended up with a low mean and variance.
+The southern euro--area member states and Ireland decreased their inflation rates almost to a level of the core countries but with a higher variance and a non--symmetric skewness.
+The Eastern European countries did reduce mean and variance significantly, especially the two countries that joined the euro--zone first.
+In the case of the Nordic non--euro countries, there is either no change at all or a clear decline in mean and variance after overcoming national crisis.
+Generally, inflation rates, which were very different in the beginning of the sample, are becoming more similar in terms of all three moments.
 
@@
 
-\begin{figure}[t!]
-<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
-par(mfrow = c(1, 2))
-plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
-lines(x$breakpoints, breaks = x$nbreaks)
-lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
-if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
-plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
-lines(x$breakpoints, breaks = x$nbreakss)
-@
-\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
-with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
-from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
-with sequence of $\mathit{LM}$~statistics,
-critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}
-
-<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
-par(mfrow = c(1, 2))
-for(i in 1:2) plot(x$model[[i]], moments = TRUE,
-  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
-  breaks = seq(-1.5, 2, by = 0.25),
-  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
-    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
-    sep = ""))
-@
-\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
-Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
-along with fitted generalized logistic probability density function, associated
-moments, and $\chi^2$ goodness-of-fit test.}
-\end{figure}
-
 Assuming a single stable set of parmaters (\ref{eq:h0}) is  not valid, as the
@@
 PIIGS  (\Sexpr{paste(colnames(hicps)[piigs], collapse = ", ")}), and the
-rest.
+other countries (\Sexpr{paste(colnames(hicps)[others], collapse = ", ")}).
 
@@
 @
-
-\begin{table}[p!]
-\begin{center}
-\footnotesize
-\begin{tabular}{llrrrlll}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
-\hline
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{No change}} \\
-\hline
-<<moments-table1, results=tex>>=
-writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
-\hline
-<<moments-table2, results=tex>>=
-writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
-\hline
-<<moments-table3, results=tex>>=
-writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Financial crisis}} \\
-\hline
-<<moments-table4, results=tex>>=
-writeLines(unlist(lapply(c(11, 19), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Non-euro}} \\
-\hline
-<<moments-table5, results=tex>>=
-writeLines(unlist(lapply(c(20, 21), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Eastern countries}} \\
-\hline
-<<moments-table6, results=tex>>=
-writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Euro area}} \\
-\hline
-<<moments-table6, results=tex>>=
-writeLines(unlist(lapply(22, format_moments)))
-@
-
-\end{tabular}
-
-\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
-distribution per segment, and ERM information for all countries under investigation.}
-\end{center}
-\end{table}
 
@@
 
-Finally, the inflation rates for the whole euro area \fixme{Is EU euro area or European union?},
-as provided by the OECD, are considered and three different regimes are found: (1)~until 
+Finally, the inflation rates for the whole euro area\footnote{This is the euro--area 17 countries series as provided by the \citet{oecd} and covers the total euro--area.
+This series is calculated as a weighted average of the HICPs of now seventeen euro countries and starts in January 1996.}, are considered and three different regimes are found: (1)~until
 1999 with both a low mean and variance, (2)~the third and final stage of monetary integration,
@@
 
+\readme{perhaps skip?} One clearly visible trend is the rise in inflation volatility in some of the euro countries (Austria, Greece, Italy, Luxembourg and Spain) and the  series for the whole euro area.
+\citet{emerson} emphasize  the fact that a high inflation rate is also more
+variable and uncertain and thus causes more relative price variability, leading
+to a less efficient price mechanism. From an economic point of view, a very volatile inflation rate will 
+likely  contribute to greater macroeconomic instability. In the literature, there is no consensus about the effects of higher 
+inflation volatility on the inflation rate. \citet{grier} find no evidence 
+that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA. However,
+\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. Naturally, an
+efficient price mechanism will benefit from inflation rates that do not diverge too much from the expectations.
+
 \fixme{some concluding remarks about the EU level bridging the current gap to the next subsection}
+In any case, the EU series as a whole -- which is already an aggregate -- does sum up the general trend, namely the low inflation rates during the late 90s, a higher level until the financial crisis and a lower mean inflation rate thereafter.
+However, this series is too highly aggregated and starts too late to be able to draw precise conclusions regarding the convergence issue.
 
@@
 
+\clearpage
 \subsection{Convergence on an aggregate level}\label{sec:aggregate}
 
-\begin{table}[t!]
-\begin{center}
-\begin{tabular}{llrrrr}
-\hline
-Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
-\hline
-
-\noalign{\medskip}
-\multicolumn{6}{l}{\emph{Core euro}} \\
-\hline
-<<cross-country-table1, echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
-@
-
-\noalign{\medskip}
-\multicolumn{6}{l}{\emph{PIIGS}} \\
-\hline
-<<cross-country-table2, echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
-@
-
-\noalign{\medskip}
-\multicolumn{6}{l}{\emph{Other}} \\
-\hline
-<<cross-country-table3, echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
-@
-
-\end{tabular}
-\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
-January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
-changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
-\end{center}
-\end{table}
-
 \fixme{intro needed}
 
+In order to both aggregate the data to a number of sub--groups but avoid too high a level of compression, as is the case of the EU series, 
+the countries are are grouped into the ``core'' of the eurozone, the PIIGS countries  and all the others. For this, we took account of possible
+breaks in the moments. These groups are chosen so as to reflect the general dynamics of European inflation rates during the time period of interest.
+
+The result of this is presented in Table~\ref{tab:results2} that shows lower quartile,
+median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
+inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}.
+
 \fixme{also ``zoom out'' more, a big picture needs to be established}
-
-Table~\ref{tab:results2} presents lower quartile,
-median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
-inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}. The countries are grouped
-into the ``core'' of the eurozone, the PIIGS countries  and all the others. For this, we took account of possible
-breaks in the moments.
-
 The time periods are chosen according to four fundamental change periods. Namely, the early 1990s, with breaks in
 Austria, Finland, France, Sweden, and UK. The late 1990s, with breaks in Greece, Italy, Luxembourg, Portugal, Czech Republic,
-Estonia, and Hungary. And also the period after the second phase of convergence, with breaks in the early 2000s (Poland, Slovakia, and Slovenia)
-and the two countries with the most visible effect of the financial crisis (Ireland and Spain).
-
-In the core-zone, there is a decrease in the lower quartile from \Sexpr{yrs[1]} to
+Estonia, and Hungary. The period after the second phase of convergence, with breaks in the early 2000s (Poland, Slovakia, and Slovenia)
+and the last period consisting of two countries with the most visible effect of the financial crisis (Ireland and Spain).
+
+In the core--zone, there is a decrease in the lower quartile from \Sexpr{yrs[1]} to
 \Sexpr{yrs[2]}. However, the interquartile range decreased by almost 20\%. The variance increased
@@
 stayed symmetric in the median, with almost an 8\% increase in the interquartile range.
+\readme{overview}Generally, there are no dramatical changes in the core--zone and after the initial stages, none of the core--countries
+had another break. This reflects the fact that their respective monetary policies were already closely connected.
 
@@
 in \Sexpr{yrs[4]}. The interquartile range did not change much too. Median skewness changed considerable 
-from highly positive to negative values. This is mostly due to the changes associated with the financial crisis, that was
-responsible for deflationary shocks. The interquartile range of skewness decreased considerably.
+from highly positive to negative values. 
+\readme{overview, skewness effects} The decline in skewness and the narrowing of the interquartile range reflects the
+fact that significant shocks to the inflation rates, which for some series were quite common in the past, have become something of a rarity. 
+Still, the PIIGS countries tend to come closer towards the core--countries on behalf of the general trend of lower values and 
+narrower bands.
 
@@
 interquartile range. The skewness decreased as well, as did the interquartile range of the skewness parameter.
+\readme{overview} The other countries started with higher mean inflation values -- since some of them are Eastern European countries -- and 
+effectually decreased the median inflation to a level much more similar to the core--zone than in \Sexpr{yrs[1]}. The bulk of changes 
+in these countries took effect after \Sexpr{yrs[2]}, as a result of both an ending of the transitionary phase (from planned to free market economies) and the intended entry
+into the EMU.
 
@@
 evidence of a convergence towards some common mean inflation rate in the countries of our sample, both within the subgroups and the three groups put together.
-This is also visible in the median values, that  diverge much less than at the beginning in \Sexpr{yrs[1]}.
+Regarding the variance, the PIIGS countries, and the core countries as well, saw an increase over time, which is mostly due to the influence of the financial
+crisis with its ``up--down--up'' turn (as depicted in Figure~\ref{fig:glmean}) leading to a higher variance.
 
 \subsection{The influence of the financial crisis}\label{sec:fcrisis}
-
-\setkeys{Gin}{width=0.8\textwidth}
-\begin{figure}[t!]
-\begin{center}
-<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
-  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
-  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
-    xlab = "Time", ylab = "Monthly inflation rates")
-  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-@
-\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
-12-month rolling GL fits (red) over time for all countries.}
-\end{center}
-\end{figure}
-\setkeys{Gin}{width=\textwidth}
 
@@
 
-More specifically, the mean inflation of the euro area \fixme{see above} from a GL fit in the 
+More specifically, the mean inflation of the euro area  from a GL fit in the 
 year before the crisis (\Sexpr{yearmon(2007 + 7/12)}--\Sexpr{yearmon(2008 + 6/12)}) was
@@
 
-\readme{start with 1) short recap of what we did, 2) conclusions}
-
 To assess changes in the dynamics of the inflation rates of a number of European countries both within
-and without the EMU, we developed a new method for estimating structural breaks resting on the 
-assumption that the inflation rates are GL distributed.
+and without the EMU, a new method for estimating structural breaks resting on the 
+assumption that the inflation rates are GL distributed has been  developed.
 
@@
 adjusted inflation rate series and that failing to account for a change in the skewness of the 
-distribution might lead to wrong conclusions concerning the dynamics.
-
-The results of \citet{holte}, whose simulations indicated that the standard deviation of the home 
-CPI inflation rate can be substantially reduced by joining a monetary union, can not be empirically validated.
-
-For the countries in the core of the eurozone, one clearly visible trend is the rise in 
-inflation volatility in almost half of the euro countries. \citet{emerson} emphasize  the fact that a high inflation rate is also more
-variable and uncertain and thus causes more relative price variability, leading
-to a less efficient price mechanism. From an economic point of view, a very volatile inflation rate will 
-likely  contribute to greater macroeconomic instability. In the literature, there is no consensus about the effects of higher 
-inflation volatility on the inflation rate. \citet{grier} find no evidence 
-that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA. However,
-\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. Naturally, an
-efficient price mechanism will benefit from inflation rates that do not diverge too much from the expectations.
-
-The good news is that mean inflation declined considerably, most notably in the Eastern European countries that clearly 
-benefited from the monetary discipline that is demanded for those countries who want to join the EMU.
+distribution might lead to potentially misleading conclusions concerning the dynamics. 
+The decline in size and frequency of mostly positive outliers in many inflation rate series 
+would have gone unnoticed without the skewness parameter to capture these changes.
+
+At the beginning of the sample, many countries did have very high inflation rates, especially the Eastern and Southern countries. 
+For the core--zone, convergence was already achieved in \Sexpr{yrs[2]}. The PIIGS countries lowered their inflation rates considerably in preparation 
+to the adoption of the euro.
+Over the time --  however -- many inflation rates declined considerably and are now much more in line with the inflation rates of the traditional low inflation countries of the core--euro zone.
+This is reflected especially well in the Eastern European countries, that lowered their high initial inflation level in the late 90s or prior to the entry into the ERM~II.
+They clearly benefited from the monetary discipline that is demanded for those countries who want to join the EMU.
 
@@
 
+\begin{figure}[htp!]
+<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
+par(mfrow = c(1, 2))
+plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
+lines(x$breakpoints, breaks = x$nbreaks)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
+plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
+lines(x$breakpoints, breaks = x$nbreakss)
+@
+\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
+with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
+from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
+with sequence of $\mathit{LM}$~statistics,
+critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[t!]
+<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
+par(mfrow = c(1, 2))
+for(i in 1:2) plot(x$model[[i]], moments = TRUE,
+  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
+  breaks = seq(-1.5, 2, by = 0.25),
+  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
+    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
+    sep = ""))
+@
+\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
+Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
+along with fitted generalized logistic probability density function, associated
+moments, and $\chi^2$ goodness-of-fit test.}
+\end{figure}
+
+\clearpage
+\begin{figure}[t!]
+\begin{center}
+<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
+  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
+  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
+    xlab = "Time", ylab = "Monthly inflation rates")
+  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+@
+\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
+12-month rolling GL fits (red) over time for all countries.}
+\end{center}
+\end{figure}
+
+\clearpage
+
 Figures~\ref{fig:nochange}--\ref{fig:eastern} depict the data underlying the results
@@
 
-\setkeys{Gin}{width=0.8\textwidth}
-\begin{figure}[p!]
+\begin{figure}[htp!]
 \begin{center}
@@
 \caption{\label{fig:nochange} No change.}
-
+\end{center}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[p!]
+\begin{center}
 <<graphs-phase1, fig=TRUE, height=4.9, width=7.5>>=
@@
 \end{figure}
+
+\clearpage
 
@@
 \caption{\label{fig:phase2} Phase~2 convergence.}
-
+\end{center}
+\end{figure}
+
+\clearpage
+\begin{figure}[p!]
+\begin{center}
 <<graphs-crisis, fig=TRUE, height=3.5, width=7.5>>=
@@
 
+\clearpage
 \begin{figure}[p!]
@@
 \caption{\label{fig:noneuro} Non-euro.}
-
+\end{center}
+\end{figure}
+
+\clearpage
+\begin{figure}[p!]
+\begin{center}
 <<graphs-eastern, fig=TRUE, height=9.1, width=7.5>>=
@@
 
+\clearpage
+\section{Tables}
+
+\begin{table}[t!]
+\begin{center}
+\footnotesize
+\begin{tabular}{llrrrlll}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{No change}} \\
+\hline
+<<moments-table1, results=tex>>=
+writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
+\hline
+<<moments-table2, results=tex>>=
+writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
+\hline
+<<moments-table3, results=tex>>=
+writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Financial crisis}} \\
+\hline
+<<moments-table4, results=tex>>=
+writeLines(unlist(lapply(c(11, 19), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Non-euro}} \\
+\hline
+<<moments-table5, results=tex>>=
+writeLines(unlist(lapply(c(20, 21), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Eastern countries}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Euro area}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(22, format_moments)))
+@
+
+\end{tabular}
+
+\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
+distribution per segment, and ERM information for all countries under investigation.}
+\end{center}
+\end{table}
+
+\clearpage
+\begin{table}[t!]
+\begin{center}
+\begin{tabular}{llrrrr}
+\hline
+Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
+\hline
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Core euro}} \\
+\hline
+<<cross-country-table1, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{PIIGS}} \\
+\hline
+<<cross-country-table2, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Other}} \\
+\hline
+<<cross-country-table3, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
+@
+
+\end{tabular}
+\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
+January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
+changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
+\end{center}
+\end{table}
+
 \end{appendix}

### 2011-06-09 pkg/paper/Paper/draft.Rnw
update draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@116 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-\author{Thomas Windberger\\Universit\"at Innsbruck \And
-        Achim Zeileis\\Universit\"at Innsbruck}
+%\author{Thomas Windberger\\Universit\"at Innsbruck \And
+%        Achim Zeileis\\Universit\"at Innsbruck}
+\author{\hfill Thomas Windberger$^a$ \hfill Achim Zeileis$^b$\thanks{Correspondence
+to: Achim Zeileis, Department of Statistics,  Universit\"at Innsbruck,
+Universit\"atsstra{\ss}e 15, 6020 Innsbruck, Austria. Tel: +43/512/507-7103. Fax: +43/512/507-2851.
+E-mail: \email{Achim.Zeileis@R-project.org}} \hfill \hfill \\
+{\small \it $^a$ Department of Statistics, Universit\"at Innsbruck, Austria} \\
+{\small \it $^b$ Department of Statistics, Universit\"at Innsbruck, Austria}}
 \Plainauthor{Thomas Windberger, Achim Zeileis}
@@
   E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
-  Telephone: +43/ 512/ 507 - 7103 \\
-  Fax: 43/ 512/ 507 - 2851 \\
+  Telephone: +43/12/507-7103 \\
+  Fax: +43/512/507-2851 \\
 }
@@
 
+\begin{figure}[t!]
+<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
+par(mfrow = c(1, 2))
+plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
+lines(x$breakpoints, breaks = x$nbreaks)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
+plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
+lines(x$breakpoints, breaks = x$nbreakss)
+@
+\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
+with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
+from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
+with sequence of $\mathit{LM}$~statistics,
+critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}
+
+<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
+par(mfrow = c(1, 2))
+for(i in 1:2) plot(x$model[[i]], moments = TRUE,
+  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
+  breaks = seq(-1.5, 2, by = 0.25),
+  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
+    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
+    sep = ""))
+@
+\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
+Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
+along with fitted generalized logistic probability density function, associated
+moments, and $\chi^2$ goodness-of-fit test.}
+\end{figure}
+
 Assuming a single stable set of parmaters (\ref{eq:h0}) is  not valid, as the
@@
 @
+
+\begin{table}[p!]
+\begin{center}
+\footnotesize
+\begin{tabular}{llrrrlll}
+\hline
+Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
+\hline
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{No change}} \\
+\hline
+<<moments-table1, results=tex>>=
+writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
+\hline
+<<moments-table2, results=tex>>=
+writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
+\hline
+<<moments-table3, results=tex>>=
+writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Financial crisis}} \\
+\hline
+<<moments-table4, results=tex>>=
+writeLines(unlist(lapply(c(11, 19), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Non-euro}} \\
+\hline
+<<moments-table5, results=tex>>=
+writeLines(unlist(lapply(c(20, 21), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Eastern countries}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
+@
+
+\noalign{\medskip}
+\multicolumn{8}{l}{\emph{Euro area}} \\
+\hline
+<<moments-table6, results=tex>>=
+writeLines(unlist(lapply(22, format_moments)))
+@
+
+\end{tabular}
+
+\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
+distribution per segment, and ERM information for all countries under investigation.}
+\end{center}
+\end{table}
 
@@
 
-\clearpage
 \subsection{Convergence on an aggregate level}\label{sec:aggregate}
+
+\begin{table}[t!]
+\begin{center}
+\begin{tabular}{llrrrr}
+\hline
+Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
+\hline
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Core euro}} \\
+\hline
+<<cross-country-table1, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{PIIGS}} \\
+\hline
+<<cross-country-table2, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Other}} \\
+\hline
+<<cross-country-table3, echo=FALSE, results=tex>>=
+writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
+@
+
+\end{tabular}
+\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
+January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
+changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
+\end{center}
+\end{table}
 
@@
 \subsection{The influence of the financial crisis}\label{sec:fcrisis}
+
+\setkeys{Gin}{width=0.8\textwidth}
+\begin{figure}[t!]
+\begin{center}
+<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
+  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
+  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
+  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
+    xlab = "Time", ylab = "Monthly inflation rates")
+  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
+@
+\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
+12-month rolling GL fits (red) over time for all countries.}
+\end{center}
+\end{figure}
+\setkeys{Gin}{width=\textwidth}
 
@@
 
-\begin{figure}[htp!]
-<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
-par(mfrow = c(1, 2))
-plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
-lines(x$breakpoints, breaks = x$nbreaks)
-lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
-if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
-plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
-lines(x$breakpoints, breaks = x$nbreakss)
-@
-\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
-with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
-from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
-with sequence of $\mathit{LM}$~statistics,
-critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[t!]
-<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
-par(mfrow = c(1, 2))
-for(i in 1:2) plot(x$model[[i]], moments = TRUE,
-  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
-  breaks = seq(-1.5, 2, by = 0.25),
-  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
-    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
-    sep = ""))
-@
-\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
-Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
-along with fitted generalized logistic probability density function, associated
-moments, and $\chi^2$ goodness-of-fit test.}
-\end{figure}
-
-\clearpage
-\begin{figure}[t!]
-\begin{center}
-<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
-  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
-  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
-    xlab = "Time", ylab = "Monthly inflation rates")
-  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
-@
-\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
-12-month rolling GL fits (red) over time for all countries.}
-\end{center}
-\end{figure}
-
-\clearpage
-
 Figures~\ref{fig:nochange}--\ref{fig:eastern} depict the data underlying the results
@@
 
-\begin{figure}[htp!]
+\setkeys{Gin}{width=0.8\textwidth}
+\begin{figure}[p!]
 \begin{center}
@@
 \caption{\label{fig:nochange} No change.}
-\end{center}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[p!]
-\begin{center}
+
 <<graphs-phase1, fig=TRUE, height=4.9, width=7.5>>=
@@
 \end{figure}
-
-\clearpage
 
@@
 \caption{\label{fig:phase2} Phase~2 convergence.}
-\end{center}
-\end{figure}
-
-\clearpage
-\begin{figure}[p!]
-\begin{center}
+
 <<graphs-crisis, fig=TRUE, height=3.5, width=7.5>>=
@@
 
-\clearpage
 \begin{figure}[p!]
@@
 \caption{\label{fig:noneuro} Non-euro.}
-\end{center}
-\end{figure}
-
-\clearpage
-\begin{figure}[p!]
-\begin{center}
+
 <<graphs-eastern, fig=TRUE, height=9.1, width=7.5>>=
@@
 
-\clearpage
-\section{Tables}
-
-\begin{table}[t!]
-\begin{center}
-\footnotesize
-\begin{tabular}{llrrrlll}
-\hline
-Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
-\hline
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{No change}} \\
-\hline
-<<moments-table1, results=tex>>=
-writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
-\hline
-<<moments-table2, results=tex>>=
-writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
-\hline
-<<moments-table3, results=tex>>=
-writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Financial crisis}} \\
-\hline
-<<moments-table4, results=tex>>=
-writeLines(unlist(lapply(c(11, 19), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Non-euro}} \\
-\hline
-<<moments-table5, results=tex>>=
-writeLines(unlist(lapply(c(20, 21), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Eastern countries}} \\
-\hline
-<<moments-table6, results=tex>>=
-writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
-@
-
-\noalign{\medskip}
-\multicolumn{8}{l}{\emph{Euro area}} \\
-\hline
-<<moments-table6, results=tex>>=
-writeLines(unlist(lapply(22, format_moments)))
-@
-
-\end{tabular}
-
-\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
-distribution per segment, and ERM information for all countries under investigation.}
-\end{center}
-\end{table}
-
-\clearpage
-\begin{table}[t!]
-\begin{center}
-\begin{tabular}{llrrrr}
-\hline
-Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
-\hline
-
-\noalign{\medskip}
-\multicolumn{6}{l}{\emph{Core euro}} \\
-\hline
-<<cross-country-table1, echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
-@
-
-\noalign{\medskip}
-\multicolumn{6}{l}{\emph{PIIGS}} \\
-\hline
-<<cross-country-table2, echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
-@
-
-\noalign{\medskip}
-\multicolumn{6}{l}{\emph{Other}} \\
-\hline
-<<cross-country-table3, echo=FALSE, results=tex>>=
-writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
-@
-
-\end{tabular}
-\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
-January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
-changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
-\end{center}
-\end{table}
-
 \end{appendix}

### 2011-06-11 pkg/paper/Paper/draft.Rnw
(close to) final touches of Z
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@117 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2011-06-06 pkg/paper/Paper/draft.Rnw
@@
         Achim Zeileis\\Universit\"at Innsbruck}
+%% \author{\hfill Thomas Windberger \hfill Achim Zeileis\thanks{Correspondence
+%% to: Achim Zeileis, Department of Statistics,  Universit\"at Innsbruck,
+%% Universit\"atsstr.~15, A-6020 Innsbruck, Austria. Tel: +43/512/507-7103. Fax: +43/512/507-2851.
+%% E-mail: \email{Achim.Zeileis@R-project.org}} \hfill \hfill \\
+%% {\small \it Department of Statistics, Universit\"at Innsbruck, Austria}}
 \Plainauthor{Thomas Windberger, Achim Zeileis}
 
+\renewcommand{\abstractname}{Summary}
 \Abstract{
-  In this paper, we test for structural breaks in the inflation dynamics of 21 European countries. To capture
-  changes in the first three moments of the inflation rates, we assumed them to be distributed according to 
-  a generalized logistic distribution. We find breaks in all but 4 countries.
-  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a change
-  towards negative skewness.
-  %% avoid "we"-constructs
-  %% "adopt" better than "assume"
+  To assess the effects of the EMU on inflation rate dynamics of
+  its member states, the inflation rate series for 21~European countries
+  are investigated for structural changes. To capture changes in mean,
+  variance, and skewness of inflation rates, a generalized logistic
+  model is adopted and complemented with structural break tests and
+  breakpoint estimation techniques. These reveal considerable differences
+  in the patterns of inflation dynamics and the structural changes
+  therein. Overall, there is a convergence towards a lower mean inflation rate
+  with reduced skewness, but accompanied by an increase in variance.
 }
@@
   E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
+  Telephone: +43/12/507-7103 \\
+  Fax: +43/512/507-2851 \\
 }
@@
 
-\readme{perhaps skip this paragraph?} The way towards the EMU basically consisted of three stages: stage~I
-(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
-convergence, and stage~III (1999--2001) the transition period, which ended with
-the introduction of the euro as legal tender. A possible fourth stage is the
-continuing integration of new member states, mostly former communist countries.
-
-Since the beginning of the European Union, the topic of a common currency
+Since the beginning of the European Union (EU), the topic of a common currency
 was a controversial issue. Although the Economic and Monetary Union (EMU) is now
@@
 settled. The controversial topics range from the question of whether or not the
-eurozone is indeed an optimal currency area \citep[as developed in:][]{Mundell:1961}, 
-all the way to the very survival of the euro in light of the budgetary problems of some of its member
-states. The effects of monetary unions on a number of macroeconomic indicators,
-with inflation being the most important one in this context, is in the center of
-an ongoing debate. This concerns the issue of short--run and steady--state
-inflation uncertainty -- in context to inflation expectations -- dealt with in
-\citet{cap}, or the degree of similarity of short--run dynamic properties of the inflation rates in EU
-countries, which is the topic of an investigation by \citet{palomba}.
-
-%Any new state that wants to join the EMU is first obliged to fulfill the
-%Maastricht criteria, which, besides setting rules for government debts and
-%interest rates, also requires the participation in the ERM~II (Exchange Rate
-%Mechanism~II) for two years, while the exchange rates towards the euro are not
-%allowed to cross the nominal band.
-
-Throughout the literature, there is still a
-considerable degree of uncertainty as to what extent the introduction of the
-euro, or monetary unions in general, affect the inflation rate.
-
-However, a deeper inquiry into the effects of monetary unions on the inflation rates of its member
-countries is in need of a method able to
-capture changes in inflation dynamics. We contribute to this field of research by
-developing a new method for testing for structural breaks in the dynamics of the
-inflation rates. The framework utilized in this paper allows testing for any 
-possible alternations  and thus helps interpreting 
-the effects of the EMU. We focus on changes in the first three moments of these 
-inflation rates, which we assume to be distributed according to a generalized logistic 
-distribution (GL). For this purpose, we build on previous research by \citet{Zeileis+Hornik:2007}.
-
-This \fixme{describe method} framework allows us to assess the evolution of mean, variance, and skewness over time.
-Our estimation technique enables the simultaneous estimation of breaks in these moments. 
-The skewness parameter turns out to be quite important to measure alterations in the
-long run inflation dynamics, as it changes considerably over the time of the
-analysis. This trend would be all but invisible using a normal distribution.
-
-\fixme{comment on results here} Overall, we find four periods during which structural breaks
-happened. Two of them in the 1990s (early and late), another one in the early 2000s, and the last period covers the financial crisis. There is also a 
-tendency towards a lower mean inflation rate, combined with a decreasing difference between
-individual inflation rates.
+eurozone is indeed an optimal currency area \citep[as developed in][]{Mundell:1961},
+all the way to the very survival of the euro in light of
+the budgetary problems of some of its member states. The effects of monetary
+unions on a number of macroeconomic indicators, with inflation being of increased
+importance, is at the center of an ongoing debate. This
+concerns the issue of short-run and steady-state inflation uncertainty -- in relation
+to inflation expectations -- as dealt with in \citet{cap}, or the degree of
+similarity of short-run dynamic properties of the inflation rates in EU
+countries, which is the topic of the investigation by \citet{palomba}. 
+Throughout the literature, there is still a considerable degree of uncertainty
+as to what extent the introduction of the euro, or monetary unions in general,
+affects the inflation rate.
+
+As the way towards the EMU essentially consisted of three stages (I:~1990--1994,
+a phase of liberalization; II: 1995--1998, a phase of convergence; III: 1999--2001,
+a transition period ending with the introduction of the euro as legal tender),
+it seems natural to ask whether this transition was accompanied by structural changes
+in the inflation dynamics in the different European countries. Furthermore, a
+possible fourth stage is the continuing integration of new member states, mostly
+former communist countries. Typically, changes in the mean level of the
+inflation rates are of most interest but there are also reasons for investigating
+changes in higher moments, such as variance and skewness. \citet{emerson} emphasize that
+a high inflation rate is usually also more variable and uncertain and thus
+causes more relative price variability, leading to a less stability and a less
+efficient price mechanism. Furthermore, \cite{jaro} finds a positive correlation
+between the level and the standard deviation of inflation.
+
+%% From an economic point of view, a very volatile inflation rate will 
+%% likely  contribute to greater macroeconomic instability. In the literature,
+%% there is no consensus about the effects of higher  inflation volatility on the
+%% inflation rate. \citet{grier} find no evidence that higher inflation
+%% uncertainty raises the average inflation rate -- at least in the case of the
+%% USA. Naturally, an efficient price mechanism will
+%% benefit from inflation rates that do not diverge too much from the expectations.
+
+%% Any new state that wants to join the EMU is first obliged to fulfill the
+%% Maastricht criteria, which, besides setting rules for government debts and
+%% interest rates, also requires the participation in the ERM~II (Exchange Rate
+%% Mechanism~II) for two years, while the exchange rates towards the euro are not
+%% allowed to cross the nominal band.
+
+Therefore, a thorough investigation of the evolution of the inflation dynamics
+within a country over time requires assessment not only of the mean level but
+also higher moments. Hence, we contribute to this field of research by
+developing a new method for testing for structural breaks based on a generalized
+logistic (GL) distribution. This provides flexible modeling of means, variances,
+and skewness, while also allowing for fat tails. Structural breaks are tested
+using score-based tests \citep{Andrews:1993,Zeileis:2005} and breakpoints are estimated
+by extending the least squares approach of \cite{bp03} to (quasi-)maximum likelihood
+estimation of GL distributions \citep[similar in spirit to the Gaussian model
+employed by][]{Zeileis+Shah+Patnaik:2010}.
+
+The empirical results show that there are different patterns and timings of
+structural breaks in the inflation dynamics of different groups of countries.
+Inflation rates in the core euro area did not change significantly after
+completion of the first convergence stage, yielding low mean and variance.
+The Southern euro area member states (and Ireland), on the other hand, managed
+to decrease the mean level of inflation in the mid-1990s, but at the price
+of higher volatility. Finally, the Eastern European countries reduced both mean
+and variance on their way towards joining the EMU. Generally, the inflation rates
+converged from very different distributions at the beginning of the sample to
+rather similar patterns at the end of the 2000s. Skewness also plays an
+important role in this convergence: Some Southern and Eastern countries
+have clear breaks from right-skewed to roughly symmetric inflation rates,
+signalling a reduction of size and frequency of outlying high inlation rates.
 
 The  remainder of this paper is
-structured as follows: Section~\ref{sec:data} presents the data,
-Section~\ref{sec:model} presents the model and the estimation techniques used,
-Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an
-example, Section~\ref{sec:results} presents the results, and
+structured as follows: Section~\ref{sec:data} briefly presents the data,
+Section~\ref{sec:model} introduces the model and the estimation techniques used,
+%% Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an example, 
+Section~\ref{sec:results} discusses the empirical results, and
 Section~\ref{sec:conclusion} concludes.
 
-\clearpage
 \section{Data}\label{sec:data}
@@
 
-\clearpage
 \section{Model}\label{sec:model}
@@
 
-The GL distribution in the parameterization employed here is also used in \citet{Shao:2002}
-% The type~I GL distribution (also known as skew-logistic distribution), as defined in \citet{Abberger:2002}
-%\citet{johnson},
-and has the log-density function\footnote{Which is similar to the type~1 GL distribution as defined in \citet{johnson}.}
+The GL distribution employed here corresponds to the type~1 GL distribution 
+from\citet{johnson}, using the parametrization and inference as in \citet{Shao:2002}. It 
+has the log-density function
 %
@@
 
-\clearpage
 \section{Results}\label{sec:results}
@@
 core-euro (\Sexpr{paste(colnames(hicps)[euro], collapse = ", ")}),
-PIIGS  (\Sexpr{paste(colnames(hicps)[piigs], collapse = ", ")}), and the
-rest.
+PIIGS  (\Sexpr{paste(colnames(hicps)[piigs[c(4, 2, 3, 1, 5)]], collapse = ", ")}), and the
+other countries (\Sexpr{paste(gsub("CzechRepublic", "Czech Republic", colnames(hicps))[others], collapse = ", ")}).
 
@@
 
-Finally, the inflation rates for the whole euro area \fixme{Is EU euro area or European union?},
-as provided by the OECD, are considered and three different regimes are found: (1)~until 
-1999 with both a low mean and variance, (2)~the third and final stage of monetary integration,
-accompanied with a significant increase in the mean, (3)~a period encompassing
-the overheating of the economy that culminated in the financial crisis, showing a big increase
-in variance and a lower mean inflation.
-
-\fixme{some concluding remarks about the EU level bridging the current gap to the next subsection}
-
-%There is an increase in mean from
-%$\Sexpr{round3(eu_mom[1, 1])}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
-%$\Sexpr{round3(eu_mom[2, 1])}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
-%$\Sexpr{round3(eu_mom[3, 1])}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) shortly before the financial crises.
-%
+Finally, the inflation rates for the whole euro area\footnote{This is the
+euro area 17~countries series as provided by the \citet{oecd},
+starting in \Sexpr{start(eu) - 1/12}. It is calculated as a weighted average of the HICPs
+of now seventeen euro countries.} are considered and
+three different regimes are found: (1)~until 1999 with both a low mean and
+variance, (2)~the third and final stage of monetary integration, accompanied
+by a significant increase in the mean, (3)~a period encompassing the
+overheating of the economy and its culmination in the financial crisis, showing a
+large increase in variance and a lower mean inflation.
 
@@
 \caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
-January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
+Jan~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
 changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
@@
 
-\fixme{intro needed}
-
-\fixme{also ``zoom out'' more, a big picture needs to be established}
-
-Table~\ref{tab:results2} presents lower quartile,
-median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
-inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}. The countries are grouped
-into the ``core'' of the eurozone, the PIIGS countries  and all the others. For this, we took account of possible
-breaks in the moments.
-
-The time periods are chosen according to four fundamental change periods. Namely, the early 1990s, with breaks in
-Austria, Finland, France, Sweden, and UK. The late 1990s, with breaks in Greece, Italy, Luxembourg, Portugal, Czech Republic,
-Estonia, and Hungary. And also the period after the second phase of convergence, with breaks in the early 2000s (Poland, Slovakia, and Slovenia)
-and the two countries with the most visible effect of the financial crisis (Ireland and Spain).
-
-In the core-zone, there is a decrease in the lower quartile from \Sexpr{yrs[1]} to
-\Sexpr{yrs[2]}. However, the interquartile range decreased by almost 20\%. The variance increased
-slightly, but also with a 15\% decrease in interquartile range. The skewness
-stayed symmetric in the median, with almost an 8\% increase in the interquartile range.
-
-The PIIGS countries significantly decreased the median inflation rates from \Sexpr{yrs[1]} to \Sexpr{yrs[4]}. 
-Whereas the interquartile range was very tight in \Sexpr{yrs[1]}, it increased in \Sexpr{yrs[2]} and again in \Sexpr{yrs[4]}.
-However, the median values are much smaller. The variance increased from \Sexpr{yrs[1]} to \Sexpr{yrs[2]} but did not change much 
-in \Sexpr{yrs[4]}. The interquartile range did not change much too. Median skewness changed considerable 
-from highly positive to negative values. This is mostly due to the changes associated with the financial crisis, that was
-responsible for deflationary shocks. The interquartile range of skewness decreased considerably.
-
-The other countries were successful in decreasing all three moments from \Sexpr{yrs[1]} to \Sexpr{yrs[3]}. Still, their median
-inflation rate is higher than those of the core countries. Also, the interquartile range of the mean decreased considerably. 
-Contrary to the euro--zone and the PIIGS countries, variance decreased also, both in terms of the median value and the
-interquartile range. The skewness decreased as well, as did the interquartile range of the skewness parameter.
-
-In context of the convergence issue, we find that, with the interquartile range of the mean of all the aggregated zones declining and getting tighter, there is  
-evidence of a convergence towards some common mean inflation rate in the countries of our sample, both within the subgroups and the three groups put together.
-This is also visible in the median values, that  diverge much less than at the beginning in \Sexpr{yrs[1]}.
+Given the results above for the individual countries, a natural question is
+whether there is convergence on an aggregate level and, if so, which pattern it
+follows. Hence, we aggregate the results from Table~\ref{tab:results1} for the
+individual countries into three groups and at four time points to reflect the
+general dynamics of European inflation rates during the time period of interest.
+The groups have been selected based on prior expectations (i.e., not based on
+the breakpoints estimation results) as the ``core'' of the euro zone
+(\Sexpr{paste(colnames(hicps)[euro], collapse = ", ")}), the PIIGS countries,
+and the rest. Similarly, the time points have been chosen as to fall ``between''
+times where transitions in one or more groups of countries could be expected:
+Jan~\Sexpr{yrs[1]} during stage~I on the way towards the EMU, Jan~\Sexpr{yrs[2]}
+during stage~II, Jan~\Sexpr{yrs[3]} after transition into the EMU but before the
+financial crisis, and Jan~\Sexpr{yrs[4]} after the crisis. The result of these
+considerations is presented in Table~\ref{tab:results2} which extracts the three
+moments for each country from Table~\ref{tab:results1} at the indicated time
+points and aggregates them within the group. Lower quartile, median, and upper
+quartile (25\%, 50\%, 75\% percentiles) are provided to assess both changes in
+location and spread across countries.
+
+For the core euro countries, there are only changes from stage~I to stage~II
+because after \Sexpr{yrs[2]} no breaks have been detected. The changes in this
+first phase did not affect mean inflation much, while variance increased
+slightly (in both quartiles) and skewness somewhat decreased (from slightly
+right- to slightly left-skewed). The heterogeneity (assessed by the
+interquartile range) decreased for all three moments of the inflation rates.
+This reflects that the monetary policies of the corresponding countries were
+already closely connected prior to stage~I and now even moved somewhat closer
+together.
+
+For the PIIGS countries, changes occured from \Sexpr{yrs[1]} to
+\Sexpr{yrs[2]} and then again during the crisis, before \Sexpr{yrs[4]}.
+In the first phase, mean inflation was almost halved, accompanied by a
+doubling of variance and a change from clearly right-skewed to approximately
+symmetric. The latter reflects that the occurence of shocks with high inflation
+rates have been successfully reduced. Moreover, the changes in the PIIGS
+countries between \Sexpr{yrs[3]} and \Sexpr{yrs[4]} lead to a further
+decrease in mean inflation and associated skewness which are both caused by
+several extremely low inflation (or even deflation) rates during the time
+following the crisis.
+
+The other countries do not exhibit any significant breakpoints during the
+mid-1990s, but many changes occur in the transition period from \Sexpr{yrs[2]}
+to \Sexpr{yrs[3]}, encompassing both the completion of transformation from
+planned to free market economies and the intended entry into the EMU.
+During this time, the countries were successful in
+decreasing all three moments of their inflation distributions while
+simultaneously decreasing the heterogeneity (as captured by the corresponding
+interquartile ranges) across countries. Before the transition all three
+moments were extremely large compared to the other two groups, while afterwards
+all moments are much closer to those of the other groups. In Jan~\Sexpr{yrs[3]},
+mean inflation is in between the levels for the core euro and PIIGS countres,
+variance is still slightly higher, and skewness is not too far from symmetry.
+
+In context of the convergence issue, we find that up to \Sexpr{yrs[3]}
+the inflation dynamics of all three groups of countries become much
+more similar, both within and across groups. The interquartile ranges
+of the means decline and become tighter for all groups and there appears to
+be convergence towards some common mean inflation rate in the
+countries of our sample. Variances also move more closely together, corresponding
+to an increase for the core euro and PIIGS countries but a clear decrease
+for the others. Right-skewness (and thus large outlying inflation rates)
+has been reduced for all three groups of countries but to different extents.
+Thus, while ending in similar inflation dynamics, the timing and patterns of
+changes in this convergence process are rather different across
+the three groups, depending on the different starting situations of the
+countries.
 
@@
 
-\clearpage
 \section{Conclusion}\label{sec:conclusion}
 
-\readme{start with 1) short recap of what we did, 2) conclusions}
-
-To assess changes in the dynamics of the inflation rates of a number of European countries both within
-and without the EMU, we developed a new method for estimating structural breaks resting on the 
-assumption that the inflation rates are GL distributed.
-
-The results suggest that a GL distribution is appropriate for the modeling of the seasonally 
-adjusted inflation rate series and that failing to account for a change in the skewness of the 
-distribution might lead to wrong conclusions concerning the dynamics.
-
-The results of \citet{holte}, whose simulations indicated that the standard deviation of the home 
-CPI inflation rate can be substantially reduced by joining a monetary union, can not be empirically validated.
-
-For the countries in the core of the eurozone, one clearly visible trend is the rise in 
-inflation volatility in almost half of the euro countries. \citet{emerson} emphasize  the fact that a high inflation rate is also more
-variable and uncertain and thus causes more relative price variability, leading
-to a less efficient price mechanism. From an economic point of view, a very volatile inflation rate will 
-likely  contribute to greater macroeconomic instability. In the literature, there is no consensus about the effects of higher 
-inflation volatility on the inflation rate. \citet{grier} find no evidence 
-that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA. However,
-\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. Naturally, an
-efficient price mechanism will benefit from inflation rates that do not diverge too much from the expectations.
-
-The good news is that mean inflation declined considerably, most notably in the Eastern European countries that clearly 
-benefited from the monetary discipline that is demanded for those countries who want to join the EMU.
-
-\clearpage
+To assess changes in the dynamics of the inflation rates of a number of European countries
+-- both within and outside the EMU -- a new method for estimating structural breaks in
+inflation rates is proposed. The method allows to test for changes in mean, variance,
+and skewness of inflation rate series by adopting a (quasi-)likelihood model based on the generalized
+logistic (GL) distribution. Thus, unlike normally distributed models, this approach
+can deal with fat tails and (potentially changing) skewness.
+
+The empirical results suggest that a GL distribution is appropriate for modeling seasonally 
+adjusted inflation rate series and is able to uncover important changes in all three moments.
+Towards the end of the 2000s, before the financial crisis, our results broadly confirm
+convergence hypotheses as the first three moments from all countries in our sample
+move closer together.
+Especially the changes in skewness -- caused by a reduction of size and frequency of outlying high inflation
+rates in some countries -- might have gone unnoticed with traditional structural change techniques,
+leading to potentially misleading conclusions concerning the dynamics.
+
+At the beginning of the sample, many PIIGS and Eastern countries did have very high inflation
+rates, while for the core euro countries, convergence was already achieved in the mid-1990s.
+The PIIGS countries lowered their inflation rates considerably in preparation to the adoption
+of the euro. The Eastern European countries lowered their high initial inflation level in the
+late 1990s or early 2000s, prior to the entry into the ERM~II. They clearly benefited from the
+monetary discipline that is demanded for those countries who want to join the EMU.
+
 \section*{Computational details}
@@
 
+\section*{Acknowledgments}
+
+We are grateful to Jesus Crespo Cuaresma for comments and discussions about
+an earlier version of the manuscript, and also to the participants of the workshop
+of the Research Platform ``Empirical and Experimental Economics'' for stimulating discussions,
+leading to substantial improvements of the paper.
+
 \bibliography{papers}
+
+\end{document}
+\SweaveOpts{eval=FALSE}
 

### 2011-06-12 pkg/paper/Paper/draft.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@118 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
   therein. Overall, there is a convergence towards a lower mean inflation rate
-  with reduced skewness, but accompanied by an increase in variance.
+  with reduced skewness, but it is accompanied by an increase in variance.
 }
@@
 a phase of liberalization; II: 1995--1998, a phase of convergence; III: 1999--2001,
-a transition period ending with the introduction of the euro as legal tender),
+a transition period ending with the replacement of the national currencies by the euro),
 it seems natural to ask whether this transition was accompanied by structural changes
@@
 former communist countries. Typically, changes in the mean level of the
-inflation rates are of most interest but there are also reasons for investigating
+inflation rates attract the widest interest, but there are also reasons for investigating
 changes in higher moments, such as variance and skewness. \citet{emerson} emphasize that
 a high inflation rate is usually also more variable and uncertain and thus
-causes more relative price variability, leading to a less stability and a less
+causes more relative price variability, leading to a less
 efficient price mechanism. Furthermore, \cite{jaro} finds a positive correlation
@@
 Therefore, a thorough investigation of the evolution of the inflation dynamics
-within a country over time requires assessment not only of the mean level but
-also higher moments. Hence, we contribute to this field of research by
+within a country over time requires assessment not only of the mean level, but
+also of higher moments. Hence, we contribute to this field of research by
 developing a new method for testing for structural breaks based on a generalized
@@
 have clear breaks from right-skewed to roughly symmetric inflation rates,
-signalling a reduction of size and frequency of outlying high inlation rates.
+signaling a reduction of size and frequency of outlying high inflation rates.
 
@@
 Cyprus, and Malta are not included since they are very small economies. 
+%% and are also not members of the OECD
 
@@
 a Commission regulation \citep[see][for details]{eurlex}.} in the HICP, resulting in
-changes of the seasonal inflation patters over time within several countries
+changes of the seasonal inflation patterns over time within several countries
 and also between the countries. See \citet{pltm} for a detailed discussion.
@@
 However, unlike many standard least squares approaches \citep[such as][]{bp03},
-not only changes in the mean level should be addressed but also changes in
+not only changes in the mean level should be addressed, but also changes in
 variance and potential skewness.
@@
 (2)~how it can be estimated (under parameter stability), (3)~how parameter stability
-can be assesed, and (4)~how breakpoints can be estimated in the presence of parameter
+can be assessed, and (4)~how breakpoints can be estimated in the presence of parameter
 instability.
@@
 The GL distribution employed here corresponds to the type~1 GL distribution 
-from\citet{johnson}, using the parametrization and inference as in \citet{Shao:2002}. It 
+from \citet{johnson}, using the parametrization and inference as in \citet{Shao:2002}. It 
 has the log-density function
@@
 bridge: $\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot)$. Based on
-this various types of test statistics can be computed
+this, various types of test statistics can be computed
 (see \citealp{Zeileis:2005} for details). In the following, the 
@@
 %
-and the appropriate $p$~values can be computed using from the corresponding
+and the appropriate $p$~values can be computed from the corresponding
 limiting distribution -- with $\mathit{efp}(t)$ replaced by $W^0(t)$.
@@
 
-## Exchange Rate Mechanism information
-## ERM~II, ERM~I (part of european monetary system expired in 1999-01-01 gave way to ERM II
-## if never entered ""right?
-## http://web.archive.org/web/20080101022356/http://lexikon.meyers.de/meyers/Europ%C3%A4isches_W%C3%A4hrungssystem
 erm <- yearmon(c(
@@
 successful in realigning its economy and introduced a number of reforms leading
-to a stable growth in recent years. The good economic performance made it
+to stable growth in recent years. The good economic performance made it
 possible for Slovenia to enter the ERM~II in June~2004 and later on to introduce
@@
 with sequence of $\mathit{LM}$~statistics,
-critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}
+critical value at 5\% level (horizontal red line), and estimated breakpoint (vertical dashed line).}
 
@@
 @
-\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
+\caption{\label{fig:slovenia2} Histogram of observed monthly inflation rates in
 Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
@@
 
-Assuming a single stable set of parmaters (\ref{eq:h0}) is  not valid, as the
+Assuming a single stable set of parameters as in (\ref{eq:h0}) is  not valid, since the
 sequence of $\mathit{LM}$~statistics (see right panel of Figure~\ref{fig:slovenia1}) clearly
 exceeds its 5\% critical value (red horizontal line), leading to a highly significant
-$p$-value \Sexpr{spval[1]}. If a quadratic spectral kernal HAC estimator \citep{Andrews:1991}
+$p$-value \Sexpr{spval[1]}. If a quadratic spectral kernel HAC estimator \citep{Andrews:1991}
 would have been used to adjust (\ref{eq:efp}), the resulting $p$-value would be somewhat
@@
 breakpoints with a minimal segment size of two years per segment. It assumes its minimum for $B = 1$
-associated with the breakpoint \Sexpr{breakdates(x$breakpoints, breaks = x$nbreaks)}, depcited
+associated with the breakpoint \Sexpr{breakdates(x$breakpoints, breaks = x$nbreaks)}, depicted
 by a vertical dashed line in both panels of Figure~\ref{fig:slovenia1}. The resulting segmented
 fitted mean (horizontal blue lines in the left panel Figure~\ref{fig:slovenia1}) shows that
-in before the break Slovenia experienced very high inflation rates but was successful in reducing
+before the break, Slovenia experienced very high inflation rates but was successful in reducing
 inflation to a much lower level afterwards.
@@
 in Slovenia that severely decreased the annual growth in the M3 monetary aggregate.
-These success of these actions ensured the Slovenian participation in
-the ERM~II only a year later and the introduction of the euro two years later.
+The success of these actions ensured the Slovenian participation in
+the ERM~II only a year later and the introduction of the euro two years afterwards.
 
@@
 As a first step towards the investigation of inflation convergence, the
-structural breaks in the inflation dynamics within each countries are
+structural breaks in the inflation dynamics within each country are
 investigated using the methods introduced in Section~\ref{sec:model}.
@@
 
+%% source for DM zone: http://ec.europa.eu/economy_finance/publications/publication6730_de.pdf
+
 Austria, Finland, and France adjusted to the requirements of the EMU during the
@@
 experienced a strong increase in its formerly very low inflation rate, both
-in terms of mean and variances, also associated with a shift in skewness towards the
+in terms of mean and variance, also associated with a shift in skewness towards the
 left. Apart from Luxembourg, the countries that had a structural break during the second 
@@
 After the changes in this group, all countries have rather similar mean inflation
-rates, closer to the levels of the previous two groups.
+rates, which are closer to the levels of the previous two groups.
 
@@
 resulting from efforts to curb in inflation by decreasing the growth of the money 
-supply. The timing of the break thus indicates the timing of the biggest
-transitionally shocks on the way towards free market economies. In Poland, this
-transition took two more years. Slovakia and Slovenia are somewhat different: On one hand,
+supply. The timing of the break thus indicates the date when the the biggest
+transitionally shocks on the way towards free market economies have been overcome. In Poland, this
+transition took two more years. Slovakia and Slovenia are somewhat different: On the one hand,
 they started with lower mean inflation rates and on the other hand, their inflation dynamics
@@
 The groups have been selected based on prior expectations (i.e., not based on
-the breakpoints estimation results) as the ``core'' of the euro zone
-(\Sexpr{paste(colnames(hicps)[euro], collapse = ", ")}), the PIIGS countries,
+the breakpoints estimation results) as the ``core'' of the euro zone, the PIIGS countries,
 and the rest. Similarly, the time points have been chosen as to fall ``between''
@@
 first phase did not affect mean inflation much, while variance increased
-slightly (in both quartiles) and skewness somewhat decreased (from slightly
+slightly (in both quartiles) and skewness decreased somewhat (from slightly
 right- to slightly left-skewed). The heterogeneity (assessed by the
-interquartile range) decreased for all three moments of the inflation rates.
+interquartile range) declined for all three moments of the inflation rates.
 This reflects that the monetary policies of the corresponding countries were
@@
 
-For the PIIGS countries, changes occured from \Sexpr{yrs[1]} to
+For the PIIGS countries, changes occurred from \Sexpr{yrs[1]} to
 \Sexpr{yrs[2]} and then again during the crisis, before \Sexpr{yrs[4]}.
@@
 doubling of variance and a change from clearly right-skewed to approximately
-symmetric. The latter reflects that the occurence of shocks with high inflation
+symmetric. The latter reflects that the occurrence of shocks with high inflation
 rates have been successfully reduced. Moreover, the changes in the PIIGS
@@
 decrease in mean inflation and associated skewness which are both caused by
-several extremely low inflation (or even deflation) rates during the time
+several extremely low inflation rates during the time
 following the crisis.
@@
 simultaneously decreasing the heterogeneity (as captured by the corresponding
-interquartile ranges) across countries. Before the transition all three
+interquartile ranges) across countries. Before the transition, all three
 moments were extremely large compared to the other two groups, while afterwards
 all moments are much closer to those of the other groups. In Jan~\Sexpr{yrs[3]},
-mean inflation is in between the levels for the core euro and PIIGS countres,
-variance is still slightly higher, and skewness is not too far from symmetry.
+mean inflation is in between the levels for the core euro and PIIGS countries,
+variance is still slightly higher, and skewness is not too far away from symmetry.
 
@@
 for the others. Right-skewness (and thus large outlying inflation rates)
-has been reduced for all three groups of countries but to different extents.
+has been reduced for all three groups of countries but to a different extent.
 Thus, while ending in similar inflation dynamics, the timing and patterns of
@@
 in most of the European countries (except Ireland and Spain).
-This may be somewhat surprising and is not enitrely correct. In fact,
-there is a rather clear effect but as it is very short-lived and not
-persistent it is not picked up by the structural break techniques.
+This may be somewhat surprising and is not entirely correct. In fact,
+there is a rather clear effect, but as it is very short-lived and not
+persistent, it is not picked up by the structural break techniques.
 
@@
 $\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
-Another way to bring out this pattern in a graphical way is
+One way to bring out this pattern in graphically is
 Figure~\ref{fig:glmean}. It depicts the inflation rates (in gray) and a 12-month rolling mean (in red),
@@
 Towards the end of the 2000s, before the financial crisis, our results broadly confirm
-convergence hypotheses as the first three moments from all countries in our sample
+convergence hypotheses, since the first three moments from all countries in our sample
 move closer together.
@@
 In France the sale periods time span is regulated by each department. In Greece there
-are also two set sales-periods in winter and summer. In Italy the sales periods
+are also two set sales-periods in winter and summer. In Italy, the sales periods
 are fixed each year by the Chambers of Commerce.  Portugal has two set sales

### 2011-06-12 pkg/paper/Paper/JAE.Rnw
Renamed "draft" to "report", deleted .pdf.
New JAE.Rnw with mimicked JAE style and increased baselinestretch.

Touch-ups in .bib.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@120 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2011-06-12 pkg/paper/Paper/draft.Rnw
@@
-\documentclass[nojss]{jss}
-\usepackage{amsfonts,amstext,amsmath}
+\documentclass[10pt,a4paper,twoside]{article}
+
+%% additional packages
+\usepackage{amsfonts,amstext,amsmath,a4wide,graphicx,color,thumbpdf}
+\usepackage{hyperref}
 \newcommand{\argmax}{\operatorname{argmax}\displaylimits}
@@
 
-%% for internal use
-\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
-\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
-
-\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}
-
-\author{Thomas Windberger\\Universit\"at Innsbruck \And
-        Achim Zeileis\\Universit\"at Innsbruck}
-%% \author{\hfill Thomas Windberger \hfill Achim Zeileis\thanks{Correspondence
-%% to: Achim Zeileis, Department of Statistics,  Universit\"at Innsbruck,
-%% Universit\"atsstr.~15, A-6020 Innsbruck, Austria. Tel: +43/512/507-7103. Fax: +43/512/507-2851.
-%% E-mail: \email{Achim.Zeileis@R-project.org}} \hfill \hfill \\
-%% {\small \it Department of Statistics, Universit\"at Innsbruck, Austria}}
-\Plainauthor{Thomas Windberger, Achim Zeileis}
-
-\renewcommand{\abstractname}{Summary}
-\Abstract{
+%% BibTeX settings
+\usepackage[authoryear,round]{natbib}
+\bibliographystyle{jae}
+\bibpunct{(}{)}{,}{a}{,}{,}
+\newcommand{\doi}[1]{\href{http://dx.doi.org/#1}{\normalfont\texttt{doi:#1}}}
+
+%% markup commands for code/software
+\let\code=\texttt
+\let\pkg=\textbf
+\let\proglang=\textsf
+\newcommand{\file}[1]{`\code{#1}'}
+\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
+
+%% paragraph formatting
+\renewcommand{\baselinestretch}{1.52}
+
+%% mimic JAE style
+\renewcommand{\section}{\secdef \mysec \mysecnn}
+\newcommand{\mysec}[2][default]{\vspace{1.7\baselineskip}%
+  \pdfbookmark[1]{#1}{Section.#1}%
+  \refstepcounter{section}%
+  \centerline{\large \thesection. \uppercase{#1}} \vspace{.5\baselineskip}}
+\newcommand{\mysecnn}[1]{\vspace{1.7\baselineskip}%
+  \centerline{\large #1} \vspace{.5\baselineskip}}
+\renewcommand{\abstractname}{\normalfont SUMMARY}
+\renewcommand{\refname}{REFERENCES}
+\renewcommand{\thetable}{\Roman{table}}
+
+%% page header (and currently no footer)
+\usepackage{fancyhdr}
+\setlength{\headheight}{15pt}
+\renewcommand{\headrulewidth}{0pt}
+\pagestyle{fancy}
+\thispagestyle{plain}
+\fancyhf{}
+\fancyhead[LE,RO]{\thepage}
+\fancyhead[CE]{{\normalsize \uppercase{T.~Windberger and A.~Zeileis}}}
+\fancyhead[CO]{{\normalsize \uppercase{Structural Breaks in Inflation Dynamics within the EMU}}} 
+% \fancyfoot[LO,LE]{\small Copyright {\copyright} 20XX John Wiley \& Sons, Ltd.}
+\fancypagestyle{plain}{
+\fancyhf{}
+% \fancyfoot[LO,LE]{\small This is a preprint of an article published in %
+%   \textit{Journal of Applied Econometrics}, \textbf{XX}(X), XXX--XXX. \\%
+%   Copyright {\copyright} 20XX John Wiley \& Sons, Ltd. \doi{10.1002/jae.XXXX} }
+}
+
+%% title information
+\title{\bf Structural Breaks in Inflation Dynamics within the European Monetary Union}
+\author{\hfill Thomas Windberger \hfill Achim Zeileis\thanks{Correspondence
+to: Achim Zeileis, Department of Statistics,  Universit\"at Innsbruck,
+Universit\"atsstr.~15, A-6020 Innsbruck, Austria. Tel: +43/512/507-7103. Fax: +43/512/507-2851.
+E-mail: \email{Achim.Zeileis@R-project.org}} \hfill \hfill \\
+{\small \it Department of Statistics, Universit\"at Innsbruck, Austria}}
+\date{}
+
+% hyperref setup
+\definecolor{Red}{rgb}{0.5,0,0}
+\definecolor{Blue}{rgb}{0,0,0.5}
+\hypersetup{%
+  pdftitle = {Structural Breaks in Inflation Dynamics within the European Monetary Union},
+  pdfsubject = {},
+  pdfkeywords = {inflation rate, structural break, EMU, generalized logistic distribution},
+  pdfauthor = {Thomas Windberger, Achim Zeileis},
+  %% change colorlinks to false for pretty printing
+  colorlinks = {true},
+  linkcolor = {Blue},
+  citecolor = {Blue},
+  urlcolor = {Red},
+  hyperindex = {true},
+  linktocpage = {true},
+}
+
+\begin{document}
+
+\maketitle
+
+\begin{abstract}
   To assess the effects of the EMU on inflation rate dynamics of
@@
   with reduced skewness, but it is accompanied by an increase in variance.
-}
-
-\Keywords{inflation rate, structural break, EMU, generalized logistic distribution}
-
-\Address{
-  Thomas Windberger, Achim Zeileis\\
-  Department of Statistics\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstra{\ss}e 15\\
-  A-6020 Innsbruck, Austria\\
-  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
-  Telephone: +43/12/507-7103 \\
-  Fax: +43/512/507-2851 \\
-}
-
-\begin{document}
+\end{abstract}
+
+\noindent {\bf Keywords:} inflation rate, structural break, EMU, generalized logistic distribution.
 
@@
 
+\renewcommand{\baselinestretch}{1.15}
 \begin{table}[p!]
@@
 \end{table}
+\renewcommand{\baselinestretch}{1.52}
 
@@
 
+\renewcommand{\baselinestretch}{1.15}
 \begin{table}[t!]
@@
 
-\section*{Computational details}
+\section*{COMPUTATIONAL DETAILS}
 
@@
 is provided in \code{demo(package = "glogis")} and the scripts can be launched easily, e.g.,
-via \code{demo("Austria", package = "glogis")} etc.
-
-\section*{Acknowledgments}
-
-We are grateful to Jesus Crespo Cuaresma for comments and discussions about
-an earlier version of the manuscript, and also to the participants of the workshop
-of the Research Platform ``Empirical and Experimental Economics'' for stimulating discussions,
-leading to substantial improvements of the paper.
-
-\bibliography{papers}
+via\linebreak \code{demo("Austria", package = "glogis")} etc.
+
+%% \section*{Acknowledgments}
+%% 
+%% We are grateful to Jesus Crespo Cuaresma for comments and discussions about
+%% an earlier version of the manuscript, and also to the participants of the workshop
+%% of the Research Platform ``Empirical and Experimental Economics'' for stimulating discussions,
+%% leading to substantial improvements of the paper.
+
+\bibliography{glogis}
 
 \end{document}
-\SweaveOpts{eval=FALSE}
-
-\clearpage
-
-\begin{appendix}
-
-\section{Data and sales periods}\label{sec:sales}
-
-In addition to Germany, Portugal, Belgium, Italy, Luxembourg, and Spain, there
-are a number of other countries where the sales periods are regulated by law.
-In France the sale periods time span is regulated by each department. In Greece there
-are also two set sales-periods in winter and summer. In Italy, the sales periods
-are fixed each year by the Chambers of Commerce.  Portugal has two set sales
-periods as well, but for two month (January and February as well as Mid-July to
-Mid-September). The source for this information is \citet{konseur}. According to
-\cite{pltm}, the official dates of the inclusion of these sales periods are:
-January 1998 for Germany and Portugal,  January 1999 for Luxembourg and Ireland and 
-January 2000 for Belgium,  while Italy and Spain introduced them in January 2001.
-
-\clearpage
-\section{Graphs}
-
-Figures~\ref{fig:nochange}--\ref{fig:eastern} depict the data underlying the results
-in Section~\ref{sec:results} along with the GL-based means from Table~\ref{tab:results1}.
-
-<<graphs-code>>=
-get_data <- function(i) {
-  rval <- do.call("merge",
-    lapply(i, function(j) results[[j]]$data))
-  colnames(rval) <- names(results)[i]
-  rval
-}
-
-get_fitted <- function(i) {
-  rval <- do.call("merge",
-    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks = results[[j]]$nbreaks, type = "mean")))
-  colnames(rval) <- names(results)[i]
-  rval
-}
-
-library("lattice")
-get_xyplot <- function(i) {
-  nam <- names(results)[i]
-  nam[nam == "CzechRepublic"] <- "Czech Republic"
-  xyplot(merge(get_data(i), get_fitted(i)),
-    screen = rep(nam, 2),
-    col = rep(c(1, 4), each = length(i))) 
-}
-@
-
-\setkeys{Gin}{width=0.8\textwidth}
-\begin{figure}[p!]
-\begin{center}
-<<graphs-nochange, fig=TRUE, height=6.3, width=7.5>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(get_xyplot(c(2, 4, 8, 14)))
-@
-\caption{\label{fig:nochange} No change.}
-
-<<graphs-phase1, fig=TRUE, height=4.9, width=7.5>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(get_xyplot(c(1, 6, 7)))
-@
-\caption{\label{fig:phase1} Phase~1 convergence.}
-\end{center}
-\end{figure}
-
-\begin{figure}[p!]
-\begin{center}
-<<graphs-phase2, fig=TRUE, height=6.3, width=7.5>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(get_xyplot(c(9, 12, 13, 16)))
-@
-\caption{\label{fig:phase2} Phase~2 convergence.}
-
-<<graphs-crisis, fig=TRUE, height=3.5, width=7.5>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(get_xyplot(c(11, 19)))
-@
-\caption{\label{fig:crisis} Financial crisis.}
-\end{center}
-\end{figure}
-
-\begin{figure}[p!]
-\begin{center}
-<<graphs-noneuro, fig=TRUE, height=3.5, width=7.5>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(get_xyplot(c(20, 21)))
-@
-\caption{\label{fig:noneuro} Non-euro.}
-
-<<graphs-eastern, fig=TRUE, height=9.1, width=7.5>>=
-trellis.par.set(theme = standard.theme(color = FALSE))
-print(get_xyplot(c(3, 5, 10, 15, 17, 18)))
-@
-\caption{\label{fig:eastern} Eastern countries.}
-\end{center}
-\end{figure}
-
-\end{appendix}
-
-\end{document}

### 2011-06-13 pkg/paper/Paper/JAE.Rnw
updated JAE
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@123 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 \caption{\label{fig:slovenia2} Histogram of observed monthly inflation rates in
-Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
+Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[2]} (right),
 along with fitted generalized logistic probability density function, associated
@@
 by a vertical dashed line in both panels of Figure~\ref{fig:slovenia1}. The resulting segmented
-fitted mean (horizontal blue lines in the left panel Figure~\ref{fig:slovenia1}) shows that
+fitted mean (horizontal blue lines in the left panel of Figure~\ref{fig:slovenia1}) shows that
 before the break, Slovenia experienced very high inflation rates but was successful in reducing
@@
 countries, the mean and the variance of their inflation rates declined in the
-late 1990s or early 2000s. Czech Republic, Estonia and Hungary experienced a break in 1997/1998
+late 1990s or early 2000s. Czech Republic, Estonia, and Hungary experienced a break in 1997/1998
 resulting from efforts to curb in inflation by decreasing the growth of the money 
@@
 mid-1990s, but many changes occur in the transition period from \Sexpr{yrs[2]}
-to \Sexpr{yrs[3]}, encompassing both the completion of transformation from
+to \Sexpr{yrs[3]}, encompassing both the completion of the transformation from
 planned to free market economies and the intended entry into the EMU.
@@
 $\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
-One way to bring out this pattern in graphically is
+One way to bring out this pattern graphically is
 Figure~\ref{fig:glmean}. It depicts the inflation rates (in gray) and a 12-month rolling mean (in red),
@@
 is implemented in the new package \pkg{glogis} that also contains the data used, both
-the raw monthly price indexes \citep[as provided by][]{oecd} and the seasonally adjusted
+the raw monthly price indexes \citep[as provided by the][]{oecd} and the seasonally adjusted
 monthly inflation rates. The seasonal adjustment was performed in \pkg{X-12-ARIMA}~0.3

### 2012-03-01 pkg/paper/Paper/EmpEc.Rnw
renamed JAE.Rnw to EmpEc.Rnw for new version
employ JSS style (instead of mimicking JAE style)

add diagnostics table (HAC, GOF, autocorrelation)



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@125 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2011-06-13 pkg/paper/Paper/JAE.Rnw
@@
-\documentclass[10pt,a4paper,twoside]{article}
-
-%% additional packages
-\usepackage{amsfonts,amstext,amsmath,a4wide,graphicx,color,thumbpdf}
-\usepackage{hyperref}
+\documentclass[nojss]{jss}
+\usepackage{amsfonts,amstext,amsmath,thumbpdf}
 \newcommand{\argmax}{\operatorname{argmax}\displaylimits}
@@
 
-%% BibTeX settings
-\usepackage[authoryear,round]{natbib}
-\bibliographystyle{jae}
-\bibpunct{(}{)}{,}{a}{,}{,}
-\newcommand{\doi}[1]{\href{http://dx.doi.org/#1}{\normalfont\texttt{doi:#1}}}
-
-%% markup commands for code/software
-\let\code=\texttt
-\let\pkg=\textbf
-\let\proglang=\textsf
-\newcommand{\file}[1]{`\code{#1}'}
-\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
-
-%% paragraph formatting
-\renewcommand{\baselinestretch}{1.52}
-
-%% mimic JAE style
-\renewcommand{\section}{\secdef \mysec \mysecnn}
-\newcommand{\mysec}[2][default]{\vspace{1.7\baselineskip}%
-  \pdfbookmark[1]{#1}{Section.#1}%
-  \refstepcounter{section}%
-  \centerline{\large \thesection. \uppercase{#1}} \vspace{.5\baselineskip}}
-\newcommand{\mysecnn}[1]{\vspace{1.7\baselineskip}%
-  \centerline{\large #1} \vspace{.5\baselineskip}}
-\renewcommand{\abstractname}{\normalfont SUMMARY}
-\renewcommand{\refname}{REFERENCES}
-\renewcommand{\thetable}{\Roman{table}}
-
-%% page header (and currently no footer)
-\usepackage{fancyhdr}
-\setlength{\headheight}{15pt}
-\renewcommand{\headrulewidth}{0pt}
-\pagestyle{fancy}
-\thispagestyle{plain}
-\fancyhf{}
-\fancyhead[LE,RO]{\thepage}
-\fancyhead[CE]{{\normalsize \uppercase{T.~Windberger and A.~Zeileis}}}
-\fancyhead[CO]{{\normalsize \uppercase{Structural Breaks in Inflation Dynamics within the EMU}}} 
-% \fancyfoot[LO,LE]{\small Copyright {\copyright} 20XX John Wiley \& Sons, Ltd.}
-\fancypagestyle{plain}{
-\fancyhf{}
-% \fancyfoot[LO,LE]{\small This is a preprint of an article published in %
-%   \textit{Journal of Applied Econometrics}, \textbf{XX}(X), XXX--XXX. \\%
-%   Copyright {\copyright} 20XX John Wiley \& Sons, Ltd. \doi{10.1002/jae.XXXX} }
-}
-
-%% title information
-\title{\bf Structural Breaks in Inflation Dynamics within the European Monetary Union}
-\author{\hfill Thomas Windberger \hfill Achim Zeileis\thanks{Correspondence
-to: Achim Zeileis, Department of Statistics,  Universit\"at Innsbruck,
-Universit\"atsstr.~15, A-6020 Innsbruck, Austria. Tel: +43/512/507-7103. Fax: +43/512/507-2851.
-E-mail: \email{Achim.Zeileis@R-project.org}} \hfill \hfill \\
-{\small \it Department of Statistics, Universit\"at Innsbruck, Austria}}
-\date{}
-
-% hyperref setup
-\definecolor{Red}{rgb}{0.5,0,0}
-\definecolor{Blue}{rgb}{0,0,0.5}
-\hypersetup{%
-  pdftitle = {Structural Breaks in Inflation Dynamics within the European Monetary Union},
-  pdfsubject = {},
-  pdfkeywords = {inflation rate, structural break, EMU, generalized logistic distribution},
-  pdfauthor = {Thomas Windberger, Achim Zeileis},
-  %% change colorlinks to false for pretty printing
-  colorlinks = {true},
-  linkcolor = {Blue},
-  citecolor = {Blue},
-  urlcolor = {Red},
-  hyperindex = {true},
-  linktocpage = {true},
-}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
+\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}
+
+\author{Thomas Windberger\\Universit\"at Innsbruck \And
+        Achim Zeileis\\Universit\"at Innsbruck}
+\Plainauthor{Thomas Windberger, Achim Zeileis}
+
+\Abstract{
   To assess the effects of the EMU on inflation rate dynamics of
@@
   with reduced skewness, but it is accompanied by an increase in variance.
-\end{abstract}
-
-\noindent {\bf Keywords:} inflation rate, structural break, EMU, generalized logistic distribution.
+}
+
+\Keywords{inflation rate, structural break, EMU, generalized logistic distribution}
+
+\Address{
+  Thomas Windberger, Achim Zeileis\\
+  Department of Statistics\\
+  Faculty of Economics and Statistics\\
+  Universit\"at Innsbruck\\
+  Universit\"atsstra{\ss}e 15\\
+  6020 Innsbruck, Austria\\
+  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
+  Telephone: +43/12/507-7103 \\
+  Fax: +43/512/507-2851 \\
+}
+
+\begin{document}
 
@@
     efp = x_efp,
+    efp_hac = gefp(x_gf, fit = NULL, vcov = vcovHAC), #FIXME# added in revision for EmpEc, should be moved to 'Analysis'
     nbreaks = x_nbreaks,
@@
 
-\renewcommand{\baselinestretch}{1.15}
 \begin{table}[p!]
-\begin{center}
+\centering
 \footnotesize
@@
 distribution per segment, and ERM information for all countries under investigation.}
-\end{center}
 \end{table}
-\renewcommand{\baselinestretch}{1.52}
+
+<<results-diagnostics>>=
+names(results)[names(results) == "CzechRepublic"] <- "Czech Rep."
+format_diag <- function(i, digits = 3) {
+  pround <- function(x) ifelse(abs(x) < 0.001, "$< 0.001$",
+    paste("$", format(round(x, digits = digits), nsmall = digits), "$", sep = ""))
+
+  rval <- cbind("", "", names(results[[i]]$model), "", "", "")
+  n <- NROW(rval)
+  if(n <= 1) rval <- rbind(rval, "")
+
+  rval[1,   1] <- if(is.character(i)) i else names(results)[i]
+  rval[1,   2] <- pround(sctest(results[[i]]$efp,     functional = supLM(0.1))$p.value)
+  rval[2,   2] <- pround(sctest(results[[i]]$efp_hac, functional = supLM(0.1))$p.value)
+  rval[1:n, 4] <- pround(sapply(results[[i]]$model, function(obj) summary(obj)$chisq.test$p.value))
+  rval[1:n, 5] <- pround(sapply(results[[i]]$model, function(obj) acf(obj$x, plot = FALSE)$acf[2]))
+  rval[1:n, 6] <- pround(sapply(results[[i]]$model, function(obj) Box.test(obj$x, type = "Ljung-Box")$p.value))
+
+  rval <- apply(rval, 1, paste, collapse = " & ")
+  rval <- paste(rval, "\\\\")
+  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
+  return(rval)
+}
+@
+
+\begin{table}[p!]
+\centering
+\footnotesize
+\begin{tabular}{lrlrrr}
+\hline
+Country & sup$\mathit{LM}$ (standard/HAC) & Segment & $\chi^2$ GOF & AR(1) & Ljung-Box \\
+\hline
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{No change}} \\
+\hline
+<<diagnostics-table1, results=tex>>=
+writeLines(unlist(lapply(c(2, 4, 8, 14), format_diag)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Phase~1 convergence}} \\
+\hline
+<<diagnostics-table2, results=tex>>=
+writeLines(unlist(lapply(c(1, 6, 7), format_diag)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Phase~2 convergence}} \\
+\hline
+<<diagnostics-table3, results=tex>>=
+writeLines(unlist(lapply(c(9, 12, 13, 16), format_diag)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Financial crisis}} \\
+\hline
+<<diagnostics-table4, results=tex>>=
+writeLines(unlist(lapply(c(11, 19), format_diag)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Non-euro}} \\
+\hline
+<<diagnostics-table5, results=tex>>=
+writeLines(unlist(lapply(c(20, 21), format_diag)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Eastern countries}} \\
+\hline
+<<diagnostics-table6, results=tex>>=
+writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_diag)))
+@
+
+\noalign{\medskip}
+\multicolumn{6}{l}{\emph{Euro area}} \\
+\hline
+<<diagnostics-table6, results=tex>>=
+writeLines(unlist(lapply(22, format_diag)))
+@
+
+\end{tabular}
+
+\caption{\label{tab:diag} Diagnostics for model fits in each country. Two sup$\mathit{LM}$ test $p$~values
+  of the full-sample model with and without HAC correction and segment-specific assessments of the model
+  fit: $\chi^2$ goodness-of-fit (GOF) $p$~value; autocorrelation at lag~1; Ljung-Box $p$~value.}
+\end{table}
 
@@
 
-\renewcommand{\baselinestretch}{1.15}
 \begin{table}[t!]
-\begin{center}
+\centering
 \begin{tabular}{llrrrr}
@@
 changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
-\end{center}
 \end{table}
-\renewcommand{\baselinestretch}{1.52}
 
@@
 \begin{figure}[t!]
-\begin{center}
+\centering
 <<glmean-plot, fig=TRUE, height=5, width=7.5>>=
@@
 12-month rolling GL fits (red) over time for all countries.}
-\end{center}
 \end{figure}
@@
 
-\section*{COMPUTATIONAL DETAILS}
+\section*{Computational details}
 

### 2012-03-18 pkg/paper/Paper/EmpEc.Rnw
updated version
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@126 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-All empirical analyses are based on seasonally adjusted inflation rates (in percent)
+All empirical analyses are based on seasonally adjusted monthly inflation rates (in percent)
 for \Sexpr{ncol(hicps)}~countries
@@
 
-Finally, Section~\ref{sec:fcrisis} focuses on the question of changes
-in the inflation dynamics during the 2008/2009 financial crisis.
+Section~\ref{sec:fcrisis} focuses on the question of changes
+in the inflation dynamics during the 2008/2009 financial crisis. Finally, the model fit and related issues will be discussed in Section~\ref{sec:fit}
 
@@
 breakpoints was selected via the LWZ criterion and subsequently all parameters
-(breakpoints and segment-specific GL parameters) were estimated by ML. No
-HAC~correction was employed as there was only negligible autocorrelation in the
-empirical scores after inclusion of the breakpoints (if any).
-In Table~\ref{tab:results1} the countries are grouped according to either a
+(breakpoints and segment-specific GL parameters) were estimated by ML. 
+The $p$~values from the sup$\mathit{LM}$ test with and without HAC correction are reported in Table~\ref{tab:diag} 
+along with further diagnostics for each of the fitted segments concerning goodness of fit and autocorrelation. 
+As autocorrelation was very low in most of the segments (typically below 0.4 and non-significant, with a few exceptions), 
+we focus on the results from the analysis without HAC correction but point out if there are noteworthy deviations. 
+In both Table~\ref{tab:results1} and~\ref{tab:diag} the countries are grouped  according to either a
 similar timing of the changes in inflation dynamics or according to economic similarities, 
@@
 the EMU did have no effect on inflation dynamics since the policies in
-these countries were already much in line with Germany and hence the ECB.
+these countries were already much in line with Germany and hence the ECB. Table~\ref{tab:diag} does not 
+suggest any problems in this segment.
 
@@
 Austria, Finland, and France adjusted to the requirements of the EMU during the
-first phase of convergence. In these countries,
+first phase of convergence. However, with HAC correction, the very short segment of France would not be
+segregated, thereby suggesting that France belongs to the group of countries which did not change at all.
+In these countries,
 the result of the adjustment process was a lower inflation rate and an overall
@@
 After the changes in this group, all countries have rather similar mean inflation
-rates, which are closer to the levels of the previous two groups.
+rates, which are closer to the levels of the previous two groups. The diagnostics for model fits in 
+this group are unproblematic.
 
@@
 In both countries, this had a strong deflationary impact, with deflation in Ireland
-over an extended period of time. In Spain, which was affected by a housing bubble of  
+over an extended period of time. For the middle segment of Spain, the Ljung-box shows that not all
+of the autocorrelation has been captured. In Spain, which was affected by a housing bubble of  
 roughly the same size as Ireland, the housing prices did not adjust so abruptly and  
@@
 into the ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
-1990s.
+1990s. With regard to the last segment for both countries, the Ljung-box suggests that the data 
+in this segment is independently distributed. Anyhow, the HAC correction would also suggest two breaks 
+for both series.
 
@@
 countries, the mean and the variance of their inflation rates declined in the
-late 1990s or early 2000s. Czech Republic, Estonia, and Hungary experienced a break in 1997/1998
+late 1990s or early 2000s. For Estonia and Slovakia, the $\chi^2$ goodness-of-fit suggests that the generalized logistic distribution
+might not be the most suitable distribution for this purpose. Also, compared to other country groups, the 
+Eastern European group shows the highest occurrence of non-captured autocorrelation. Nevertheless, compared to other 
+inflation rates, these series show the most dynamic behaviour.
+
+Czech Republic, Estonia, and Hungary experienced a break in 1997/1998
 resulting from efforts to curb in inflation by decreasing the growth of the money 

### 2012-03-19 pkg/paper/Paper/EmpEc.Rnw
small touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@127 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 
-Section~\ref{sec:fcrisis} focuses on the question of changes
-in the inflation dynamics during the 2008/2009 financial crisis. Finally, the model fit and related issues will be discussed in Section~\ref{sec:fit}
+Finally, Section~\ref{sec:fcrisis} focuses on the question of changes
+in the inflation dynamics during the 2008/2009 financial crisis.
 
@@
 these countries were already much in line with Germany and hence the ECB. Table~\ref{tab:diag} does not 
-suggest any problems in this segment.
+suggest any problems in this group of countries.
 
@@
 Austria, Finland, and France adjusted to the requirements of the EMU during the
-first phase of convergence. However, with HAC correction, the very short segment of France would not be
-segregated, thereby suggesting that France belongs to the group of countries which did not change at all.
-In these countries,
-the result of the adjustment process was a lower inflation rate and an overall
-decrease in the variance.
+first phase of convergence. In these countries, the result of the adjustment process
+was a lower inflation rate and an overall decrease in the variance.
+Additionally, if a HAC correction is used, the change for France would be nonsignificant,
+suggesting that France may also belong to the group of countries which did not change at all.
 
@@
 After the changes in this group, all countries have rather similar mean inflation
-rates, which are closer to the levels of the previous two groups. The diagnostics for model fits in 
-this group are unproblematic.
+rates, which are closer to the levels of the previous two groups. The diagnostics 
+from Table~\ref{tab:diag} indicate good model fits with low autocorrelations.
 
@@
 In both countries, this had a strong deflationary impact, with deflation in Ireland
-over an extended period of time. For the middle segment of Spain, the Ljung-box shows that not all
-of the autocorrelation has been captured. In Spain, which was affected by a housing bubble of  
+over an extended period of time. In Spain, which was affected by a housing bubble of  
 roughly the same size as Ireland, the housing prices did not adjust so abruptly and  
-so strongly.
+so strongly.  The model fit diagnostics appear to be reasonable except for the middle
+segment of Spain, where the Ljung-box shows that not all of the autocorrelation has been captured.
 
@@
 into the ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
-1990s. With regard to the last segment for both countries, the Ljung-box suggests that the data 
-in this segment is independently distributed. Anyhow, the HAC correction would also suggest two breaks 
-for both series.
+1990s. With regard to the last segment for both countries, the Ljung-box suggests that 
+the autocorrelation, albeit low around 0.16, is significant. However, both the standard
+and the HAC-corrected sup$\mathit{LM}$ lead to highly significant evidence for structural
+breaks.
 
@@
 countries, the mean and the variance of their inflation rates declined in the
-late 1990s or early 2000s. For Estonia and Slovakia, the $\chi^2$ goodness-of-fit suggests that the generalized logistic distribution
-might not be the most suitable distribution for this purpose. Also, compared to other country groups, the 
-Eastern European group shows the highest occurrence of non-captured autocorrelation. Nevertheless, compared to other 
-inflation rates, these series show the most dynamic behaviour.
-
+late 1990s or early 2000s. 
 Czech Republic, Estonia, and Hungary experienced a break in 1997/1998
@@
 they started with lower mean inflation rates and on the other hand, their inflation dynamics
-both changed roughly one year prior to their entry into the ERM~II.
+both changed roughly one year prior to their entry into the ERM~II. Thus,
+compared to other inflation rates, these series show the most dynamic behaviour. However,
+the models also do not fit as well as for the previous groups of countries.
+There is the highest occurrence of non-captured autocorrelation and for Estonia
+and Slovakia the $\chi^2$ goodness-of-fit test suggests that the generalized logistic distribution
+might not fit sufficiently well in the first segments.
 

### 2013-07-24 pkg/paper/Paper/EmpEc.Rnw
- changed graphics to employ gray levels etc. instead of colors - Ljung-box -> Ljung-Box test
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@128 b679b2f7-6539-45d0-a605-5ddcb976cde9

@@
 lines(x$breakpoints, breaks = x$nbreaks)
-lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
+lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), lwd = 2)
 if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
-plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
+
+plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics", ylim = c(0, 40), boundary = FALSE)
+abline(h = supLM(0.1)$computeCritval(0.05, 3), col = "darkgray", lwd = 1.5)
 lines(x$breakpoints, breaks = x$nbreakss)
@@
 with sequence of $\mathit{LM}$~statistics,
-critical value at 5\% level (horizontal red line), and estimated breakpoint (vertical dashed line).}
+critical value at 5\% level (horizontal gray line), and estimated breakpoint (vertical dashed line).}
 
@@
 for(i in 1:2) plot(x$model[[i]], moments = TRUE,
+  col = "black", lwd = 1.5,
   xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
@@
 so strongly.  The model fit diagnostics appear to be reasonable except for the middle
-segment of Spain, where the Ljung-box shows that not all of the autocorrelation has been captured.
+segment of Spain, where the Ljung-Box test shows that not all of the autocorrelation has been captured.
 
@@
 into the ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
-1990s. With regard to the last segment for both countries, the Ljung-box suggests that 
+1990s. With regard to the last segment for both countries, the Ljung-Box test suggests that 
 the autocorrelation, albeit low around 0.16, is significant. However, both the standard
@@
 <<glmean-plot, fig=TRUE, height=5, width=7.5>>=
-  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
-  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
+gray_red <- rgb(c(0.2, 0.8), c(0.2, 0.0), c(0.2, 0.0), alpha = 0.3)
+gray_red <- rgb(c(0.2, 0.0), c(0.2, 0.0), c(0.2, 0.0), alpha = c(0.3, 0.5)) ## actually gray-darkgray now
 plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
@@
 @
-\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
-12-month rolling GL fits (red) over time for all countries.}
+\caption{\label{fig:glmean} Observed monthly inflation rates (light gray) and estimated means from
+12-month rolling GL fits (dark red) over time for all countries.}
 \end{figure}

### 2014-09-07 pkg/paper/Paper/paper.Rnw
preparation for new release accompanying EEE paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@131 b679b2f7-6539-45d0-a605-5ddcb976cde9

--- 2013-07-24 pkg/paper/Paper/EmpEc.Rnw
@@
-\documentclass[nojss]{jss}
+\documentclass[nojss,noheadings]{jss}
 \usepackage{amsfonts,amstext,amsmath,thumbpdf}
@@
   6020 Innsbruck, Austria\\
-  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
-  Telephone: +43/12/507-7103 \\
-  Fax: +43/512/507-2851 \\
+  E-mail: \email{Achim.Zeileis@R-project.org} \\
+  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
+}
+
+%% publication info and copyright notice
+\usepackage{fancyhdr}
+\setlength{\headheight}{15pt}
+\renewcommand{\headrulewidth}{0pt}
+\pagestyle{fancy}
+\thispagestyle{plain}
+\fancyhf{}
+\fancyhead[RO,LE]{\thepage}
+\fancyhead[CO]{\textit{\normalsize Thomas Windberger, Achim Zeileis}}
+\fancyhead[CE]{\textit{\normalsize Structural Breaks in Inflation Dynamics within the European Monetary Union}}
+\fancyfoot[LO,LE]{\small Copyright {\copyright} 2014 M.E. Sharpe, Inc.}
+\fancypagestyle{plain}{
+\fancyhf{}
+\fancyfoot[LO,LE]{\small This is a preprint of an article published in \textit{Eastern European Economics}, \textbf{52}(3), 67--89.
+  \doi{10.2753/EEE0012-8775520304}.
+  Copyright~{\copyright} 2014 M.E. Sharpe, Inc.
+  }
 }
@@
 The null hypothesis of parameter stability (\ref{eq:h0}) can be assessed using the empirical
-scores $s(y_i | \hat \phi)$ as measures of model deviation \citep{Zeileis+Hornik:2007} from the
-model fit under $H_0$. Systematic deviations over time from the full sample estimates $\hat \phi$
+scores $s(y_i | \hat \phi)$ as measures of model deviation from the model fit under $H_0$
+\citep[as in][]{Zeileis+Hornik:2007}. Systematic deviations over time from the full sample estimates $\hat \phi$
 can then be captured by cumulative sums of the empirical scores in an empirical fluctuation
@@
 lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), lwd = 2)
-if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
+if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks), col = "darkgray")
 
@@
 @
-\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
-with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
-from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
-with sequence of $\mathit{LM}$~statistics,
-critical value at 5\% level (horizontal gray line), and estimated breakpoint (vertical dashed line).}
+\caption{\label{fig:slovenia1} 
+  Inflation rate series for Slovenia with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)})
+  and associated sequence of LM statistics
+  Left: Inflation rate series with breakpoint estimate (vertical dashed line),
+  associated confidence interval (gray, at the bottom), and fitted mean from generalized
+  logistic distribution (bold horizontal lines). Right: Corresponding sup$\mathit{LM}$ test
+  with sequence of LM statistics, critical value at 5\% level (horizontal gray
+  line), and estimated breakpoint (vertical dashed line).
+}
 
@@
 by a vertical dashed line in both panels of Figure~\ref{fig:slovenia1}. The resulting segmented
-fitted mean (horizontal blue lines in the left panel of Figure~\ref{fig:slovenia1}) shows that
+fitted mean (horizontal bold lines in the left panel of Figure~\ref{fig:slovenia1}) shows that
 before the break, Slovenia experienced very high inflation rates but was successful in reducing
@@
 
-Austria, Finland, and France adjusted to the requirements of the EMU during the
+The second group of countries, Austria, Finland, and France, adjusted to the requirements of the EMU during the
 first phase of convergence. In these countries, the result of the adjustment process
@@
 
-The next group consists of two countries that, after a very short participation in the
+The fifth group consists of two countries that, after a very short participation in the
 ERM, decided not to be part of it and hence also not to participate in the EMU:
@@
 traced back to economic crises: the currency crisis of the United Kingdom cumulating
-into the ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
+in ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
 1990s. With regard to the last segment for both countries, the Ljung-Box test suggests that 
@@
 Finally, the inflation rates for the whole euro area\footnote{This is the
-euro area 17~countries series as provided by the \citet{oecd},
+EU17 series as provided by the \citet{oecd},
 starting in \Sexpr{start(eu) - 1/12}. It is calculated as a weighted average of the HICPs
-of now seventeen euro countries.} are considered and
+of the then seventeen euro countries.} are considered and
 three different regimes are found: (1)~until 1999 with both a low mean and
@@
 gray_red <- rgb(c(0.2, 0.8), c(0.2, 0.0), c(0.2, 0.0), alpha = 0.3)
-gray_red <- rgb(c(0.2, 0.0), c(0.2, 0.0), c(0.2, 0.0), alpha = c(0.3, 0.5)) ## actually gray-darkgray now
+gray_red <- rgb(c(0.2, 0.0), c(0.2, 0.0), c(0.2, 0.0), alpha = c(0.2, 0.5)) ## actually gray-darkgray now
 plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
@@
 Our results were obtained using 
-\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} \citep{R} with the packages
-\pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]},
-\pkg{strucchange}~\Sexpr{packageDescription("strucchange")["Version"]} \citep{Zeileis+Leisch+Hornik:2002}, and
-\pkg{fxregime}~\Sexpr{packageDescription("fxregime")["Version"]} \citep{Zeileis+Shah+Patnaik:2010},
+\proglang{R}~3.0.0 %\Sexpr{paste(R.Version()[6:7], collapse = ".")}
+\citep{R} with the packages
+\pkg{glogis}~0.1-0, %\Sexpr{packageDescription("glogis")["Version"]},
+\pkg{strucchange}~1.4-7 %\Sexpr{packageDescription("strucchange")["Version"]}
+\citep{Zeileis+Leisch+Hornik:2002}, and
+\pkg{fxregime}~1.0-2 %\Sexpr{packageDescription("fxregime")["Version"]}
+\citep{Zeileis+Shah+Patnaik:2010},
 all of which are available under the General Public License (GPL) from the
